### 2023-02-20 paper/main.tex
checkin

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%%
%% IMPORTANT NOTICE:%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf,review, anonymous]{acmart}
\acmConference[ISSTA 2023]{ACM SIGSOFT International Symposium on Software Testing and Analysis}{17-21 July, 2023}{Seattle, USA}

\usepackage{code}
\usepackage{graphicx}
\usepackage{balance}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{listings}
\usepackage{booktabs}
\usepackage{subfig}
\usepackage{url}
\usepackage{comment}
\usepackage{xcolor}
\usepackage{xspace}

\definecolor{dkgreen}{rgb}{0,0.5,0}
\definecolor{dkred}{rgb}{0.5,0,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}

\lstdefinestyle{javastyle} {
language=Java,
basicstyle=\ttfamily\bfseries\footnotesize,
  morekeywords={virtualinvoke},
  keywordstyle=\color{blue},
  ndkeywordstyle=\color{red},
  commentstyle=\color{dkred},
  stringstyle=\color{dkgreen},
  numbers=left,
  breaklines=true,
  numberstyle=\ttfamily\footnotesize\color{gray},
  stepnumber=1,
  numbersep=10pt,
  backgroundcolor=\color{white},
  tabsize=4,
  showspaces=false,
  showstringspaces=false,
  xleftmargin=.23in
}
\lstset{style=javastyle}

%%% The following is specific to Onward! '21 and the paper
%%% 'Let a Thousand Flowers Bloom: On the Uses of Diversity in Software Testing'
%%% by Alex Groce.
%%%

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Program Mutation is Randomized Program Transformation: A
  Universal Approach to Mutant Generation}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
%\author{Alex Groce}
%\affiliation{\institution{Northern Arizona University}\country{United States}}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.

%% Table shortcuts
\newcommand{\mr}[2]{\multirow{#1}{*}{#2}}
\newcommand{\mc}[3]{\multicolumn{#1}{#2}{#3}}
\newcommand{\um}{\texttt{universalmutator}\xspace}
%% comments
\newcommand{\clg}[1]{\textcolor{blue}{#1}}
\newcommand{\adg}[1]{\textcolor{purple}{#1}}
\newcommand{\kj}[1]{\textcolor{olive}{#1}}

%% numbers
\newcommand{\averageprojvariance}{402}
\newcommand{\averagevariance}{604}
\newcommand{\outliertotalfiles}{26}
\newcommand{\outliertestissues}{12}
\newcommand{\outlierumissues}{7}
\newcommand{\outlierunclear}{7}
\newcommand{\allcorr}{0.7479}
\newcommand{\covcorr}{0.2066}
\newcommand{\allrsquared}{0.573}
\newcommand{\allr}{0.757}
\newcommand{\covrsquared}{0.021}
\newcommand{\covr}{0.145}

\begin{abstract}
While mutation testing has been a topic of academic interest for
decades, it is only recently that ``real-world'' developers, including
industry leaders such as Google and Meta, have adopted mutation
testing.  In this paper we propose a new approach to the development of mutation
testing tools, and in particular the core challenge of
\emph{generating mutants}.  Current practice tends towards two
limited approaches to mutation generation: mutants are either (1)
generated at the bytecode/IR level, and thus neither human readable
nor adaptable to source-level features of languages or projects, or
(2) generated at the source level by language-specific tools that are
hard to write and maintain, and in fact are often abandoned by both
developers and users.  We propose instead that source-level mutation
generation is a special case of \emph{program transformation} in
general, and that adopting this approach allows for a single tool that
can effectively generate source-level mutants for essentially
\emph{any} programming language. Furthermore, by using \emph{parser
  parser combinators} many of the seeming limitations of an
any-language approach can be overcome, without the need to parse
specific languages.
\end{abstract}

\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10010940.10010992.10010998.10011001</concept_id>
<concept_desc>Software and its engineering~Dynamic analysis</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10011074.10011099.10011102.10011103</concept_id>
<concept_desc>Software and its engineering~Software testing and debugging</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Dynamic analysis}
\ccsdesc[500]{Software and its engineering~Software testing and debugging}

\maketitle

\section{Introduction}

\section{Related Work}

There is a large body of work studying mutation testing and its relationship 
with defects, and an extensive literature on test adequacy measured by
code coverage that dates almost to the beginnings of software
engineering; however, we are
unaware of work analyzing the relationship
between mutation score and coverage in terms similar to ours.   While
there is foundational work on the oracle/test
distinction~\cite{StaatsOracle} at a theoretical level, most work assumes a basic framework of  trying to determine
correlation of mutation testing or coverage alone with fault
detection and/or coverage with mutation score~\cite{papadakis2018mutation,PapadakisStudy,ThierryStudy}.  In
such work, the oracle/test distinction is not considered in the
light of whether code is ``more executed'' or ``more checked'' in a
test effort.  Perhaps the
closest approach, which does not involve mutants at all, is the notion
of \emph{checked coverage}~\cite{ZellerCheckedCov}, which, however, still results in a
single score, as does pseudotesting~\cite{pseudotestingorig,
  pseudotestingstudy, descartes}.  Further, checked coverage is
inherently unable to identify buggy assertions that do not in practice
detect mutants, if there is any dynamic flow
from coverage to the faulty assertion.
%\textbf{Relevant Mutation Testing Studies}

%Petrovic et al. \cite{PetrovicTestingPractices} study the relationship between
%mutants and faults, while describing how developers
%have been using mutation testing at Google. They find that mutation testing has a 70\% coupling with high priority
%defects. This helps support our assertion that higher mutation scores are strongly correlated with higher quality oracles, with
%mutation testing being a viable method of catching defects before they occur. Combining the results of with our study regarding the
%oracle gap provides information for practitioners about both the value of mutation testing and when they should be running mutation testing on their code base.

\begin{comment}
Delgado-Pérez et al. \cite{DelgadoPerezCaseStudy} also performed mutation testing on 2,509 mutants over 15 functions. Their study 
primarily focused on analyzing equivalent mutants and the computational cost of finding these mutants.
They agree with the hypothesis that mutation testing is a better predictor of faults than a
traditional coverage based approach. This large scale study of mutation is similar to our Java study of a large number of
mutants, however the analysis and results differ in that they were analyzing equivalent mutants, while we are examining the
correlation between mutation score and coverage through the lens of the oracle gap. Additionally, our study is over a much 
larger sample size, with over 30 Java projects, that total have 463,417 lines of code, making it more appropriate for extrapolating
generalizable correlations.
\end{comment}

Just et al. \cite{JustMutationFault} found that 73\% of real-world faults can be associated with common mutation
operators. Beller et al. \cite{BellerFacebookMutation} examined how to feasibly
implement mutation testing at Facebook. These studies support the notion that
lagging mutation score indicates lagging fault \emph{detection}
capability even in the presence of coverage, as also suggested by work
by Sina et al. on whether automatically generated unit tests help find bugs~\cite{DoGenerated}.
Smith et al. \cite{SmithCoverageMutation} briefly discussed the
relationship between coverage and mutation score, but 
examined only two open source projects. Li et al. \cite{LiCoverageMutation} also examined when it is best to use various testing methods, including mutation testing. Rather than suggesting
that developers should \emph{either} use mutation score or coverage, we suggest that they should use both metrics, as they
complement each other.  Recent industrial uses of
mutants~\cite{PetrovicMutationGoogle,BellerFacebookMutation} do not
include gap analysis, but informally hint that large-scale users may
be amenable to the type of analysis we provide, given how they use
mutants to pinpoint test issues now.  Finally, it is interesting to consider that
large positive covered oracle gap can be alternatively thought of as
the presence of many \emph{stubborn} mutants in a file~\cite{papadakis2018mutant}.

\section{Conclusion and Implications}

\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}

\end{document}

### 2023-02-21 paper/main.tex
promises
ok

@@
 \documentclass[sigconf,review, anonymous]{acmart}
-\acmConference[ISSTA 2023]{ACM SIGSOFT International Symposium on Software Testing and Analysis}{17-21 July, 2023}{Seattle, USA}
+\acmConference[ICSE 2024]{46th International Conference on Software Engineering}{April 2024}{Lisbon, Portugal}
 
@@
 
-%%% The following is specific to Onward! '21 and the paper
-%%% 'Let a Thousand Flowers Bloom: On the Uses of Diversity in Software Testing'
-%%% by Alex Groce.
-%%%
-
-%%
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
-
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-%%\citestyle{acmauthoryear}
-
-%%
-%% end of the preamble, start of the body of the document source.
-\begin{document}
-
-%%
 %% The "title" command has an optional parameter,
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Program Mutation is Randomized Program Transformation: A
+\title{Program Mutation is Just Program Transformation: \\ A
   Universal Approach to Mutant Generation}
@@
 
+\begin{document}
+
 \begin{abstract}
@@
 any-language approach can be overcome, without the need to parse
-specific languages.
+specific languages.  We compare the universal program transformation
+approach to mutation to existing tools, and demonstrate the advantages
+of using parser-parser combinators to improve on a simple regular-expression
+based approach to generation.  The ease of building new tools is
+further demonstrated by showing that the implementation of a recently
+published mutation tool for Coq can be duplicated using only XX
+transformation rules, without parsing the Coq language.
 \end{abstract}
@@
 
-There is a large body of work studying mutation testing and its relationship 
-with defects, and an extensive literature on test adequacy measured by
-code coverage that dates almost to the beginnings of software
-engineering; however, we are
-unaware of work analyzing the relationship
-between mutation score and coverage in terms similar to ours.   While
-there is foundational work on the oracle/test
-distinction~\cite{StaatsOracle} at a theoretical level, most work assumes a basic framework of  trying to determine
-correlation of mutation testing or coverage alone with fault
-detection and/or coverage with mutation score~\cite{papadakis2018mutation,PapadakisStudy,ThierryStudy}.  In
-such work, the oracle/test distinction is not considered in the
-light of whether code is ``more executed'' or ``more checked'' in a
-test effort.  Perhaps the
-closest approach, which does not involve mutants at all, is the notion
-of \emph{checked coverage}~\cite{ZellerCheckedCov}, which, however, still results in a
-single score, as does pseudotesting~\cite{pseudotestingorig,
-  pseudotestingstudy, descartes}.  Further, checked coverage is
-inherently unable to identify buggy assertions that do not in practice
-detect mutants, if there is any dynamic flow
-from coverage to the faulty assertion.
-%\textbf{Relevant Mutation Testing Studies}
+Many approaches have been proposed to tackle the \emph{computational} cost of mutation, including weak-mutation, 
+meta-mutation, mutation-sampling, and predicting which mutants will be
+killed~\cite{offuttMutant1996,
+  untch1993mutation,KaufmanFAKAJ2022,zhang2016pmt}.  Approaches to reducing the cost of
+mutation analysis were categorized as \textit{do smarter}, \textit{do
+faster}, and \textit{do fewer} by Offutt et al.~\cite{offutt2001mutation}.
+The \textit{do smarter} approaches include space-time trade-offs, weak
+mutation analysis, and parallelization of mutation analysis. The \textit{do
+faster} approaches include mutant schema generation, code patching, and
+other methods to make mutants run faster. Finally, the
+\textit{do fewer} approaches try to reduce the number of mutants examined,
+and include selective mutation and mutant sampling.
 
-%Petrovic et al. \cite{PetrovicTestingPractices} study the relationship between
-%mutants and faults, while describing how developers
-%have been using mutation testing at Google. They find that mutation testing has a 70\% coupling with high priority
-%defects. This helps support our assertion that higher mutation scores are strongly correlated with higher quality oracles, with
-%mutation testing being a viable method of catching defects before they occur. Combining the results of with our study regarding the
-%oracle gap provides information for practitioners about both the value of mutation testing and when they should be running mutation testing on their code base.
+None of these approaches focus on the cost in \emph{human} time to
+develop and maintain mutation testing tools.  In fact, the complexity
+and sophistication of some of these approaches imposes a daunting
+barrier to those who would develop ``good'' mutation tools for a new
+language.  However, arguably the most powerful and generalizable
+methods for reducing the cost of mutation, such as random sampling~\cite{GopinathSampleSize,gopinath2017mutation} and
+predictive mutation testing~\cite{zhang2016pmt,kim2022predictive}, are independent of the generation of
+mutants and so applicable to any language.
 
-\begin{comment}
-Delgado-Pérez et al. \cite{DelgadoPerezCaseStudy} also performed mutation testing on 2,509 mutants over 15 functions. Their study 
-primarily focused on analyzing equivalent mutants and the computational cost of finding these mutants.
-They agree with the hypothesis that mutation testing is a better predictor of faults than a
-traditional coverage based approach. This large scale study of mutation is similar to our Java study of a large number of
-mutants, however the analysis and results differ in that they were analyzing equivalent mutants, while we are examining the
-correlation between mutation score and coverage through the lens of the oracle gap. Additionally, our study is over a much 
-larger sample size, with over 30 Java projects, that total have 463,417 lines of code, making it more appropriate for extrapolating
-generalizable correlations.
-\end{comment}
+Hariri et al. compared C mutation approaches at the source and
+compiler IR levels~\cite{CompareSrcBinary} and found that overall
+source level mutation was better, producing fewer mutants overall, but
+matching the IR approach in the important measures of surface and
+minimal mutants and overall mutation score.  Numerous studies compare
+Java mutation tools~\cite{MajorPIT,gopinath2017does}, including a
+recent article for a more general audience in Communications of the
+ACM~\cite{CommACMJavaTool} (perhaps showing the growing interest in
+practical mutation testing).  This paper showed that users
+considered active maintenance, support for a variety of testing
+frameworks, and support for recent Java versions as the most important
+features in a Java mutation tool.  The approach proposed in this paper
+by its nature tends to promise all three of these key factors without
+imposing onerous burdens on maintainers.
 
-Just et al. \cite{JustMutationFault} found that 73\% of real-world faults can be associated with common mutation
-operators. Beller et al. \cite{BellerFacebookMutation} examined how to feasibly
-implement mutation testing at Facebook. These studies support the notion that
-lagging mutation score indicates lagging fault \emph{detection}
-capability even in the presence of coverage, as also suggested by work
-by Sina et al. on whether automatically generated unit tests help find bugs~\cite{DoGenerated}.
-Smith et al. \cite{SmithCoverageMutation} briefly discussed the
-relationship between coverage and mutation score, but 
-examined only two open source projects. Li et al. \cite{LiCoverageMutation} also examined when it is best to use various testing methods, including mutation testing. Rather than suggesting
-that developers should \emph{either} use mutation score or coverage, we suggest that they should use both metrics, as they
-complement each other.  Recent industrial uses of
-mutants~\cite{PetrovicMutationGoogle,BellerFacebookMutation} do not
-include gap analysis, but informally hint that large-scale users may
-be amenable to the type of analysis we provide, given how they use
-mutants to pinpoint test issues now.  Finally, it is interesting to consider that
-large positive covered oracle gap can be alternatively thought of as
-the presence of many \emph{stubborn} mutants in a file~\cite{papadakis2018mutant}.
+Finally, the practical use of mutation testing at Google argues that
+support for a variety of languages is critical, and that the approach
+we take is more than sufficient.  Google has used its
+substantial resources to provide mutation generation for C++, Java,
+Python, Javascript, Go, Typescript, and Common
+Lisp~\cite{PetrovicMutationGoogle}, all of which we also support.  The
+operators Google uses for these
+languages (based on the original Mothra operators~\cite{offutt1996experimental}) are a subset of those provided by our implementation, with the caveat
+that statement block removal is only supported when using Comby;
+regular expression mutation is limited to single-line statement
+deletion.  For Lisp-like languages, identifying statements vs. value-returning
+function calls makes use of block deletion impractical, but it can be
+supported as well.  The majority of these operators are implemented in
+a few lines as universal rules for all languages; only specializing
+logical operators and implementing block/statement deletion even requires
+language identification.
 

### 2023-02-27 paper/main.tex
note we rock vs google
spacing
more text

@@
 
+\section{The Universal Source-Based Approach}
+
+\subsection{Incremental Mutation}
+
+One obvious concern with any-language source-based mutation is that every time the
+program changes, the cost of invalid mutations must be paid again.
+However, the locality of mutants and source changes in fact means this
+is seldom required.
+
+The {\tt git merge-file} utility takes a base file and two modified
+versions of that file, and produces an automatic merge, using the
+usual git merge resolution algorithm.    Because a mutant is a (very
+small) source change, this means that in practice updating a mutant to
+reflect even large changes to a source file is cost-free, and if the
+original mutant was valid, the new mutant will usually also be valid.
+New mutant generation is only required for new lines of code and
+modified lines of code.  To the extent development is incremental,
+therefore, mutation generation can also be incremental.
+
+\section{Using Parser Parser Combinators Instead of Regular Expressions}
+
+\section{Experimental Evaluatoni}
+
+\subsection{Universal Source-Based Mutation vs. Previous Approaches}
+
+\subsection{Regular Expressions vs. Parser Parser Combinators}
+
 \section{Related Work}
@@
 regular expression mutation is limited to single-line statement
-deletion.  For Lisp-like languages, identifying statements vs. value-returning
-function calls makes use of block deletion impractical, but it can be
-supported as well.  The majority of these operators are implemented in
+deletion.  For Lisp-like languages, the difficulty of identifying statements vs. value-returning
+function calls makes use of block deletion somewhat impractical, but
+still easy to implement.  The majority of the Google Mothra operators
+can be implemented in
 a few lines as universal rules for all languages; only specializing
 logical operators and implementing block/statement deletion even requires
-language identification.
+language identification.  Unlike Google's mutation infrastructure, our tool is open source, easily extensible, and not tied to any particular development environment.
 
-\section{Conclusion and Implications}
+\section{Conclusions and Future Work}
 

### 2023-02-28 paper/main.tex
describe the basic approach

@@
 
+The key insight of this paper is that \emph{most mutation operators
+  proposed in the literature do not in fact require parsing of source
+  code.}  Consider, for example, one of the most commonly used
+mutation operators, replacement of arithmetic operations.  We do not
+need to parse a program containing the string {\tt ``x + y''} in order to
+replace that string with {\tt ``x - y''}.  Instead, this change is
+guaranteed to be effected if we simply search the program, represented
+as a string, for all occurrences of  {\tt ``+''}, and produce for each
+such occurrence, a mutant where tha character is replaced by {\tt
+  ``-''}.  We can do the same for each pair of arithmetic operations,
+and thus trivially implement a large fraction of the classic Mothra~\cite{offutt1996experimental}
+mutation operators.
+
+The obvious objection to this simplistic approach is that unless we
+parse the code to identify arithmetic expressions, some of these
+replacements will be invalid.  E.g., in a C++ program we will produce
+mutants like {\tt x-+} replacing {\tt x++}.  However, most of these
+instances can be avoided by slightly more judicious choice of search
+target: instances of {\tt ``+''} where the preceeding or following
+character is not another {\tt ``+''}, for instance.
+
+Of course, in some cases this will still leave us with invalid
+mutants.  However, rejecting such mutants is easy: the truly invalid
+mutants will fail to compile.  Of course, we must pay the price of
+checking each mutant to make sure that it does compile, at generation
+time.
+
+However, consider what has been gained by paying this cost.  We have
+replaced a complex implementation, that requires parsing source code,
+with an implementation that only requires string search-and-replace.
+One can imagine an implementation of the above approach to arithmetic
+operation replacement being implemented in as little as 10-20 lines of
+Python code, including a call to a compiler to check for invalid
+mutants.  Moreover, defining new mutation operators is not a matter of
+understanding a representation of a parsed program, but is essentially
+a matter of providing two strings, where the second is a proposed
+replacement for the first.  Instead of code for arithmetic operator
+replacement, we can imagine the above hypothetical Python program
+reading in a file of thus expressed mutation operators, and
+implementing an unbounded number of operators in a minimalist fashion.
+
+Of course, simple string replacement is not general enough to handle
+many interesting mutations.  Consider the classic case of statement
+deletion, one of the most widely used and important mutation
+operators.  What string can we replace with another string in order to
+delete a statement in a C program?  We would need an ``operator'' for
+each possible C statement.  However, if we allow the use of
+\emph{regular expression match and replace pairs} in place of simple
+strings, we can easily express this operator, with some precision:
+
+\begin{verbatim}
+(^\s*)(\S+[^{}]+.*)\n ==> \1/*\2*/\n
+\end{verbatim}
+
+In Python regular expression syntax, this expresses the replacement of
+strings with a possibly-empty amount of white space followed by a
+character pattern indicating a line of source code, with the same
+source code, preserving indentation, turned into a comment.
+
+\subsection{Hierarchies of Operator Applicability}
+
 \subsection{Incremental Mutation}

### 2023-03-02 paper/main.tex
more text

@@
 deletion, one of the most widely used and important mutation
-operators.  What string can we replace with another string in order to
+operators~\cite{deng2013empirical}.  What string can we replace with another string in order to
 delete a statement in a C program?  We would need an ``operator'' for
@@
 source code, preserving indentation, turned into a comment.
+
+Using regular expressions to express mutants has many benefits.  While
+much more expressive than simple string replacement, regular
+expressions are familliar to most developers.  In principle,
+development of complex regular expressions for operators can even be
+performed by simply providing examples~\cite{bartoli2014automatic}  or
+describing the operator in natural
+language~\cite{zhong2018generating}.  Implementing a mutation
+generation tool for a language by providing the equivalent of our
+Python program above with a list of regular expression matches and
+replacements is clearly much easier than writing a mutation tool from
+scratch, and much easier to maintain and extend.
 
@@
 \section{Using Parser Parser Combinators Instead of Regular Expressions}
+
+\section{Implementation}
 

### 2023-03-20 paper/main.tex
remove coq
heirarchy

@@
 of using parser-parser combinators to improve on a simple regular-expression
-based approach to generation.  The ease of building new tools is
-further demonstrated by showing that the implementation of a recently
-published mutation tool for Coq can be duplicated using only XX
-transformation rules, without parsing the Coq language.
+based approach to generation.  
 \end{abstract}
@@
 
+Implementing mutation generation by providing a set of regular-expression
+based textual transformations for each target language, and using a
+single engine to apply rules to source files is certainly less
+difficult in terms of development and maintenance effort than building
+a tool for each language around a parser for that language.  However,
+even this approach ignores one of the primary benefits of a universal,
+text-based approach.  Consider the case of replacing instances of {\tt
+  ``+''}  with {\tt ``-''}.  This mutation is not tied to any
+particular language, but applies to \emph{almost all programming
+  languages in use.}  Many of the most widely used mutation operators
+in the literature are of this universal nature.  Others are not quite
+so widely applicable, but still apply to a large number of specific
+languages.  E.g., many programming languages share C's logical
+operators, though important exception such as Python and Lisp-family
+languages do not.
+
+Implementing a tool for a language $L$ therefore, generally need not
+require writing rules for all mutation operators to be applied to $L$
+programs, but instead can proceed by process of 1) identifying $L$'s
+place in a \emph{hierarchy} of language \emph{famillies}, and then 2)
+identifying additional operators needed for $L$ itself.  The first of
+these tasks is often trivial, as in practice there are only a few
+basic syntactic forms for languages, at the level needed to describe
+transformation rules for operators.  In fact, ignoring the second step
+will often provide ``good-enough'' mutation testing, in that, for
+example, most of the Mothra rules and statement deletion can be
+entirely handled without descending to the specific-language level at all.
+
+As an
+example, consider mutation of Java code.  Many Java mutants can be
+generated by the kind of ``universal'' rules (e.g., arithmetic
+operator replacements) considered above.  Other mutation operators
+suitable for Java source are provided by considering the set of
+``C-like'' languages that use the logical operators and control
+constructs common to e.g.,  Java, C, and C++ ({{\tt if}, {\tt while},
+  etc.).  Statement deletion at the line level can also be implemented
+  by using the common comment notation for such languages.
+  Implementing Java mutation, given universal and C-like rules, may
+  require no more than a handful of additional rules.
+
+  Figure~\ref{fig:rules} shows examples of universal, C-like, and
+  language specific rules for a few languages, taken directly from our
+  implementation.
+
 \subsection{Incremental Mutation}

### 2023-03-21 paper/main.tex
Alex first stab at intro

@@
 
+Mutation testing, though introduced in the late
+1970s~\cite{demillo1978hints,mathur2012foundations,demillo1978hints},
+has largely been a topic of academic interest rather than widespread
+practice until quite recently.  However, with the adoption of mutation
+testing by bellwether software industry companies including
+Google~\cite{GoogleMut}, Meta~\cite{BellerFacebookMutation}, and
+Amazon~\cite{AmazonMut}, interest in using mutation testing for
+real-world projects has grown widely.
+
+Interest in mutation testing has become widespread enough in the open
+source community that a popular (more than 300 stars on GitHub)
+repository, ``Awesome Mutation Testing,'' lists more than 40 tools,
+covering almost twenty languages:
+\url{https://github.com/theofidry/awesome-mutation-testing}.  Some of
+these tools are the products of academic research, but many are
+essentially the hobby projects of developers interested in enabling
+mutation testing for their favorite language.
+
+However, the same website demonstrates the existence of a sad history
+of mutation testing tools: mutation tools have a tendency to be
+developed for a while and then abandoned.   The page listing abandoned
+mutation testing tools
+(\url{https://github.com/theofidry/awesome-mutation-testing/blob/master/abandoned.md})
+lists almost half as many tools as are under active maintenance, and
+these tools cover approximately half as many languages.  Each of these
+tools required considerable development time and effort, and was
+popular enough at some point to be worth noting in the ``Awesome
+Mutation Testing'' resource.  Many of the abandoned tools have GitHub
+repositories (some archived) with more than 20 stars, and so may well
+have had users who at some point faced the reality of no future bug
+fixes or changes to address changes to the programming language being
+mutated\footnote{While this is a minor concern for languages such as C
+  that are relatively stable, it means that a mutation tool may fail
+  to parse C++ 2020, and is particularly problematic for somewhat
+  fast-moving languages such as Rust.}.  Of course, many open source
+projects are abandoned, and academic software has a long history of
+being available and supported only so long as a graduate student is
+working on a thesis or a particular grant is being funded.  Still,
+there appears to be a significant risk in relying on any mutation tools other
+than the most widely used and well-supported, such as PIT.
+
+...
+
+In this paper, we argue that developing a large number of mutation
+testing tools, each requiring its own development community, workflow
+for mutation, and effort to keep up with language changes, is
+unnecessary.  Mutation testing, and in particular, the core problem,
+\emph{mutation generation} is simply a particular instance of the
+problem of \emph{program transformations}.  Moreover,
+mutation testing can be treated as an almost entirely \emph{syntactic}
+transformation problem; in fact, in their classic introduction to
+software testing, Ammann and Offutt refer to mutation testing as
+``syntax-based testing'' ~\cite{ammann2016introduction}.  While, e.g., refactoring may need to
+``understand'' a program to some extent, mutation can often operate
+with essentially no context beyond a single line of code.
+
+We further propose that in fact, for the most part, mutation
+generation need not even ``know'' the syntax of a target language, in
+terms of a complete grammar.  Instead, mutation testing can operate to
+a large extent in a \emph{universal} manner, where programs are
+transformed at the level of \emph{patterns of characters}.
+
+The contributions of this paper are:
+
+\begin{enumerate}
+  \item The description of a novel approach to mutation generation
+    based on generalized description of multi-lingual program transformations
+    that allows a single tool to handle mutant generation for
+    essentially all programming languages.  This approach also makes
+    mutation testing easy to extend for new languages and even allows
+    developers to write project-specific rules easily without
+    having to modify a mutation tool's implementation.
+
+          \item A proposal to use parser-parser combinators to improve the
+      efficiency and expressive power of that approach, and an
+      evaluation of the gains thus achieved.
+
+      \item A comparison of an implementation of mutation generation
+       based on this approach with existing tools for four important
+       programming languages.  The single-language tools range from
+       approximately 3,800 LOC to nearly 60,000 LOC, and each handle
+       one programming language.  Our tool, which provides comparable
+       core mutation testing functionality, supports over a dozen
+       languages using only about 2200 LOC of Python and less than 500
+       lines of rules defining mutation operators.
+
+\end{enumerate}
+
 \section{The Universal Source-Based Approach}
@@
 
-\section{Experimental Evaluatoni}
+\section{Experimental Evaluation}
+
+\begin{itemize}
+
+  \item{RQ1:}  How does the use of parser-parser combinators modify 
+    the efficiency (in terms of invalid mutants) and effectiveness (in
+    terms of mutation score and equivalent mutants) of the universal
+    approach to mutation generation?
+
+\item{RQ2:}  How does the univeral approach, using regular expressions
+  or parser-combinator combinator rules, compare to existing
+  single-language mutation tools, in terms of number of generated
+  mutants, mutation scores, equivalent mutants, and other evaluation measures used in the literature?
+
+  \end{itemize}
+
+\subsection{Regular Expressions vs. Parser Parser Combinators}
 
 \subsection{Universal Source-Based Mutation vs. Previous Approaches}
-
-\subsection{Regular Expressions vs. Parser Parser Combinators}
 

### 2023-03-21 paper/main.tex
more intro

@@
 for mutation, and effort to keep up with language changes, is
-unnecessary.  Mutation testing, and in particular, the core problem,
-\emph{mutation generation} is simply a particular instance of the
-problem of \emph{program transformations}.  Moreover,
+unnecessary.  The core problem of mutation testing,
+\emph{mutant generation}, is simply a particular instance of the
+problem of \emph{program transformation}.  Program transformation
+considers the problem of taking a program, $P$, and producing a
+program, $P'$ that is the result of applying some function $f$ to $P$
+(i.e., $P' = f(P)$).  In mutant generation, we
+consider a family of functions $f$ where each $f_{o,loc}$ is an
+instantiation of a \emph{mutation operator} $o$  (as classically, if
+often informally, defined in the literature) at a particular program
+location $loc$ in $P$.  The inclusion of $loc$ in the specification of
+$f$ indicates that the problem is a specialization of general
+program transformation in that all classic mutation operators we are
+aware of are \emph{local}: they only affect a small portion of a
+program, genearlly a contiguous string of characters or tokens.
+
+Moreover, as this suggests,
 mutation testing can be treated as an almost entirely \emph{syntactic}
@@
 
-We further propose that in fact, for the most part, mutation
+We further propose that in fact, for the most part, mutant
 generation need not even ``know'' the syntax of a target language, in
@@
 a large extent in a \emph{universal} manner, where programs are
-transformed at the level of \emph{patterns of characters}.
-
-The contributions of this paper are:
+transformed at the level of \emph{patterns of characters}.  That is,
+each $o$ and $loc$ can be defined without reference to, e.g., the
+context-free grammar of a programming language, but only with respect
+to a position in $P$ treated as a string, and a transformation on a
+string defined in terms of, e.g., regular expressions with capture and
+replacement.  Treating mutant generation as a problem of (textual, local)
+program transformation provides two major benefits:
 
 \begin{enumerate}
-  \item The description of a novel approach to mutation generation
+  \item A single tool can provide mutant generation for almost any
+    programming language, without needing to parse that language, and
+    rules for one language can be re-used for another language.  This
+    reduces the development and maintenance effort for mutation
+    testing tools by orders of magnitude.  Moreover, it removes the
+    need for many   maintenance activities related to parsing code;
+    such changes need to be made only when a change to the language
+    would introduce new mutation operators or modify existing ones.
+    The local nature of operators tends to make the second case in
+    particular very rare.
+    \item The existing literature on multi-lingual or language
+      agnostic program transformation can be applied to the problem of
+      mutant generation, making it easy to express mutation operators
+      that operate across many languages, and efficient to generate
+      mutants.
+      \end{enumerate}
+
+The contributions of this paper are therefore:
+
+\begin{enumerate}
+  \item The description of a novel approach to mutant generation
     based on generalized description of multi-lingual program transformations
@@
 
-      \item A comparison of an implementation of mutation generation
+      \item A comparison of an implementation of mutant generation
        based on this approach with existing tools for four important
        programming languages.  The single-language tools range from
-       approximately 3,800 LOC to nearly 60,000 LOC, and each handle
+       approximately 3,800 LOC to nearly 60,000 LOC, and, we emphasize, each handle
        one programming language.  Our tool, which provides comparable
        core mutation testing functionality, supports over a dozen
-       languages using only about 2200 LOC of Python and less than 500
-       lines of rules defining mutation operators.
+       languages using only about 2,200 LOC of Python and less than 500
+       lines for rules defining mutation operators over a dozen
+       specific languages (and which can effectively mutate an
+       essentially unlimited number of other languages).
 
@@
 
-Implementing mutation generation by providing a set of regular-expression
+Implementing mutant generation by providing a set of regular-expression
 based textual transformations for each target language, and using a
@@
 modified lines of code.  To the extent development is incremental,
-therefore, mutation generation can also be incremental.
+therefore, mutant generation can also be incremental.
 
@@
     terms of mutation score and equivalent mutants) of the universal
-    approach to mutation generation?
+    approach to mutant generation?
 
@@
 we take is more than sufficient.  Google has used its
-substantial resources to provide mutation generation for C++, Java,
+substantial resources to provide mutant generation for C++, Java,
 Python, Javascript, Go, Typescript, and Common

### 2023-03-22 paper/main.tex
tce note
redacted cite for um paper

@@
 
+Our approach also makes introducing advanced techniques such as
+Trivial Compiler Equivalence (TCE) into a mutation workflow easy.
+While we support TCE for some languages inside the tool, a user can
+always write a simple script to check...
+
 \section{Experimental Evaluation}
@@
 logical operators and implementing block/statement deletion even requires
-language identification.  Unlike Google's mutation infrastructure, our tool is open source, easily extensible, and not tied to any particular development environment.
+language identification.  Unlike Google's mutation infrastructure, our
+tool is open source, easily extensible, and not tied to any particular
+development environment.
+
+The approach reported in this paper was briefly presented in part in a
+tool paper (citation redacted for blinding purposes) that introduced
+the idea of using regular expressions to generate mutants for multiple
+languages, and provided an evaluation against PIT for a single toy
+Java program.  The present paper provides the full idea and context,
+provides empirical data for universal mutant generation on real world
+programs over four languages,
+and introduces the idea of using parser-parser combinators to improve
+on regular expression-based generation.
 

### 2023-03-23 paper/main.tex

@@
 \usepackage{xspace}
+
+\usepackage{tabularx} % for tables with adjustable column width
+\usepackage{makecell} % for formatting table cell contents
 
@@
 
+% ---------------- JAVA table --------------------------
+\begin{table}[hbtp]
+\centering
+\caption{JAVA (Regex vs. Comby)}
+\label{tab:table_java1}
+\resizebox{\columnwidth}{!}{%
+
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+ &  & \textbf{Valid} & \textbf{Invalid} & \textbf{Redundant} & \textbf{Valid \%}  \\ \hline
+\multirow{2}{*}{pebble} & Regex & 86 & 928 & 0 & 8.48\%  \\\cline{2-6}
+    & Comby & 85 & 667 & 0 & 11.30\% \\ \hline
+\multirow{2}{*}{spring-hateoas} & Regex & 248 & 2231 & 0 & 10.00\% \\\cline{2-6}
+    & Comby & 176 & 1212 & 0 & 12.68\% \\ \hline
+\multirow{2}{*}{javacc} & Regex & 542 & 1346 & 0 & 28.70\% \\\cline{2-6}
+    & Comby & 503 & 782 & 0 & 39.14\% \\ \hline
+\multirow{2}{*}{coffee-gb} & Regex & 224 & 1703 & 0 & 11.63\% \\\cline{2-6}
+    & Comby & 174 & 624 & 0 & 21.80\% \\ \hline
+\multirow{2}{*}{egads} & Regex & 624 & 1830 & 0 & 25.43\% \\\cline{2-6}
+    & Comby & 622 & 566 & 0 & 52.36\% \\ \hline
+\multirow{2}{*}{apk-parser} & Regex & 351 & 1232 & 0 & 22.17\% \\\cline{2-6}
+    & Comby & 338 & 866 & 0 & 28.07\% \\ \hline
+
+\end{tabular}
+
+}
+\end{table}
+
+% ---------------- Python table --------------------------
+
+\begin{table}[hbtp]
+\centering
+\caption{Python (Regex vs. Comby)}
+\label{tab:table_python1}
+\resizebox{\columnwidth}{!}{%
+
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+ &  & \textbf{Valid} & \textbf{Invalid} & \textbf{Redundant} & \textbf{Valid \%}  \\ \hline
+\multirow{2}{*}{keract} & Comby & 1655 & 1600 &	729 & 41.53\%  \\\cline{2-6}
+    & Comby & 1471 & 1063 & 533 & 47.96\% \\ \hline
+\multirow{2}{*}{dtw} & Comby & 973 & 566 & 567 & 46.20\% \\\cline{2-6}
+    & Comby & 866 & 330 & 639 &	47.19\% \\ \hline
+\multirow{2}{*}{notion-sdk-py} & Comby & 196 & 760 & 82 & 18.89\% \\\cline{2-6}
+    & Comby & 196 & 701 & 122 & 19.23\% \\ \hline
+
+\multirow{2}{*}{atlassian-python-api} & Comby & 302	& 1002 & 207 &	19.99\% \\\cline{2-6}
+    & Comby & 333 & 635 & 351 & 25.25\% \\ \hline
+\multirow{2}{*}{ESD} & Comby & 1859	& 2767 & 1642 & 29.46\% \\\cline{2-6}
+    & Comby & 1712 & 1548 &	614 & 44.18\% \\ \hline
+\multirow{2}{*}{fiber} & Comby & 493 & 909 & 163 & 31.08\% \\\cline{2-6}
+    & Comby & 458 & 767	& 241 & 31.24\% \\ \hline
+
+\end{tabular}
+
+}
+\end{table}
+
 \subsection{Universal Source-Based Mutation vs. Previous Approaches}
+
+% ---------------- JAVA table --------------------------
+
+\begin{table}[htbp]
+\centering
+\caption{JAVA (UniversalMutator vs. PIT vs. LittleDarwin)}
+\label{tab:table_java2}
+\resizebox{\columnwidth}{!}{
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+ &  & \textbf{Killed} & \textbf{Not Killed} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{Equivalent \%} \\ \hline
+
+ \multirow{4}{*}{\small pebble} & Regex & 70 & 16 & 0.81 & - \\\cline{2-6}
+    & Comby & 68 & 17 & 0.80 & - \\\cline{2-6}
+    & PIT & 19 & 1 & 0.95 & - \\ \cline{2-6}
+    & LittleDarwin & 8 & 0 & 1.00 & - \\ \hline
+\multirow{3}{*}{\small spring-hateoas} & Regex & 82 & 166 & 0.33 & - \\\cline{2-6}
+    & Comby & 65 & 111 & 0.37 & - \\ \cline{2-6}
+    & PIT & 82 & 32 & 0.28 & - \\ \cline{2-6}
+    & LittleDarwin & 15 & 23 & 0.39 & - \\ \hline
+\multirow{3}{*}{\small javacc} & Regex & 123 & 419 & 0.23 & - \\\cline{2-6}
+    & Comby & 103 & 400 & 0.21 & - \\ \cline{2-6}
+    & PIT & 30 & 113 & 0.21 & -\\ \cline{2-6}
+    & LittleDarwin & 64 & 18 & 0.78 & - \\ \hline
+\multirow{3}{*}{\small coffee-gb} & Regex & 48 & 176 & 0.21 & - \\\cline{2-6}
+    & Comby & 40 & 134 & 0.23 & - \\ \cline{2-6}
+    & PIT & 32 & 4 & 0.89 & - \\ \cline{2-6}
+    & LittleDarwin & 11 & 18 & 0.38 & -\\ \hline
+\multirow{3}{*}{\small egads} & Regex & 442 & 182 & 0.71 & - \\\cline{2-6}
+    & Comby & 422 & 200 & 0.68 & - \\ \cline{2-6}
+    & PIT & 87 & 20 & 0.81 & - \\ \cline{2-6}
+    & LittleDarwin & 74 & 11 & 0.87 & - \\ \hline
+\multirow{3}{*}{\small apk-parser} & Regex & 26 & 325 & 0.07 & \\\cline{2-6}
+    & Comby & 26 & 312 & 0.08 & -\\ \cline{2-6}
+    & PIT & 14 & 75  & 0.16 & -\\ \cline{2-6}
+    & LittleDarwin & 3 & 30 & 0.09 & -\\ \hline
+
+\multirow{12}{*}{Overall} & \multirow{3}{*}{Regex} & \multicolumn{2}{c|}{Mean} & 0.39  & -\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.28 & - \\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.27 & - \\\cline{2-6}
+
+ & \multirow{3}{*}{Comby} & \multicolumn{2}{c|}{Mean} & 0.40 & -\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.30 & - \\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.26 & - \\\cline{2-6}
+
+ & \multirow{3}{*}{PIT} & \multicolumn{2}{c|}{Mean} & 0.55 & -\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.55 & - \\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.34 & - \\\cline{2-6}
+
+ & \multirow{3}{*}{LittleDarwin} & \multicolumn{2}{c|}{Mean} & 0.59 & -\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.59 & - \\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.32 & - \\ \hline
+
+\end{tabular}
+}
+\end{table}
+
+% ----------------  Python table -------------------------------
+
+\begin{table}[htbp]
+\centering
+\caption{Python (UniversalMutator vs. Mutmut vs. CosmicRay)}
+\label{tab:table_python2}
+\resizebox{\columnwidth}{!}{
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+ &  & \textbf{Killed} & \textbf{Not Killed} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{Equivalent \%} \\ \hline
+
+ \multirow{4}{*}{\small keract} & Regex & 463 & 1192 & 0.28 & - \\\cline{2-6}
+    & Comby & 383 & 1088 & 0.26 & - \\\cline{2-6}
+    & Mutmut & 349 & 133 & 0.72 & - \\ \cline{2-6}
+    & CosmicRay & 250 & 671 & 0.27 & - \\ \hline
+\multirow{3}{*}{\small dtw} & Regex & 506 & 467 & 0.52 & - \\\cline{2-6}
+    & Comby & 475 & 391 & 0.55 & - \\ \cline{2-6}
+    & Mutmut & 118 & 152 & 0.44 & - \\ \cline{2-6}
+    & CosmicRay & 328 & 385 & 0.46 & - \\ \hline
+\multirow{3}{*}{\small notion-sdk-py} & Regex & 112 & 84 & 0.57 & - \\\cline{2-6}
+    & Comby & 118 & 78 & 0.60 & - \\ \cline{2-6}
+    & Mutmut & 41 & 22 & 0.65 & -\\ \cline{2-6}
+    & CosmicRay & 23 & 29 & 0.44 & - \\ \hline
+
+\multirow{3}{*}{\makecell{\small atlassian-python \\ api}} & Regex & 193 & 109 & 0.64 & - \\\cline{2-6}
+    & Comby & 207 & 126 & 0.62 & - \\ \cline{2-6}
+    & Mutmut & 123 & 51 & 0.71 & - \\ \cline{2-6}
+    & CosmicRay & 78 & 28 & 0.74 & - \\ \hline
+\multirow{3}{*}{\small ESD} & Regex & 132 & 1727 & 0.07 & - \\\cline{2-6}
+    & Comby & 121 & 1591 & 0.07 & - \\ \cline{2-6}
+    & Mutmut & (57+0) & 663 & 0.08 & - \\ \cline{2-6}
+    & CosmicRay & 63 & 950 & 0.06 & - \\ \hline
+\multirow{3}{*}{\small fiber} & Regex & 307 & 186 & 0.62 & \\\cline{2-6}
+    & Comby & 291 & 167 & 0.64 & -\\ \cline{2-6}
+    & Mutmut & 127 & 95 & 0.57 & -\\ \cline{2-6}
+    & CosmicRay & 259 & 61 & 0.81 & -\\ \hline
+
+\multirow{12}{*}{Overall} & \multirow{3}{*}{Regex} & \multicolumn{2}{c|}{Mean} & 0.45  & -\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.55 & - \\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.21 & - \\\cline{2-6}
+
+ & \multirow{3}{*}{Comby} & \multicolumn{2}{c|}{Mean} & 0.46 & -\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.58 & - \\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.22 & - \\\cline{2-6}
+
+ & \multirow{3}{*}{Mutmut} & \multicolumn{2}{c|}{Mean} & 0.53 & -\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.61 & - \\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.22 & - \\\cline{2-6}
+
+ & \multirow{3}{*}{CosmicRay} & \multicolumn{2}{c|}{Mean} & 0.46 & -\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.45 & - \\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.26 & - \\ \hline
+
+\end{tabular}
+}
+\end{table}
 

### 2023-03-25 paper/main.tex
Remove redundant column
Lowercase JAVA
add mean, median, std dev for valid/invalid tables
add tables for rust

@@
 
-% ---------------- JAVA table --------------------------
+% ---------------- JAVA table UM Only--------------------------
 \begin{table}[hbtp]
 \centering
-\caption{JAVA (Regex vs. Comby)}
+\caption{Java (Regex vs. Comby)}
 \label{tab:table_java1}
@@
 \hline
- &  & \textbf{Valid} & \textbf{Invalid} & \textbf{Redundant} & \textbf{Valid \%}  \\ \hline
-\multirow{2}{*}{pebble} & Regex & 86 & 928 & 0 & 8.48\%  \\\cline{2-6}
-    & Comby & 85 & 667 & 0 & 11.30\% \\ \hline
-\multirow{2}{*}{spring-hateoas} & Regex & 248 & 2231 & 0 & 10.00\% \\\cline{2-6}
-    & Comby & 176 & 1212 & 0 & 12.68\% \\ \hline
-\multirow{2}{*}{javacc} & Regex & 542 & 1346 & 0 & 28.70\% \\\cline{2-6}
-    & Comby & 503 & 782 & 0 & 39.14\% \\ \hline
-\multirow{2}{*}{coffee-gb} & Regex & 224 & 1703 & 0 & 11.63\% \\\cline{2-6}
-    & Comby & 174 & 624 & 0 & 21.80\% \\ \hline
-\multirow{2}{*}{egads} & Regex & 624 & 1830 & 0 & 25.43\% \\\cline{2-6}
-    & Comby & 622 & 566 & 0 & 52.36\% \\ \hline
-\multirow{2}{*}{apk-parser} & Regex & 351 & 1232 & 0 & 22.17\% \\\cline{2-6}
-    & Comby & 338 & 866 & 0 & 28.07\% \\ \hline
+ & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%}  \\ \hline
+\multirow{2}{*}{pebble} & \multicolumn{2}{c|}{Regex} & 86 & 928 & 8.48\%  \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 85 & 667 & 11.30\% \\ \hline
+\multirow{2}{*}{spring-hateoas} & \multicolumn{2}{c|}{Regex} & 248 & 2231 & 10.00\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 172 & 1212 & 12.43\% \\ \hline
+\multirow{2}{*}{javacc} & \multicolumn{2}{c|}{Regex} & 542 & 1346 & 28.70\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 494 & 782 & 38.71\% \\ \hline
+
+\multirow{2}{*}{coffee-gb} & \multicolumn{2}{c|}{Regex} & 224 & 1703 & 11.63\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 174 & 624 & 21.80\% \\ \hline
+\multirow{2}{*}{egads} & \multicolumn{2}{c|}{Regex} & 624 & 1830 & 25.43\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 597 & 566 & 51.33\% \\ \hline
+\multirow{2}{*}{apk-parser} & \multicolumn{2}{c|}{Regex} & 351 & 1232 & 22.17\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 333 & 866 & 27.77\% \\ \hline
+
+\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 345.83 & 1545.00  & 17.74\%\\\cline{3-6}
+    &   &  Median & 299.50 & 1524.50 & 16.90\%  \\\cline{3-6}
+    &   &  Std. Dev & 203.86 & 467.98 & 8.73\%  \\\cline{2-6}
+
+ & \multirow{3}{*}{Comby} & Mean & 309.17 & 786.17 & 27.22\% \\\cline{3-6}
+    &   &  Median & 253.50 & 724.50 & 24.79\%  \\\cline{3-6}
+    &   &  Std. Dev & 202.44 & 235.24  & 15.59\%  \\\hline
 
@@
 
-% ---------------- Python table --------------------------
+% ---------------- Python table UM Only --------------------------
 
@@
 
+\begin{tabular}{|c|c|c|c|c|c|c|}
+\hline
+ & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Redundant} & \textbf{Valid \%}  \\ \hline
+\multirow{2}{*}{keract} & \multicolumn{2}{c|}{Regex} & 1655 & 1600 &	729 & 41.53\%  \\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 1471 & 1063 & 533 & 47.96\% \\ \hline
+\multirow{2}{*}{dtw} & \multicolumn{2}{c|}{Regex} & 973 & 566 & 567 & 46.20\% \\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 866 & 330 & 639 &	47.19\% \\ \hline
+\multirow{2}{*}{notion-sdk-py} & \multicolumn{2}{c|}{Regex} & 196 & 760 & 82 & 18.89\% \\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 196 & 701 & 122 & 19.23\% \\ \hline
+
+\multirow{2}{*}{{\makecell{atlassian-python \\ api}}} & \multicolumn{2}{c|}{Regex} & 302	& 1002 & 207 &	19.99\% \\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 333 & 635 & 351 & 25.25\% \\ \hline
+\multirow{2}{*}{ESD} & \multicolumn{2}{c|}{Regex} & 1859	& 2767 & 1642 & 29.46\% \\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 1712 & 1548 &	614 & 44.18\% \\ \hline
+\multirow{2}{*}{fiber} & \multicolumn{2}{c|}{Regex} & 493 & 909 & 163 & 31.08\% \\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 458 & 767	& 241 & 31.24\% \\ \hline
+
+\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 913.00 & 1267.33 & 565.00 & 31.19\%   \\\cline{3-7}
+    &   &  Median & 733.00 & 955.50 & 387.00 & 30.27\%  \\\cline{3-7}
+    &   &  Std. Dev & 708.93 & 813.35 & 584.79 & 11.06\%  \\\cline{2-7}
+
+ & \multirow{3}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% \\\cline{3-7}
+    &   &  Median & 662.00 & 734.00 & 442.00 & 37.71\%  \\\cline{3-7}
+    &   &  Std. Dev & 628.79 & 418.99 & 211.62 & 12.28\%  \\\hline
+
+\end{tabular}
+
+}
+\end{table}
+
+% ---------------- Rust table UM Only --------------------------
+\begin{table}[hbtp]
+\centering
+\caption{Rust (Regex vs. Comby)}
+\label{tab:table_rust1}
+\resizebox{\columnwidth}{!}{%
+
 \begin{tabular}{|c|c|c|c|c|c|}
 \hline
- &  & \textbf{Valid} & \textbf{Invalid} & \textbf{Redundant} & \textbf{Valid \%}  \\ \hline
-\multirow{2}{*}{keract} & Comby & 1655 & 1600 &	729 & 41.53\%  \\\cline{2-6}
-    & Comby & 1471 & 1063 & 533 & 47.96\% \\ \hline
-\multirow{2}{*}{dtw} & Comby & 973 & 566 & 567 & 46.20\% \\\cline{2-6}
-    & Comby & 866 & 330 & 639 &	47.19\% \\ \hline
-\multirow{2}{*}{notion-sdk-py} & Comby & 196 & 760 & 82 & 18.89\% \\\cline{2-6}
-    & Comby & 196 & 701 & 122 & 19.23\% \\ \hline
+ & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%}  \\ \hline
+\multirow{2}{*}{cargo release} & \multicolumn{2}{c|}{Regex} & 851 &	1163 & 42.25\%  \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 1690 & 711 & 70.39\% \\ \hline
+\multirow{2}{*}{passerine} & \multicolumn{2}{c|}{Regex} & 213 & 2081 & 9.29\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 173 & 1518 & 10.23\% \\ \hline
+\multirow{2}{*}{typos} & \multicolumn{2}{c|}{Regex} & 386 & 1534 & 20.10\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 366 & 1199 &	 23.39\% \\ \hline
 
-\multirow{2}{*}{atlassian-python-api} & Comby & 302	& 1002 & 207 &	19.99\% \\\cline{2-6}
-    & Comby & 333 & 635 & 351 & 25.25\% \\ \hline
-\multirow{2}{*}{ESD} & Comby & 1859	& 2767 & 1642 & 29.46\% \\\cline{2-6}
-    & Comby & 1712 & 1548 &	614 & 44.18\% \\ \hline
-\multirow{2}{*}{fiber} & Comby & 493 & 909 & 163 & 31.08\% \\\cline{2-6}
-    & Comby & 458 & 767	& 241 & 31.24\% \\ \hline
+\multirow{2}{*}{ord} & \multicolumn{2}{c|}{Regex} & 762	& 2096 & 26.66\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 643 & 1669 & 27.81\% \\ \hline
+\multirow{2}{*}{bazuka} & \multicolumn{2}{c|}{Regex} & 451 & 1103 & 29.02\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 402 & 920 & 30.41\% \\ \hline
+\multirow{2}{*}{strum} & \multicolumn{2}{c|}{Regex} & 51 & 377	& 11.92\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 50 & 318	& 13.51\% \\ \hline
+
+\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 452.33 & 1392.33	& 23.21\%\\\cline{3-6}
+    &   &  Median & 418.50 & 1348.50 & 23.38\%  \\\cline{3-6}
+    &   &  Std. Dev & 309.05 & 656.76 & 12.16\%  \\\cline{2-6}
+
+ & \multirow{3}{*}{Comby} & Mean & 554.00 & 1055.83	& 29.29\% \\\cline{3-6}
+    &   &  Median & 384.00 & 1059.50	& 25.60\%  \\\cline{3-6}
+    &   &  Std. Dev & 592.63 & 508.30 & 21.63\%  \\\hline
 
@@
 
-% ---------------- JAVA table --------------------------
+% ---------------- JAVA table (UM vs Others) --------------------------
 
@@
 \centering
-\caption{JAVA (UniversalMutator vs. PIT vs. LittleDarwin)}
+\caption{Java (UniversalMutator vs. PIT vs. LittleDarwin)}
 \label{tab:table_java2}
@@
     & LittleDarwin & 8 & 0 & 1.00 & - \\ \hline
-\multirow{3}{*}{\small spring-hateoas} & Regex & 82 & 166 & 0.33 & - \\\cline{2-6}
+\multirow{4}{*}{\small spring-hateoas} & Regex & 82 & 166 & 0.33 & - \\\cline{2-6}
     & Comby & 65 & 111 & 0.37 & - \\ \cline{2-6}
@@
     & LittleDarwin & 15 & 23 & 0.39 & - \\ \hline
-\multirow{3}{*}{\small javacc} & Regex & 123 & 419 & 0.23 & - \\\cline{2-6}
+\multirow{4}{*}{\small javacc} & Regex & 123 & 419 & 0.23 & - \\\cline{2-6}
     & Comby & 103 & 400 & 0.21 & - \\ \cline{2-6}
@@
     & LittleDarwin & 64 & 18 & 0.78 & - \\ \hline
-\multirow{3}{*}{\small coffee-gb} & Regex & 48 & 176 & 0.21 & - \\\cline{2-6}
+\multirow{4}{*}{\small coffee-gb} & Regex & 48 & 176 & 0.21 & - \\\cline{2-6}
     & Comby & 40 & 134 & 0.23 & - \\ \cline{2-6}
@@
     & LittleDarwin & 11 & 18 & 0.38 & -\\ \hline
-\multirow{3}{*}{\small egads} & Regex & 442 & 182 & 0.71 & - \\\cline{2-6}
-    & Comby & 422 & 200 & 0.68 & - \\ \cline{2-6}
+\multirow{4}{*}{\small egads} & Regex & 442 & 182 & 0.71 & - \\\cline{2-6}
+    & Comby & 422 & 200 & 0.71 & - \\ \cline{2-6}
     & PIT & 87 & 20 & 0.81 & - \\ \cline{2-6}
     & LittleDarwin & 74 & 11 & 0.87 & - \\ \hline
-\multirow{3}{*}{\small apk-parser} & Regex & 26 & 325 & 0.07 & \\\cline{2-6}
+\multirow{4}{*}{\small apk-parser} & Regex & 26 & 325 & 0.07 & \\\cline{2-6}
     & Comby & 26 & 312 & 0.08 & -\\ \cline{2-6}
@@
     &   &  \multicolumn{2}{c|}{Median} & 0.28 & - \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.27 & - \\\cline{2-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.30 & - \\\cline{2-6}
 
@@
     &   &  \multicolumn{2}{c|}{Median} & 0.30 & - \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.26 & - \\\cline{2-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.29 & - \\\cline{2-6}
 
@@
     &   &  \multicolumn{2}{c|}{Median} & 0.55 & - \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.34 & - \\\cline{2-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.37 & - \\\cline{2-6}
 
@@
     &   &  \multicolumn{2}{c|}{Median} & 0.59 & - \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.32 & - \\ \hline
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.35 & - \\ \hline
 
@@
 
-% ----------------  Python table -------------------------------
+% ----------------  Python table (UM vs Others) -------------------------------
 
@@
     & Comby & 121 & 1591 & 0.07 & - \\ \cline{2-6}
-    & Mutmut & (57+0) & 663 & 0.08 & - \\ \cline{2-6}
+    & Mutmut & 57 & 663 & 0.08 & - \\ \cline{2-6}
     & CosmicRay & 63 & 950 & 0.06 & - \\ \hline
@@
     &   &  \multicolumn{2}{c|}{Median} & 0.55 & - \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.21 & - \\\cline{2-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.23 & - \\\cline{2-6}
 
@@
     &   &  \multicolumn{2}{c|}{Median} & 0.58 & - \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.22 & - \\\cline{2-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.24 & - \\\cline{2-6}
 
@@
     &   &  \multicolumn{2}{c|}{Median} & 0.61 & - \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.22 & - \\\cline{2-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.24 & - \\\cline{2-6}
 
@@
     &   &  \multicolumn{2}{c|}{Median} & 0.45 & - \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.26 & - \\ \hline
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.28 & - \\ \hline
+
+\end{tabular}
+}
+\end{table}
+
+% ---------------- Rust table (UM vs Others)--------------------------
+
+\begin{table}[htbp]
+\centering
+\caption{Rust (UniversalMutator vs. Cargo-Mutants)}
+\label{tab:table_rust2}
+\resizebox{\columnwidth}{!}{
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+ &  & \textbf{Killed} & \textbf{Not Killed} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{Equivalent \%} \\ \hline
+
+ \multirow{3}{*}{\small cargo release} & Regex & 721 & 130 & 0.85 & - \\\cline{2-6}
+    & Comby & 1344 & 346 & 0.80 & - \\\cline{2-6}
+    & Cargo-Mutants & 2	& 3 & 0.40 & - \\ \hline
+\multirow{3}{*}{\small passerine} & Regex & 122 & 91 & 0.57 & - \\\cline{2-6}
+    & Comby & 90 & 83 & 0.52 & - \\ \cline{2-6}
+    & Cargo-Mutants & 0	& 0 & 0 & - \\ \hline
+\multirow{3}{*}{\small typos} & Regex & 295 & 91 & 0.76 & - \\\cline{2-6}
+    & Comby & 283 & 83 & 0.77 & - \\ \cline{2-6}
+    & Cargo-Mutants & 5 & 1 & 0.83 & -\\ \hline
+
+\multirow{3}{*}{\small ord} & Regex & 558 & 204 & 0.73 & - \\\cline{2-6}
+    & Comby & 533 & 110 & 0.83 & - \\ \cline{2-6}
+    & Cargo-Mutants & 6 & 0 & 1.00 & - \\ \hline
+\multirow{3}{*}{\small bazuka} & Regex & 281 & 170 & 0.62 & - \\\cline{2-6}
+    & Comby & 291 & 111 & 0.73 & - \\ \cline{2-6}
+    & Cargo-Mutants & 0 & 0 & 0 & - \\ \hline
+\multirow{3}{*}{\small strum} & Regex & 8 & 43 & 0.16 & \\\cline{2-6}
+    & Comby & 11 & 39 & 0.22 & -\\ \cline{2-6}
+    & Cargo-Mutants & 2 & 0 & 1.00 & -\\ \hline
+
+\multirow{9}{*}{Overall} & \multirow{3}{*}{Regex} & \multicolumn{2}{c|}{Mean} & 0.62  & -\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.68 & - \\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.24 & - \\\cline{2-6}
+
+ & \multirow{3}{*}{Comby} & \multicolumn{2}{c|}{Mean} & 0.64 & -\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.75 & - \\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.25 & - \\\cline{2-6}
+
+ & \multirow{3}{*}{Cargo-Mutants} & \multicolumn{2}{c|}{Mean} & 0.54 & -\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.62 & - \\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.47 & - \\ \hline
 

### 2023-03-27 paper/main.tex
add tables for c++
section 3 bit

@@
 
-\section{Using Parser Parser Combinators Instead of Regular Expressions}
+\section{Beyond Regular Expressions with Parser Parser Combinators}
+% Better to frame this as "Beyond regex with PPC" than "PPC instead of RE" 
+% instead of RE" because RE pattern matching can be implemented by PPCs,
+% they are just pattern matching functions that can be restricted to lexical
+% RE-style matching too, and, e.g., comby embeds PCRE regex matching support,
+% as a parser combinator). In this sense, PPCs _really are_ a better/more
+% general/more universal matching engine abstraction at the expressivity level
+% where we care about mutating source syntax with CFG properties.
+
+We implement a lightweight parser-driven approach that attunes mutation
+operators to language-general syntax-aware transformations. For example, our
+operators can ensure syntactically well-formed outputs (e.g., transformations
+must preserve balanced parentheses or braces). More generally, operators are
+sensitive to coarse-grained context-free grammar properties. These operators go
+beyond the expressivity of regular expressions (which are typically limited to
+lexical transformations) to generate \emph{varied} and \emph{valid}
+programs during mutation testing.
+
+Parser parser combinators~\cite{vanTonderPPC} is a recent approach enabling
+syntax-transformation for multiple languages, and (we show) an apt choice for
+mutation testing operators. This key idea is to replace code constructs
+common to many languages (e.g., multi-line code blocks delineated by braces)
+with simple declarative patterns. Using parser combinators for coarse syntax
+matching avoids complexity of integrating heterogeneous rewrite tools for
+mutation testing (e.g., one for Java, one for C++) and human effort to
+implement language-specific mutation operators (e.g., where the user has to
+learn to write operators for each language tool). The mechanization and
+expressive properties of parser parser combinators is covered in prior
+literature~\cite{vanTonderPPC}; in this paper we demonstrate that their application is
+uniquely effective for universal source-based mutation testing. In particular,
+build mutation operators with \texttt{comby},\footnote{https://comby.dev} the tool that implements parser parser
+combinators to declaratively match and rewrite source code syntax.
+
+Comby is especially suited for matching multi-line code blocks and
+disambiguating code from comments and strings---constructs that otherwise
+confound and complicate regular expression patterns. Comby supports over 40
+languages and implements a generic parser for additional languages to recognize
+these syntactic constructs. The following is a brief overview of Comby
+definitions and syntax for mutation operators implemented in our approach:
+
+\begin{itemize}
+
+\item \texttt{\small:[hole]} syntax matches source code assigned to a variable \emph{hole}.
+Holes match all characters (including whitespace) lazily up to its suffix (analogous to \texttt{\small.*?} in regex), but only \emph{within} its level of balanced delimiters.
+For example, \texttt{\small\{:[hole]\}} will match all characters inside balanced braces.
+By default, parentheses and square brackets are also treated as balanced delimiters, as applicable to most languages.
+
+\item \texttt{\small ":[x]"} matches the body of a well-quoted string.
+Unlike \texttt{\small :[x]} (without quotes), the quoted variety implies that a data string may be any value, including strings that contain unbalanced parentheses, like \texttt{\small "item)"}.
+
+\item \texttt{\small:[hole:e]} matches words like \texttt{\small hello} and contiguous well-balanced expression-like syntax, like \texttt{\small print("hello world")} or \texttt{\small (a + b)}.
+It stops matching at whitespace boundaries, and so does not match a string like \texttt{\small a + b}.
+It also does not match unbalanced code syntax like \texttt{\small foo)} in typical languages where expressions are expected to be well-balanced.
+
+\item \texttt{\small :[[hole]]} matches alphanumeric characters in source code (similar to \texttt{\small\textbackslash w+} in regex).
+
+\item When a variable \emph{hole} occurs multiple times in a match template, the matched matched values be equal.
+
+\item Non-whitespace characters are matched literally.
+
+\item Contiguous whitespace (e.g., spaces and newline) match contiguous whitespace in the source code.
+That is, match templates are sensitive to the \emph{presence} of whitespace, but \emph{not} the exact layout of spaces (i.e., the number of spaces may not correspond exactly between match template and source code).
+
+\item Matching is insensitive to comments. Comments are parsed as whitespace when matching non-hole syntax in the match template.
+
+\end{itemize}
+
+Comby provides mechanisms to further define custom match syntax and behavior
+beyond these standard patterns.\footnote{https://comby.dev/docs/advanced-usage\#custom-language-definitions} \texttt{Comby} patterns can also embed regular expressions, subsuming their
+expressive power to enable regex-based mutation operators intermixed with
+context-free syntax properties.\footnote{https://comby.dev/docs/basic-usage\#using-regular-expressions}
 
@@
 \subsection{Regular Expressions vs. Parser Parser Combinators}
+
+% ---------------- C++ table UM Only--------------------------
+\begin{table}[hbtp]
+\centering
+\caption{C++ (Regex vs. Comby)}
+\label{tab:table_cpp1}
+\resizebox{\columnwidth}{!}{%
+
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+ & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%}  \\ \hline
+\multirow{2}{*}{ompl} & \multicolumn{2}{c|}{Regex} & 595 & 1938 & 23.49\%  \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 564 & 1610 & 25.94\% \\ \hline
+\multirow{2}{*}{libgraphqlparser} & \multicolumn{2}{c|}{Regex} & 157 & 584 & 21.19\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 131 & 309 & 29.77\% \\ \hline
+\multirow{2}{*}{rethinkdb\_rebirth} & \multicolumn{2}{c|}{Regex} & 545 & 1701 & 24.27\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 492 & 1517 & 24.49\% \\ \hline
+
+\multirow{2}{*}{xoreos} & \multicolumn{2}{c|}{Regex} & 2419 & 1457 & 62.41\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 2280 & 1017 & 69.15\% \\ \hline
+\multirow{2}{*}{libxmljs} & \multicolumn{2}{c|}{Regex} & 881 & 3108 & 22.09\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 790 & 3218 & 19.71\% \\ \hline
+\multirow{2}{*}{tiny-differentiable-simulator} & \multicolumn{2}{c|}{Regex} & 747 & 535 & 58.27\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 770 & 444 & 63.43\% \\ \hline
+
+\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 890.67 &	1553.83 & 35.28\%\\\cline{3-6}
+    &   &  Median & 671.00 & 1579.00 & 23.88\%  \\\cline{3-6}
+    &   &  Std. Dev & 787.62 & 956.31 & 19.48\%  \\\cline{2-6}
+
+ & \multirow{3}{*}{Comby} & Mean & 837.83 & 1352.50 & 38.75\% \\\cline{3-6}
+    &   &  Median & 667.00 & 1267.00 & 27.86\%  \\\cline{3-6}
+    &   &  Std. Dev & 745.78 & 1058.33	& 21.66\%  \\\hline
+
+\end{tabular}
+}
+
+\end{table}
 
@@
 \subsection{Universal Source-Based Mutation vs. Previous Approaches}
+
+% ---------------- C++ table (UM vs Others)--------------------------
+
+\begin{table}[htbp]
+\centering
+\caption{C++ (UniversalMutator vs. Dextool)}
+\label{tab:table_cpp2}
+\resizebox{\columnwidth}{!}{
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+ &  & \textbf{Killed} & \textbf{Not Killed} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{Equivalent \%} \\ \hline
+
+ \multirow{3}{*}{\small ompl} & Regex & 29 & 566 & 0.05 & 30\% \\\cline{2-6}
+    & Comby & 41 & 523 & 0.07 & 40\% \\\cline{2-6}
+    & Dextool & 23 & 255 & 0.08 & N/A \\ \hline
+\multirow{3}{*}{\small libgraphqlparser} & Regex & 142 & 15 & 0.90 & 30\% \\\cline{2-6}
+    & Comby & 116 & 15 & 0.89 & 20\% \\ \cline{2-6}
+    & Dextool & 50 & 2 & 0.96 & N/A \\ \hline
+\multirow{3}{*}{\small rethinkdb\_rebirth} & Regex & 65 & 480 & 0.12 & 10\% \\\cline{2-6}
+    & Comby & 49 & 443 & 0.10 & 20\% \\ \cline{2-6}
+    & Dextool & 9 & 16 & 0.36 & N/A \\ \hline
+
+\multirow{3}{*}{\small xoreos} & Regex & 33 & 2386 & 0.01 & 0\% \\\cline{2-6}
+    & Comby & 33 & 2247 & 0.01 & 10\% \\ \cline{2-6}
+    & Dextool & 8 & 406 & 0.02 & N/A \\ \hline
+\multirow{3}{*}{\small libxmljs} & Regex & 547 & 334 & 0.62 & 10\% \\\cline{2-6}
+    & Comby & 507 & 283 & 0.64 & 10\% \\ \cline{2-6}
+    & Dextool & 383 & 447 & 0.46 & N/A \\ \hline
+\multirow{3}{*}{\small tiny-differentiable-simulator} & Regex & 38 & 709 & 0.05 & 0\% \\\cline{2-6}
+    & Comby & 32 & 738 & 0.04 & 0\% \\ \cline{2-6}
+    & Dextool & 13 & 48 & 0.21 & N/A\\ \hline
+
+\multirow{9}{*}{Overall} & \multirow{3}{*}{Regex} & \multicolumn{2}{c|}{Mean} & 0.29  & 13.33\%\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.09 & 10.00\% \\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.38 & 13.66\% \\\cline{2-6}
+
+ & \multirow{3}{*}{Comby} & \multicolumn{2}{c|}{Mean} & 0.29 & 16.67\% \\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.09 & 15.00\% \\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.37 & 13.66\% \\\cline{2-6}
+
+ & \multirow{3}{*}{Cargo-Mutants} & \multicolumn{2}{c|}{Mean} & 0.33 & N/A\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.29 & N/A \\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.38 & N/A \\ \hline
+
+\end{tabular}
+}
+\end{table}
 

### 2023-03-27 paper/main.tex
Cleaning up tables

@@
 
-\begin{tabular}{|c|c|c|c|c|c|}
+\begin{tabular}{|c|c|c|r|r|r|}
 \hline
  & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%}  \\ \hline
-\multirow{2}{*}{ompl} & \multicolumn{2}{c|}{Regex} & 595 & 1938 & 23.49\%  \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 564 & 1610 & 25.94\% \\ \hline
-\multirow{2}{*}{libgraphqlparser} & \multicolumn{2}{c|}{Regex} & 157 & 584 & 21.19\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 131 & 309 & 29.77\% \\ \hline
-\multirow{2}{*}{rethinkdb\_rebirth} & \multicolumn{2}{c|}{Regex} & 545 & 1701 & 24.27\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 492 & 1517 & 24.49\% \\ \hline
+\multirow{2}{*}{ompl} & \multicolumn{2}{c|}{Regex} & 595 \hspace{8pt} & 1938 \hspace{8pt} & 23.49\%  \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 564 \hspace{8pt} & 1610 \hspace{8pt} & 25.94\% \\ \hline
+\multirow{2}{*}{libgraphqlparser} & \multicolumn{2}{c|}{Regex} & 157 \hspace{8pt} & 584 \hspace{8pt} & 21.19\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 131 \hspace{8pt} & 309 \hspace{8pt} & 29.77\% \\ \hline
+\multirow{2}{*}{rethinkdb\_rebirth} & \multicolumn{2}{c|}{Regex} & 545 \hspace{8pt} & 1701 \hspace{8pt} & 24.27\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 492 \hspace{8pt} & 1517 \hspace{8pt} & 24.49\% \\ \hline
 
-\multirow{2}{*}{xoreos} & \multicolumn{2}{c|}{Regex} & 2419 & 1457 & 62.41\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 2280 & 1017 & 69.15\% \\ \hline
-\multirow{2}{*}{libxmljs} & \multicolumn{2}{c|}{Regex} & 881 & 3108 & 22.09\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 790 & 3218 & 19.71\% \\ \hline
-\multirow{2}{*}{tiny-differentiable-simulator} & \multicolumn{2}{c|}{Regex} & 747 & 535 & 58.27\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 770 & 444 & 63.43\% \\ \hline
+\multirow{2}{*}{xoreos} & \multicolumn{2}{c|}{Regex} & 2419 \hspace{8pt} & 1457 \hspace{8pt} & 62.41\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 2280 \hspace{8pt} & 1017 \hspace{8pt} & 69.15\% \\ \hline
+\multirow{2}{*}{libxmljs} & \multicolumn{2}{c|}{Regex} & 881 \hspace{8pt} & 3108 \hspace{8pt} & 22.09\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 790 \hspace{8pt} & 3218 \hspace{8pt} & 19.71\% \\ \hline
+\multirow{2}{*}{tiny-differentiable-simulator} & \multicolumn{2}{c|}{Regex} & 747 \hspace{8pt} & 535 \hspace{8pt} & 58.27\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 770 \hspace{8pt} & 444 \hspace{8pt} & 63.43\% \\ \hline
 
@@
 
+\begin{table}[hbtp]
+  \centering
+  \caption{Overview of Mutation Testing Tools. universalmutator has singificantly fewer lines of code than any of the single language tools.}
+  \label{tab:mutationtools}
+
+  \begin{tabular}{|c|c|r|}
+  \hline
+  \textbf{Language} & \textbf{Tool} & \textbf{LOC}  \\
+  \hline
+  \multirow{2}{*}{Python}  & Mutmut &	3870  \\\cline{2-3}
+      & cosmic-ray & 4599 \\ \hline
+  \multirow{2}{*}{Java}  & PIT &	59577  \\\cline{2-3}
+      & LittleDarwin & 22359 \\ \hline
+      Rust & Cargo-Mutants & 7020 \\\hline
+      C++ & Dextool & 38611 \\\hline
+      ALL & universalmutator & 2244 \\\hline
+  \end{tabular}
+  \end{table}
+
 % ---------------- JAVA table UM Only--------------------------
@@
 
-\begin{tabular}{|c|c|c|c|c|c|}
+\begin{tabular}{|c|c|c|r|r|r|}
 \hline
  & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%}  \\ \hline
-\multirow{2}{*}{pebble} & \multicolumn{2}{c|}{Regex} & 86 & 928 & 8.48\%  \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 85 & 667 & 11.30\% \\ \hline
-\multirow{2}{*}{spring-hateoas} & \multicolumn{2}{c|}{Regex} & 248 & 2231 & 10.00\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 172 & 1212 & 12.43\% \\ \hline
-\multirow{2}{*}{javacc} & \multicolumn{2}{c|}{Regex} & 542 & 1346 & 28.70\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 494 & 782 & 38.71\% \\ \hline
+\multirow{2}{*}{pebble} & \multicolumn{2}{c|}{Regex} & 86 \hspace{8pt} & 928 \hspace{8pt} & 8.48\%  \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 85 \hspace{8pt} & 667 \hspace{8pt} & 11.30\% \\ \hline
+\multirow{2}{*}{spring-hateoas} & \multicolumn{2}{c|}{Regex} & 248 \hspace{8pt} & 2231 \hspace{8pt} & 10.00\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 172 \hspace{8pt} & 1212 \hspace{8pt} & 12.43\% \\ \hline
+\multirow{2}{*}{javacc} & \multicolumn{2}{c|}{Regex} & 542 \hspace{8pt} & 1346 \hspace{8pt} & 28.70\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 494 \hspace{8pt} & 782 \hspace{8pt} & 38.71\% \\ \hline
 
-\multirow{2}{*}{coffee-gb} & \multicolumn{2}{c|}{Regex} & 224 & 1703 & 11.63\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 174 & 624 & 21.80\% \\ \hline
-\multirow{2}{*}{egads} & \multicolumn{2}{c|}{Regex} & 624 & 1830 & 25.43\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 597 & 566 & 51.33\% \\ \hline
-\multirow{2}{*}{apk-parser} & \multicolumn{2}{c|}{Regex} & 351 & 1232 & 22.17\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 333 & 866 & 27.77\% \\ \hline
+\multirow{2}{*}{coffee-gb} & \multicolumn{2}{c|}{Regex} & 224 \hspace{8pt} & 1703 \hspace{8pt} & 11.63\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 174 \hspace{8pt} & 624 \hspace{8pt} & 21.80\% \\ \hline
+\multirow{2}{*}{egads} & \multicolumn{2}{c|}{Regex} & 624 \hspace{8pt} & 1830 \hspace{8pt} & 25.43\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 597 \hspace{8pt} & 566 \hspace{8pt} & 51.33\% \\ \hline
+\multirow{2}{*}{apk-parser} & \multicolumn{2}{c|}{Regex} & 351 \hspace{8pt} & 1232 \hspace{8pt} & 22.17\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 333 \hspace{8pt} & 866 \hspace{8pt} & 27.77\% \\ \hline
 
@@
 
-\begin{tabular}{|c|c|c|c|c|c|c|}
+\begin{tabular}{|c|c|c|r|r|r|r|}
 \hline
  & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Redundant} & \textbf{Valid \%}  \\ \hline
-\multirow{2}{*}{keract} & \multicolumn{2}{c|}{Regex} & 1655 & 1600 &	729 & 41.53\%  \\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 1471 & 1063 & 533 & 47.96\% \\ \hline
-\multirow{2}{*}{dtw} & \multicolumn{2}{c|}{Regex} & 973 & 566 & 567 & 46.20\% \\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 866 & 330 & 639 &	47.19\% \\ \hline
-\multirow{2}{*}{notion-sdk-py} & \multicolumn{2}{c|}{Regex} & 196 & 760 & 82 & 18.89\% \\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 196 & 701 & 122 & 19.23\% \\ \hline
+\multirow{2}{*}{keract} & \multicolumn{2}{c|}{Regex} & 1655 \hspace{8pt} & 1600 \hspace{8pt} &	729 \hspace{8pt} & 41.53\%  \\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 1471 \hspace{8pt} & 1063 \hspace{8pt} & 533 \hspace{8pt} & 47.96\% \\ \hline
+\multirow{2}{*}{dtw} & \multicolumn{2}{c|}{Regex} & 973 \hspace{8pt} & 566 \hspace{8pt} & 567 \hspace{8pt} & 46.20\% \\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 866 \hspace{8pt} & 330 \hspace{8pt} & 639 \hspace{8pt} &	47.19\% \\ \hline
+\multirow{2}{*}{notion-sdk-py} & \multicolumn{2}{c|}{Regex} & 196 \hspace{8pt} & 760 \hspace{8pt} & 82 \hspace{8pt} & 18.89\% \\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 196 \hspace{8pt} & 701 \hspace{8pt} & 122 \hspace{8pt} & 19.23\% \\ \hline
 
-\multirow{2}{*}{{\makecell{atlassian-python \\ api}}} & \multicolumn{2}{c|}{Regex} & 302	& 1002 & 207 &	19.99\% \\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 333 & 635 & 351 & 25.25\% \\ \hline
-\multirow{2}{*}{ESD} & \multicolumn{2}{c|}{Regex} & 1859	& 2767 & 1642 & 29.46\% \\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 1712 & 1548 &	614 & 44.18\% \\ \hline
-\multirow{2}{*}{fiber} & \multicolumn{2}{c|}{Regex} & 493 & 909 & 163 & 31.08\% \\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 458 & 767	& 241 & 31.24\% \\ \hline
+\multirow{2}{*}{{\makecell{atlassian-python \\ api}}} & \multicolumn{2}{c|}{Regex} & 302 \hspace{8pt}	& 1002 \hspace{8pt} & 207 \hspace{8pt} &	19.99\% \\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 333 \hspace{8pt} & 635 \hspace{8pt} & 351 \hspace{8pt} & 25.25\% \\ \hline
+\multirow{2}{*}{ESD} & \multicolumn{2}{c|}{Regex} & 1859 \hspace{8pt}	& 2767 \hspace{8pt} & 1642 \hspace{8pt} & 29.46\% \\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 1712 \hspace{8pt} & 1548 \hspace{8pt} &	614 \hspace{8pt} & 44.18\% \\ \hline
+\multirow{2}{*}{fiber} & \multicolumn{2}{c|}{Regex} & 493 \hspace{8pt} & 909 \hspace{8pt} & 163 \hspace{8pt} & 31.08\% \\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 458 \hspace{8pt} & 767 \hspace{8pt}	& 241 \hspace{8pt} & 31.24\% \\ \hline
 
@@
 
-\begin{tabular}{|c|c|c|c|c|c|}
+\begin{tabular}{|c|c|c|r|r|r|}
 \hline
  & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%}  \\ \hline
-\multirow{2}{*}{cargo release} & \multicolumn{2}{c|}{Regex} & 851 &	1163 & 42.25\%  \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 1690 & 711 & 70.39\% \\ \hline
-\multirow{2}{*}{passerine} & \multicolumn{2}{c|}{Regex} & 213 & 2081 & 9.29\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 173 & 1518 & 10.23\% \\ \hline
-\multirow{2}{*}{typos} & \multicolumn{2}{c|}{Regex} & 386 & 1534 & 20.10\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 366 & 1199 &	 23.39\% \\ \hline
+\multirow{2}{*}{cargo release} & \multicolumn{2}{c|}{Regex} & 851 \hspace{8pt} &	1163 \hspace{8pt} & 42.25\%  \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 1690 \hspace{8pt} & 711 \hspace{8pt} & 70.39\% \\ \hline
+\multirow{2}{*}{passerine} & \multicolumn{2}{c|}{Regex} & 213 \hspace{8pt} & 2081 \hspace{8pt} & 9.29\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 173 \hspace{8pt} & 1518 \hspace{8pt} & 10.23\% \\ \hline
+\multirow{2}{*}{typos} & \multicolumn{2}{c|}{Regex} & 386 \hspace{8pt} & 1534 \hspace{8pt} & 20.10\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 366 \hspace{8pt} & 1199 \hspace{8pt} &	 23.39\% \\ \hline
 
-\multirow{2}{*}{ord} & \multicolumn{2}{c|}{Regex} & 762	& 2096 & 26.66\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 643 & 1669 & 27.81\% \\ \hline
-\multirow{2}{*}{bazuka} & \multicolumn{2}{c|}{Regex} & 451 & 1103 & 29.02\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 402 & 920 & 30.41\% \\ \hline
-\multirow{2}{*}{strum} & \multicolumn{2}{c|}{Regex} & 51 & 377	& 11.92\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 50 & 318	& 13.51\% \\ \hline
+\multirow{2}{*}{ord} & \multicolumn{2}{c|}{Regex} & 762 \hspace{8pt}	& 2096 \hspace{8pt} & 26.66\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 643 \hspace{8pt} & 1669 \hspace{8pt} & 27.81\% \\ \hline
+\multirow{2}{*}{bazuka} & \multicolumn{2}{c|}{Regex} & 451 \hspace{8pt} & 1103 \hspace{8pt} & 29.02\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 402 \hspace{8pt} & 920 \hspace{8pt} & 30.41\% \\ \hline
+\multirow{2}{*}{strum} & \multicolumn{2}{c|}{Regex} & 51 \hspace{8pt} & 377 \hspace{8pt}	& 11.92\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 50 \hspace{8pt} & 318 \hspace{8pt}	& 13.51\% \\ \hline
 
@@
 \resizebox{\columnwidth}{!}{
-\begin{tabular}{|c|c|c|c|c|c|}
+\begin{tabular}{|c|c|r|r|r|r|}
 \hline

### 2023-03-27 paper/main.tex
haskell example and other edits

--- 2023-03-27 paper/main.tex
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Program Mutation is Just Program Transformation: \\ A
-  Universal Approach to Mutant Generation}
+\title{Syntax is all You Need: A Universal-Language\\Approach to Mutant Generation}
 
@@
 
-...
+Moreover, newer languages, including ones with considerable interest
+in the development community, often lack mutation testing tools.  Even
+languages with a long history may long lack a working mutation testing
+tool; e.g., it was not until 2014 that Haskell had such a tool
+\cite{mucheck}.  Furthermore, that tool may no longer  be
+supported (the last commits or package updates were in 2015), and it
+is not clear that a successor tool has appeared in the ensuing years.
+
+A great deal of research effort has been devoted to devising clever
+schemes for reducing the computational burden of mutation testing over
+the years; in practice, however, the recent adoption of mutation
+testing arguably relies more on the availability of vast (cloud) computing
+resources than any particular approach from the literature, beyond
+heuristic restrictions on the mutants used (e.g., one per line, or
+only some siimple operators).   On the other hand, almost no
+research has addressed perhaps the most central issue of
+mutation testing: in the absence of a working mutation testing
+tool for the language in which one is developing software, one cannot
+obtain the benefits of mutation testing.  The absence of such tools is
+due to the high human cost of developing and maintaining mutation
+testing tools.  Ideally, mutation testing would be available for
+\emph{any} newly proposed programming language, as soon as that
+language has any community of users at all.  In the present
+state-of-the-art, however, it may be years before even rudimentary
+tools appear, and they may ``disappear'' shortly there-after.
 
@@
 testing tools, each requiring its own development community, workflow
-for mutation, and effort to keep up with language changes, is
+for mutation, and effort to keep up with language changes, is fundamentally
 unnecessary.  The core problem of mutation testing,
@@
       that operate across many languages, and efficient to generate
-      mutants.
+      valid mutants even in the absence of language-specific
+      development effort.
       \end{enumerate}
@@
     essentially all programming languages.  This approach also makes
-    mutation testing easy to extend for new languages and even allows
+    mutation testing almost trivial to extend to most new languages and even allows
     developers to write project-specific rules easily without
@@
 \end{itemize}
+
+In addition to these core research questions, evaluated over six files
+from six real-world projects for four languages, we also present a
+case study, applying our implementation to a program in a language for
+which it has no rules, Haskell.
 
@@
 
+\subsection{Haskell Case Study}
+
+\begin{figure}
+\begin{code}
+ qsort :: [ Int] -> [ Int ]
+ qsort [] = []
+ qsort (x: xs ) = qsort l ++ [x] ++ qsort r
+     where l = filter (< x) xs
+           r = filter ( >= x) xs
+\end{code}
+\caption{Example from MuCheck Paper}
+\label{fig;haskell}
+\end{figure}
+
+Figure~\ref{fig:haskell} shows a simple Haskell program, the
+motivating example in the paper presenting the MuCheck \cite{mucheck}
+mutation testing tool for Haskell.  As noted in the introduction, to our
+knowledge MuCheck is the first (and perhaps only) mutation tool for
+Haskell code.  MuCheck has not been updated since 2015.
+To see how our approach fares when mutating code in a language for
+which 1) it has no specific rules and 2) the assumption of a simple
+C-like syntax does not hold, we applied our implementation to this example.
+
+Using regular expressions, and only the universal rules, our tool
+generates 10 valid mutants and 34 invalid mutants of the code (22.73\%
+valid mutants).  Comby produces an identical 10 valid mutants, but 65
+invalid mutants (13.33\% validity).  These mutants produce a mutation
+score of 0.8 for the two properties defined in the original MuCheck
+paper.  In other words, like MuCheck, our tool is able to
+determine that the properties provided are insufficient.  In
+particular, while they check idempotence and that the result is
+sorted, they do not check that the result of {\tt qsort} is a permutaton of the input
+list.  The MuCheck paper only provides mutation scores for the two
+properties independently.  For idempotence, MuCheck gives a mutation
+score of 0.84, and our implementation yields a score of 0.8.  Checking
+only sortedness, MuCheck yields a score of 0.61, while our approach
+produces an even lower score of 0.4.  Our lower
+scores are not due to equivalent mutants: adding a property to check
+that the sorted list is a permutation of the original list, all
+mutants generated by our approach can be killed.
+
+Of course, our implementation does not provide the small set of Haskell-specific
+mutations provided by MuCheck (e.g., type-aware function replacement).  It nonetheless produces clearly useful
+mutants, even for a language arguably radically different from those
+for which the universal rules were developed.  In short, our approach
+likely provides users of a new language working, if sub-optimal, mutation testing,
+\emph{even if no effort has been expended to define rules for the
+  language at all}.
+
+\section{Discussion}
+
 \section{Related Work}

### 2023-03-27 paper/main.tex
bib, capitals

@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Syntax is all You Need: A Universal-Language\\Approach to Mutant Generation}
+\title{Syntax Is All You Need: A Universal-Language\\Approach to Mutant Generation}
 
@@
 
+       \item A small case study showing that our approach produces
+         useful results, even without the definition of
+         language-specific rules, for Haskell, a language arguably
+         well outside the language paradigms we considered when
+         developing the universal mutation rules.
+
 \end{enumerate}
 
-\section{The Universal Source-Based Approach}
+\section{The Universal Language Source-Based Approach}
 
@@
 with simple declarative patterns. Using parser combinators for coarse syntax
-matching avoids complexity of integrating heterogeneous rewrite tools for
+matching avoids the complexity of integrating heterogeneous rewrite tools for
 mutation testing (e.g., one for Java, one for C++) and human effort to
@@
 literature~\cite{vanTonderPPC}; in this paper we demonstrate that their application is
-uniquely effective for universal source-based mutation testing. In particular,
+uniquely effective for universal source-based mutation testing. In
+particular, we
 build mutation operators with \texttt{comby},\footnote{https://comby.dev} the tool that implements parser parser

### 2023-03-27 paper/main.tex
small stuff

@@
 approach to mutation to existing tools, and demonstrate the advantages
-of using parser-parser combinators to improve on a simple regular-expression
+of using parser parser combinators to improve on a simple regular-expression
 based approach to generation.  
@@
   to parse C++ 2020, and is particularly problematic for somewhat
-  fast-moving languages such as Rust.}.  Of course, many open source
-projects are abandoned, and academic software has a long history of
-being available and supported only so long as a graduate student is
-working on a thesis or a particular grant is being funded.  Still,
-there appears to be a significant risk in relying on any mutation tools other
+  fast-moving languages such as Rust.}.  There
+appears to be a significant risk in relying on any mutation tools other
 than the most widely used and well-supported, such as PIT.
@@
 
-          \item A proposal to use parser-parser combinators to improve the
+          \item A proposal to use parser parser combinators to improve the
       efficiency and expressive power of that approach, and an
@@
 
-  Figure~\ref{fig:rules} shows examples of universal, C-like, and
-  language specific rules for a few languages, taken directly from our
-  implementation.
+\begin{table}[hbtp]
+\centering
+\caption{Rule Examples}
+\label{tab:rules}
+\resizebox{\columnwidth}{!}{%
+
+  \begin{tabular}{|l|l|l|}
+    \hline
+    Universal & C-Like & Language-Specific \\
+    \hline
+{\tt < ==> >} & else ==> & synchronized ==> \\ 
+{\tt < ==> ==} & {\tt if (\textbackslash(.*\textbackslash)) ==> if (!\textbackslash 1)} &
+                                                                 {\tt (\^{}\textbackslash s*)(\textbackslash S+.*\textbackslash n) ==> \textbackslash 1pass}\\
+{\tt < ==> <=} & {\tt if (\textbackslash(.*\textbackslash)) ==> if (1==1)} & any\_of ==> all\_of\\
+\hline
+
+\end{tabular}
+}
+
+\end{table}
+
+  Table~\ref{tab:rules} shows examples of universal, C-like, and
+  language-specific rules for a few languages, taken directly from our
+  implementation.  In all cases, the form of a rule is: {\tt regexp
+    ==> replacement}, using Python regular expression and replacement syntax.
+
+  The first column shows universal rules applied to
+  all languages, here drawn from the large set of transformations for
+  less-than comparisons.  The second column shows somewhat more
+  complex rules, for removing elses (which always leaves valid code in
+  C-like languages, but with the else clause always executing) and for
+  replacing if conditions with, respectively, a negation and a
+  constantly true expression.  Finally, the last column shows
+  language-specific rules, for, respectively, Java, Python (statement
+  deletion, including a {\tt pass} and respecting indentation level),
+  and C++.  The last rule was contributed by a user using our tool to
+  perform mutation testing on the bitcoin core implementation.  For a
+  C++ program, all of the rules in the first two columns will be
+  applied, as well as the last rule in the third column.  For a Python
+  program, however, only the first column of rules and the second rule
+  in the last column apply, as Python is not sufficiently ``C-like''.
 
@@
 
-  \item{RQ1:}  How does the use of parser-parser combinators modify 
+  \item{RQ1:}  How does the use of parser parser combinators modify 
     the efficiency (in terms of invalid mutants) and effectiveness (in
@@
 \item{RQ2:}  How does the univeral approach, using regular expressions
-  or parser-combinator combinator rules, compare to existing
+  or parser parser combinator rules, compare to existing
   single-language mutation tools, in terms of number of generated
@@
       C++ & Dextool & 38611 \\\hline
-      ALL & universalmutator & 2244 \\\hline
+      ALL & universalmutator & 2680 \\\hline
   \end{tabular}
@@
 \end{code}
+\label{fig;haskell}
 \caption{Example from MuCheck Paper}
-\label{fig;haskell}
 \end{figure}
@@
 programs over four languages,
-and introduces the idea of using parser-parser combinators to improve
+and introduces the idea of using parser parser combinators to improve
 on regular expression-based generation.

### 2023-03-27 paper/main.tex
alex stop for now

@@
 tool; e.g., it was not until 2014 that Haskell had such a tool
-\cite{mucheck}.  Furthermore, that tool may no longer  be
-supported (the last commits or package updates were in 2015), and it
-is not clear that a successor tool has appeared in the ensuing years.
+\cite{mucheck}.  Furthermore, that tool is no longer
+supported\footnote{The last commits or package updates were in 2015, and the
+primary author confirms that changes to {\tt haskell-src-ext} since
+introduced render it unusable at present.}, and it
+appears that no successor tool has appeared in the ensuing
+years\footnote{There is FitSpec \cite{FitSpec}, which is maintained,
+  but FitSpec does not perform traditional mutation testing or produce
+  source code (or bytecode) mutants; it rather
+  performs black-box type-aware mutation of function return values.}.
+Haskell developers (until now, see below) have no option for
+performing true mutation testing.
 
@@
     would introduce new mutation operators or modify existing ones.
-    The local nature of operators tends to make the second case in
-    particular very rare.
+    The local nature of operators tends to make the second case very rare.
     \item The existing literature on multi-lingual or language
@@
     \hline
-{\tt < ==> >} & else ==> & synchronized ==> \\ 
+{\tt < ==> >} &{\tt  else ==>} & {\tt synchronized ==>} \\ 
 {\tt < ==> ==} & {\tt if (\textbackslash(.*\textbackslash)) ==> if (!\textbackslash 1)} &
                                                                  {\tt (\^{}\textbackslash s*)(\textbackslash S+.*\textbackslash n) ==> \textbackslash 1pass}\\
-{\tt < ==> <=} & {\tt if (\textbackslash(.*\textbackslash)) ==> if (1==1)} & any\_of ==> all\_of\\
+{\tt < ==> <=} & {\tt if (\textbackslash(.*\textbackslash)) ==> if
+                 (1==1)} & {\tt any\_of ==> all\_of}\\
 \hline
@@
 \begin{figure}
+
   \begin{code}
@@
 \end{code}
-\label{fig;haskell}
 \caption{Example from MuCheck Paper}
+\label{fig:haskell}
+
 \end{figure}
@@
 valid mutants).  Comby produces an identical 10 valid mutants, but 65
-invalid mutants (13.33\% validity).  These mutants produce a mutation
+invalid mutants (13.33\% validity).  Both tools produce a large number
+of invalid mutants here due to including rules such as adding {\tt
+  break} and {\tt continue} that are sufficiently widespread in
+applicability to be included in our ``universal'' set but do not apply
+to Haskell; we believe that comby's awareness of Haskell syntax in
+this case actually gives it \emph{more} opportunities to apply such
+inapplicable mutants.
+
+These mutants produce a mutation
 score of 0.8 for the two properties defined in the original MuCheck

### 2023-03-28 paper/main.tex
bit of text

@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Syntax Is All You Need: A Universal-Language\\Approach to Mutant Generation}
+\title{Syntax Is All You Need: A Universal-Language Approach to Mutant Generation}
 
@@
 
-Our approach also makes introducing advanced techniques such as
-Trivial Compiler Equivalence (TCE) into a mutation workflow easy.
-While we support TCE for some languages inside the tool, a user can
-always write a simple script to check...
+We implemented our approach in a mutation testing tool, written in
+about 2,200 lines of Python.  The rules defining mutation operators
+for a set of languages including C, C++, Python, Java, Lisp, Swift,
+Rust, Solidity, Fe, Vyper, JavaScript, and Go, required another 436
+lines; this is the total for comby and regexp-based rules, so each
+version requires only 218 lines of rule definitions.
+
+While many tools provide a GUI front-end, or integrate to a particular
+testing library, we chose to focus on an easily configured CLI,
+allowing a user to specify the commmands used to build mutants or
+run tests.  While less convenient in the case where there is a good
+fit between a tool's integrations and a project's environment, we
+found this flexibility useful even when limiting considerations to a
+single language.  For example, the widely used PIT system is strongly tied to
+certain versions of JUnit, and mutating projects using a different
+testing infrastructure, or an older JUnit, can be difficult.  Our
+approach also is suitable for use in CI and other automated environments.
+
+For the most part we expect users to provide build commands, but do
+support automatic compilation and comparison of mutants for Python,
+Solidity, Fe, Vyper, Swift, Rust, and Java, including Trivial Compiler
+Equivalence \cite{TCE} checks for redundancy of mutants.  These
+default handlers can always be over-ridden if they do not work with a
+particular build setup.
+
+Our tool has been available as an open source project on GitHub and
+available for installation via Python pip for
+some time, has received contributions from users not involved in
+our project, and been adopted by at least one very large project as
+the officially recommended mutation tool.  We omit details for
+purposes of blinding.
 
@@
 
-In addition to these core research questions, evaluated over six files
-from six real-world projects for four languages, we also present a
+Both research questions were addressed by generating and running
+mutants for a set of four programming languages: C++, Java, Python,
+and Rust.  We picked six projects with 1) a large number of GitHub
+stars and 2) an easy-to-execute, passing test suite for each language,
+and then chose the largest source file below 20KB in size from each project to mutate.  Limiting
+mutation to one large but not extremely large file made is easier to compare generated mutants and
+manually inspect for mutant equivalency.  The only bias to file
+selection was size, which we do not believe should affect our results
+significantly.
+
+We additionally present a small
 case study, applying our implementation to a program in a language for
-which it has no rules, Haskell.
+which it has \emph{no} rules, Haskell.
 
@@
 
+% ---------------- JAVA table UM Only--------------------------
 \begin{table}[hbtp]
 \centering
-  \caption{Overview of Mutation Testing Tools. universalmutator has singificantly fewer lines of code than any of the single language tools.}
+\caption{Java (Regex vs. Comby)}
+\label{tab:table_java1}
+\resizebox{\columnwidth}{!}{%
+
+\begin{tabular}{|c|c|c|r|r|r|}
+\hline
+ & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%}  \\ \hline
+\multirow{2}{*}{pebble} & \multicolumn{2}{c|}{Regex} & 86 \hspace{8pt} & 928 \hspace{8pt} & 8.48\%  \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 85 \hspace{8pt} & 667 \hspace{8pt} & 11.30\% \\ \hline
+\multirow{2}{*}{spring-hateoas} & \multicolumn{2}{c|}{Regex} & 248 \hspace{8pt} & 2231 \hspace{8pt} & 10.00\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 172 \hspace{8pt} & 1212 \hspace{8pt} & 12.43\% \\ \hline
+\multirow{2}{*}{javacc} & \multicolumn{2}{c|}{Regex} & 542 \hspace{8pt} & 1346 \hspace{8pt} & 28.70\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 494 \hspace{8pt} & 782 \hspace{8pt} & 38.71\% \\ \hline
+
+\multirow{2}{*}{coffee-gb} & \multicolumn{2}{c|}{Regex} & 224 \hspace{8pt} & 1703 \hspace{8pt} & 11.63\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 174 \hspace{8pt} & 624 \hspace{8pt} & 21.80\% \\ \hline
+\multirow{2}{*}{egads} & \multicolumn{2}{c|}{Regex} & 624 \hspace{8pt} & 1830 \hspace{8pt} & 25.43\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 597 \hspace{8pt} & 566 \hspace{8pt} & 51.33\% \\ \hline
+\multirow{2}{*}{apk-parser} & \multicolumn{2}{c|}{Regex} & 351 \hspace{8pt} & 1232 \hspace{8pt} & 22.17\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 333 \hspace{8pt} & 866 \hspace{8pt} & 27.77\% \\ \hline
+
+\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 345.83 & 1545.00  & 17.74\%\\\cline{3-6}
+    &   &  Median & 299.50 & 1524.50 & 16.90\%  \\\cline{3-6}
+    &   &  Std. Dev & 203.86 & 467.98 & 8.73\%  \\\cline{2-6}
+
+ & \multirow{3}{*}{Comby} & Mean & 309.17 & 786.17 & 27.22\% \\\cline{3-6}
+    &   &  Median & 253.50 & 724.50 & 24.79\%  \\\cline{3-6}
+    &   &  Std. Dev & 202.44 & 235.24  & 15.59\%  \\\hline
+
+\end{tabular}
+
+}
+\end{table}
+
+% ---------------- Python table UM Only --------------------------
+
+\begin{table}[hbtp]
+\centering
+\caption{Python (Regex vs. Comby)}
+\label{tab:table_python1}
+\resizebox{\columnwidth}{!}{%
+
+\begin{tabular}{|c|c|c|r|r|r|r|}
+\hline
+ & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Redundant} & \textbf{Valid \%}  \\ \hline
+\multirow{2}{*}{keract} & \multicolumn{2}{c|}{Regex} & 1655 \hspace{8pt} & 1600 \hspace{8pt} &	729 \hspace{8pt} & 41.53\%  \\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 1471 \hspace{8pt} & 1063 \hspace{8pt} & 533 \hspace{8pt} & 47.96\% \\ \hline
+\multirow{2}{*}{dtw} & \multicolumn{2}{c|}{Regex} & 973 \hspace{8pt} & 566 \hspace{8pt} & 567 \hspace{8pt} & 46.20\% \\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 866 \hspace{8pt} & 330 \hspace{8pt} & 639 \hspace{8pt} &	47.19\% \\ \hline
+\multirow{2}{*}{notion-sdk-py} & \multicolumn{2}{c|}{Regex} & 196 \hspace{8pt} & 760 \hspace{8pt} & 82 \hspace{8pt} & 18.89\% \\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 196 \hspace{8pt} & 701 \hspace{8pt} & 122 \hspace{8pt} & 19.23\% \\ \hline
+
+\multirow{2}{*}{{\makecell{atlassian-python \\ api}}} & \multicolumn{2}{c|}{Regex} & 302 \hspace{8pt}	& 1002 \hspace{8pt} & 207 \hspace{8pt} &	19.99\% \\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 333 \hspace{8pt} & 635 \hspace{8pt} & 351 \hspace{8pt} & 25.25\% \\ \hline
+\multirow{2}{*}{ESD} & \multicolumn{2}{c|}{Regex} & 1859 \hspace{8pt}	& 2767 \hspace{8pt} & 1642 \hspace{8pt} & 29.46\% \\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 1712 \hspace{8pt} & 1548 \hspace{8pt} &	614 \hspace{8pt} & 44.18\% \\ \hline
+\multirow{2}{*}{fiber} & \multicolumn{2}{c|}{Regex} & 493 \hspace{8pt} & 909 \hspace{8pt} & 163 \hspace{8pt} & 31.08\% \\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 458 \hspace{8pt} & 767 \hspace{8pt}	& 241 \hspace{8pt} & 31.24\% \\ \hline
+
+\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 913.00 & 1267.33 & 565.00 & 31.19\%   \\\cline{3-7}
+    &   &  Median & 733.00 & 955.50 & 387.00 & 30.27\%  \\\cline{3-7}
+    &   &  Std. Dev & 708.93 & 813.35 & 584.79 & 11.06\%  \\\cline{2-7}
+
+ & \multirow{3}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% \\\cline{3-7}
+    &   &  Median & 662.00 & 734.00 & 442.00 & 37.71\%  \\\cline{3-7}
+    &   &  Std. Dev & 628.79 & 418.99 & 211.62 & 12.28\%  \\\hline
+
+\end{tabular}
+
+}
+\end{table}
+
+% ---------------- Rust table UM Only --------------------------
+\begin{table}[hbtp]
+\centering
+\caption{Rust (Regex vs. Comby)}
+\label{tab:table_rust1}
+\resizebox{\columnwidth}{!}{%
+
+\begin{tabular}{|c|c|c|r|r|r|}
+\hline
+ & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%}  \\ \hline
+\multirow{2}{*}{cargo release} & \multicolumn{2}{c|}{Regex} & 851 \hspace{8pt} &	1163 \hspace{8pt} & 42.25\%  \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 1690 \hspace{8pt} & 711 \hspace{8pt} & 70.39\% \\ \hline
+\multirow{2}{*}{passerine} & \multicolumn{2}{c|}{Regex} & 213 \hspace{8pt} & 2081 \hspace{8pt} & 9.29\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 173 \hspace{8pt} & 1518 \hspace{8pt} & 10.23\% \\ \hline
+\multirow{2}{*}{typos} & \multicolumn{2}{c|}{Regex} & 386 \hspace{8pt} & 1534 \hspace{8pt} & 20.10\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 366 \hspace{8pt} & 1199 \hspace{8pt} &	 23.39\% \\ \hline
+
+\multirow{2}{*}{ord} & \multicolumn{2}{c|}{Regex} & 762 \hspace{8pt}	& 2096 \hspace{8pt} & 26.66\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 643 \hspace{8pt} & 1669 \hspace{8pt} & 27.81\% \\ \hline
+\multirow{2}{*}{bazuka} & \multicolumn{2}{c|}{Regex} & 451 \hspace{8pt} & 1103 \hspace{8pt} & 29.02\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 402 \hspace{8pt} & 920 \hspace{8pt} & 30.41\% \\ \hline
+\multirow{2}{*}{strum} & \multicolumn{2}{c|}{Regex} & 51 \hspace{8pt} & 377 \hspace{8pt}	& 11.92\% \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 50 \hspace{8pt} & 318 \hspace{8pt}	& 13.51\% \\ \hline
+
+\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 452.33 & 1392.33	& 23.21\%\\\cline{3-6}
+    &   &  Median & 418.50 & 1348.50 & 23.38\%  \\\cline{3-6}
+    &   &  Std. Dev & 309.05 & 656.76 & 12.16\%  \\\cline{2-6}
+
+ & \multirow{3}{*}{Comby} & Mean & 554.00 & 1055.83	& 29.29\% \\\cline{3-6}
+    &   &  Median & 384.00 & 1059.50	& 25.60\%  \\\cline{3-6}
+    &   &  Std. Dev & 592.63 & 508.30 & 21.63\%  \\\hline
+
+\end{tabular}
+
+}
+\end{table}
+
+\subsection{Universal Source-Based Mutation vs. Previous Approaches}
+
+\begin{table}[hbtp]
+  \centering
+  \caption{Overview of Mutation Testing Tools.}
   \label{tab:mutationtools}
@@
 
-% ---------------- JAVA table UM Only--------------------------
-\begin{table}[hbtp]
-\centering
-\caption{Java (Regex vs. Comby)}
-\label{tab:table_java1}
-\resizebox{\columnwidth}{!}{%
-
-\begin{tabular}{|c|c|c|r|r|r|}
-\hline
- & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%}  \\ \hline
-\multirow{2}{*}{pebble} & \multicolumn{2}{c|}{Regex} & 86 \hspace{8pt} & 928 \hspace{8pt} & 8.48\%  \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 85 \hspace{8pt} & 667 \hspace{8pt} & 11.30\% \\ \hline
-\multirow{2}{*}{spring-hateoas} & \multicolumn{2}{c|}{Regex} & 248 \hspace{8pt} & 2231 \hspace{8pt} & 10.00\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 172 \hspace{8pt} & 1212 \hspace{8pt} & 12.43\% \\ \hline
-\multirow{2}{*}{javacc} & \multicolumn{2}{c|}{Regex} & 542 \hspace{8pt} & 1346 \hspace{8pt} & 28.70\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 494 \hspace{8pt} & 782 \hspace{8pt} & 38.71\% \\ \hline
-
-\multirow{2}{*}{coffee-gb} & \multicolumn{2}{c|}{Regex} & 224 \hspace{8pt} & 1703 \hspace{8pt} & 11.63\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 174 \hspace{8pt} & 624 \hspace{8pt} & 21.80\% \\ \hline
-\multirow{2}{*}{egads} & \multicolumn{2}{c|}{Regex} & 624 \hspace{8pt} & 1830 \hspace{8pt} & 25.43\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 597 \hspace{8pt} & 566 \hspace{8pt} & 51.33\% \\ \hline
-\multirow{2}{*}{apk-parser} & \multicolumn{2}{c|}{Regex} & 351 \hspace{8pt} & 1232 \hspace{8pt} & 22.17\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 333 \hspace{8pt} & 866 \hspace{8pt} & 27.77\% \\ \hline
-
-\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 345.83 & 1545.00  & 17.74\%\\\cline{3-6}
-    &   &  Median & 299.50 & 1524.50 & 16.90\%  \\\cline{3-6}
-    &   &  Std. Dev & 203.86 & 467.98 & 8.73\%  \\\cline{2-6}
-
- & \multirow{3}{*}{Comby} & Mean & 309.17 & 786.17 & 27.22\% \\\cline{3-6}
-    &   &  Median & 253.50 & 724.50 & 24.79\%  \\\cline{3-6}
-    &   &  Std. Dev & 202.44 & 235.24  & 15.59\%  \\\hline
-
-\end{tabular}
-
-}
-\end{table}
-
-% ---------------- Python table UM Only --------------------------
-
-\begin{table}[hbtp]
-\centering
-\caption{Python (Regex vs. Comby)}
-\label{tab:table_python1}
-\resizebox{\columnwidth}{!}{%
-
-\begin{tabular}{|c|c|c|r|r|r|r|}
-\hline
- & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Redundant} & \textbf{Valid \%}  \\ \hline
-\multirow{2}{*}{keract} & \multicolumn{2}{c|}{Regex} & 1655 \hspace{8pt} & 1600 \hspace{8pt} &	729 \hspace{8pt} & 41.53\%  \\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 1471 \hspace{8pt} & 1063 \hspace{8pt} & 533 \hspace{8pt} & 47.96\% \\ \hline
-\multirow{2}{*}{dtw} & \multicolumn{2}{c|}{Regex} & 973 \hspace{8pt} & 566 \hspace{8pt} & 567 \hspace{8pt} & 46.20\% \\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 866 \hspace{8pt} & 330 \hspace{8pt} & 639 \hspace{8pt} &	47.19\% \\ \hline
-\multirow{2}{*}{notion-sdk-py} & \multicolumn{2}{c|}{Regex} & 196 \hspace{8pt} & 760 \hspace{8pt} & 82 \hspace{8pt} & 18.89\% \\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 196 \hspace{8pt} & 701 \hspace{8pt} & 122 \hspace{8pt} & 19.23\% \\ \hline
-
-\multirow{2}{*}{{\makecell{atlassian-python \\ api}}} & \multicolumn{2}{c|}{Regex} & 302 \hspace{8pt}	& 1002 \hspace{8pt} & 207 \hspace{8pt} &	19.99\% \\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 333 \hspace{8pt} & 635 \hspace{8pt} & 351 \hspace{8pt} & 25.25\% \\ \hline
-\multirow{2}{*}{ESD} & \multicolumn{2}{c|}{Regex} & 1859 \hspace{8pt}	& 2767 \hspace{8pt} & 1642 \hspace{8pt} & 29.46\% \\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 1712 \hspace{8pt} & 1548 \hspace{8pt} &	614 \hspace{8pt} & 44.18\% \\ \hline
-\multirow{2}{*}{fiber} & \multicolumn{2}{c|}{Regex} & 493 \hspace{8pt} & 909 \hspace{8pt} & 163 \hspace{8pt} & 31.08\% \\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 458 \hspace{8pt} & 767 \hspace{8pt}	& 241 \hspace{8pt} & 31.24\% \\ \hline
-
-\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 913.00 & 1267.33 & 565.00 & 31.19\%   \\\cline{3-7}
-    &   &  Median & 733.00 & 955.50 & 387.00 & 30.27\%  \\\cline{3-7}
-    &   &  Std. Dev & 708.93 & 813.35 & 584.79 & 11.06\%  \\\cline{2-7}
-
- & \multirow{3}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% \\\cline{3-7}
-    &   &  Median & 662.00 & 734.00 & 442.00 & 37.71\%  \\\cline{3-7}
-    &   &  Std. Dev & 628.79 & 418.99 & 211.62 & 12.28\%  \\\hline
-
-\end{tabular}
-
-}
-\end{table}
-
-% ---------------- Rust table UM Only --------------------------
-\begin{table}[hbtp]
-\centering
-\caption{Rust (Regex vs. Comby)}
-\label{tab:table_rust1}
-\resizebox{\columnwidth}{!}{%
-
-\begin{tabular}{|c|c|c|r|r|r|}
-\hline
- & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%}  \\ \hline
-\multirow{2}{*}{cargo release} & \multicolumn{2}{c|}{Regex} & 851 \hspace{8pt} &	1163 \hspace{8pt} & 42.25\%  \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 1690 \hspace{8pt} & 711 \hspace{8pt} & 70.39\% \\ \hline
-\multirow{2}{*}{passerine} & \multicolumn{2}{c|}{Regex} & 213 \hspace{8pt} & 2081 \hspace{8pt} & 9.29\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 173 \hspace{8pt} & 1518 \hspace{8pt} & 10.23\% \\ \hline
-\multirow{2}{*}{typos} & \multicolumn{2}{c|}{Regex} & 386 \hspace{8pt} & 1534 \hspace{8pt} & 20.10\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 366 \hspace{8pt} & 1199 \hspace{8pt} &	 23.39\% \\ \hline
-
-\multirow{2}{*}{ord} & \multicolumn{2}{c|}{Regex} & 762 \hspace{8pt}	& 2096 \hspace{8pt} & 26.66\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 643 \hspace{8pt} & 1669 \hspace{8pt} & 27.81\% \\ \hline
-\multirow{2}{*}{bazuka} & \multicolumn{2}{c|}{Regex} & 451 \hspace{8pt} & 1103 \hspace{8pt} & 29.02\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 402 \hspace{8pt} & 920 \hspace{8pt} & 30.41\% \\ \hline
-\multirow{2}{*}{strum} & \multicolumn{2}{c|}{Regex} & 51 \hspace{8pt} & 377 \hspace{8pt}	& 11.92\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 50 \hspace{8pt} & 318 \hspace{8pt}	& 13.51\% \\ \hline
-
-\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 452.33 & 1392.33	& 23.21\%\\\cline{3-6}
-    &   &  Median & 418.50 & 1348.50 & 23.38\%  \\\cline{3-6}
-    &   &  Std. Dev & 309.05 & 656.76 & 12.16\%  \\\cline{2-6}
-
- & \multirow{3}{*}{Comby} & Mean & 554.00 & 1055.83	& 29.29\% \\\cline{3-6}
-    &   &  Median & 384.00 & 1059.50	& 25.60\%  \\\cline{3-6}
-    &   &  Std. Dev & 592.63 & 508.30 & 21.63\%  \\\hline
-
-\end{tabular}
-
-}
-\end{table}
-
-\subsection{Universal Source-Based Mutation vs. Previous Approaches}
+Table \ref{tab:mutationtools} shows the set of tools compared.  We
+attempted to identify well-maintained and commonly used tools for each
+language chosen.
 

### 2023-03-28 paper/main.tex
add equivalent % for java, python, rust

@@
 
- & \multirow{3}{*}{Cargo-Mutants} & \multicolumn{2}{c|}{Mean} & 0.33 & N/A\\\cline{3-6}
+ & \multirow{3}{*}{Dextool} & \multicolumn{2}{c|}{Mean} & 0.33 & N/A\\\cline{3-6}
     &   &  \multicolumn{2}{c|}{Median} & 0.29 & N/A \\\cline{3-6}
@@
 
- \multirow{4}{*}{\small pebble} & Regex & 70 & 16 & 0.81 & - \\\cline{2-6}
-    & Comby & 68 & 17 & 0.80 & - \\\cline{2-6}
-    & PIT & 19 & 1 & 0.95 & - \\ \cline{2-6}
-    & LittleDarwin & 8 & 0 & 1.00 & - \\ \hline
-\multirow{4}{*}{\small spring-hateoas} & Regex & 82 & 166 & 0.33 & - \\\cline{2-6}
-    & Comby & 65 & 111 & 0.37 & - \\ \cline{2-6}
-    & PIT & 82 & 32 & 0.28 & - \\ \cline{2-6}
-    & LittleDarwin & 15 & 23 & 0.39 & - \\ \hline
-\multirow{4}{*}{\small javacc} & Regex & 123 & 419 & 0.23 & - \\\cline{2-6}
-    & Comby & 103 & 400 & 0.21 & - \\ \cline{2-6}
-    & PIT & 30 & 113 & 0.21 & -\\ \cline{2-6}
-    & LittleDarwin & 64 & 18 & 0.78 & - \\ \hline
-\multirow{4}{*}{\small coffee-gb} & Regex & 48 & 176 & 0.21 & - \\\cline{2-6}
-    & Comby & 40 & 134 & 0.23 & - \\ \cline{2-6}
-    & PIT & 32 & 4 & 0.89 & - \\ \cline{2-6}
-    & LittleDarwin & 11 & 18 & 0.38 & -\\ \hline
-\multirow{4}{*}{\small egads} & Regex & 442 & 182 & 0.71 & - \\\cline{2-6}
-    & Comby & 422 & 200 & 0.71 & - \\ \cline{2-6}
-    & PIT & 87 & 20 & 0.81 & - \\ \cline{2-6}
-    & LittleDarwin & 74 & 11 & 0.87 & - \\ \hline
-\multirow{4}{*}{\small apk-parser} & Regex & 26 & 325 & 0.07 & \\\cline{2-6}
-    & Comby & 26 & 312 & 0.08 & -\\ \cline{2-6}
-    & PIT & 14 & 75  & 0.16 & -\\ \cline{2-6}
-    & LittleDarwin & 3 & 30 & 0.09 & -\\ \hline
-
-\multirow{12}{*}{Overall} & \multirow{3}{*}{Regex} & \multicolumn{2}{c|}{Mean} & 0.39  & -\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.28 & - \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.30 & - \\\cline{2-6}
-
- & \multirow{3}{*}{Comby} & \multicolumn{2}{c|}{Mean} & 0.40 & -\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.30 & - \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.29 & - \\\cline{2-6}
-
- & \multirow{3}{*}{PIT} & \multicolumn{2}{c|}{Mean} & 0.55 & -\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.55 & - \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.37 & - \\\cline{2-6}
-
- & \multirow{3}{*}{LittleDarwin} & \multicolumn{2}{c|}{Mean} & 0.59 & -\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.59 & - \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.35 & - \\ \hline
+ \multirow{4}{*}{\small pebble} & Regex & 70 & 16 & 0.81 & 50\% \\\cline{2-6}
+    & Comby & 68 & 17 & 0.80 & 20\% \\\cline{2-6}
+    & PIT & 19 & 1 & 0.95 & N/A \\ \cline{2-6}
+    & LittleDarwin & 8 & 0 & 1.00 & 0\% \\ \hline
+\multirow{4}{*}{\small spring-hateoas} & Regex & 82 & 166 & 0.33 & 0\% \\\cline{2-6}
+    & Comby & 65 & 111 & 0.37 & 0\% \\ \cline{2-6}
+    & PIT & 82 & 32 & 0.28 & N/A\% \\ \cline{2-6}
+    & LittleDarwin & 15 & 23 & 0.39 & 0\% \\ \hline
+\multirow{4}{*}{\small javacc} & Regex & 123 & 419 & 0.23 & 0\% \\\cline{2-6}
+    & Comby & 103 & 400 & 0.21 & 0\% \\ \cline{2-6}
+    & PIT & 30 & 113 & 0.21 & N/A\\ \cline{2-6}
+    & LittleDarwin & 64 & 18 & 0.78 & 0\%\\ \hline
+\multirow{4}{*}{\small coffee-gb} & Regex & 48 & 176 & 0.21 & 0\%\\\cline{2-6}
+    & Comby & 40 & 134 & 0.23 & 0\% \\ \cline{2-6}
+    & PIT & 32 & 4 & 0.89 & N/A \\ \cline{2-6}
+    & LittleDarwin & 11 & 18 & 0.38 & 0\%\\ \hline
+\multirow{4}{*}{\small egads} & Regex & 442 & 182 & 0.71 & 20\% \\\cline{2-6}
+    & Comby & 422 & 200 & 0.71 & 10\% \\ \cline{2-6}
+    & PIT & 87 & 20 & 0.81 & N/A \\ \cline{2-6}
+    & LittleDarwin & 74 & 11 & 0.87 & 0\% \\ \hline
+\multirow{4}{*}{\small apk-parser} & Regex & 26 & 325 & 0.07 & 10\%\\\cline{2-6}
+    & Comby & 26 & 312 & 0.08 & 0\%\\ \cline{2-6}
+    & PIT & 14 & 75  & 0.16 & N/A\\ \cline{2-6}
+    & LittleDarwin & 3 & 30 & 0.09 & 0\%\\ \hline
+
+\multirow{12}{*}{Overall} & \multirow{3}{*}{Regex} & \multicolumn{2}{c|}{Mean} & 0.39  & 13.33\%\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.28 & 5.00\% \\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.30 & 19.66\% \\\cline{2-6}
+
+ & \multirow{3}{*}{Comby} & \multicolumn{2}{c|}{Mean} & 0.40 & 5.00\%\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.30 & 0.00\% \\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.29 & 8.37\%\\\cline{2-6}
+
+ & \multirow{3}{*}{PIT} & \multicolumn{2}{c|}{Mean} & 0.55 & N/A\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.55 & N/A \\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.37 & N/A \\\cline{2-6}
+
+ & \multirow{3}{*}{LittleDarwin} & \multicolumn{2}{c|}{Mean} & 0.59 & 0.00\%\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.59 & 0.00\%\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.35 & 0.00\%\\ \hline
 
@@
 
- \multirow{4}{*}{\small keract} & Regex & 463 & 1192 & 0.28 & - \\\cline{2-6}
-    & Comby & 383 & 1088 & 0.26 & - \\\cline{2-6}
-    & Mutmut & 349 & 133 & 0.72 & - \\ \cline{2-6}
-    & CosmicRay & 250 & 671 & 0.27 & - \\ \hline
-\multirow{3}{*}{\small dtw} & Regex & 506 & 467 & 0.52 & - \\\cline{2-6}
-    & Comby & 475 & 391 & 0.55 & - \\ \cline{2-6}
-    & Mutmut & 118 & 152 & 0.44 & - \\ \cline{2-6}
-    & CosmicRay & 328 & 385 & 0.46 & - \\ \hline
-\multirow{3}{*}{\small notion-sdk-py} & Regex & 112 & 84 & 0.57 & - \\\cline{2-6}
-    & Comby & 118 & 78 & 0.60 & - \\ \cline{2-6}
-    & Mutmut & 41 & 22 & 0.65 & -\\ \cline{2-6}
-    & CosmicRay & 23 & 29 & 0.44 & - \\ \hline
-
-\multirow{3}{*}{\makecell{\small atlassian-python \\ api}} & Regex & 193 & 109 & 0.64 & - \\\cline{2-6}
-    & Comby & 207 & 126 & 0.62 & - \\ \cline{2-6}
-    & Mutmut & 123 & 51 & 0.71 & - \\ \cline{2-6}
-    & CosmicRay & 78 & 28 & 0.74 & - \\ \hline
-\multirow{3}{*}{\small ESD} & Regex & 132 & 1727 & 0.07 & - \\\cline{2-6}
-    & Comby & 121 & 1591 & 0.07 & - \\ \cline{2-6}
-    & Mutmut & 57 & 663 & 0.08 & - \\ \cline{2-6}
-    & CosmicRay & 63 & 950 & 0.06 & - \\ \hline
-\multirow{3}{*}{\small fiber} & Regex & 307 & 186 & 0.62 & \\\cline{2-6}
-    & Comby & 291 & 167 & 0.64 & -\\ \cline{2-6}
-    & Mutmut & 127 & 95 & 0.57 & -\\ \cline{2-6}
-    & CosmicRay & 259 & 61 & 0.81 & -\\ \hline
-
-\multirow{12}{*}{Overall} & \multirow{3}{*}{Regex} & \multicolumn{2}{c|}{Mean} & 0.45  & -\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.55 & - \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.23 & - \\\cline{2-6}
-
- & \multirow{3}{*}{Comby} & \multicolumn{2}{c|}{Mean} & 0.46 & -\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.58 & - \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.24 & - \\\cline{2-6}
-
- & \multirow{3}{*}{Mutmut} & \multicolumn{2}{c|}{Mean} & 0.53 & -\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.61 & - \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.24 & - \\\cline{2-6}
-
- & \multirow{3}{*}{CosmicRay} & \multicolumn{2}{c|}{Mean} & 0.46 & -\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.45 & - \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.28 & - \\ \hline
+ \multirow{4}{*}{\small keract} & Regex & 463 & 1192 & 0.28 & 0\% \\\cline{2-6}
+    & Comby & 383 & 1088 & 0.26 & 10\% \\\cline{2-6}
+    & Mutmut & 349 & 133 & 0.72 & 10\% \\ \cline{2-6}
+    & CosmicRay & 250 & 671 & 0.27 & 10\% \\ \hline
+\multirow{3}{*}{\small dtw} & Regex & 506 & 467 & 0.52 & 10\% \\\cline{2-6}
+    & Comby & 475 & 391 & 0.55 & 10\% \\ \cline{2-6}
+    & Mutmut & 118 & 152 & 0.44 & 0\% \\ \cline{2-6}
+    & CosmicRay & 328 & 385 & 0.46 & 20\% \\ \hline
+\multirow{3}{*}{\small notion-sdk-py} & Regex & 112 & 84 & 0.57 & 10\% \\\cline{2-6}
+    & Comby & 118 & 78 & 0.60 & 20\% \\ \cline{2-6}
+    & Mutmut & 41 & 22 & 0.65 & 0\%\\ \cline{2-6}
+    & CosmicRay & 23 & 29 & 0.44 & 0\%\\ \hline
+
+\multirow{3}{*}{\makecell{\small atlassian-python \\ api}} & Regex & 193 & 109 & 0.64 & 20\%\\\cline{2-6}
+    & Comby & 207 & 126 & 0.62 & 40\% \\ \cline{2-6}
+    & Mutmut & 123 & 51 & 0.71 & 10\% \\ \cline{2-6}
+    & CosmicRay & 78 & 28 & 0.74 & 10\% \\ \hline
+\multirow{3}{*}{\small ESD} & Regex & 132 & 1727 & 0.07 & 30\%\\\cline{2-6}
+    & Comby & 121 & 1591 & 0.07 & 10\%\\ \cline{2-6}
+    & Mutmut & 57 & 663 & 0.08 & 0\%\\ \cline{2-6}
+    & CosmicRay & 63 & 950 & 0.06 & 10\%\\ \hline
+\multirow{3}{*}{\small fiber} & Regex & 307 & 186 & 0.62 & 10\%\\\cline{2-6}
+    & Comby & 291 & 167 & 0.64 & 20\%\\ \cline{2-6}
+    & Mutmut & 127 & 95 & 0.57 & 40\%\\ \cline{2-6}
+    & CosmicRay & 259 & 61 & 0.81 & 0\%\\ \hline
+
+\multirow{12}{*}{Overall} & \multirow{3}{*}{Regex} & \multicolumn{2}{c|}{Mean} & 0.45  & 13.00\%\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.55 & 10.00\%\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.23 & 10.33\%\\\cline{2-6}
+
+ & \multirow{3}{*}{Comby} & \multicolumn{2}{c|}{Mean} & 0.46 & 18.00\%\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.58 & 15.00\%\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.24 & 11.69\%\\\cline{2-6}
+
+ & \multirow{3}{*}{Mutmut} & \multicolumn{2}{c|}{Mean} & 0.53 & 10.00\%\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.61 & 5.00\%\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.24 & 15.49\%\\\cline{2-6}
+
+ & \multirow{3}{*}{CosmicRay} & \multicolumn{2}{c|}{Mean} & 0.46 & 8.33\%\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.45 & 10.00\%\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.28 & 7.53\%\\ \hline
 
@@
 
- \multirow{3}{*}{\small cargo release} & Regex & 721 & 130 & 0.85 & - \\\cline{2-6}
-    & Comby & 1344 & 346 & 0.80 & - \\\cline{2-6}
-    & Cargo-Mutants & 2	& 3 & 0.40 & - \\ \hline
-\multirow{3}{*}{\small passerine} & Regex & 122 & 91 & 0.57 & - \\\cline{2-6}
-    & Comby & 90 & 83 & 0.52 & - \\ \cline{2-6}
-    & Cargo-Mutants & 0	& 0 & 0 & - \\ \hline
-\multirow{3}{*}{\small typos} & Regex & 295 & 91 & 0.76 & - \\\cline{2-6}
-    & Comby & 283 & 83 & 0.77 & - \\ \cline{2-6}
-    & Cargo-Mutants & 5 & 1 & 0.83 & -\\ \hline
-
-\multirow{3}{*}{\small ord} & Regex & 558 & 204 & 0.73 & - \\\cline{2-6}
-    & Comby & 533 & 110 & 0.83 & - \\ \cline{2-6}
-    & Cargo-Mutants & 6 & 0 & 1.00 & - \\ \hline
-\multirow{3}{*}{\small bazuka} & Regex & 281 & 170 & 0.62 & - \\\cline{2-6}
-    & Comby & 291 & 111 & 0.73 & - \\ \cline{2-6}
-    & Cargo-Mutants & 0 & 0 & 0 & - \\ \hline
-\multirow{3}{*}{\small strum} & Regex & 8 & 43 & 0.16 & \\\cline{2-6}
-    & Comby & 11 & 39 & 0.22 & -\\ \cline{2-6}
-    & Cargo-Mutants & 2 & 0 & 1.00 & -\\ \hline
-
-\multirow{9}{*}{Overall} & \multirow{3}{*}{Regex} & \multicolumn{2}{c|}{Mean} & 0.62  & -\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.68 & - \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.24 & - \\\cline{2-6}
-
- & \multirow{3}{*}{Comby} & \multicolumn{2}{c|}{Mean} & 0.64 & -\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.75 & - \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.25 & - \\\cline{2-6}
-
- & \multirow{3}{*}{Cargo-Mutants} & \multicolumn{2}{c|}{Mean} & 0.54 & -\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.62 & - \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.47 & - \\ \hline
+ \multirow{3}{*}{\small cargo release} & Regex & 721 & 130 & 0.85 & 0\% \\\cline{2-6}
+    & Comby & 1344 & 346 & 0.80 & 0\% \\\cline{2-6}
+    & Cargo-Mutants & 2	& 3 & 0.40 & 0\% \\ \hline
+\multirow{3}{*}{\small passerine} & Regex & 122 & 91 & 0.57 & 20\% \\\cline{2-6}
+    & Comby & 90 & 83 & 0.52 & 50\% \\ \cline{2-6}
+    & Cargo-Mutants & 0	& 0 & 0 & 0\% \\ \hline
+\multirow{3}{*}{\small typos} & Regex & 295 & 91 & 0.76 & 40\% \\\cline{2-6}
+    & Comby & 283 & 83 & 0.77 & 20\%\\ \cline{2-6}
+    & Cargo-Mutants & 5 & 1 & 0.83 & 0\%\\ \hline
+
+\multirow{3}{*}{\small ord} & Regex & 558 & 204 & 0.73 & 10\% \\\cline{2-6}
+    & Comby & 533 & 110 & 0.83 & 20\% \\ \cline{2-6}
+    & Cargo-Mutants & 6 & 0 & 1.00 & 0\% \\ \hline
+\multirow{3}{*}{\small bazuka} & Regex & 281 & 170 & 0.62 & 40\% \\\cline{2-6}
+    & Comby & 291 & 111 & 0.73 & 0\% \\ \cline{2-6}
+    & Cargo-Mutants & 0 & 0 & 0 & 0\% \\ \hline
+\multirow{3}{*}{\small strum} & Regex & 8 & 43 & 0.16 & 10\%\\\cline{2-6}
+    & Comby & 11 & 39 & 0.22 & 40\%\\ \cline{2-6}
+    & Cargo-Mutants & 2 & 0 & 1.00 & 0\%\\ \hline
+
+\multirow{9}{*}{Overall} & \multirow{3}{*}{Regex} & \multicolumn{2}{c|}{Mean} & 0.62  & 20.00\%\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.68 & 15.00\%\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.24 & 16.73\% \\\cline{2-6}
+
+ & \multirow{3}{*}{Comby} & \multicolumn{2}{c|}{Mean} & 0.64 & 22.00\%\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.75 & 20.00\% \\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.25 & 20.41\% \\\cline{2-6}
+
+ & \multirow{3}{*}{Cargo-Mutants} & \multicolumn{2}{c|}{Mean} & 0.54 & 0.00\%\\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Median} & 0.62 & 0.00\% \\\cline{3-6}
+    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.47 & 0.00\%\\ \hline
 

### 2023-03-28 paper/main.tex
alex so far

@@
 lines; this is the total for comby and regexp-based rules, so each
-version requires only 218 lines of rule definitions.
+version requires only 218 lines of rule definitions.  The tool makes
+use of parser parser combinators via the {\tt comby-python} module.
 
@@
 testing infrastructure, or an older JUnit, can be difficult.  Our
-approach also is suitable for use in CI and other automated environments.
+approach also is suitable for use in CI and other automated
+environments.
+
+The tool attempts to automatically identify which rule sets to use in
+mutating a source file, but can also be instructed to run fewer or
+more rule sets, and allows users to supply their own custom rule sets.
 
@@
 default handlers can always be over-ridden if they do not work with a
-particular build setup.
-
-Our tool has been available as an open source project on GitHub and
+particular build setup (for Python, the approach almost always works,
+but in the case of multi-file projects in other languages, it ususally
+requires some manual setup).
+
+The tool also provides a number of utilities, including mechanisms for
+incorporating code coverage into mutant selection, integration with a
+Python automated testing tool, prioritization of unkilled mutants to
+show dis-similar mutants early in a ranking (similar to the FPF
+approach to identifying unique bugs found in fuzzing
+\cite{10.1145/2491956.2462173}), and tools for pruning a class of
+mutants based on some criteria (e.g., removing all mutants of logging
+statements from a set of generated mutants).
+
+Our tool has been available as an open source project on GitHub (with
+over 100 stars to date) and
 available for installation via Python pip for
 some time, has received contributions from users not involved in
-our project, and been adopted by at least one very large project as
-the officially recommended mutation tool.  We omit details for
+our project, and has been adopted by at least one flagship project
+(over 19K stars) as
+its officially recommended mutation tool.  We omit details for
 purposes of blinding.
@@
 mutants for a set of four programming languages: C++, Java, Python,
-and Rust.  We picked six projects with 1) a large number of GitHub
-stars and 2) an easy-to-execute, passing test suite for each language,
-and then chose the largest source file below 20KB in size from each project to mutate.  Limiting
-mutation to one large but not extremely large file made is easier to compare generated mutants and
-manually inspect for mutant equivalency.  The only bias to file
-selection was size, which we do not believe should affect our results
-significantly.
+and Rust.  The criteria for language selection was as follows: C++,
+Java, and Python are among the most widely used programming languages
+at present.  Furthermore, C++, unlike C, can be notoriously hard to
+parse, making development of mutation tools that use an AST and
+properly consider mutants of, e.g., code inside a template,
+difficult.  Java is the language in which mutation testing has been
+studied most extensively in recent literature, and PIT is almost
+certainly the most popular and widely used mutation tool.  Python long
+suffered from a lack of working and maintained mutation tools, despite
+a large number of abandoned efforts, though at present it is served by
+multiple maintained, AST-based tools.  Finally, Rust is an ``emerging
+language'' with few tools targeting it.
+
+To evaluate both RQ1 and RQ2 we assembled a set of 24 source files,
+six from each language.  We first selected, using the GitHub search
+API and manual examination, six GitHub
+projects for each language meeting a set of criteria:
+
+\begin{enumerate}
+\item The project must have a minimum of 1,000 stars.
+  \item The percentage of source languages other than the target
+    language must be small.
+  \item There must be an executable test suite where all tests pass.
+    \item The project must be relatively easy to build, without a
+      large set of external dependencies.
+      \item In addition, for Java, we required use of maven.
+      \end{enumerate}
+
+      Files were then selected by picking for each project, the
+      largest source file that was less than 20KB in size.  If that
+      file was no covered by the test suite, the next closest to 20KB in order
+      was chosen.  Finally, if no source file under 20KB existed or
+      was covered by tests, the smallest file larger than 20KB covered
+      by tests was selected.
 
@@
 case study, applying our implementation to a program in a language for
-which it has \emph{no} rules, Haskell.
+which it has \emph{no} rules, Haskell.  The case study concerns a
+fairly trivial example, but was chosen due to being the only program
+for which data was available as to the effectiveness of the tool we
+compared against (since that tool no longer works).
 
 \subsection{Regular Expressions vs. Parser Parser Combinators}
+
+While the primary advantages of using parser parser combinators as a
+basis for mutant generation are in expressiveness and ease of use,
+particularly rule readability, a second advantage is a reduction in
+the number of invalid mutants generated.
+
+Tables~\ref{tab:table_cpp1}-\ref{tab:table_rust1} compare generated
+mutants for ``Regex'' and ``Comby'' modes of our tool.  Invalid
+mutants are mutants that the tool produces, but which fail to
+compile.  The cost of each such mutant is usually small,
+since compilation generally fails early for trivial reasons (e.g.,
+{\tt break} or {\tt continue} outside a loop), 
+but reducing the number of such mutants directly reduces the cost of
+initial mutant generation, and generally indicates a better
+``understanding'' of the source being mutated.
 
@@
 \end{table}
+
+The mean and median percentage of mutants that are valid is higher for
+Comby for all languages.  On average, even for a single modestly sized
+file, using parser combinator combinators saves 100-600 failed
+compilations.
+
+The number of valid mutants is usually
+slightly smaller for Comby, but Comby actually produces a larger mean
+(but not median) number of mutants for Rust, due to one project where
+Comby produces a much larger number of valid mutants.  As discussed 
+below, mutation scores and mutant equivalence rates are similar for
+both modes.
+
+In sum, in addition to providing a major advantage in expressiveness
+(the ability to easily define complex multi-line mutations that
+preserve structure, e.g. going beyond statement deletion to block
+deletion), parser parser combinators offer a notable gain in the
+efficiency of initial mutant generation, with no apparent loss of overall effectiveness.
 
@@
     C++ & Dextool & 38611 \\\hline
-      ALL & universalmutator & 2680 \\\hline
+    \hline
+      ALL & Our Tool & 2680 \\\hline
   \end{tabular}
@@
 Table \ref{tab:mutationtools} shows the set of tools compared.  We
-attempted to identify well-maintained and commonly used tools for each
-language chosen.
+attempted to identify well-maintained and commonly used open
+source tools for each
+language chosen.  The tools cover a number of paradigms; for the most
+part we did not choose bytecode based tools, since these often make it
+difficult or impossible to mutate a single source file.  PIT, however,
+provides this functionality and is perhaps the paradigmatic bytecode
+based mutation tool.  Only Dextool, for C++, implements the
+\emph{mutant schema} \cite{untch1993mutation} approach often advocated
+in the research literature, where a single metaprogram is produced
+and, e.g., environment variabes are used to control which particular
+version of a program actually executes.  Using a schema has the major
+advantage of dramatically reducing the time required to compile
+mutants.  We speculate, however, that the fact that the approach is
+somewhat complex to implement for some kinds of mutants, and
+furthermore fails if even one invalid mutant is generated, has
+prevented most developers of non-academic tools from adopting it.
+
+The most striking thing about this table, and a strong support for
+the central argument of this paper, is that our implementation, which
+covers \emph{all the languages considered plus a number of additional
+languages}, is \emph{the smallest of the tools by a margin of over 1,000 LOC},
+and can be extended to handle additional languages with minimal
+effort.  Each currently supported language requires about 15 rules on
+average, though these must be implemented as both Regex and Comby
+rules if both modes are to be used for the language.
+
+The central question is: does using a universal approach produce
+notably less effective mutation testing?  Or, to put it another way,
+does the substantial effort required to develop and maintain a
+mutation tool for a single language pay off in obvious advantages in
+mutation testing results?
+Tables~\ref{tab:table_cpp2}-\ref{tab:table_rust2} show the results of
+mutation testing for the generated mutants for our tool, in both Regex
+and Comby modes, and the above-listed tools for each of the four
+studied languages.
+
+We estimated mutant equivalency when possible by examining a sample of
+10 randomly selected unkilled mutants for each file.  For some tools,
+human inspection of the individual unkilled mutants was not feasible.
 
@@
 \centering
-\caption{C++ (UniversalMutator vs. Dextool)}
+\caption{C++ (Our Implementation vs. Dextool)}
 \label{tab:table_cpp2}
@@
 \centering
-\caption{Java (UniversalMutator vs. PIT vs. LittleDarwin)}
+\caption{Java (Our Implementation vs. PIT vs. LittleDarwin)}
 \label{tab:table_java2}
@@
 \centering
-\caption{Python (UniversalMutator vs. Mutmut vs. CosmicRay)}
+\caption{Python (Our Implementation vs. Mutmut vs. CosmicRay)}
 \label{tab:table_python2}
@@
 \centering
-\caption{Rust (UniversalMutator vs. Cargo-Mutants)}
+\caption{Rust (Our Implementation vs. Cargo-Mutants)}
 \label{tab:table_rust2}
@@
 \end{table}
+
+Previous studies comparing mutation tools for Java
+\cite{MajorPIT,gopinath2017does,CommACMJavaTool} found that results
+for individual projects often varied considerably across tools without
+indicating a clear advantage for one tool over another.  Our primary
+criteria for comparing tools is whether there is an overall pattern of
+clearly less-effective mutation for our approach.  Below, we discuss
+the question for each language separately.  Each section also
+addresses briefly the question of whether Regex and Comby modes
+perform differently in terms of mutation score and equivalent mutants,
+which is relevant to RQ1.
+
+\subsubsection{C++}
+
+We found only one tool that allowed single-file mutation and worked
+reliably for C++.  Dextool generally produced fewer mutants than
+either mode of our tool, and for some projects produced an extremely
+small number of mutants.  For example, for the {\tt
+  tiny-differentiable-simulator} project, Dextool only produced a
+total of 61 mutants, and gave a much higher mutation score than our
+two modes.  Our tools produced more than 600 additional unkilled
+mutants, and none of the inspected mutants were equivalent.
+Similarly, Dextool only found 2 unkilled mutants for the {\tt
+  libgraphqlparser} project.  We doubt that the set of test weaknesses
+in the code is adequately represented by such a small number of
+mutants when our tools were able to produce significantly more
+unkilled mutants.  Even assuming Dextool's mutants are never
+equivalent (highly unlikely) and supposing that a third of our
+generated mutants are equivalent, it seems likely that Dextool is
+missing important mutants in some cases.  Overall, mean mutation
+scores were similar for Dextool and both modes of our tool, while
+median mutation scores were much lower for our tools, which produced
+many more unkilled mutants for some projects.  For C++, there seems to
+be no reason to suspect that the universal approach is producing
+sub-par mutation testing.
+
+Comparing Regex and Comby modes, mutation scores are very similar, as
+are estimated equivalency rates, though Comby produces a somewhat
+larger number and percentage of equivalent mutants.
+
+\subsubsection{Java}
+
+\subsubsection{Python}
+
+\subsubsection{Rust}
 

### 2023-03-28 paper/main.tex
Adding URLs

--- 2023-03-28 paper/main.tex
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Syntax Is All You Need: A Universal-Language Approach to Mutant Generation}
+\title{Syntax Is All You Need: A Universal-Language\\Approach to Mutant Generation}
 
@@
 approach to mutation to existing tools, and demonstrate the advantages
-of using parser parser combinators to improve on a simple regular-expression
+of using parser-parser combinators to improve on a simple regular-expression
 based approach to generation.  
@@
   to parse C++ 2020, and is particularly problematic for somewhat
-  fast-moving languages such as Rust.}.  There
-appears to be a significant risk in relying on any mutation tools other
+  fast-moving languages such as Rust.}.  Of course, many open source
+projects are abandoned, and academic software has a long history of
+being available and supported only so long as a graduate student is
+working on a thesis or a particular grant is being funded.  Still,
+there appears to be a significant risk in relying on any mutation tools other
 than the most widely used and well-supported, such as PIT.
@@
 tool; e.g., it was not until 2014 that Haskell had such a tool
-\cite{mucheck}.  Furthermore, that tool is no longer
-supported\footnote{The last commits or package updates were in 2015, and the
-primary author confirms that changes to {\tt haskell-src-ext} since
-introduced render it unusable at present.}, and it
-appears that no successor tool has appeared in the ensuing
-years\footnote{There is FitSpec \cite{FitSpec}, which is maintained,
-  but FitSpec does not perform traditional mutation testing or produce
-  source code (or bytecode) mutants; it rather
-  performs black-box type-aware mutation of function return values.}.
-Haskell developers (until now, see below) have no option for
-performing true mutation testing.
+\cite{mucheck}.  Furthermore, that tool may no longer  be
+supported (the last commits or package updates were in 2015), and it
+is not clear that a successor tool has appeared in the ensuing years.
 
@@
     would introduce new mutation operators or modify existing ones.
-    The local nature of operators tends to make the second case very rare.
+    The local nature of operators tends to make the second case in
+    particular very rare.
     \item The existing literature on multi-lingual or language
@@
 
-          \item A proposal to use parser parser combinators to improve the
+          \item A proposal to use parser-parser combinators to improve the
       efficiency and expressive power of that approach, and an
@@
 
-\begin{table}[hbtp]
-\centering
-\caption{Rule Examples}
-\label{tab:rules}
-\resizebox{\columnwidth}{!}{%
-
-  \begin{tabular}{|l|l|l|}
-    \hline
-    Universal & C-Like & Language-Specific \\
-    \hline
-{\tt < ==> >} &{\tt  else ==>} & {\tt synchronized ==>} \\ 
-{\tt < ==> ==} & {\tt if (\textbackslash(.*\textbackslash)) ==> if (!\textbackslash 1)} &
-                                                                 {\tt (\^{}\textbackslash s*)(\textbackslash S+.*\textbackslash n) ==> \textbackslash 1pass}\\
-{\tt < ==> <=} & {\tt if (\textbackslash(.*\textbackslash)) ==> if
-                 (1==1)} & {\tt any\_of ==> all\_of}\\
-\hline
-
-\end{tabular}
-}
-
-\end{table}
-
-  Table~\ref{tab:rules} shows examples of universal, C-like, and
-  language-specific rules for a few languages, taken directly from our
-  implementation.  In all cases, the form of a rule is: {\tt regexp
-    ==> replacement}, using Python regular expression and replacement syntax.
-
-  The first column shows universal rules applied to
-  all languages, here drawn from the large set of transformations for
-  less-than comparisons.  The second column shows somewhat more
-  complex rules, for removing elses (which always leaves valid code in
-  C-like languages, but with the else clause always executing) and for
-  replacing if conditions with, respectively, a negation and a
-  constantly true expression.  Finally, the last column shows
-  language-specific rules, for, respectively, Java, Python (statement
-  deletion, including a {\tt pass} and respecting indentation level),
-  and C++.  The last rule was contributed by a user using our tool to
-  perform mutation testing on the bitcoin core implementation.  For a
-  C++ program, all of the rules in the first two columns will be
-  applied, as well as the last rule in the third column.  For a Python
-  program, however, only the first column of rules and the second rule
-  in the last column apply, as Python is not sufficiently ``C-like''.
+  Figure~\ref{fig:rules} shows examples of universal, C-like, and
+  language specific rules for a few languages, taken directly from our
+  implementation.
 
@@
 
-We implemented our approach in a mutation testing tool, written in
-about 2,200 lines of Python.  The rules defining mutation operators
-for a set of languages including C, C++, Python, Java, Lisp, Swift,
-Rust, Solidity, Fe, Vyper, JavaScript, and Go, required another 436
-lines; this is the total for comby and regexp-based rules, so each
-version requires only 218 lines of rule definitions.
-
-While many tools provide a GUI front-end, or integrate to a particular
-testing library, we chose to focus on an easily configured CLI,
-allowing a user to specify the commmands used to build mutants or
-run tests.  While less convenient in the case where there is a good
-fit between a tool's integrations and a project's environment, we
-found this flexibility useful even when limiting considerations to a
-single language.  For example, the widely used PIT system is strongly tied to
-certain versions of JUnit, and mutating projects using a different
-testing infrastructure, or an older JUnit, can be difficult.  Our
-approach also is suitable for use in CI and other automated environments.
-
-For the most part we expect users to provide build commands, but do
-support automatic compilation and comparison of mutants for Python,
-Solidity, Fe, Vyper, Swift, Rust, and Java, including Trivial Compiler
-Equivalence \cite{TCE} checks for redundancy of mutants.  These
-default handlers can always be over-ridden if they do not work with a
-particular build setup.
-
-Our tool has been available as an open source project on GitHub and
-available for installation via Python pip for
-some time, has received contributions from users not involved in
-our project, and been adopted by at least one very large project as
-the officially recommended mutation tool.  We omit details for
-purposes of blinding.
+Our approach also makes introducing advanced techniques such as
+Trivial Compiler Equivalence (TCE) into a mutation workflow easy.
+While we support TCE for some languages inside the tool, a user can
+always write a simple script to check...
 
@@
 
-  \item{RQ1:}  How does the use of parser parser combinators modify 
+  \item{RQ1:}  How does the use of parser-parser combinators modify 
     the efficiency (in terms of invalid mutants) and effectiveness (in
@@
 \item{RQ2:}  How does the univeral approach, using regular expressions
-  or parser parser combinator rules, compare to existing
+  or parser-combinator combinator rules, compare to existing
   single-language mutation tools, in terms of number of generated
@@
 
-Both research questions were addressed by generating and running
-mutants for a set of four programming languages: C++, Java, Python,
-and Rust.  We picked six projects with 1) a large number of GitHub
-stars and 2) an easy-to-execute, passing test suite for each language,
-and then chose the largest source file below 20KB in size from each project to mutate.  Limiting
-mutation to one large but not extremely large file made is easier to compare generated mutants and
-manually inspect for mutant equivalency.  The only bias to file
-selection was size, which we do not believe should affect our results
-significantly.
-
-We additionally present a small
+In addition to these core research questions, evaluated over six files
+from six real-world projects for four languages, we also present a
 case study, applying our implementation to a program in a language for
-which it has \emph{no} rules, Haskell.
+which it has no rules, Haskell.
 
@@
 
+\begin{table*}[hbtp]
+  \centering
+  \caption{Overview of Mutation Testing Tools. universalmutator has singificantly fewer lines of code than any of the single language tools.}
+  \label{tab:mutationtools}
+
+  \begin{tabular}{|c|c|c|r|}
+  \hline
+  \textbf{Language} & \textbf{Tool} & \textbf{URL} & \textbf{LOC}  \\
+  \hline
+  \multirow{2}{*}{Python}  & Mutmut & https://github.com/boxed/mutmut & 3870  \\\cline{2-4}
+      & cosmic-ray & https://github.com/sixty-north/cosmic-ray & 4599 \\ \hline
+  \multirow{2}{*}{Java}  & PIT & https://github.com/hcoles/pitest & 59577  \\\cline{2-4}
+      & LittleDarwin & https://github.com/aliparsai/LittleDarwin & 22359 \\ \hline
+      Rust & Cargo-Mutants & https://github.com/sourcefrog/cargo-mutants & 7020 \\\hline
+      C++ & Dextool & https://github.com/joakim-brannstrom/dextool & 38611 \\\hline
+      ALL & universalmutator & https://github.com/agroce/universalmutator & 2244 \\\hline
+  \end{tabular}
+  \end{table*}
+
 % ---------------- JAVA table UM Only--------------------------
@@
 \subsection{Universal Source-Based Mutation vs. Previous Approaches}
-
-\begin{table}[hbtp]
-  \centering
-  \caption{Overview of Mutation Testing Tools.}
-  \label{tab:mutationtools}
-
-  \begin{tabular}{|c|c|r|}
-  \hline
-  \textbf{Language} & \textbf{Tool} & \textbf{LOC}  \\
-  \hline
-  \multirow{2}{*}{Python}  & Mutmut &	3870  \\\cline{2-3}
-      & cosmic-ray & 4599 \\ \hline
-  \multirow{2}{*}{Java}  & PIT &	59577  \\\cline{2-3}
-      & LittleDarwin & 22359 \\ \hline
-      Rust & Cargo-Mutants & 7020 \\\hline
-      C++ & Dextool & 38611 \\\hline
-      ALL & universalmutator & 2680 \\\hline
-  \end{tabular}
-\end{table}
-
-Table \ref{tab:mutationtools} shows the set of tools compared.  We
-attempted to identify well-maintained and commonly used tools for each
-language chosen.
 
@@
 \begin{figure}
-
 \begin{code}
@@
 \caption{Example from MuCheck Paper}
-\label{fig:haskell}
-
+\label{fig;haskell}
 \end{figure}
@@
 valid mutants).  Comby produces an identical 10 valid mutants, but 65
-invalid mutants (13.33\% validity).  Both tools produce a large number
-of invalid mutants here due to including rules such as adding {\tt
-  break} and {\tt continue} that are sufficiently widespread in
-applicability to be included in our ``universal'' set but do not apply
-to Haskell; we believe that comby's awareness of Haskell syntax in
-this case actually gives it \emph{more} opportunities to apply such
-inapplicable mutants.
-
-These mutants produce a mutation
+invalid mutants (13.33\% validity).  These mutants produce a mutation
 score of 0.8 for the two properties defined in the original MuCheck
@@
 programs over four languages,
-and introduces the idea of using parser parser combinators to improve
+and introduces the idea of using parser-parser combinators to improve
 on regular expression-based generation.

### 2023-03-28 paper/main.tex

--- 2023-03-28 paper/main.tex
@@
 
-\begin{table}[hbtp]
+\begin{table*}[hbtp]
     \centering
-  \caption{Overview of Mutation Testing Tools.}
+    \caption{Overview of Mutation Testing Tools. universalmutator has singificantly fewer lines of code than any of the single language tools.}
     \label{tab:mutationtools}
 
-  \begin{tabular}{|c|c|r|}
+    \begin{tabular}{|c|c|c|r|}
     \hline
-  \textbf{Language} & \textbf{Tool} & \textbf{LOC}  \\
+    \textbf{Language} & \textbf{Tool} & \textbf{URL} & \textbf{LOC}  \\
     \hline
-  \multirow{2}{*}{Python}  & Mutmut &	3870  \\\cline{2-3}
-      & cosmic-ray & 4599 \\ \hline
-  \multirow{2}{*}{Java}  & PIT &	59577  \\\cline{2-3}
-      & LittleDarwin & 22359 \\ \hline
-      Rust & Cargo-Mutants & 7020 \\\hline
-    C++ & Dextool & 38611 \\\hline
-    \hline
-      ALL & Our Tool & 2680 \\\hline
+    \multirow{2}{*}{Python}  & Mutmut & https://github.com/boxed/mutmut & 3870  \\\cline{2-4}
+        & cosmic-ray & https://github.com/sixty-north/cosmic-ray & 4599 \\ \hline
+    \multirow{2}{*}{Java}  & PIT & https://github.com/hcoles/pitest & 59577  \\\cline{2-4}
+        & LittleDarwin & https://github.com/aliparsai/LittleDarwin & 22359 \\ \hline
+        Rust & Cargo-Mutants & https://github.com/sourcefrog/cargo-mutants & 7020 \\\hline
+        C++ & Dextool & https://github.com/joakim-brannstrom/dextool & 38611 \\\hline
+        ALL & universalmutator & https://github.com/agroce/universalmutator & 2244 \\\hline
     \end{tabular}
-\end{table}
+    \end{table*}
 

### 2023-03-28 paper/main.tex
alex exp done

--- 2023-03-28 paper/main.tex
@@
 
+For Python, where TCE based on Python bytecode was possible, the
+number of redundant (equivalent to the original bytecode, or
+equivalent to the bytecode of an already-generated mutant) mutants was
+overall similar for both Regex and Comby, though varied widely by file.
+
 In sum, in addition to providing a major advantage in expressiveness
@@
 
+We do not report the time taken for mutation generation and testing,
+since due to only Dextools using schema (and PIT using bytecode mutants), most tools took similar
+amounts of time to compile mutants, and mutant execution time was
+generally similar, except for tools that produced very few mutants.
+We expect that our approach is generally slower for initial mutant
+generation, due to a higher number of invalid mutants, but that this
+cost is relatively small compared to the time required to compile and
+execute valid mutants for almost all projects.
+
 We estimated mutant equivalency when possible by examining a sample of
@@
 human inspection of the individual unkilled mutants was not feasible.
+However, we were of course able to estimtate the percent of unkilled
+mutants that were unkillable for our two modes for all files, and
+these numbers can be put in the perspective of rates reported in the
+literature.  Reports of equivalent mutant percentages in the
+literature sometimes report the percent of \emph{all} mutants
+(including killed mutants, which obviously cannot be equivalent) that
+are equivalent.  We follow Schuler and Zeller \cite{EquivMut} in
+reporting the percentage of \emph{unkilled} mutants that were found to
+be equivalent.  Schuler and Zeller found about 45\% of unkilled
+mutants of Java programs to be equivalent.  Other reports of lower
+rates, e.g, 9\% \cite{offutt1997automatically} or 7\% \cite{TCE}, report the percent of
+\emph{all} mutants, and so translate to much higher rates of
+equivalency for unkilled mutants.  We therefore suggest that as a
+fraction of unkilled mutants, a rate of 30\% or fewer equivalent
+mutants is good, and a rate below 10\% is very good.
 
@@
 are estimated equivalency rates, though Comby produces a somewhat
-larger number and percentage of equivalent mutants.
+larger number and percentage of equivalent mutants.  Both modes
+produce relatively low mean percentages of equivalent mutants.
 
@@
 
+For Java, we compare to PIT, arguably the standard for mutation
+testing tools in real-world usage (and a tool very frequently used in
+the literature) as well as LittleDarwin, a source-based tool.  Our
+tool produces lower but comparable mutation scores, and usually
+acceptable or excellent equivalency rates.  LittleDarwin produces
+\emph{no} equivalent mutants, but is a highly conservative tool, only
+implementing a small set of operators, and not including statement
+deletion.  Our tools always produced more non-equivalent unkilled
+mutants than PIT by a large margin.
+
+Mutation scores for both modes of our tool were again very similar.
+For Java, Comby shows a significant advantage in terms of equivalent
+mutants, with very low equivalency rates compared to Regex.
+
 \subsubsection{Python}
 
+Our comparison of Python mutation tools showed the largest degree of
+similarity between all tools.  Muation scores and equivalency rates
+were all, given the large variance expected from previous studies of
+Java tools, quite similar.  Again, there is no clear disadvantage to
+using our approach, and a large advantage in terms of development and
+maintenance ease.  The similarity extends to the two modes of our
+tool, though Comby had a somewhat higher equivalency rate than the
+other tools, including the Regexp mode of our tool.  We discuss
+possible reasons for this below.
+
 \subsubsection{Rust}
+
+Finally, for Rust, we observed that the Cargo-Mutants tool simply does
+not generate many mutants.  Mutation scores were broadly similar to
+ours, though lower, and none of the generated mutants we examined were
+equivalent, but for one project Cargo-Mutants did not produce
+\emph{any} mutants even though our tool was able to produce a large
+number of non-equivalent unkilled mutants.  Again, behavior of the two
+modes of our tool was similar, with both having acceptable equivalency
+rates and similar mutation scores.
+
+\subsubsection{Discussion: Summary of Tool Comparisons}
+
+Our tools produce somewhat lower mutation scores than other tools, and
+higher (though still generally good) equivalency rates than some
+tools, but clearly fall into the same broad similarity of behavior
+reported for high-quality mutation tools in the literature comparing
+Java mutation tools.   The tools with very low or zero equivalency rates, furthermore,
+are very conservative in mutation generation, and often simply don't
+produce many mutants.
+
+We speculate that our somewhat lower scores and possibly higher
+equivalenc rates derive from the simple fact that we generate more
+mutants than most tools.  The ease of defining mutation operators,
+plus our interest in generating large numbers of mutants and
+prioritizing the most interesting, to make mutation testing useful for
+projects with extremely high quality test suites, or even using formal
+verification, has encouraged a ``let a thousand flowers bloom''
+approach to choosing mutation operators.  For example, our approach
+provides many more control-flow mutations than most tools, due to our
+observation that test suites sometimes fail to check for loop
+execution counts other than one and zero.  Few other tools introduce
+{\tt break} and {\tt continue} statements into programs, or transform
+one into the other.  These mutants are often equivalent, but also
+sometimes expose subtle weaknesses in a test suite, in our
+experience.  In general, we take our philosophy of mutation from the
+proposal by Gopinath et al. \cite{Limits,gopinath2017mutation} that
+expanding the set of mutation operations and classes of such
+operations, in the long run, is more important than using only a small
+set of ``good'' operators, in that the end-goal of mutation testing is
+to identify weaknesses in test suites, not simply to estimate a single score.
+
+The price paid for this
+expansiveness in allowing ways to detect weaknessess in testing is a
+larger number of mutants to execute, and a somewhat larger number of
+equivalent mutants.  However, for our tool, this observation is not a
+fundamental limitation or forced choice.  We propose as future work to
+split the operators for each level of the hierarchy into two sets, one
+set of core operators similar to (but larger than) those provided by
+LittleDarwin or Cargo-Mutants, and a set of ``aggressive'' operators
+encompassing our full set of operators, a very simple matter due to
+our approach to mutant generation.
 
@@
   language at all}.
-
-\section{Discussion}
 

### 2023-03-28 paper/main.tex
add loc for projects

@@
 
-\begin{tabular}{|c|c|c|r|r|r|}
+\begin{tabular}{|c|c|c|r|r|r|c|}
 \hline
- & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%}  \\ \hline
-\multirow{2}{*}{ompl} & \multicolumn{2}{c|}{Regex} & 595 \hspace{8pt} & 1938 \hspace{8pt} & 23.49\%  \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 564 \hspace{8pt} & 1610 \hspace{8pt} & 25.94\% \\ \hline
-\multirow{2}{*}{libgraphqlparser} & \multicolumn{2}{c|}{Regex} & 157 \hspace{8pt} & 584 \hspace{8pt} & 21.19\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 131 \hspace{8pt} & 309 \hspace{8pt} & 29.77\% \\ \hline
-\multirow{2}{*}{rethinkdb\_rebirth} & \multicolumn{2}{c|}{Regex} & 545 \hspace{8pt} & 1701 \hspace{8pt} & 24.27\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 492 \hspace{8pt} & 1517 \hspace{8pt} & 24.49\% \\ \hline
-
-\multirow{2}{*}{xoreos} & \multicolumn{2}{c|}{Regex} & 2419 \hspace{8pt} & 1457 \hspace{8pt} & 62.41\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 2280 \hspace{8pt} & 1017 \hspace{8pt} & 69.15\% \\ \hline
-\multirow{2}{*}{libxmljs} & \multicolumn{2}{c|}{Regex} & 881 \hspace{8pt} & 3108 \hspace{8pt} & 22.09\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 790 \hspace{8pt} & 3218 \hspace{8pt} & 19.71\% \\ \hline
-\multirow{2}{*}{tiny-differentiable-simulator} & \multicolumn{2}{c|}{Regex} & 747 \hspace{8pt} & 535 \hspace{8pt} & 58.27\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 770 \hspace{8pt} & 444 \hspace{8pt} & 63.43\% \\ \hline
-
-\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 890.67 &	1553.83 & 35.28\%\\\cline{3-6}
-    &   &  Median & 671.00 & 1579.00 & 23.88\%  \\\cline{3-6}
-    &   &  Std. Dev & 787.62 & 956.31 & 19.48\%  \\\cline{2-6}
-
- & \multirow{3}{*}{Comby} & Mean & 837.83 & 1352.50 & 38.75\% \\\cline{3-6}
-    &   &  Median & 667.00 & 1267.00 & 27.86\%  \\\cline{3-6}
-    &   &  Std. Dev & 745.78 & 1058.33	& 21.66\%  \\\hline
+ & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%} & \textbf{LOC} \\ \hline
+\multirow{2}{*}{ompl} & \multicolumn{2}{c|}{Regex} & 595 \hspace{8pt} & 1938 \hspace{8pt} & 23.49\% & \multirow{2}{*}{525} \\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 564 \hspace{8pt} & 1610 \hspace{8pt} & 25.94\% &  \\ \hline
+\multirow{2}{*}{libgraphqlparser} & \multicolumn{2}{c|}{Regex} & 157 \hspace{8pt} & 584 \hspace{8pt} & 21.19\% & \multirow{2}{*}{137}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 131 \hspace{8pt} & 309 \hspace{8pt} & 29.77\% & \\ \hline
+\multirow{2}{*}{rethinkdb\_rebirth} & \multicolumn{2}{c|}{Regex} & 545 \hspace{8pt} & 1701 \hspace{8pt} & 24.27\% & \multirow{2}{*}{471}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 492 \hspace{8pt} & 1517 \hspace{8pt} & 24.49\% & \\ \hline
+
+\multirow{2}{*}{xoreos} & \multicolumn{2}{c|}{Regex} & 2419 \hspace{8pt} & 1457 \hspace{8pt} & 62.41\% & \multirow{2}{*}{732}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 2280 \hspace{8pt} & 1017 \hspace{8pt} & 69.15\% & \\ \hline
+\multirow{2}{*}{libxmljs} & \multicolumn{2}{c|}{Regex} & 881 \hspace{8pt} & 3108 \hspace{8pt} & 22.09\% & \multirow{2}{*}{630}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 790 \hspace{8pt} & 3218 \hspace{8pt} & 19.71\% & \\ \hline
+\multirow{2}{*}{tiny-differentiable-simulator} & \multicolumn{2}{c|}{Regex} & 747 \hspace{8pt} & 535 \hspace{8pt} & 58.27\% & \multirow{2}{*}{195}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 770 \hspace{8pt} & 444 \hspace{8pt} & 63.43\% & \\ \hline
+
+\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 890.67 &	1553.83 & 35.28\% & \\\cline{3-6}
+    &   &  Median & 671.00 & 1579.00 & 23.88\%  & \\\cline{3-6}
+    &   &  Std. Dev & 787.62 & 956.31 & 19.48\% & \\\cline{2-6}
+
+ & \multirow{3}{*}{Comby} & Mean & 837.83 & 1352.50 & 38.75\% & \\\cline{3-6}
+    &   &  Median & 667.00 & 1267.00 & 27.86\% & \\\cline{3-6}
+    &   &  Std. Dev & 745.78 & 1058.33	& 21.66\% & \\\hline
 
@@
 
-\begin{tabular}{|c|c|c|r|r|r|}
+\begin{tabular}{|c|c|c|r|r|r|c|}
 \hline
- & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%}  \\ \hline
-\multirow{2}{*}{pebble} & \multicolumn{2}{c|}{Regex} & 86 \hspace{8pt} & 928 \hspace{8pt} & 8.48\%  \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 85 \hspace{8pt} & 667 \hspace{8pt} & 11.30\% \\ \hline
-\multirow{2}{*}{spring-hateoas} & \multicolumn{2}{c|}{Regex} & 248 \hspace{8pt} & 2231 \hspace{8pt} & 10.00\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 172 \hspace{8pt} & 1212 \hspace{8pt} & 12.43\% \\ \hline
-\multirow{2}{*}{javacc} & \multicolumn{2}{c|}{Regex} & 542 \hspace{8pt} & 1346 \hspace{8pt} & 28.70\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 494 \hspace{8pt} & 782 \hspace{8pt} & 38.71\% \\ \hline
-
-\multirow{2}{*}{coffee-gb} & \multicolumn{2}{c|}{Regex} & 224 \hspace{8pt} & 1703 \hspace{8pt} & 11.63\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 174 \hspace{8pt} & 624 \hspace{8pt} & 21.80\% \\ \hline
-\multirow{2}{*}{egads} & \multicolumn{2}{c|}{Regex} & 624 \hspace{8pt} & 1830 \hspace{8pt} & 25.43\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 597 \hspace{8pt} & 566 \hspace{8pt} & 51.33\% \\ \hline
-\multirow{2}{*}{apk-parser} & \multicolumn{2}{c|}{Regex} & 351 \hspace{8pt} & 1232 \hspace{8pt} & 22.17\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 333 \hspace{8pt} & 866 \hspace{8pt} & 27.77\% \\ \hline
-
-\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 345.83 & 1545.00  & 17.74\%\\\cline{3-6}
-    &   &  Median & 299.50 & 1524.50 & 16.90\%  \\\cline{3-6}
-    &   &  Std. Dev & 203.86 & 467.98 & 8.73\%  \\\cline{2-6}
-
- & \multirow{3}{*}{Comby} & Mean & 309.17 & 786.17 & 27.22\% \\\cline{3-6}
-    &   &  Median & 253.50 & 724.50 & 24.79\%  \\\cline{3-6}
-    &   &  Std. Dev & 202.44 & 235.24  & 15.59\%  \\\hline
+ & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%} & \textbf{LOC}\\ \hline
+\multirow{2}{*}{pebble} & \multicolumn{2}{c|}{Regex} & 86 \hspace{8pt} & 928 \hspace{8pt} & 8.48\%  & \multirow{2}{*}{279}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 85 \hspace{8pt} & 667 \hspace{8pt} & 11.30\% & \\ \hline
+\multirow{2}{*}{spring-hateoas} & \multicolumn{2}{c|}{Regex} & 248 \hspace{8pt} & 2231 \hspace{8pt} & 10.00\% & \multirow{2}{*}{397}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 172 \hspace{8pt} & 1212 \hspace{8pt} & 12.43\% & \\ \hline
+\multirow{2}{*}{javacc} & \multicolumn{2}{c|}{Regex} & 542 \hspace{8pt} & 1346 \hspace{8pt} & 28.70\% & \multirow{2}{*}{363}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 494 \hspace{8pt} & 782 \hspace{8pt} & 38.71\% & \\ \hline
+
+\multirow{2}{*}{coffee-gb} & \multicolumn{2}{c|}{Regex} & 224 \hspace{8pt} & 1703 \hspace{8pt} & 11.63\% & \multirow{2}{*}{463}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 174 \hspace{8pt} & 624 \hspace{8pt} & 21.80\% & \\ \hline
+\multirow{2}{*}{egads} & \multicolumn{2}{c|}{Regex} & 624 \hspace{8pt} & 1830 \hspace{8pt} & 25.43\% & \multirow{2}{*}{252}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 597 \hspace{8pt} & 566 \hspace{8pt} & 51.33\% & \\ \hline
+\multirow{2}{*}{apk-parser} & \multicolumn{2}{c|}{Regex} & 351 \hspace{8pt} & 1232 \hspace{8pt} & 22.17\% & \multirow{2}{*}{324}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 333 \hspace{8pt} & 866 \hspace{8pt} & 27.77\% & \\ \hline
+
+\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 345.83 & 1545.00  & 17.74\%& \\\cline{3-6}
+    &   &  Median & 299.50 & 1524.50 & 16.90\%  & \\\cline{3-6}
+    &   &  Std. Dev & 203.86 & 467.98 & 8.73\%  & \\\cline{2-6}
+
+ & \multirow{3}{*}{Comby} & Mean & 309.17 & 786.17 & 27.22\% & \\\cline{3-6}
+    &   &  Median & 253.50 & 724.50 & 24.79\%  & \\\cline{3-6}
+    &   &  Std. Dev & 202.44 & 235.24  & 15.59\%  & \\\hline
 
@@
 
-\begin{tabular}{|c|c|c|r|r|r|r|}
+\begin{tabular}{|c|c|c|r|r|r|r|c|}
 \hline
- & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Redundant} & \textbf{Valid \%}  \\ \hline
-\multirow{2}{*}{keract} & \multicolumn{2}{c|}{Regex} & 1655 \hspace{8pt} & 1600 \hspace{8pt} &	729 \hspace{8pt} & 41.53\%  \\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 1471 \hspace{8pt} & 1063 \hspace{8pt} & 533 \hspace{8pt} & 47.96\% \\ \hline
-\multirow{2}{*}{dtw} & \multicolumn{2}{c|}{Regex} & 973 \hspace{8pt} & 566 \hspace{8pt} & 567 \hspace{8pt} & 46.20\% \\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 866 \hspace{8pt} & 330 \hspace{8pt} & 639 \hspace{8pt} &	47.19\% \\ \hline
-\multirow{2}{*}{notion-sdk-py} & \multicolumn{2}{c|}{Regex} & 196 \hspace{8pt} & 760 \hspace{8pt} & 82 \hspace{8pt} & 18.89\% \\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 196 \hspace{8pt} & 701 \hspace{8pt} & 122 \hspace{8pt} & 19.23\% \\ \hline
-
-\multirow{2}{*}{{\makecell{atlassian-python \\ api}}} & \multicolumn{2}{c|}{Regex} & 302 \hspace{8pt}	& 1002 \hspace{8pt} & 207 \hspace{8pt} &	19.99\% \\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 333 \hspace{8pt} & 635 \hspace{8pt} & 351 \hspace{8pt} & 25.25\% \\ \hline
-\multirow{2}{*}{ESD} & \multicolumn{2}{c|}{Regex} & 1859 \hspace{8pt}	& 2767 \hspace{8pt} & 1642 \hspace{8pt} & 29.46\% \\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 1712 \hspace{8pt} & 1548 \hspace{8pt} &	614 \hspace{8pt} & 44.18\% \\ \hline
-\multirow{2}{*}{fiber} & \multicolumn{2}{c|}{Regex} & 493 \hspace{8pt} & 909 \hspace{8pt} & 163 \hspace{8pt} & 31.08\% \\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 458 \hspace{8pt} & 767 \hspace{8pt}	& 241 \hspace{8pt} & 31.24\% \\ \hline
-
-\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 913.00 & 1267.33 & 565.00 & 31.19\%   \\\cline{3-7}
-    &   &  Median & 733.00 & 955.50 & 387.00 & 30.27\%  \\\cline{3-7}
-    &   &  Std. Dev & 708.93 & 813.35 & 584.79 & 11.06\%  \\\cline{2-7}
-
- & \multirow{3}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% \\\cline{3-7}
-    &   &  Median & 662.00 & 734.00 & 442.00 & 37.71\%  \\\cline{3-7}
-    &   &  Std. Dev & 628.79 & 418.99 & 211.62 & 12.28\%  \\\hline
+ & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Redundant} & \textbf{Valid \%} & \textbf{LOC} \\ \hline
+\multirow{2}{*}{keract} & \multicolumn{2}{c|}{Regex} & 1655 \hspace{8pt} & 1600 \hspace{8pt} &	729 \hspace{8pt} & 41.53\%  & \multirow{2}{*}{402}\\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 1471 \hspace{8pt} & 1063 \hspace{8pt} & 533 \hspace{8pt} & 47.96\% & \\ \hline
+\multirow{2}{*}{dtw} & \multicolumn{2}{c|}{Regex} & 973 \hspace{8pt} & 566 \hspace{8pt} & 567 \hspace{8pt} & 46.20\% & \multirow{2}{*}{118}\\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 866 \hspace{8pt} & 330 \hspace{8pt} & 639 \hspace{8pt} &	47.19\% & \\ \hline
+\multirow{2}{*}{notion-sdk-py} & \multicolumn{2}{c|}{Regex} & 196 \hspace{8pt} & 760 \hspace{8pt} & 82 \hspace{8pt} & 18.89\% & \multirow{2}{*}{189}\\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 196 \hspace{8pt} & 701 \hspace{8pt} & 122 \hspace{8pt} & 19.23\% & \\ \hline
+
+\multirow{2}{*}{{\makecell{atlassian-python \\ api}}} & \multicolumn{2}{c|}{Regex} & 302 \hspace{8pt}	& 1002 \hspace{8pt} & 207 \hspace{8pt} &	19.99\% & \multirow{2}{*}{236}\\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 333 \hspace{8pt} & 635 \hspace{8pt} & 351 \hspace{8pt} & 25.25\% & \\ \hline
+\multirow{2}{*}{ESD} & \multicolumn{2}{c|}{Regex} & 1859 \hspace{8pt}	& 2767 \hspace{8pt} & 1642 \hspace{8pt} & 29.46\% & \multirow{2}{*}{698}\\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 1712 \hspace{8pt} & 1548 \hspace{8pt} &	614 \hspace{8pt} & 44.18\% & \\ \hline
+\multirow{2}{*}{fiber} & \multicolumn{2}{c|}{Regex} & 493 \hspace{8pt} & 909 \hspace{8pt} & 163 \hspace{8pt} & 31.08\% & \multirow{2}{*}{358}\\\cline{2-7}
+    & \multicolumn{2}{c|}{Comby} & 458 \hspace{8pt} & 767 \hspace{8pt}	& 241 \hspace{8pt} & 31.24\% & \\ \hline
+
+\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 913.00 & 1267.33 & 565.00 & 31.19\%  & \\\cline{3-7}
+    &   &  Median & 733.00 & 955.50 & 387.00 & 30.27\%  & \\\cline{3-7}
+    &   &  Std. Dev & 708.93 & 813.35 & 584.79 & 11.06\%  & \\\cline{2-7}
+
+ & \multirow{3}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% & \\\cline{3-7}
+    &   &  Median & 662.00 & 734.00 & 442.00 & 37.71\%  & \\\cline{3-7}
+    &   &  Std. Dev & 628.79 & 418.99 & 211.62 & 12.28\%  & \\\hline
 
@@
 
-\begin{tabular}{|c|c|c|r|r|r|}
+\begin{tabular}{|c|c|c|r|r|r|c|}
 \hline
- & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%}  \\ \hline
-\multirow{2}{*}{cargo release} & \multicolumn{2}{c|}{Regex} & 851 \hspace{8pt} &	1163 \hspace{8pt} & 42.25\%  \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 1690 \hspace{8pt} & 711 \hspace{8pt} & 70.39\% \\ \hline
-\multirow{2}{*}{passerine} & \multicolumn{2}{c|}{Regex} & 213 \hspace{8pt} & 2081 \hspace{8pt} & 9.29\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 173 \hspace{8pt} & 1518 \hspace{8pt} & 10.23\% \\ \hline
-\multirow{2}{*}{typos} & \multicolumn{2}{c|}{Regex} & 386 \hspace{8pt} & 1534 \hspace{8pt} & 20.10\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 366 \hspace{8pt} & 1199 \hspace{8pt} &	 23.39\% \\ \hline
-
-\multirow{2}{*}{ord} & \multicolumn{2}{c|}{Regex} & 762 \hspace{8pt}	& 2096 \hspace{8pt} & 26.66\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 643 \hspace{8pt} & 1669 \hspace{8pt} & 27.81\% \\ \hline
-\multirow{2}{*}{bazuka} & \multicolumn{2}{c|}{Regex} & 451 \hspace{8pt} & 1103 \hspace{8pt} & 29.02\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 402 \hspace{8pt} & 920 \hspace{8pt} & 30.41\% \\ \hline
-\multirow{2}{*}{strum} & \multicolumn{2}{c|}{Regex} & 51 \hspace{8pt} & 377 \hspace{8pt}	& 11.92\% \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 50 \hspace{8pt} & 318 \hspace{8pt}	& 13.51\% \\ \hline
-
-\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 452.33 & 1392.33	& 23.21\%\\\cline{3-6}
-    &   &  Median & 418.50 & 1348.50 & 23.38\%  \\\cline{3-6}
-    &   &  Std. Dev & 309.05 & 656.76 & 12.16\%  \\\cline{2-6}
-
- & \multirow{3}{*}{Comby} & Mean & 554.00 & 1055.83	& 29.29\% \\\cline{3-6}
-    &   &  Median & 384.00 & 1059.50	& 25.60\%  \\\cline{3-6}
-    &   &  Std. Dev & 592.63 & 508.30 & 21.63\%  \\\hline
+ & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%} &  \textbf{LOC} \\ \hline
+\multirow{2}{*}{cargo release} & \multicolumn{2}{c|}{Regex} & 851 \hspace{8pt} &	1163 \hspace{8pt} & 42.25\%  & \multirow{2}{*}{346}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 1690 \hspace{8pt} & 711 \hspace{8pt} & 70.39\% & \\ \hline
+\multirow{2}{*}{passerine} & \multicolumn{2}{c|}{Regex} & 213 \hspace{8pt} & 2081 \hspace{8pt} & 9.29\% & \multirow{2}{*}{376}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 173 \hspace{8pt} & 1518 \hspace{8pt} & 10.23\% & \\ \hline
+\multirow{2}{*}{typos} & \multicolumn{2}{c|}{Regex} & 386 \hspace{8pt} & 1534 \hspace{8pt} & 20.10\% & \multirow{2}{*}{463}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 366 \hspace{8pt} & 1199 \hspace{8pt} &	 23.39\% & \\ \hline
+
+\multirow{2}{*}{ord} & \multicolumn{2}{c|}{Regex} & 762 \hspace{8pt}	& 2096 \hspace{8pt} & 26.66\% & \multirow{2}{*}{533}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 643 \hspace{8pt} & 1669 \hspace{8pt} & 27.81\% & \\ \hline
+\multirow{2}{*}{bazuka} & \multicolumn{2}{c|}{Regex} & 451 \hspace{8pt} & 1103 \hspace{8pt} & 29.02\% & \multirow{2}{*}{285}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 402 \hspace{8pt} & 920 \hspace{8pt} & 30.41\% & \\ \hline
+\multirow{2}{*}{strum} & \multicolumn{2}{c|}{Regex} & 51 \hspace{8pt} & 377 \hspace{8pt}	& 11.92\% & \multirow{2}{*}{90}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 50 \hspace{8pt} & 318 \hspace{8pt}	& 13.51\% & \\ \hline
+
+\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 452.33 & 1392.33	& 23.21\% & \\\cline{3-6}
+    &   &  Median & 418.50 & 1348.50 & 23.38\%  & \\\cline{3-6}
+    &   &  Std. Dev & 309.05 & 656.76 & 12.16\%  & \\\cline{2-6}
+
+ & \multirow{3}{*}{Comby} & Mean & 554.00 & 1055.83	& 29.29\% & \\\cline{3-6}
+    &   &  Median & 384.00 & 1059.50	& 25.60\%  & \\\cline{3-6}
+    &   &  Std. Dev & 592.63 & 508.30 & 21.63\%  & \\\hline
 
@@
     & Comby & 65 & 111 & 0.37 & 0\% \\ \cline{2-6}
-    & PIT & 82 & 32 & 0.28 & N/A\% \\ \cline{2-6}
+    & PIT & 82 & 32 & 0.28 & N/A \\ \cline{2-6}
     & LittleDarwin & 15 & 23 & 0.39 & 0\% \\ \hline

### 2023-03-28 paper/main.tex
could submit this, but needs claire's eyes

@@
 of using parser parser combinators to improve on a simple regular-expression
-based approach to generation.  
+based approach to generation.  Finally, we show that our approach
+can provide effective mutant generation even for a language for which
+it lacks any language-specific operators, and that is not very similar
+in syntax to any language it has been applied to previously.
 \end{abstract}
@@
 appears that no successor tool has appeared in the ensuing
-years\footnote{There is FitSpec \cite{FitSpec}, which is maintained,
+years\footnote{There is FitSpec~\cite{FitSpec}, which is maintained,
   but FitSpec does not perform traditional mutation testing or produce
@@
 Solidity, Fe, Vyper, Swift, Rust, and Java, including Trivial Compiler
-Equivalence \cite{TCE} checks for redundancy of mutants.  These
+Equivalence~\cite{TCE} checks for redundancy of mutants.  These
 default handlers can always be over-ridden if they do not work with a
@@
 based mutation tool.  Only Dextool, for C++, implements the
-\emph{mutant schema} \cite{untch1993mutation} approach often advocated
+\emph{mutant schema}~\cite{untch1993mutation} approach often advocated
 in the research literature, where a single metaprogram is produced
@@
 (including killed mutants, which obviously cannot be equivalent) that
-are equivalent.  We follow Schuler and Zeller \cite{EquivMut} in
+are equivalent.  We follow Schuler and Zeller~\cite{EquivMut} in
 reporting the percentage of \emph{unkilled} mutants that were found to
@@
 mutants of Java programs to be equivalent.  Other reports of lower
-rates, e.g, 9\% \cite{offutt1997automatically} or 7\% \cite{TCE}, report the percent of
+rates, e.g, 9\%~\cite{offutt1997automatically} or 7\%~\cite{TCE}, report the percent of
 \emph{all} mutants, and so translate to much higher rates of
@@
 are very conservative in mutation generation, and often simply don't
-produce many mutants.
+produce many mutants.  In fact, inspection of the non-comment source
+lines as measured by CLOC shown in Tables~\ref{tab:table_java1} and
+\ref{tab:table_java2} and Tables~\ref{tab:table_rust1} and
+\ref{tab:table_rust2} shows that for many programs, LittleDarwin and
+Cargo-Mutants generated \emph{much less than one mutant per actual
+  line code}.  Such a conservative approach is likely to produce few
+or no actionable (unkilled and non-equivalent) mutants for a program
+with a very good, high-coverage test suite.  But such suites are
+arguably the ones where mutation testing offers the most promise, by
+allowing the detection of very subtle remaining holes in tests for
+high-criticality software.
 
@@
 experience.  In general, we take our philosophy of mutation from the
-proposal by Gopinath et al. \cite{Limits,gopinath2017mutation} that
+proposal by Gopinath et al.~\cite{Limits,gopinath2017mutation} that
 expanding the set of mutation operations and classes of such
@@
 Figure~\ref{fig:haskell} shows a simple Haskell program, the
-motivating example in the paper presenting the MuCheck \cite{mucheck}
+motivating example in the paper presenting the MuCheck~\cite{mucheck}
 mutation testing tool for Haskell.  As noted in the introduction, to our
@@
   language at all}.
+
+\section{Threats to Validity}
+
+Our study only examined 24 total source files, from 24 different projects, across four different
+languages.  None of these files were very large, though they were of
+sizes common to many source files in projects across GitHub, and none
+were less than about 100 non-comment lines of code.  More importantly,
+there is no simple objective way to determine if a mutation testing
+tool is ``as good as'' another tool, in an even approximate sense.
+Previous studies of the topic have considered that so long as a tool
+generates a significant number of non-equivalent unkilled mutants, and
+produces a mutation score not extremely differnt than for other tools,
+it is a useful mutation testing tool.  Further, there is no
+established acceptable tradeoff between generating more mutants and
+accepting a higher cost of compilation and execution for mutants.
 
@@
 programs over four languages,
-and introduces the idea of using parser parser combinators to improve
-on regular expression-based generation.
+and introduces the idea of using parser parser combinators to go
+beyond regular expression-based generation.
 
 \section{Conclusions and Future Work}
+
+We conclude that, by using a hierarchical, source-based approach,
+either relying on a pure-text regular expression definition of
+mutation operators or a richer parser parser combinator transformation
+definition, it is possibly to implement effective, easily extensible
+mutation testing for essentially \emph{all} commonly used programming
+languages, and many non-so-commonly-used programming languages, in
+fewer lines of code than any single-language tool we examined.  This
+approach fundamentally relies on the insight that program mutation is
+just a restricted instance of the general problem of program
+transformation, where transformations are highly syntactic (rather
+than semantics-aware) and highly localized.
+
+As future work, we plan to give users more control over which mutation
+operators are used, with controls common to multiple languages.  We
+add that our approach in general seems to be a good fit to modern
+large software projects, which are often
+multilingual~\cite{Multilingual}: a single mutation tool can be used
+for an entire project.
 

### 2023-03-28 paper/main.tex
remove last universalmutator mention for blinding

@@
     \centering
-    \caption{Overview of Mutation Testing Tools. universalmutator has singificantly fewer lines of code than any of the single language tools.}
+    \caption{Overview of Mutation Testing Tools. Our tool has singificantly fewer lines of code than any of the single language tools.}
     \label{tab:mutationtools}
@@
         C++ & Dextool & https://github.com/joakim-brannstrom/dextool & 38611 \\\hline
-        ALL & universalmutator & https://github.com/agroce/universalmutator & 2244 \\\hline
+        ALL & Our Tool & removed for blinding & 2244 \\\hline
     \end{tabular}

### 2023-03-29 paper/main.tex
some edits, comments in intro

@@
 1970s~\cite{demillo1978hints,mathur2012foundations,demillo1978hints},
-has largely been a topic of academic interest rather than widespread
-practice until quite recently.  However, with the adoption of mutation
+has long been a topic of academic interest.  With the recent adoption of mutation
 testing by bellwether software industry companies including
@@
 Amazon~\cite{AmazonMut}, interest in using mutation testing for
-real-world projects has grown widely.
-
-Interest in mutation testing has become widespread enough in the open
+real-world projects has grown widely in practice. 
+%
+Indeed, interest in mutation testing has become widespread enough in the open
 source community that a popular (more than 300 stars on GitHub)
@@
 \url{https://github.com/theofidry/awesome-mutation-testing}.  Some of
-these tools are the products of academic research, but many are
+these tools are the products of academic research; others are
 essentially the hobby projects of developers interested in enabling
@@
 
-However, the same website demonstrates the existence of a sad history
-of mutation testing tools: mutation tools have a tendency to be
-developed for a while and then abandoned.   The page listing abandoned
-mutation testing tools
-(\url{https://github.com/theofidry/awesome-mutation-testing/blob/master/abandoned.md})
-lists almost half as many tools as are under active maintenance, and
-these tools cover approximately half as many languages.  Each of these
-tools required considerable development time and effort, and was
-popular enough at some point to be worth noting in the ``Awesome
-Mutation Testing'' resource.  Many of the abandoned tools have GitHub
-repositories (some archived) with more than 20 stars, and so may well
-have had users who at some point faced the reality of no future bug
-fixes or changes to address changes to the programming language being
-mutated\footnote{While this is a minor concern for languages such as C
-  that are relatively stable, it means that a mutation tool may fail
-  to parse C++ 2020, and is particularly problematic for somewhat
-  fast-moving languages such as Rust.}.  There
-appears to be a significant risk in relying on any mutation tools other
-than the most widely used and well-supported, such as PIT.
-
-Moreover, newer languages, including ones with considerable interest
+This diversity of tools highlights the growth of widespread interest in mutation
+testing.  However, it also demonstrates various ways that bringing mutation testing to
+typically polyglot modern projects is cumbersome. Each language (or IR, at
+minimum) requires its own underlying mutation and testing framework.
+\clg{Alex, can something be said about 
+  how many LOC the tools we look at are on average, this early?}
+
+Newer languages, including ones with considerable interest
 in the development community, often lack mutation testing tools.  Even
 languages with a long history may long lack a working mutation testing
-tool; e.g., it was not until 2014 that Haskell had such a tool
-\cite{mucheck}.  Furthermore, that tool is no longer
-supported\footnote{The last commits or package updates were in 2015, and the
+tool; e.g., it was not until 2014 that Haskell had such a tool~\cite{mucheck}. 
+%
+Furthermore, the Haskell mutation tool is no longer
+supported,\footnote{The last commits or package updates were in 2015, and the
 primary author confirms that changes to {\tt haskell-src-ext} since
-introduced render it unusable at present.}, and it
+introduced render it unusable at present.} and it
 appears that no successor tool has appeared in the ensuing
@@
 
+Indeed, abandonment is a feature of the sad history
+of many mutation testing tools, exacerbating the challenges of its uptake.  The page listing abandoned
+mutation testing tools
+(\url{https://github.com/theofidry/awesome-mutation-testing/blob/master/abandoned.md})
+lists almost half as many tools as are under active maintenance. Each tool represents 
+considerable development effort, and was
+popular enough to be noted in the ``Awesome
+Mutation Testing'' resource. Many of the abandoned repositories have more than
+20 stars, suggesting inconvenienced users, especially for tools targeting
+evolving languages. While this is a minor concern for languages such as C
+that are relatively stable, it means that a mutation tool may fail
+to parse C++ 2020, and is particularly problematic for 
+faster-moving languages like Rust.
+
+%There
+%appears to be a significant risk in relying on any mutation tools other
+%than the most widely used and well-supported, such as PIT.
 A great deal of research effort has been devoted to devising clever
 schemes for reducing the computational burden of mutation testing over
-the years; in practice, however, the recent adoption of mutation
+the years;\clg{a few references?} in practice, however, the recent adoption of mutation
 testing arguably relies more on the availability of vast (cloud) computing
@@
 due to the high human cost of developing and maintaining mutation
+\clg{OK I don't mind this argument, BUT it kind of seems in conflict with the
+  argument above that says there are a lot of tools and mutation testing is
+  popular.  I think it might be worth shifting the emphasis of this paragraph to
+  be to motivate source-level mutation testing over IR, because that's missing
+  so far and would be good to motivate.  Note that it's also an argument made in
+  the abstract, which I think is a particularly good summary of the paper.}
 testing tools.  Ideally, mutation testing would be available for

### 2023-03-29 paper/main.tex
long-ass comments

@@
 
-\section{The Universal Language Source-Based Approach}
+\section{A Universal Language Syntax-Based Approach}
 
@@
 such occurrence, a mutant where tha character is replaced by {\tt
-  ``-''}.  We can do the same for each pair of arithmetic operations,
-and thus trivially implement a large fraction of the classic Mothra~\cite{offutt1996experimental}
+  ``-''}.  Many languages express arithmetic as infix operators between
+expressions; if expressed syntactically, mutations for them can be expressed
+once, trivially implementing a large fraction of the classic Mothra~\cite{offutt1996experimental}
 mutation operators.
@@
 instances can be avoided by slightly more judicious choice of search
-target: instances of {\tt ``+''} where the preceeding or following
-character is not another {\tt ``+''}, for instance.
-
-Of course, in some cases this will still leave us with invalid
-mutants.  However, rejecting such mutants is easy: the truly invalid
-mutants will fail to compile.  Of course, we must pay the price of
-checking each mutant to make sure that it does compile, at generation
+target, e.g., instances of {\tt ``+''} where the preceeding or following
+character is not another {\tt ``+''}.  This may still in some cases produce
+invalid mutants, but rejecting such mutants is easy, since they will fail
+to compile.  Of course, we pay the price of
+compiling each mutant at generation
 time.
 
-However, consider what has been gained by paying this cost.  We have
-replaced a complex implementation, that requires parsing source code,
-with an implementation that only requires string search-and-replace.
-One can imagine an implementation of the above approach to arithmetic
-operation replacement being implemented in as little as 10-20 lines of
-Python code, including a call to a compiler to check for invalid
-mutants.  Moreover, defining new mutation operators is not a matter of
-understanding a representation of a parsed program, but is essentially
+On the positive side, this argues that we can replace a
+complex implementation requiring parsing and analysis with
+withone that only requires string search-and-replace.
+Such an implementation on face can be done using much less code. 
+% It also allows for simpler definitions of 
+% One can imagine an implementation of the above approach to arithmetic
+% operation replacement being implemented in as little as 10-20 lines of
+% Python code, including a call to a compiler to check for invalid
+%mutants.  
+Moreover, defining new mutation operators is not a matter of
+understanding a parsed program representation, but is essentially
 a matter of providing two strings, where the second is a proposed
-replacement for the first.  Instead of code for arithmetic operator
-replacement, we can imagine the above hypothetical Python program
-reading in a file of thus expressed mutation operators, and
-implementing an unbounded number of operators in a minimalist fashion.
-
-Of course, simple string replacement is not general enough to handle
+replacement for the first.  This simplifies operator specification from
+implementation of
+tree transformations into a minimal list. 
+%Instead of code for arithmetic operator
+%replacement, we can imagine the above hypothetical Python program
+%reading in a file of thus expressed mutation operators, and
+%implementing an unbounded number of operators in a minimalist fashion.
+
+Of course, simple string replacement cannot handle
 many interesting mutations.  Consider the classic case of statement
 deletion, one of the most widely used and important mutation
-operators~\cite{deng2013empirical}.  What string can we replace with another string in order to
-delete a statement in a C program?  We would need an ``operator'' for
-each possible C statement.  However, if we allow the use of
-\emph{regular expression match and replace pairs} in place of simple
-strings, we can easily express this operator, with some precision:
+operators~\cite{deng2013empirical}. This cannot be concisely specified with string pairs. 
+%  What string can we replace with another string in order to
+% delete a statement in a C program?  We would need an ``operator'' for
+% each possible C statement. 
+However, if we allow the use of
+\emph{regular expression match and replace pairs}, we can express this operator, with some precision:
 
@@
 
+\clg{I know we're walking a line on UM anonymity but I think we should cite it
+  here, just in the third person.}
 Using regular expressions to express mutants has many benefits.  While
@@
 describing the operator in natural
-language~\cite{zhong2018generating}.  Implementing a mutation
-generation tool for a language by providing the equivalent of our
-Python program above with a list of regular expression matches and
-replacements is clearly much easier than writing a mutation tool from
-scratch, and much easier to maintain and extend.
+language~\cite{zhong2018generating}.  
+\clg{Just as a note (feel free to read and delete): I tried to cut out our
+  hypothetical Python program because I read it and was like OK well then why
+  didn't you write it? Or show me if it's so short? You know? And I felt like
+  the point could be made without it. I think reference to UM here does the
+  trick as well really}
+These approaches speak to the potential of string manipulation-based approaches
+for mutation testing, without the need for program parsing or analysis.
+
+% generation tool for a language by providing the equivalent of our
+% Python program above with a list of regular expression matches and
+% replacements is clearly much easier than writing a mutation tool from
+% scratch, and much easier to maintain and extend. 
+% CLG: This claim feels strong to me even though I know you think it's true
+% because you did it. ;-)
 
@@
 
-Implementing mutant generation by providing a set of regular-expression
-based textual transformations for each target language, and using a
-single engine to apply rules to source files is certainly less
-difficult in terms of development and maintenance effort than building
-a tool for each language around a parser for that language.  However,
-even this approach ignores one of the primary benefits of a universal,
-text-based approach.  Consider the case of replacing instances of {\tt
+Regular expressions suggest a mechanism for textual transformation customized to
+multiple target languages, which could then share a single engine to apply rules
+to source files without the complexity of parsing. 
+However, we observe an additional property of mutation operators and mutation
+testing that speaks to the potential of a language-agnostic approach: many of the
+most widely used mutation operators in the literature are \emph{universal}.  For
+example, consider the case of replacing instances of {\tt
   ``+''}  with {\tt ``-''}.  This mutation is not tied to any
 particular language, but applies to \emph{almost all programming
-  languages in use.}  Many of the most widely used mutation operators
-in the literature are of this universal nature.  Others are not quite
+  languages in use.}  Others are not quite
 so widely applicable, but still apply to a large number of specific
@@
 programs, but instead can proceed by process of 1) identifying $L$'s
-place in a \emph{hierarchy} of language \emph{famillies}, and then 2)
+place in a \emph{hierarchy} of language \emph{families}, and then 2)
 identifying additional operators needed for $L$ itself.  The first of
@@
 As an
-example, consider mutation of Java code.  Many Java mutants can be
+example, consider mutation of Java code.  Many typical Java mutants can be
 generated by the kind of ``universal'' rules (e.g., arithmetic
@@
 
+\clg{OK this is confusing and I don't know how to fix it.  Two problems.  (1) we
+  haven't actually introduced a tool yet so talking about something from its
+  implementation is confusing.  We do have kind of a hypothetical tool, though
+  I've been removing specific references to it, because it wasn't clear enough
+  that it was supposed to represent our ACTUAL implementation and it was
+  confusing me.  If we'd like it to introduce an actual implementation, then we
+  might undo my changes there, but make it clear we're introducing an
+  implementation and not a thought experiment.  Problem (2) is that a lot of
+  this paper (and even the intro/abstract to date) is about using parser
+  combinators for this, not regular expressions.  So talking about using regular
+  expressions in our not-introduced implementation when alter we'll be talking
+  about parser combinators as a contribution is also confusing.  Beyond
+  ``concretize our hypothetical tool into introduicng actual tool, other
+  solutions I can think of are (1) forward ref a section of an implementation and expand slightly this
+  sentence to be like ``taken directly from our implementation of a regular
+  expression approach to syntactic mutation (Section FOO)'', or (2) be more abstract
+  about it here, like ``Table X shows examples of universal rules...we discuss
+  them further and how such rules are used in our implementation in Section
+  FOO.''.  Maybe those are the same I dunno.}
 Table~\ref{tab:rules} shows examples of universal, C-like, and
@@
 
+\clg{OK.  I think maybe ??? This section works better as a subsection in the
+  previous section, which can build to ``beyond...'' Then this text can/should
+  start with a TINY bit more transition from the regular expression thing we've
+  discussed so far to motivate comby ``instead'' (even though it's not really
+  instead.)  This is more a point for discussion than a unilateral decree, but
+  it feels like it flows better as a complete story.  Otherwise it's not super
+  clear what the point of Section 2 was, if Section 3 is the real thing we're
+  proposing? Is it? On that note, if we do this, I think we can/should move
+  ``Incremental Mutation'' down to implementation even if we didn't do it, just
+  as a point of discussion like ``oh btw we could do it if we wanted it's
+  easy.'' Otherwise it kind of disrupts the flow AND hierararies of operators
+  leads REALLY REALLY nicely into the parse parser combinator approach.}
+\clg{Either way regardless this needs a better starting transition haha.}
 We implement a lightweight parser-driven approach that attunes mutation

### 2023-03-29 paper/main.tex
at least one question

@@
 language'' with few tools targeting it.
+\clg{Kinda want a table (or a pointer to tables) with projects per language.
+  Definitely want a pointer to at least Table 6, with the tools we compare to.}
 
@@
 ``understanding'' of the source being mutated.
+\clg{What are LOC in the tables?}
 
@@
     \centering
-    \caption{Overview of Mutation Testing Tools. Our tool has singificantly fewer lines of code than any of the single language tools.}
+    \caption{Overview of Mutation Testing Tools. Our tool has significantly fewer lines of code than any of the single language tools.}
     \label{tab:mutationtools}
 
-    \begin{tabular}{|c|c|c|r|}
+    \begin{tabular}{|l|l|l|r|}
     \hline

### 2023-03-29 paper/main.tex
on compile time

@@
 ``understanding'' of the source being mutated.
-\clg{What are LOC in the tables?}
 
@@
 cost is relatively small compared to the time required to compile and
-execute valid mutants for almost all projects.
+execute valid mutants for almost all projects.  An informal study of
+large Java projects by the author of the PIT gradle
+plugin\footnote{\url{https://solidsoft.wordpress.com/2017/09/19/how-fast-or-slow-mutation-testing-really-is/}}
+  suggests that even a successful compile of a typical project takes
+  somewhere between one-sixth to one-third the time required for
+  executing tests, and failed compiles are usually much
+  faster than successful ones.
 

### 2023-03-29 paper/main.tex
playing with tables

--- 2023-03-29 paper/main.tex
@@
 \label{tab:table_cpp1}
-\resizebox{\columnwidth}{!}{%
-
-\begin{tabular}{|c|c|c|r|r|r|c|}
-\hline
- & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%} & \textbf{LOC} \\ \hline
-\multirow{2}{*}{ompl} & \multicolumn{2}{c|}{Regex} & 595 \hspace{8pt} & 1938 \hspace{8pt} & 23.49\% & \multirow{2}{*}{525} \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 564 \hspace{8pt} & 1610 \hspace{8pt} & 25.94\% &  \\ \hline
-\multirow{2}{*}{libgraphqlparser} & \multicolumn{2}{c|}{Regex} & 157 \hspace{8pt} & 584 \hspace{8pt} & 21.19\% & \multirow{2}{*}{137}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 131 \hspace{8pt} & 309 \hspace{8pt} & 29.77\% & \\ \hline
-\multirow{2}{*}{rethinkdb\_rebirth} & \multicolumn{2}{c|}{Regex} & 545 \hspace{8pt} & 1701 \hspace{8pt} & 24.27\% & \multirow{2}{*}{471}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 492 \hspace{8pt} & 1517 \hspace{8pt} & 24.49\% & \\ \hline
-
-\multirow{2}{*}{xoreos} & \multicolumn{2}{c|}{Regex} & 2419 \hspace{8pt} & 1457 \hspace{8pt} & 62.41\% & \multirow{2}{*}{732}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 2280 \hspace{8pt} & 1017 \hspace{8pt} & 69.15\% & \\ \hline
-\multirow{2}{*}{libxmljs} & \multicolumn{2}{c|}{Regex} & 881 \hspace{8pt} & 3108 \hspace{8pt} & 22.09\% & \multirow{2}{*}{630}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 790 \hspace{8pt} & 3218 \hspace{8pt} & 19.71\% & \\ \hline
-\multirow{2}{*}{tiny-differentiable-simulator} & \multicolumn{2}{c|}{Regex} & 747 \hspace{8pt} & 535 \hspace{8pt} & 58.27\% & \multirow{2}{*}{195}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 770 \hspace{8pt} & 444 \hspace{8pt} & 63.43\% & \\ \hline
-
-\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 890.67 &	1553.83 & 35.28\% & \\\cline{3-6}
-    &   &  Median & 671.00 & 1579.00 & 23.88\%  & \\\cline{3-6}
-    &   &  Std. Dev & 787.62 & 956.31 & 19.48\% & \\\cline{2-6}
-
- & \multirow{3}{*}{Comby} & Mean & 837.83 & 1352.50 & 38.75\% & \\\cline{3-6}
-    &   &  Median & 667.00 & 1267.00 & 27.86\% & \\\cline{3-6}
-    &   &  Std. Dev & 745.78 & 1058.33	& 21.66\% & \\\hline
-
+\begin{tabular}{l|r|cr|cr}
+\toprule
+                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
+\textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
+ompl & 525 & $\frac{595}{2533}$ & 23.49\% &  $\frac{564}{2127}$  & 25.94\%   \\[1ex]
+libgraphqlparser& 137 & $\frac{157}{741}$ & 21.19\%  & $\frac{131}{440}$ & 29.77\% \\[1ex]
+rethinkdb\_rebirth &471 &  $\frac{545}{2246}$ & 24.27\% & $\frac{492}{2009}$ & 24.49\% \\[1ex]
+xoreos & 732  & $\frac{2419}{3876}$  & 62.41\% & $\frac{2280}{3297}$ & 69.15\% \\[1ex]
+libxmljs &630 &  $\frac{881}{3989}$ & 22.09\% & $\frac{790}{4008}$ & 19.71\% \\
+tiny-differentiable &  \multirow{2}{*}{195} & \multirow{2}{*}{$\frac{747}{1282}$} & \multirow{2}{*}{58.27\%} & \multirow{2}{*}{$\frac{770}{1214}$} & \multirow{2}{*}{63.43\%} \\
+-simulator          & & & & \\[1ex]
+\midrule
+\mc{6}{c}{\textbf{Overall}} \\
+                   &    & \mc{1}{c}{Valid}  & \mc{1}{c}{Invalid} & \mc{1}{c}{\%} \\\midrule
+ \multirow{3}{*}{Regex} & Mean & 890.67 &  \mc{1}{r}{1553.83} & 35.28\% \\
+                       &  Med. & 671.00 & \mc{1}{r}{1579.00} & 23.88\%   \\
+                       &  Std.Dev & 787.62 & \mc{1}{r}{956.31} & 19.48\%  \\\midrule
+ \multirow{3}{*}{Comby} & Mean & 837.83 & \mc{1}{r}{1352.50} & 38.75\%  \\
+                        &  Med. & 667.00 & \mc{1}{r}{1267.00} & 27.86\%  \\
+                        &  Std.Dev & 745.78 & \mc{1}{r}{1058.33}& 21.66\%  \\\hline
 \end{tabular}
-}
+
+\end{table}
+
+% % ---------------- C++ table UM Only--------------------------
+% \begin{table}[hbtp]
+% \centering
+% \caption{C++ (Regex vs. Comby)}
+% \label{tab:table_cpp1}
+% \resizebox{\columnwidth}{!}{%
+
+% \begin{tabular}{|c|c|c|r|r|r|c|}
+% \toprule
+%  & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%} & \textbf{LOC} \\ \hline
+% \multirow{2}{*}{ompl} & \multicolumn{2}{c|}{Regex} & 595 \hspace{8pt} & 1938 \hspace{8pt} & 23.49\% & \multirow{2}{*}{525} \\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 564 \hspace{8pt} & 1610 \hspace{8pt} & 25.94\% &  \\ \hline
+% \multirow{2}{*}{libgraphqlparser} & \multicolumn{2}{c|}{Regex} & 157 \hspace{8pt} & 584 \hspace{8pt} & 21.19\% & \multirow{2}{*}{137}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 131 \hspace{8pt} & 309 \hspace{8pt} & 29.77\% & \\ \hline
+% \multirow{2}{*}{rethinkdb\_rebirth} & \multicolumn{2}{c|}{Regex} & 545 \hspace{8pt} & 1701 \hspace{8pt} & 24.27\% & \multirow{2}{*}{471}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 492 \hspace{8pt} & 1517 \hspace{8pt} & 24.49\% & \\ \hline
+
+% \multirow{2}{*}{xoreos} & \multicolumn{2}{c|}{Regex} & 2419 \hspace{8pt} & 1457 \hspace{8pt} & 62.41\% & \multirow{2}{*}{732}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 2280 \hspace{8pt} & 1017 \hspace{8pt} & 69.15\% & \\ \hline
+% \multirow{2}{*}{libxmljs} & \multicolumn{2}{c|}{Regex} & 881 \hspace{8pt} & 3108 \hspace{8pt} & 22.09\% & \multirow{2}{*}{630}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 790 \hspace{8pt} & 3218 \hspace{8pt} & 19.71\% & \\ \hline
+% \multirow{2}{*}{tiny-differentiable-simulator} & \multicolumn{2}{c|}{Regex} & 747 \hspace{8pt} & 535 \hspace{8pt} & 58.27\% & \multirow{2}{*}{195}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 770 \hspace{8pt} & 444 \hspace{8pt} & 63.43\% & \\ \hline
+
+% \multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 890.67 &	1553.83 & 35.28\% & \\\cline{3-6}
+%     &   &  Median & 671.00 & 1579.00 & 23.88\%  & \\\cline{3-6}
+%     &   &  Std. Dev & 787.62 & 956.31 & 19.48\% & \\\cline{2-6}
+
+%  & \multirow{3}{*}{Comby} & Mean & 837.83 & 1352.50 & 38.75\% & \\\cline{3-6}
+%     &   &  Median & 667.00 & 1267.00 & 27.86\% & \\\cline{3-6}
+%     &   &  Std. Dev & 745.78 & 1058.33	& 21.66\% & \\\hline
+
+% \end{tabular}
+% }
+
+% \end{table}
+
+% ---------------- JAVA table UM Only--------------------------
+\begin{table}[hbtp]
+\centering
+\caption{Java (Regex vs. Comby)}
+\label{tab:table_java1}
+
+\begin{tabular}{l|r|cr|cr}
+\toprule
+                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
+\textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
+pebble         & 279 & $\frac{86}{1014}$ & 8.48\% & $\frac{85}{752}$ & 11.30\%  \\[1ex]
+spring-hateoas & 397 & $\frac{248}{2479}$ & 10.00\% & $\frac{172}{1384}$ & 12.43\%  \\[1ex]
+javacc         & 363 & $\frac{542}{1888}$ & 28.70\% & $\frac{494}{1281}$ & 38.71\%  \\[1ex]
+coffee-gb      & 463 & $\frac{224}{1927}$ & 11.63\% & $\frac{174}{798}$ & 21.80\% \\[1ex]
+egads          & 252 & $\frac{624}{2427}$ & 25.43\% & $\frac{597}{1163}$ & 51.33\%  \\[1ex]
+apk-parser     & 324 & $\frac{351}{1583}$ & 22.17\% & $\frac{333}{1199}$ & 27.77\%  \\[1ex]
+
+% \multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 345.83 & 1545.00  & 17.74\%& \\\cline{3-6}
+%     &   &  Median & 299.50 & 1524.50 & 16.90\%  & \\\cline{3-6}
+%     &   &  Std. Dev & 203.86 & 467.98 & 8.73\%  & \\\cline{2-6}
+
+%  & \multirow{3}{*}{Comby} & Mean & 309.17 & 786.17 & 27.22\% & \\\cline{3-6}
+%     &   &  Median & 253.50 & 724.50 & 24.79\%  & \\\cline{3-6}
+%     &   &  Std. Dev & 202.44 & 235.24  & 15.59\%  & \\\hline
+
+\end{tabular}
 

### 2023-03-29 paper/main.tex
with subtables, maybe not worth the space

@@
 \usepackage{booktabs}
-\usepackage{subfig}
+%\usepackage{subfig}
+\usepackage{float}
+\usepackage{subcaption}
 \usepackage{url}
@@
 \centering
-\caption{C++ (Regex vs. Comby)}
+\caption{Regex vs. Comby for C++ (top) and Java (bottom)}
+{\begin{subtable}[h]{\columnwidth}
+%\caption{C++ (Regex vs. Comby)}
 \label{tab:table_cpp1}
-\begin{tabular}{l|r|cr|cr}
+\begin{tabularx}{\columnwidth}{X|r|cr|cr}
 \toprule
-                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
+                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\[1ex]
 \textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
@@
 xoreos & 732  & $\frac{2419}{3876}$  & 62.41\% & $\frac{2280}{3297}$ & 69.15\% \\[1ex]
-libxmljs &630 &  $\frac{881}{3989}$ & 22.09\% & $\frac{790}{4008}$ & 19.71\% \\
-tiny-differentiable &  \multirow{2}{*}{195} & \multirow{2}{*}{$\frac{747}{1282}$} & \multirow{2}{*}{58.27\%} & \multirow{2}{*}{$\frac{770}{1214}$} & \multirow{2}{*}{63.43\%} \\
--simulator          & & & & \\[1ex]
-\midrule
-\mc{6}{c}{\textbf{Overall}} \\
-                   &    & \mc{1}{c}{Valid}  & \mc{1}{c}{Invalid} & \mc{1}{c}{\%} \\\midrule
- \multirow{3}{*}{Regex} & Mean & 890.67 &  \mc{1}{r}{1553.83} & 35.28\% \\
-                       &  Med. & 671.00 & \mc{1}{r}{1579.00} & 23.88\%   \\
-                       &  Std.Dev & 787.62 & \mc{1}{r}{956.31} & 19.48\%  \\\midrule
- \multirow{3}{*}{Comby} & Mean & 837.83 & \mc{1}{r}{1352.50} & 38.75\%  \\
-                        &  Med. & 667.00 & \mc{1}{r}{1267.00} & 27.86\%  \\
-                        &  Std.Dev & 745.78 & \mc{1}{r}{1058.33}& 21.66\%  \\\hline
-\end{tabular}
-
+libxmljs &630 &  $\frac{881}{3989}$ & 22.09\% & $\frac{790}{4008}$ & 19.71\% \\[1ex]
+tiny-diff.-simulator &  195 & $\frac{747}{1282}$ &58.27\% &$\frac{770}{1214}$ & 63.43\% \\[1ex]
+\bottomrule
+\end{tabularx}
+\begin{tabularx}{\columnwidth}{Xllrrr}
+                 & &       & Valid  & Invalid  & \% \\\midrule
+\multirow{6}{*}{\textbf{Overall}} & \multirow{3}{*}{Regex} & Mean     & 890.67 &  1553.83 & 35.28\% \\
+                         & & Med.    & 671.00 & 1579.00  & 23.88\%   \\
+                         & & Std.Dev. & 787.62 & 956.31   & 19.48\%  \\\cline{2-6}
+ & \multirow{3}{*}{Comby}  & Mean     & 837.83 & 1352.50 & 38.75\%  \\
+                        & & Med.    & 667.00 & 1267.00 & 27.86\%  \\
+                        & & Std.Dev. & 745.78 & 1058.33& 21.66\%  \\\bottomrule
+\end{tabularx}
+\end{subtable}}
+\vspace{1ex}
+
+{\begin{subtable}[h]{\columnwidth}
+%\caption{Java }
+\label{tab:table_java1}
+\begin{tabularx}{\columnwidth}{X|r|cr|cr}
+\toprule
+                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
+\textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
+pebble         & 279 & $\frac{86}{1014}$ & 8.48\% & $\frac{85}{752}$ & 11.30\%  \\[1ex]
+spring-hateoas & 397 & $\frac{248}{2479}$ & 10.00\% & $\frac{172}{1384}$ & 12.43\%  \\[1ex]
+javacc         & 363 & $\frac{542}{1888}$ & 28.70\% & $\frac{494}{1281}$ & 38.71\%  \\[1ex]
+coffee-gb      & 463 & $\frac{224}{1927}$ & 11.63\% & $\frac{174}{798}$ & 21.80\% \\[1ex]
+egads          & 252 & $\frac{624}{2427}$ & 25.43\% & $\frac{597}{1163}$ & 51.33\%  \\[1ex]
+apk-parser     & 324 & $\frac{351}{1583}$ & 22.17\% & $\frac{333}{1199}$ & 27.77\%  \\[1ex]
+\end{tabularx}
+\begin{tabularx}{\columnwidth}{Xllrrr}
+              & & & Valid  & Invalid  & \% \\\midrule
+\multirow{6}{*}{\textbf{Overall}} & \mr{3}{Regex} & Mean     &  345.83 & 1545.00  & 17.74\% \\
+              &   & Med.     & 299.50 & 1524.50 & 16.90\%  \\
+              &   & Std.Dev. & 203.86 & 467.98 & 8.73\%  \\\cline{2-6}
+& \mr{3}{Comby} & Mean     & 309.17 & 786.17 & 27.22\% \\
+              &   & Med.     & 253.50 & 724.50 & 24.79\% \\
+              &   & Std.Dev. & 202.44 & 235.24  & 15.59\% \\\bottomrule
+\end{tabularx}
+\end{subtable}}
 \end{table}
@@
 \centering
-\caption{Java (Regex vs. Comby)}
-\label{tab:table_java1}
-
-\begin{tabular}{l|r|cr|cr}
-\toprule
-                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
-\textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
-pebble         & 279 & $\frac{86}{1014}$ & 8.48\% & $\frac{85}{752}$ & 11.30\%  \\[1ex]
-spring-hateoas & 397 & $\frac{248}{2479}$ & 10.00\% & $\frac{172}{1384}$ & 12.43\%  \\[1ex]
-javacc         & 363 & $\frac{542}{1888}$ & 28.70\% & $\frac{494}{1281}$ & 38.71\%  \\[1ex]
-coffee-gb      & 463 & $\frac{224}{1927}$ & 11.63\% & $\frac{174}{798}$ & 21.80\% \\[1ex]
-egads          & 252 & $\frac{624}{2427}$ & 25.43\% & $\frac{597}{1163}$ & 51.33\%  \\[1ex]
-apk-parser     & 324 & $\frac{351}{1583}$ & 22.17\% & $\frac{333}{1199}$ & 27.77\%  \\[1ex]
-
-% \multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 345.83 & 1545.00  & 17.74\%& \\\cline{3-6}
-%     &   &  Median & 299.50 & 1524.50 & 16.90\%  & \\\cline{3-6}
-%     &   &  Std. Dev & 203.86 & 467.98 & 8.73\%  & \\\cline{2-6}
-
-%  & \multirow{3}{*}{Comby} & Mean & 309.17 & 786.17 & 27.22\% & \\\cline{3-6}
-%     &   &  Median & 253.50 & 724.50 & 24.79\%  & \\\cline{3-6}
-%     &   &  Std. Dev & 202.44 & 235.24  & 15.59\%  & \\\hline
-
-\end{tabular}
-
 \end{table}

### 2023-03-29 paper/main.tex
clg fixes

--- 2023-03-29 paper/main.tex
@@
 minimum) requires its own underlying mutation and testing framework.
-\clg{Alex, can something be said about 
-  how many LOC the tools we look at are on average, this early?}
+Mutation tools are not trivial to develop; the smallest
+single-language tools we examined were nearly 4KLOC, and the mean size
+of a mutation testing tool was over 20K non-comment source lines.
 
@@
 schemes for reducing the computational burden of mutation testing over
-the years;\clg{a few references?} in practice, however, the recent adoption of mutation
+the years (e.g., ~\cite{offutt2001mutation,
+  untch1993mutation,KaufmanFAKAJ2022}); in practice, however, the recent adoption of mutation
 testing arguably relies more on the availability of vast (cloud) computing
@@
 due to the high human cost of developing and maintaining mutation
-\clg{OK I don't mind this argument, BUT it kind of seems in conflict with the
-  argument above that says there are a lot of tools and mutation testing is
-  popular.  I think it might be worth shifting the emphasis of this paragraph to
-  be to motivate source-level mutation testing over IR, because that's missing
-  so far and would be good to motivate.  Note that it's also an argument made in
-  the abstract, which I think is a particularly good summary of the paper.}
-testing tools.  Ideally, mutation testing would be available for
+testing tools.  In particular, new languages may have unstable
+grammars, and few tools to aid developers in parsing and manipulating
+source code.   Ideally, mutation testing would be available for
 \emph{any} newly proposed programming language, as soon as that
@@
 state-of-the-art, however, it may be years before even rudimentary
-tools appear, and they may ``disappear'' shortly there-after.
+tools appear, and they may ``disappear'' shortly there-after.  In
+fact, the difficulties of parsing code make bytecode or IL-level
+mutation a popular recent choice in tool development, despite the advantages of source-based mutation for
+users of mutation testing~\cite{CompareSrcBinary}, including the
+ability to read and understand mutants (which is often essential if
+one aims to add a test to kill a mutant) and to define mutants in
+terms of source-level language (or even project/library)-specific
+features.  Moreover, even this arguably less desirable bytecode mutation
+approach is not available for languages that do not compile to, e.g.,
+Java or LLVM bytecode, such as Haskell.
 
@@
 
-\clg{I know we're walking a line on UM anonymity but I think we should cite it
-  here, just in the third person.}
-Using regular expressions to express mutants has many benefits.  While
+Using regular expressions to express mutants has many benefits, and
+has been demonstrated to be possible, with a relatively small code base~\cite{regexpMut}.  While
 much more expressive than simple string replacement, regular
@@
 language~\cite{zhong2018generating}.  
-\clg{Just as a note (feel free to read and delete): I tried to cut out our
-  hypothetical Python program because I read it and was like OK well then why
-  didn't you write it? Or show me if it's so short? You know? And I felt like
-  the point could be made without it. I think reference to UM here does the
-  trick as well really}
 These approaches speak to the potential of string manipulation-based approaches
@@
 \centering
-\caption{Rule Examples}
+\caption{Regular Expression Rule Examples}
 \label{tab:rules}
@@
 
-\clg{OK this is confusing and I don't know how to fix it.  Two problems.  (1) we
-  haven't actually introduced a tool yet so talking about something from its
-  implementation is confusing.  We do have kind of a hypothetical tool, though
-  I've been removing specific references to it, because it wasn't clear enough
-  that it was supposed to represent our ACTUAL implementation and it was
-  confusing me.  If we'd like it to introduce an actual implementation, then we
-  might undo my changes there, but make it clear we're introducing an
-  implementation and not a thought experiment.  Problem (2) is that a lot of
-  this paper (and even the intro/abstract to date) is about using parser
-  combinators for this, not regular expressions.  So talking about using regular
-  expressions in our not-introduced implementation when alter we'll be talking
-  about parser combinators as a contribution is also confusing.  Beyond
-  ``concretize our hypothetical tool into introduicng actual tool, other
-  solutions I can think of are (1) forward ref a section of an implementation and expand slightly this
-  sentence to be like ``taken directly from our implementation of a regular
-  expression approach to syntactic mutation (Section FOO)'', or (2) be more abstract
-  about it here, like ``Table X shows examples of universal rules...we discuss
-  them further and how such rules are used in our implementation in Section
-  FOO.''.  Maybe those are the same I dunno.}
 Table~\ref{tab:rules} shows examples of universal, C-like, and
 language-specific rules for a few languages, taken directly from our
-implementation.  In all cases, the form of a rule is: {\tt regexp
-  ==> replacement}, using Python regular expression and replacement syntax.
+implementation of regular-expression based operators (see Section \ref{sec:imp}).  In all cases, the form of a rule is: {\tt regexp
+  ==> replacement}, using Python regular expression and replacement
+syntax.  While these rules are non-trivial to write and read, we show
+below that using parser parser combinators often simplifies and
+clarifies such rules.
 
@@
 
-\subsection{Incremental Mutation}
-
-One obvious concern with any-language source-based mutation is that every time the
-program changes, the cost of invalid mutations must be paid again.
-However, the locality of mutants and source changes in fact means this
-is seldom required.
-
-The {\tt git merge-file} utility takes a base file and two modified
-versions of that file, and produces an automatic merge, using the
-usual git merge resolution algorithm.    Because a mutant is a (very
-small) source change, this means that in practice updating a mutant to
-reflect even large changes to a source file is cost-free, and if the
-original mutant was valid, the new mutant will usually also be valid.
-New mutant generation is only required for new lines of code and
-modified lines of code.  To the extent development is incremental,
-therefore, mutant generation can also be incremental.
-
-\section{Beyond Regular Expressions with Parser Parser Combinators}
+\subsection{Beyond Regular Expressions with Parser Parser Combinators}
 % Better to frame this as "Beyond regex with PPC" than "PPC instead of RE" 
@@
 
-\clg{OK.  I think maybe ??? This section works better as a subsection in the
-  previous section, which can build to ``beyond...'' Then this text can/should
-  start with a TINY bit more transition from the regular expression thing we've
-  discussed so far to motivate comby ``instead'' (even though it's not really
-  instead.)  This is more a point for discussion than a unilateral decree, but
-  it feels like it flows better as a complete story.  Otherwise it's not super
-  clear what the point of Section 2 was, if Section 3 is the real thing we're
-  proposing? Is it? On that note, if we do this, I think we can/should move
-  ``Incremental Mutation'' down to implementation even if we didn't do it, just
-  as a point of discussion like ``oh btw we could do it if we wanted it's
-  easy.'' Otherwise it kind of disrupts the flow AND hierararies of operators
-  leads REALLY REALLY nicely into the parse parser combinator approach.}
-\clg{Either way regardless this needs a better starting transition haha.}
-We implement a lightweight parser-driven approach that attunes mutation
-operators to language-general syntax-aware transformations. For example, our
-operators can ensure syntactically well-formed outputs (e.g., transformations
-must preserve balanced parentheses or braces). More generally, operators are
-sensitive to coarse-grained context-free grammar properties. These operators go
-beyond the expressivity of regular expressions (which are typically limited to
-lexical transformations) to generate \emph{varied} and \emph{valid}
-programs during mutation testing.
+While regular expressions have been shown to be capable of driving mutant generation, we
+propose going beyond regular expressions using a lightweight
+parser-driven approach that attunes mutation operators to
+language-general syntax-aware transformations. Such operators can
+perform tasks that go beyond regular transforms and ensure
+syntactically well-formed outputs (e.g., transformations must preserve
+balanced parentheses or braces). More generally, operators can be made
+sensitive to coarse-grained context-free grammar properties. In short,
+these
+operators go beyond the expressivity of regular expressions (which are
+typically limited to lexical transformations) to generate more
+\emph{varied} and more \emph{valid} mutants.
 
@@
 particular, we
-build mutation operators with \texttt{comby},\footnote{https://comby.dev} the tool that implements parser parser
+build mutation operators with Comby\footnote{https://comby.dev}, the tool that implements parser parser
 combinators to declaratively match and rewrite source code syntax.
@@
 languages and implements a generic parser for additional languages to recognize
-these syntactic constructs. The following is a brief overview of Comby
+their syntactic constructs. The following is a brief overview of Comby
 definitions and syntax for mutation operators implemented in our approach:
@@
 
+Table~\ref{tab:crules} shows how the rules in Table~\ref{tab:rules}
+look, when implemented as Comby patterns.  While some simple rules are
+unchanged, others are either more succinct or provide more context as
+to applicability (e.g., only less than between valid holes should be
+mutated).  This shows that a
+further major benefit of using Comby: once a user has learned
+the simple concepts above, it is often much easier to read---and
+especially \emph{write}---mutation
+operator definitions in Comby than using regular expressions.  The
+added precision comes not only ``for free'' but in fact yields a
+benefit for the ``end-user-programmer'' of mutations.
+
 Comby provides mechanisms to further define custom match syntax and behavior
-beyond these standard patterns.\footnote{https://comby.dev/docs/advanced-usage\#custom-language-definitions} \texttt{Comby} patterns can also embed regular expressions, subsuming their
+beyond these standard patterns.\footnote{https://comby.dev/docs/advanced-usage\#custom-language-definitions} Comby patterns can also embed regular expressions, subsuming their
 expressive power to enable regex-based mutation operators intermixed with
-context-free syntax properties.\footnote{https://comby.dev/docs/basic-usage\#using-regular-expressions}
+context-free syntax
+properties\footnote{https://comby.dev/docs/basic-usage\#using-regular-expressions}.
+
+\begin{table}[hbtp]
+\centering
+\caption{Parser Parser Combinator Rule Examples}
+\label{tab:crules}
+\resizebox{\columnwidth}{!}{%
+
+  \begin{tabular}{|l|l|l|}
+    \hline
+    Universal & C-Like & Language-Specific \\
+    \hline
+{\tt :[a]<:[b] ==> :[a]>:[b]} &{\tt  else ==>} & {\tt synchronized ==>} \\ 
+{\tt :[a]<:[b] ==> :[a]==:[b]} & {\tt if (:[cond]) ==> if (!(:[cond]))} &
+                                                                 {\tt
+                                                                          :[ s]:[expr]:[lf~[\textbackslash n]] ==> :[s]pass:[lf]}\\
+{\tt :[a]<:[b] ==> :[a]<=:[b]} & {\tt if (:[cond]) ==> if (1==1)} & {\tt any\_of ==> all\_of}\\
+\hline
+
+\end{tabular}
+}
+\end{table}
 
 \section{Implementation}
+\label{sec:imp}
 
@@
 Rust, Solidity, Fe, Vyper, JavaScript, and Go, required another 436
-lines; this is the total for comby and regexp-based rules, so each
+lines; this is the total for Comby and regexp-based rules, so each
 version requires only 218 lines of rule definitions.  The tool makes
@@
 purposes of blinding.
+
+\subsection{Incremental Mutation}
+
+One obvious concern with any-language source-based mutation is that every time the
+program changes, the cost of invalid mutants must be paid again.
+However, the locality of mutants and source changes in fact means this
+is seldom required.
+
+The {\tt git merge-file} utility takes a base file and two modified
+versions of that file, and produces an automatic merge, using the
+usual git merge resolution algorithm.    Because a mutant is a (very
+small) source change, this means that in practice updating a mutant to
+reflect even large changes to a source file is cost-free, and if the
+original mutant was valid, the new mutant will usually also be valid.
+New mutant generation is only required for new lines of code and
+modified lines of code.  To the extent development is incremental,
+therefore, mutant generation can also be incremental.  Our
+implementation allows the use of {\tt git merge-file} to control the
+modification of most mutants, and uses our mechanism for
+selective mutant generation based on coverage to force mutation where
+there is a conflict in a ``mutant patch.''
 
@@
 multiple maintained, AST-based tools.  Finally, Rust is an ``emerging
-language'' with few tools targeting it.
-\clg{Kinda want a table (or a pointer to tables) with projects per language.
-  Definitely want a pointer to at least Table 6, with the tools we compare to.}
+language'' with few tools targeting it.  Table~\ref{tab:mutationtools}
+gives an overview of tools we identified for each language as working tools
+capable of single-file mutant generation.
 
@@
 Tables~\ref{tab:table_cpp1}-\ref{tab:table_rust1} compare generated
-mutants for ``Regex'' and ``Comby'' modes of our tool.  Invalid
+mutants for Regex and Comby modes of our tool.  Invalid
 mutants are mutants that the tool produces, but which fail to

### 2023-03-29 paper/main.tex
combined C++ and java tables nicer spacing bigger font, but a lot of space on summary stats making me have sadz

--- 2023-03-29 paper/main.tex
@@
 \caption{Regex vs. Comby for C++ (top) and Java (bottom)}
-{\begin{subtable}[h]{\columnwidth}
-%\caption{C++ (Regex vs. Comby)}
-\label{tab:table_cpp1}
+\label{tab:table_cpp1java}
 \begin{tabularx}{\columnwidth}{X|r|cr|cr}
 \toprule
+\mc{6}{c}{\textbf{C++}} \\\midrule
                  &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\[1ex]
@@
  \end{tabularx}
-\end{subtable}}
-\vspace{1ex}
-
-{\begin{subtable}[h]{\columnwidth}
-%\caption{Java }
-\label{tab:table_java1}
 \begin{tabularx}{\columnwidth}{X|r|cr|cr}
 \toprule
+\mc{6}{c}{\textbf{Java}} \\\midrule
                  &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
@@
  egads          & 252 & $\frac{624}{2427}$ & 25.43\% & $\frac{597}{1163}$ & 51.33\%  \\[1ex]
-apk-parser     & 324 & $\frac{351}{1583}$ & 22.17\% & $\frac{333}{1199}$ & 27.77\%  \\[1ex]
+ apk-parser     & 324 & $\frac{351}{1583}$ & 22.17\% & $\frac{333}{1199}$ & 27.77\%  \\[1ex]\midrule
 \end{tabularx}
@@
 \end{tabularx}
-\end{subtable}}
 \end{table}
@@
 
-% ---------------- JAVA table UM Only--------------------------
-\begin{table}[hbtp]
-\centering
-\end{table}
-
-% ---------------- JAVA table UM Only--------------------------
-\begin{table}[hbtp]
-\centering
-\caption{Java (Regex vs. Comby)}
-\label{tab:table_java1}
-\resizebox{\columnwidth}{!}{%
-
-\begin{tabular}{|c|c|c|r|r|r|c|}
-\hline
- & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%} & \textbf{LOC}\\ \hline
-\multirow{2}{*}{pebble} & \multicolumn{2}{c|}{Regex} & 86 \hspace{8pt} & 928 \hspace{8pt} & 8.48\%  & \multirow{2}{*}{279}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 85 \hspace{8pt} & 667 \hspace{8pt} & 11.30\% & \\ \hline
-\multirow{2}{*}{spring-hateoas} & \multicolumn{2}{c|}{Regex} & 248 \hspace{8pt} & 2231 \hspace{8pt} & 10.00\% & \multirow{2}{*}{397}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 172 \hspace{8pt} & 1212 \hspace{8pt} & 12.43\% & \\ \hline
-\multirow{2}{*}{javacc} & \multicolumn{2}{c|}{Regex} & 542 \hspace{8pt} & 1346 \hspace{8pt} & 28.70\% & \multirow{2}{*}{363}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 494 \hspace{8pt} & 782 \hspace{8pt} & 38.71\% & \\ \hline
-
-\multirow{2}{*}{coffee-gb} & \multicolumn{2}{c|}{Regex} & 224 \hspace{8pt} & 1703 \hspace{8pt} & 11.63\% & \multirow{2}{*}{463}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 174 \hspace{8pt} & 624 \hspace{8pt} & 21.80\% & \\ \hline
-\multirow{2}{*}{egads} & \multicolumn{2}{c|}{Regex} & 624 \hspace{8pt} & 1830 \hspace{8pt} & 25.43\% & \multirow{2}{*}{252}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 597 \hspace{8pt} & 566 \hspace{8pt} & 51.33\% & \\ \hline
-\multirow{2}{*}{apk-parser} & \multicolumn{2}{c|}{Regex} & 351 \hspace{8pt} & 1232 \hspace{8pt} & 22.17\% & \multirow{2}{*}{324}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 333 \hspace{8pt} & 866 \hspace{8pt} & 27.77\% & \\ \hline
-
-\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 345.83 & 1545.00  & 17.74\%& \\\cline{3-6}
-    &   &  Median & 299.50 & 1524.50 & 16.90\%  & \\\cline{3-6}
-    &   &  Std. Dev & 203.86 & 467.98 & 8.73\%  & \\\cline{2-6}
-
- & \multirow{3}{*}{Comby} & Mean & 309.17 & 786.17 & 27.22\% & \\\cline{3-6}
-    &   &  Median & 253.50 & 724.50 & 24.79\%  & \\\cline{3-6}
-    &   &  Std. Dev & 202.44 & 235.24  & 15.59\%  & \\\hline
-
-\end{tabular}
-
-}
-\end{table}
+% % ---------------- JAVA table UM Only--------------------------
+% \begin{table}[hbtp]
+% \centering
+% \caption{Java (Regex vs. Comby)}
+% \label{tab:table_java1}
+% \resizebox{\columnwidth}{!}{%
+
+% \begin{tabular}{|c|c|c|r|r|r|c|}
+% \hline
+%  & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%} & \textbf{LOC}\\ \hline
+% \multirow{2}{*}{pebble} & \multicolumn{2}{c|}{Regex} & 86 \hspace{8pt} & 928 \hspace{8pt} & 8.48\%  & \multirow{2}{*}{279}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 85 \hspace{8pt} & 667 \hspace{8pt} & 11.30\% & \\ \hline
+% \multirow{2}{*}{spring-hateoas} & \multicolumn{2}{c|}{Regex} & 248 \hspace{8pt} & 2231 \hspace{8pt} & 10.00\% & \multirow{2}{*}{397}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 172 \hspace{8pt} & 1212 \hspace{8pt} & 12.43\% & \\ \hline
+% \multirow{2}{*}{javacc} & \multicolumn{2}{c|}{Regex} & 542 \hspace{8pt} & 1346 \hspace{8pt} & 28.70\% & \multirow{2}{*}{363}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 494 \hspace{8pt} & 782 \hspace{8pt} & 38.71\% & \\ \hline
+
+% \multirow{2}{*}{coffee-gb} & \multicolumn{2}{c|}{Regex} & 224 \hspace{8pt} & 1703 \hspace{8pt} & 11.63\% & \multirow{2}{*}{463}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 174 \hspace{8pt} & 624 \hspace{8pt} & 21.80\% & \\ \hline
+% \multirow{2}{*}{egads} & \multicolumn{2}{c|}{Regex} & 624 \hspace{8pt} & 1830 \hspace{8pt} & 25.43\% & \multirow{2}{*}{252}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 597 \hspace{8pt} & 566 \hspace{8pt} & 51.33\% & \\ \hline
+% \multirow{2}{*}{apk-parser} & \multicolumn{2}{c|}{Regex} & 351 \hspace{8pt} & 1232 \hspace{8pt} & 22.17\% & \multirow{2}{*}{324}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 333 \hspace{8pt} & 866 \hspace{8pt} & 27.77\% & \\ \hline
+
+% \multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 345.83 & 1545.00  & 17.74\%& \\\cline{3-6}
+%     &   &  Median & 299.50 & 1524.50 & 16.90\%  & \\\cline{3-6}
+%     &   &  Std. Dev & 203.86 & 467.98 & 8.73\%  & \\\cline{2-6}
+
+%  & \multirow{3}{*}{Comby} & Mean & 309.17 & 786.17 & 27.22\% & \\\cline{3-6}
+%     &   &  Median & 253.50 & 724.50 & 24.79\%  & \\\cline{3-6}
+%     &   &  Std. Dev & 202.44 & 235.24  & 15.59\%  & \\\hline
+
+% \end{tabular}
+
+% }
+% \end{table}
 

### 2023-03-29 paper/main.tex

--- 2023-03-29 paper/main.tex
@@
 \centering
-\caption{C++ (Regex vs. Comby)}
-\label{tab:table_cpp1}
-\resizebox{\columnwidth}{!}{%
-
-\begin{tabular}{|c|c|c|r|r|r|c|}
-\hline
- & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%} & \textbf{LOC} \\ \hline
-\multirow{2}{*}{ompl} & \multicolumn{2}{c|}{Regex} & 595 \hspace{8pt} & 1938 \hspace{8pt} & 23.49\% & \multirow{2}{*}{525} \\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 564 \hspace{8pt} & 1610 \hspace{8pt} & 25.94\% &  \\ \hline
-\multirow{2}{*}{libgraphqlparser} & \multicolumn{2}{c|}{Regex} & 157 \hspace{8pt} & 584 \hspace{8pt} & 21.19\% & \multirow{2}{*}{137}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 131 \hspace{8pt} & 309 \hspace{8pt} & 29.77\% & \\ \hline
-\multirow{2}{*}{rethinkdb\_rebirth} & \multicolumn{2}{c|}{Regex} & 545 \hspace{8pt} & 1701 \hspace{8pt} & 24.27\% & \multirow{2}{*}{471}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 492 \hspace{8pt} & 1517 \hspace{8pt} & 24.49\% & \\ \hline
-
-\multirow{2}{*}{xoreos} & \multicolumn{2}{c|}{Regex} & 2419 \hspace{8pt} & 1457 \hspace{8pt} & 62.41\% & \multirow{2}{*}{732}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 2280 \hspace{8pt} & 1017 \hspace{8pt} & 69.15\% & \\ \hline
-\multirow{2}{*}{libxmljs} & \multicolumn{2}{c|}{Regex} & 881 \hspace{8pt} & 3108 \hspace{8pt} & 22.09\% & \multirow{2}{*}{630}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 790 \hspace{8pt} & 3218 \hspace{8pt} & 19.71\% & \\ \hline
-\multirow{2}{*}{tiny-differentiable-simulator} & \multicolumn{2}{c|}{Regex} & 747 \hspace{8pt} & 535 \hspace{8pt} & 58.27\% & \multirow{2}{*}{195}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 770 \hspace{8pt} & 444 \hspace{8pt} & 63.43\% & \\ \hline
-
-\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 890.67 &	1553.83 & 35.28\% & \\\cline{3-6}
-    &   &  Median & 671.00 & 1579.00 & 23.88\%  & \\\cline{3-6}
-    &   &  Std. Dev & 787.62 & 956.31 & 19.48\% & \\\cline{2-6}
-
- & \multirow{3}{*}{Comby} & Mean & 837.83 & 1352.50 & 38.75\% & \\\cline{3-6}
-    &   &  Median & 667.00 & 1267.00 & 27.86\% & \\\cline{3-6}
-    &   &  Std. Dev & 745.78 & 1058.33	& 21.66\% & \\\hline
-
-\end{tabular}
-}
-
+\caption{Regex vs. Comby for C++ (top) and Java (bottom)}
+\label{tab:table_cpp1java}
+\begin{tabularx}{\columnwidth}{X|r|cr|cr}
+\toprule
+\mc{6}{c}{\textbf{C++}} \\\midrule
+                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\[1ex]
+\textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
+ompl & 525 & $\frac{595}{2533}$ & 23.49\% &  $\frac{564}{2127}$  & 25.94\%   \\[1ex]
+libgraphqlparser& 137 & $\frac{157}{741}$ & 21.19\%  & $\frac{131}{440}$ & 29.77\% \\[1ex]
+rethinkdb\_rebirth &471 &  $\frac{545}{2246}$ & 24.27\% & $\frac{492}{2009}$ & 24.49\% \\[1ex]
+xoreos & 732  & $\frac{2419}{3876}$  & 62.41\% & $\frac{2280}{3297}$ & 69.15\% \\[1ex]
+libxmljs &630 &  $\frac{881}{3989}$ & 22.09\% & $\frac{790}{4008}$ & 19.71\% \\[1ex]
+tiny-diff.-simulator &  195 & $\frac{747}{1282}$ &58.27\% &$\frac{770}{1214}$ & 63.43\% \\[1ex]
+\bottomrule
+\end{tabularx}
+ \begin{tabularx}{\columnwidth}{Xllrrr}
+                  & &       & Valid  & Invalid  & \% \\\midrule
+ \multirow{6}{*}{\textbf{Overall}} & \multirow{3}{*}{Regex} & Mean     & 890.67 &  1553.83 & 35.28\% \\
+                          & & Med.    & 671.00 & 1579.00  & 23.88\%   \\
+                          & & Std.Dev. & 787.62 & 956.31   & 19.48\%  \\\cline{2-6}
+  & \multirow{3}{*}{Comby}  & Mean     & 837.83 & 1352.50 & 38.75\%  \\
+                         & & Med.    & 667.00 & 1267.00 & 27.86\%  \\
+                         & & Std.Dev. & 745.78 & 1058.33& 21.66\%  \\\bottomrule
+ \end{tabularx}
+\begin{tabularx}{\columnwidth}{X|r|cr|cr}
+\toprule
+\mc{6}{c}{\textbf{Java}} \\\midrule
+                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
+\textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
+ pebble         & 279 & $\frac{86}{1014}$ & 8.48\% & $\frac{85}{752}$ & 11.30\%  \\[1ex]
+ spring-hateoas & 397 & $\frac{248}{2479}$ & 10.00\% & $\frac{172}{1384}$ & 12.43\%  \\[1ex]
+ javacc         & 363 & $\frac{542}{1888}$ & 28.70\% & $\frac{494}{1281}$ & 38.71\%  \\[1ex]
+ coffee-gb      & 463 & $\frac{224}{1927}$ & 11.63\% & $\frac{174}{798}$ & 21.80\% \\[1ex]
+ egads          & 252 & $\frac{624}{2427}$ & 25.43\% & $\frac{597}{1163}$ & 51.33\%  \\[1ex]
+ apk-parser     & 324 & $\frac{351}{1583}$ & 22.17\% & $\frac{333}{1199}$ & 27.77\%  \\[1ex]\midrule
+\end{tabularx}
+\begin{tabularx}{\columnwidth}{Xllrrr}
+              & & & Valid  & Invalid  & \% \\\midrule
+\multirow{6}{*}{\textbf{Overall}} & \mr{3}{Regex} & Mean     &  345.83 & 1545.00  & 17.74\% \\
+              &   & Med.     & 299.50 & 1524.50 & 16.90\%  \\
+              &   & Std.Dev. & 203.86 & 467.98 & 8.73\%  \\\cline{2-6}
+& \mr{3}{Comby} & Mean     & 309.17 & 786.17 & 27.22\% \\
+              &   & Med.     & 253.50 & 724.50 & 24.79\% \\
+              &   & Std.Dev. & 202.44 & 235.24  & 15.59\% \\\bottomrule
+\end{tabularx}
 \end{table}
 
-% ---------------- JAVA table UM Only--------------------------
-\begin{table}[hbtp]
-\centering
-\caption{Java (Regex vs. Comby)}
-\label{tab:table_java1}
-\resizebox{\columnwidth}{!}{%
-
-\begin{tabular}{|c|c|c|r|r|r|c|}
-\hline
- & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%} & \textbf{LOC}\\ \hline
-\multirow{2}{*}{pebble} & \multicolumn{2}{c|}{Regex} & 86 \hspace{8pt} & 928 \hspace{8pt} & 8.48\%  & \multirow{2}{*}{279}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 85 \hspace{8pt} & 667 \hspace{8pt} & 11.30\% & \\ \hline
-\multirow{2}{*}{spring-hateoas} & \multicolumn{2}{c|}{Regex} & 248 \hspace{8pt} & 2231 \hspace{8pt} & 10.00\% & \multirow{2}{*}{397}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 172 \hspace{8pt} & 1212 \hspace{8pt} & 12.43\% & \\ \hline
-\multirow{2}{*}{javacc} & \multicolumn{2}{c|}{Regex} & 542 \hspace{8pt} & 1346 \hspace{8pt} & 28.70\% & \multirow{2}{*}{363}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 494 \hspace{8pt} & 782 \hspace{8pt} & 38.71\% & \\ \hline
-
-\multirow{2}{*}{coffee-gb} & \multicolumn{2}{c|}{Regex} & 224 \hspace{8pt} & 1703 \hspace{8pt} & 11.63\% & \multirow{2}{*}{463}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 174 \hspace{8pt} & 624 \hspace{8pt} & 21.80\% & \\ \hline
-\multirow{2}{*}{egads} & \multicolumn{2}{c|}{Regex} & 624 \hspace{8pt} & 1830 \hspace{8pt} & 25.43\% & \multirow{2}{*}{252}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 597 \hspace{8pt} & 566 \hspace{8pt} & 51.33\% & \\ \hline
-\multirow{2}{*}{apk-parser} & \multicolumn{2}{c|}{Regex} & 351 \hspace{8pt} & 1232 \hspace{8pt} & 22.17\% & \multirow{2}{*}{324}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 333 \hspace{8pt} & 866 \hspace{8pt} & 27.77\% & \\ \hline
-
-\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 345.83 & 1545.00  & 17.74\%& \\\cline{3-6}
-    &   &  Median & 299.50 & 1524.50 & 16.90\%  & \\\cline{3-6}
-    &   &  Std. Dev & 203.86 & 467.98 & 8.73\%  & \\\cline{2-6}
-
- & \multirow{3}{*}{Comby} & Mean & 309.17 & 786.17 & 27.22\% & \\\cline{3-6}
-    &   &  Median & 253.50 & 724.50 & 24.79\%  & \\\cline{3-6}
-    &   &  Std. Dev & 202.44 & 235.24  & 15.59\%  & \\\hline
-
-\end{tabular}
-
-}
-\end{table}
+% % ---------------- C++ table UM Only--------------------------
+% \begin{table}[hbtp]
+% \centering
+% \caption{C++ (Regex vs. Comby)}
+% \label{tab:table_cpp1}
+% \resizebox{\columnwidth}{!}{%
+
+% \begin{tabular}{|c|c|c|r|r|r|c|}
+% \toprule
+%  & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%} & \textbf{LOC} \\ \hline
+% \multirow{2}{*}{ompl} & \multicolumn{2}{c|}{Regex} & 595 \hspace{8pt} & 1938 \hspace{8pt} & 23.49\% & \multirow{2}{*}{525} \\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 564 \hspace{8pt} & 1610 \hspace{8pt} & 25.94\% &  \\ \hline
+% \multirow{2}{*}{libgraphqlparser} & \multicolumn{2}{c|}{Regex} & 157 \hspace{8pt} & 584 \hspace{8pt} & 21.19\% & \multirow{2}{*}{137}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 131 \hspace{8pt} & 309 \hspace{8pt} & 29.77\% & \\ \hline
+% \multirow{2}{*}{rethinkdb\_rebirth} & \multicolumn{2}{c|}{Regex} & 545 \hspace{8pt} & 1701 \hspace{8pt} & 24.27\% & \multirow{2}{*}{471}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 492 \hspace{8pt} & 1517 \hspace{8pt} & 24.49\% & \\ \hline
+
+% \multirow{2}{*}{xoreos} & \multicolumn{2}{c|}{Regex} & 2419 \hspace{8pt} & 1457 \hspace{8pt} & 62.41\% & \multirow{2}{*}{732}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 2280 \hspace{8pt} & 1017 \hspace{8pt} & 69.15\% & \\ \hline
+% \multirow{2}{*}{libxmljs} & \multicolumn{2}{c|}{Regex} & 881 \hspace{8pt} & 3108 \hspace{8pt} & 22.09\% & \multirow{2}{*}{630}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 790 \hspace{8pt} & 3218 \hspace{8pt} & 19.71\% & \\ \hline
+% \multirow{2}{*}{tiny-differentiable-simulator} & \multicolumn{2}{c|}{Regex} & 747 \hspace{8pt} & 535 \hspace{8pt} & 58.27\% & \multirow{2}{*}{195}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 770 \hspace{8pt} & 444 \hspace{8pt} & 63.43\% & \\ \hline
+
+% \multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 890.67 &	1553.83 & 35.28\% & \\\cline{3-6}
+%     &   &  Median & 671.00 & 1579.00 & 23.88\%  & \\\cline{3-6}
+%     &   &  Std. Dev & 787.62 & 956.31 & 19.48\% & \\\cline{2-6}
+
+%  & \multirow{3}{*}{Comby} & Mean & 837.83 & 1352.50 & 38.75\% & \\\cline{3-6}
+%     &   &  Median & 667.00 & 1267.00 & 27.86\% & \\\cline{3-6}
+%     &   &  Std. Dev & 745.78 & 1058.33	& 21.66\% & \\\hline
+
+% \end{tabular}
+% }
+
+% \end{table}
+
+% % ---------------- JAVA table UM Only--------------------------
+% \begin{table}[hbtp]
+% \centering
+% \caption{Java (Regex vs. Comby)}
+% \label{tab:table_java1}
+% \resizebox{\columnwidth}{!}{%
+
+% \begin{tabular}{|c|c|c|r|r|r|c|}
+% \hline
+%  & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%} & \textbf{LOC}\\ \hline
+% \multirow{2}{*}{pebble} & \multicolumn{2}{c|}{Regex} & 86 \hspace{8pt} & 928 \hspace{8pt} & 8.48\%  & \multirow{2}{*}{279}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 85 \hspace{8pt} & 667 \hspace{8pt} & 11.30\% & \\ \hline
+% \multirow{2}{*}{spring-hateoas} & \multicolumn{2}{c|}{Regex} & 248 \hspace{8pt} & 2231 \hspace{8pt} & 10.00\% & \multirow{2}{*}{397}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 172 \hspace{8pt} & 1212 \hspace{8pt} & 12.43\% & \\ \hline
+% \multirow{2}{*}{javacc} & \multicolumn{2}{c|}{Regex} & 542 \hspace{8pt} & 1346 \hspace{8pt} & 28.70\% & \multirow{2}{*}{363}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 494 \hspace{8pt} & 782 \hspace{8pt} & 38.71\% & \\ \hline
+
+% \multirow{2}{*}{coffee-gb} & \multicolumn{2}{c|}{Regex} & 224 \hspace{8pt} & 1703 \hspace{8pt} & 11.63\% & \multirow{2}{*}{463}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 174 \hspace{8pt} & 624 \hspace{8pt} & 21.80\% & \\ \hline
+% \multirow{2}{*}{egads} & \multicolumn{2}{c|}{Regex} & 624 \hspace{8pt} & 1830 \hspace{8pt} & 25.43\% & \multirow{2}{*}{252}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 597 \hspace{8pt} & 566 \hspace{8pt} & 51.33\% & \\ \hline
+% \multirow{2}{*}{apk-parser} & \multicolumn{2}{c|}{Regex} & 351 \hspace{8pt} & 1232 \hspace{8pt} & 22.17\% & \multirow{2}{*}{324}\\\cline{2-6}
+%     & \multicolumn{2}{c|}{Comby} & 333 \hspace{8pt} & 866 \hspace{8pt} & 27.77\% & \\ \hline
+
+% \multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 345.83 & 1545.00  & 17.74\%& \\\cline{3-6}
+%     &   &  Median & 299.50 & 1524.50 & 16.90\%  & \\\cline{3-6}
+%     &   &  Std. Dev & 203.86 & 467.98 & 8.73\%  & \\\cline{2-6}
+
+%  & \multirow{3}{*}{Comby} & Mean & 309.17 & 786.17 & 27.22\% & \\\cline{3-6}
+%     &   &  Median & 253.50 & 724.50 & 24.79\%  & \\\cline{3-6}
+%     &   &  Std. Dev & 202.44 & 235.24  & 15.59\%  & \\\hline
+
+% \end{tabular}
+
+% }
+% \end{table}
 

### 2023-03-29 paper/main.tex
empty rust table for Kush

@@
 \begin{tabularx}{\columnwidth}{X|r|cr|cr}
-\toprule
+\toprule\toprule
 \mc{6}{c}{\textbf{C++}} \\\midrule
@@
  \multirow{6}{*}{\textbf{Overall}} & \multirow{3}{*}{Regex} & Mean     & 890.67 &  1553.83 & 35.28\% \\
-                          & & Med.    & 671.00 & 1579.00  & 23.88\%   \\
+%                          & & Med.    & 671.00 & 1579.00  & 23.88\%   \\
                           & & Std.Dev. & 787.62 & 956.31   & 19.48\%  \\\cline{2-6}
   & \multirow{3}{*}{Comby}  & Mean     & 837.83 & 1352.50 & 38.75\%  \\
-                         & & Med.    & 667.00 & 1267.00 & 27.86\%  \\
+%                         & & Med.    & 667.00 & 1267.00 & 27.86\%  \\
                          & & Std.Dev. & 745.78 & 1058.33& 21.66\%  \\\bottomrule
@@
 \multirow{6}{*}{\textbf{Overall}} & \mr{3}{Regex} & Mean     &  345.83 & 1545.00  & 17.74\% \\
-              &   & Med.     & 299.50 & 1524.50 & 16.90\%  \\
+%              &   & Med.     & 299.50 & 1524.50 & 16.90\%  \\
               &   & Std.Dev. & 203.86 & 467.98 & 8.73\%  \\\cline{2-6}
 & \mr{3}{Comby} & Mean     & 309.17 & 786.17 & 27.22\% \\
-              &   & Med.     & 253.50 & 724.50 & 24.79\% \\
+%              &   & Med.     & 253.50 & 724.50 & 24.79\% \\
               &   & Std.Dev. & 202.44 & 235.24  & 15.59\% \\\bottomrule
@@
 
+\iffalse
+\begin{table}[hbtp]
+\centering
+\caption{Regex vs. Comby for Python++, including redundant mutants.}
+\label{tab:table_pyrust}
+\begin{tabularx}{\columnwidth}{X|r|rcr|rcr}
+\toprule\toprule
+\mc{6}{c}{\textbf{Python++}} \\\midrule
+                 &                 & \multicolumn{3}{c|}{\textbf{Regex}}             &\multicolumn{3}{c}{\textbf{Comby}}  \\[1ex]
+\textbf{Project} & \textbf{LOC}    & redundant & $\frac{valid}{total}$ & \textbf{\%} & redundant & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
+keract           & 402             & 729       & $\frac{1655}{3984}$   & 41.53\%     & 533       & $\frac{1471}{3067}$   & 47.96\% \\[1ex]
+dtw              & 118             & 567       & $\frac{973}{2106}$    & 46.20\%     & 639       & $\frac{866}{1836}$    & 47.19\% & \\[1ex]
+notion-sdk-py    & 189             & 82        & $\frac{196}{1038}$    & 18.89\%     & 122       & $\frac{196}{1019}$    & 19.23\% & \\[1ex]
+{\makecell{atlassian-python \\ api}} & 236 & $\frac{302}{1304}$  & 207  &	19.99\% & \multirow{2}{*}{236}\\\cline{2-7}
+       & \multicolumn{2}{c|}{Comby} & 333  & 635  & 351  & 25.25\% & \\ \hline
+ESD    & 698 & $\frac{1859}{4626}$  & 1642  & 29.46\% & \multirow{2}{*}{698}\\\cline{2-7}
+       & \multicolumn{2}{c|}{Comby} & 1712  & 1548  &	614  & 44.18\% & \\ \hline
+fiber  & 358 & $\frac{493}{1402}$  & 163  & 31.08\% & \multirow{2}{*}{358}\\\cline{2-7}
+       & \multicolumn{2}{c|}{Comby} & 458  & 767 	& 241  & 31.24\% & \\ \hline
+
+\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 913.00 & 1267.33 & 565.00 & 31.19\%  & \\\cline{3-7}
+    &   &  Std. Dev & 708.93 & 813.35 & 584.79 & 11.06\%  & \\\cline{2-7}
+
+ & \multirow{3}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% & \\\cline{3-7}
+    &   &  Std. Dev & 628.79 & 418.99 & 211.62 & 12.28\%  & \\\hline
+\end{tabularx}
+\end{table}
+\fi
 % ---------------- Python table UM Only --------------------------
@@
 
+\end{tabular}
+
+}
+\end{table}
+
+% ---------------- Rust table UM Only --------------------------
+\begin{table}[hbtp]
+\centering
+{\small
+\caption{Rust (Regex vs. Comby)}
+\label{tab:table_rust1}
+\begin{tabularx}{\columnwidth}{X|r|cr|cr}
+\toprule
+                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
+\textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
+cargo release    & ...    & & & \\[1ex]
+\end{tabularx}
+ \begin{tabularx}{\columnwidth}{Xllrrr}
+               & & & Valid  & Invalid  & \% \\\midrule
+ \multirow{6}{*}{\textbf{Overall}} & \mr{3}{Regex} & Mean     &   &  & \\
+               &   & Std.Dev. & &  &   \\\cline{2-6}
+ & \mr{3}{Comby} & Mean     &  &  &  \\
+ %              &   & Med.     & 253.50 & 724.50 & 24.79\% \\
+               &   & Std.Dev. & &  &   \\\bottomrule
+\end{tabularx}}
+\end{table}
+
+\begin{table}[hbtp]
+\centering
+\caption{Rust (Regex vs. Comby)}
+\label{tab:table_rust1}
+\resizebox{\columnwidth}{!}{%
+\begin{tabular}{|c|c|c|r|r|r|c|}
+\hline
+ & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%} &  \textbf{LOC} \\ \hline
+\multirow{2}{*}{cargo release} & \multicolumn{2}{c|}{Regex} & 851 \hspace{8pt} &	1163 \hspace{8pt} & 42.25\%  & \multirow{2}{*}{346}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 1690 \hspace{8pt} & 711 \hspace{8pt} & 70.39\% & \\ \hline
+\multirow{2}{*}{passerine} & \multicolumn{2}{c|}{Regex} & 213 \hspace{8pt} & 2081 \hspace{8pt} & 9.29\% & \multirow{2}{*}{376}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 173 \hspace{8pt} & 1518 \hspace{8pt} & 10.23\% & \\ \hline
+\multirow{2}{*}{typos} & \multicolumn{2}{c|}{Regex} & 386 \hspace{8pt} & 1534 \hspace{8pt} & 20.10\% & \multirow{2}{*}{463}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 366 \hspace{8pt} & 1199 \hspace{8pt} &	 23.39\% & \\ \hline
+
+\multirow{2}{*}{ord} & \multicolumn{2}{c|}{Regex} & 762 \hspace{8pt}	& 2096 \hspace{8pt} & 26.66\% & \multirow{2}{*}{533}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 643 \hspace{8pt} & 1669 \hspace{8pt} & 27.81\% & \\ \hline
+\multirow{2}{*}{bazuka} & \multicolumn{2}{c|}{Regex} & 451 \hspace{8pt} & 1103 \hspace{8pt} & 29.02\% & \multirow{2}{*}{285}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 402 \hspace{8pt} & 920 \hspace{8pt} & 30.41\% & \\ \hline
+\multirow{2}{*}{strum} & \multicolumn{2}{c|}{Regex} & 51 \hspace{8pt} & 377 \hspace{8pt}	& 11.92\% & \multirow{2}{*}{90}\\\cline{2-6}
+    & \multicolumn{2}{c|}{Comby} & 50 \hspace{8pt} & 318 \hspace{8pt}	& 13.51\% & \\ \hline
+
+\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 452.33 & 1392.33	& 23.21\% & \\\cline{3-6}
+    &   &  Median & 418.50 & 1348.50 & 23.38\%  & \\\cline{3-6}
+    &   &  Std. Dev & 309.05 & 656.76 & 12.16\%  & \\\cline{2-6}
+
+ & \multirow{3}{*}{Comby} & Mean & 554.00 & 1055.83	& 29.29\% & \\\cline{3-6}
+    &   &  Median & 384.00 & 1059.50	& 25.60\%  & \\\cline{3-6}
+    &   &  Std. Dev & 592.63 & 508.30 & 21.63\%  & \\\hline
 \end{tabular}

### 2023-03-29 paper/main.tex
starter python for regex vs comby

@@
 
-\iffalse
 \begin{table}[hbtp]
 \centering
-\caption{Regex vs. Comby for Python++, including redundant mutants.}
+{\small
+\caption{Regex vs. Comby for Python, including redundant mutants.}
 \label{tab:table_pyrust}
@@
 \toprule\toprule
-\mc{6}{c}{\textbf{Python++}} \\\midrule
                  &                 & \multicolumn{3}{c|}{\textbf{Regex}}             &\multicolumn{3}{c}{\textbf{Comby}}  \\[1ex]
-\textbf{Project} & \textbf{LOC}    & redundant & $\frac{valid}{total}$ & \textbf{\%} & redundant & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
+\textbf{Project} & \textbf{LOC}    & red.  & $\frac{valid}{total}$ & \textbf{\%} & red. & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
 keract           & 402             & 729       & $\frac{1655}{3984}$   & 41.53\%     & 533       & $\frac{1471}{3067}$   & 47.96\% \\[1ex]
-dtw              & 118             & 567       & $\frac{973}{2106}$    & 46.20\%     & 639       & $\frac{866}{1836}$    & 47.19\% & \\[1ex]
-notion-sdk-py    & 189             & 82        & $\frac{196}{1038}$    & 18.89\%     & 122       & $\frac{196}{1019}$    & 19.23\% & \\[1ex]
-{\makecell{atlassian-python \\ api}} & 236 & $\frac{302}{1304}$  & 207  &	19.99\% & \multirow{2}{*}{236}\\\cline{2-7}
-       & \multicolumn{2}{c|}{Comby} & 333  & 635  & 351  & 25.25\% & \\ \hline
-ESD    & 698 & $\frac{1859}{4626}$  & 1642  & 29.46\% & \multirow{2}{*}{698}\\\cline{2-7}
-       & \multicolumn{2}{c|}{Comby} & 1712  & 1548  &	614  & 44.18\% & \\ \hline
-fiber  & 358 & $\frac{493}{1402}$  & 163  & 31.08\% & \multirow{2}{*}{358}\\\cline{2-7}
-       & \multicolumn{2}{c|}{Comby} & 458  & 767 	& 241  & 31.24\% & \\ \hline
-
-\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 913.00 & 1267.33 & 565.00 & 31.19\%  & \\\cline{3-7}
-    &   &  Std. Dev & 708.93 & 813.35 & 584.79 & 11.06\%  & \\\cline{2-7}
-
- & \multirow{3}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% & \\\cline{3-7}
-    &   &  Std. Dev & 628.79 & 418.99 & 211.62 & 12.28\%  & \\\hline
+dtw              & ... \\
+%
 \end{tabularx}
+\begin{tabularx}{\columnwidth}{Xllrrrrrr}
+                  && &  Red.     & Valid  & Invalid  & \% \\\midrule
+\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean &  565.00 & 913.00 & 1267.33 & 31.19\%  & \\
+    &   &  Std. Dev & 584.79  & 708.93 & 813.35 & 11.06\%  \\
+ & \multirow{3}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% \\
+  &   &  Std. Dev &  211.62 & 628.79 & 418.99 & 12.28\%   \\
+\end{tabularx}}
 \end{table}
-\fi
+
 % ---------------- Python table UM Only --------------------------
@@
 % ---------------- C++ table (UM vs Others)--------------------------
+\iffalse
+\begin{table}[htbp]
+\centering
+\caption{C++ (Our Implementation vs. Dextool)}
+\label{tab:table_cpp2}
+\resizebox{\columnwidth}{!}{
+\begin{tabular}{|c|c|r|r|r|r|}
+\hline
+ & \multicolumn{3}{c|}{Comby}  & \multicolumn{3}{c|}{Dextool} \\ 
+ & $\frac{killed}{total}$ & \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Equiv.} & $\frac{killed}{total}$ & \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Equiv.} \\[1ex]\midrule
+ \multirow{3}{*}{\small ompl} & Comby & 41 & 523 & 0.07 & 40\% \\\cline{2-6}
+    & Dextool & 23 & 255 & 0.08 & N/A \\ \hline
+%\multirow{3}{*}{\small libgraphqlparser} & Comby & 116 & 15 & 0.89 & 20\% \\ \cline{2-6}
+%    & Dextool & 50 & 2 & 0.96 & N/A \\ \hline
+%\multirow{3}{*}{\small rethinkdb\_rebirth} & Comby & 49 & 443 & 0.10 & 20\% \\ \cline{2-6}
+%    & Dextool & 9 & 16 & 0.36 & N/A \\ \hline
+
+%\multirow{3}{*}{\small xoreos} & Comby & 33 & 2247 & 0.01 & 10\% \\ \cline{2-6}
+%    & Dextool & 8 & 406 & 0.02 & N/A \\ \hline
+%\multirow{3}{*}{\small libxmljs}  & Comby & 507 & 283 & 0.64 & 10\% \\ \cline{2-6}
+%    & Dextool & 383 & 447 & 0.46 & N/A \\ \hline
+%\multirow{3}{*}{\small tiny-differentiable-simulator} & Comby & 32 & 738 & 0.04 & 0\% \\ \cline{2-6}
+%    & Dextool & 13 & 48 & 0.21 & N/A\\ \hline
+%    
+%\multirow{9}{*}{Overall}\multirow{3}{*}{Comby} & \multicolumn{2}{c|}{Mean} & 0.29 & 16.67\% \\\cline{3-6}
+%    &   &  \multicolumn{2}{c|}{Median} & 0.09 & 15.00\% \\\cline{3-6}
+%    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.37 & 13.66\% \\\cline{2-6}
+%
+% & \multirow{3}{*}{Dextool} & \multicolumn{2}{c|}{Mean} & 0.33 & N/A\\\cline{3-6}
+%    &   &  \multicolumn{2}{c|}{Median} & 0.29 & N/A \\\cline{3-6}
+%    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.38 & N/A \\ \hline
+\end{tabular}
+}
+\end{table}
+\fi
 

### 2023-03-29 paper/main.tex
new table for C++ us vs. dextool

@@
 % ---------------- C++ table (UM vs Others)--------------------------
-\iffalse
 \begin{table}[htbp]
 \centering
+{\small
 \caption{C++ (Our Implementation vs. Dextool)}
 \label{tab:table_cpp2}
-\resizebox{\columnwidth}{!}{
-\begin{tabular}{|c|c|r|r|r|r|}
+\begin{tabularx}{\columnwidth}{X|crr|crr}
 \hline
- & \multicolumn{3}{c|}{Comby}  & \multicolumn{3}{c|}{Dextool} \\ 
- & $\frac{killed}{total}$ & \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Equiv.} & $\frac{killed}{total}$ & \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Equiv.} \\[1ex]\midrule
- \multirow{3}{*}{\small ompl} & Comby & 41 & 523 & 0.07 & 40\% \\\cline{2-6}
-    & Dextool & 23 & 255 & 0.08 & N/A \\ \hline
+        & \multicolumn{3}{c|}{Comby}  & \multicolumn{3}{c}{Dextool} \\ 
+Project & $\frac{killed}{total}$ & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & $\frac{killed}{total}$ & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} \\[1ex]\midrule
+ ompl & $\frac{41}{564}$ & 0.07 & 40\% & $\frac{23}{278}$ & 0.08 & N/A \\[1ex]
 %\multirow{3}{*}{\small libgraphqlparser} & Comby & 116 & 15 & 0.89 & 20\% \\ \cline{2-6}
@@
 %    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.38 & N/A \\ \hline
-\end{tabular}
+\end{tabularx}
 }
 \end{table}
-\fi
 

### 2023-03-29 paper/main.tex
alex done

--- 2023-03-29 paper/main.tex
@@
 
+It is true that this approach to mutant generation does pay a cost in
+terms of generating invalid mutants that must be rejected at the
+compilation stage of the mutation testing process.  However, as
+observed by Pizzoleto et al. in their systematic review of mutation
+testing~\cite{pizzoleto2019systematic} cost reduction approaches, mutation testing consists of four
+stages: 1) execution of a test suite on the un-mutated program, 2)
+generation of mutants, 3) execution of the test suite on the mutants,
+and 4) analysis (by a human, usually) of the mutants.  It is only steps 3
+and 4 that are ``very costly,'' as Pizzoletto et al. note.  Our approach increases the costs of
+the second step only, and, as we discuss below, can help reduce
+costs of steps 3 and 4.
+
 The contributions of this paper are therefore:
@@
 \begin{tabularx}{\columnwidth}{X|r|cr|cr}
-\toprule\toprule
+\toprule
 \mc{6}{c}{\textbf{C++}} \\\midrule
@@
  \multirow{6}{*}{\textbf{Overall}} & \multirow{3}{*}{Regex} & Mean     & 890.67 &  1553.83 & 35.28\% \\
-%                          & & Med.    & 671.00 & 1579.00  & 23.88\%   \\
+                          & & Med.    & 671.00 & 1579.00  & 23.88\%   \\
                           & & Std.Dev. & 787.62 & 956.31   & 19.48\%  \\\cline{2-6}
   & \multirow{3}{*}{Comby}  & Mean     & 837.83 & 1352.50 & 38.75\%  \\
-%                         & & Med.    & 667.00 & 1267.00 & 27.86\%  \\
+                         & & Med.    & 667.00 & 1267.00 & 27.86\%  \\
                          & & Std.Dev. & 745.78 & 1058.33& 21.66\%  \\\bottomrule
@@
 \multirow{6}{*}{\textbf{Overall}} & \mr{3}{Regex} & Mean     &  345.83 & 1545.00  & 17.74\% \\
-%              &   & Med.     & 299.50 & 1524.50 & 16.90\%  \\
+              &   & Med.     & 299.50 & 1524.50 & 16.90\%  \\
               &   & Std.Dev. & 203.86 & 467.98 & 8.73\%  \\\cline{2-6}
 & \mr{3}{Comby} & Mean     & 309.17 & 786.17 & 27.22\% \\
-%              &   & Med.     & 253.50 & 724.50 & 24.79\% \\
+              &   & Med.     & 253.50 & 724.50 & 24.79\% \\
               &   & Std.Dev. & 202.44 & 235.24  & 15.59\% \\\bottomrule
@@
 
-\iffalse
-\begin{table}[hbtp]
-\centering
-\caption{Regex vs. Comby for Python++, including redundant mutants.}
-\label{tab:table_pyrust}
-\begin{tabularx}{\columnwidth}{X|r|rcr|rcr}
-\toprule\toprule
-\mc{6}{c}{\textbf{Python++}} \\\midrule
-                 &                 & \multicolumn{3}{c|}{\textbf{Regex}}             &\multicolumn{3}{c}{\textbf{Comby}}  \\[1ex]
-\textbf{Project} & \textbf{LOC}    & redundant & $\frac{valid}{total}$ & \textbf{\%} & redundant & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
-keract           & 402             & 729       & $\frac{1655}{3984}$   & 41.53\%     & 533       & $\frac{1471}{3067}$   & 47.96\% \\[1ex]
-dtw              & 118             & 567       & $\frac{973}{2106}$    & 46.20\%     & 639       & $\frac{866}{1836}$    & 47.19\% & \\[1ex]
-notion-sdk-py    & 189             & 82        & $\frac{196}{1038}$    & 18.89\%     & 122       & $\frac{196}{1019}$    & 19.23\% & \\[1ex]
-{\makecell{atlassian-python \\ api}} & 236 & $\frac{302}{1304}$  & 207  &	19.99\% & \multirow{2}{*}{236}\\\cline{2-7}
-       & \multicolumn{2}{c|}{Comby} & 333  & 635  & 351  & 25.25\% & \\ \hline
-ESD    & 698 & $\frac{1859}{4626}$  & 1642  & 29.46\% & \multirow{2}{*}{698}\\\cline{2-7}
-       & \multicolumn{2}{c|}{Comby} & 1712  & 1548  &	614  & 44.18\% & \\ \hline
-fiber  & 358 & $\frac{493}{1402}$  & 163  & 31.08\% & \multirow{2}{*}{358}\\\cline{2-7}
-       & \multicolumn{2}{c|}{Comby} & 458  & 767 	& 241  & 31.24\% & \\ \hline
-
-\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 913.00 & 1267.33 & 565.00 & 31.19\%  & \\\cline{3-7}
-    &   &  Std. Dev & 708.93 & 813.35 & 584.79 & 11.06\%  & \\\cline{2-7}
-
- & \multirow{3}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% & \\\cline{3-7}
-    &   &  Std. Dev & 628.79 & 418.99 & 211.62 & 12.28\%  & \\\hline
-\end{tabularx}
-\end{table}
-\fi
 % ---------------- Python table UM Only --------------------------
@@
 \centering
-{\small
-\caption{Rust (Regex vs. Comby)}
-\label{tab:table_rust1}
-\begin{tabularx}{\columnwidth}{X|r|cr|cr}
-\toprule
-                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
-\textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
-cargo release    & ...    & & & \\[1ex]
-\end{tabularx}
- \begin{tabularx}{\columnwidth}{Xllrrr}
-               & & & Valid  & Invalid  & \% \\\midrule
- \multirow{6}{*}{\textbf{Overall}} & \mr{3}{Regex} & Mean     &   &  & \\
-               &   & Std.Dev. & &  &   \\\cline{2-6}
- & \mr{3}{Comby} & Mean     &  &  &  \\
- %              &   & Med.     & 253.50 & 724.50 & 24.79\% \\
-               &   & Std.Dev. & &  &   \\\bottomrule
-\end{tabularx}}
-\end{table}
-
-\begin{table}[hbtp]
-\centering
 \caption{Rust (Regex vs. Comby)}
@@
 \resizebox{\columnwidth}{!}{%
+
 \begin{tabular}{|c|c|c|r|r|r|c|}
@@
     &   &  Std. Dev & 592.63 & 508.30 & 21.63\%  & \\\hline
+
 \end{tabular}
@@
 
-% ---------------- Rust table UM Only --------------------------
-\begin{table}[hbtp]
-\centering
-\caption{Rust (Regex vs. Comby)}
-\label{tab:table_rust1}
-\resizebox{\columnwidth}{!}{%
-
-\begin{tabular}{|c|c|c|r|r|r|c|}
-\hline
- & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%} &  \textbf{LOC} \\ \hline
-\multirow{2}{*}{cargo release} & \multicolumn{2}{c|}{Regex} & 851 \hspace{8pt} &	1163 \hspace{8pt} & 42.25\%  & \multirow{2}{*}{346}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 1690 \hspace{8pt} & 711 \hspace{8pt} & 70.39\% & \\ \hline
-\multirow{2}{*}{passerine} & \multicolumn{2}{c|}{Regex} & 213 \hspace{8pt} & 2081 \hspace{8pt} & 9.29\% & \multirow{2}{*}{376}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 173 \hspace{8pt} & 1518 \hspace{8pt} & 10.23\% & \\ \hline
-\multirow{2}{*}{typos} & \multicolumn{2}{c|}{Regex} & 386 \hspace{8pt} & 1534 \hspace{8pt} & 20.10\% & \multirow{2}{*}{463}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 366 \hspace{8pt} & 1199 \hspace{8pt} &	 23.39\% & \\ \hline
-
-\multirow{2}{*}{ord} & \multicolumn{2}{c|}{Regex} & 762 \hspace{8pt}	& 2096 \hspace{8pt} & 26.66\% & \multirow{2}{*}{533}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 643 \hspace{8pt} & 1669 \hspace{8pt} & 27.81\% & \\ \hline
-\multirow{2}{*}{bazuka} & \multicolumn{2}{c|}{Regex} & 451 \hspace{8pt} & 1103 \hspace{8pt} & 29.02\% & \multirow{2}{*}{285}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 402 \hspace{8pt} & 920 \hspace{8pt} & 30.41\% & \\ \hline
-\multirow{2}{*}{strum} & \multicolumn{2}{c|}{Regex} & 51 \hspace{8pt} & 377 \hspace{8pt}	& 11.92\% & \multirow{2}{*}{90}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 50 \hspace{8pt} & 318 \hspace{8pt}	& 13.51\% & \\ \hline
-
-\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 452.33 & 1392.33	& 23.21\% & \\\cline{3-6}
-    &   &  Median & 418.50 & 1348.50 & 23.38\%  & \\\cline{3-6}
-    &   &  Std. Dev & 309.05 & 656.76 & 12.16\%  & \\\cline{2-6}
-
- & \multirow{3}{*}{Comby} & Mean & 554.00 & 1055.83	& 29.29\% & \\\cline{3-6}
-    &   &  Median & 384.00 & 1059.50	& 25.60\%  & \\\cline{3-6}
-    &   &  Std. Dev & 592.63 & 508.30 & 21.63\%  & \\\hline
-
-\end{tabular}
-
-}
-\end{table}
-
-The mean and median percentage of mutants that are valid is higher for
+The mean (and median) percentage of mutants that are valid is higher for
 Comby for all languages.  On average, even for a single modestly sized
@@
 
-We estimated mutant equivalency when possible by examining a sample of
+We estimated mutant equivalence when possible by examining a sample of
 10 randomly selected unkilled mutants for each file.  For some tools,
@@
 \emph{all} mutants, and so translate to much higher rates of
-equivalency for unkilled mutants.  We therefore suggest that as a
+equivalence for unkilled mutants.  We therefore suggest that as a
 fraction of unkilled mutants, a rate of 30\% or fewer equivalent
@@
 criteria for comparing tools is whether there is an overall pattern of
-clearly less-effective mutation for our approach.  Below, we discuss
-the question for each language separately.  Each section also
-addresses briefly the question of whether Regex and Comby modes
-perform differently in terms of mutation score and equivalent mutants,
-which is relevant to RQ1.
+clearly less-effective mutation for our approach.  We only report
+results for Comby, as Comby and Regexp approaches produced nearly
+identical mutation scores, and had similar equivalence rates.  In
+fact, Regexp equivalence rates were somewhat lower over all than those
+shown here.  Full results for Regexp mutation execution are included
+in our replication package.
 
@@
 missing important mutants in some cases.  Overall, mean mutation
-scores were similar for Dextool and both modes of our tool, while
-median mutation scores were much lower for our tools, which produced
-many more unkilled mutants for some projects.  For C++, there seems to
+scores were similar for Dextool and our tool, while
+median mutation scores were much lower for our tool, which produced
+many more unkilled mutants for some projects.  While our equivalence
+rate was somewhat high for one project, it was otherwise very good
+compared to results in the literature.  For C++, there seems to
 be no reason to suspect that the universal approach is producing
 sub-par mutation testing.
-
-Comparing Regex and Comby modes, mutation scores are very similar, as
-are estimated equivalency rates, though Comby produces a somewhat
-larger number and percentage of equivalent mutants.  Both modes
-produce relatively low mean percentages of equivalent mutants.
 
@@
 tool produces lower but comparable mutation scores, and usually
-acceptable or excellent equivalency rates.  LittleDarwin produces
+acceptable or excellent equivalence rates.  LittleDarwin produces
 \emph{no} equivalent mutants, but is a highly conservative tool, only
@@
 deletion.  Our tools always produced more non-equivalent unkilled
-mutants than PIT by a large margin.
-
-Mutation scores for both modes of our tool were again very similar.
-For Java, Comby shows a significant advantage in terms of equivalent
-mutants, with very low equivalency rates compared to Regex.
+mutants than PIT by a large margin.  Our equivalence rates were
+uniformly low, with a median of \emph{no} equivalent mutants.
 
@@
 Our comparison of Python mutation tools showed the largest degree of
-similarity between all tools.  Muation scores and equivalency rates
+similarity between all tools.  Muation scores and equivalence rates
 were all, given the large variance expected from previous studies of
@@
 using our approach, and a large advantage in terms of development and
-maintenance ease.  The similarity extends to the two modes of our
-tool, though Comby had a somewhat higher equivalency rate than the
-other tools, including the Regexp mode of our tool.  We discuss
+maintenance ease.  Comby had a somewhat higher equivalence rate than the
+other tools (including the Regexp mode of our tool).  We discuss
 possible reasons for this below.
@@
 \emph{any} mutants even though our tool was able to produce a large
-number of non-equivalent unkilled mutants.  Again, behavior of the two
-modes of our tool was similar, with both having acceptable equivalency
-rates and similar mutation scores.
+number of non-equivalent unkilled mutants.  Our tool did produce a
+very large percentage of equivalent mutants for one project, and high
+percentage for another project, but these numbers are still well
+within the range of expected results in past inspections of
+equivalent mutants.
 
@@
 Our tools produce somewhat lower mutation scores than other tools, and
-higher (though still generally good) equivalency rates than some
+higher (though still generally good) equivalence rates than some
 tools, but clearly fall into the same broad similarity of behavior
 reported for high-quality mutation tools in the literature comparing
-Java mutation tools.   The tools with very low or zero equivalency rates, furthermore,
+Java mutation tools.   The tools with very low or zero equivalence rates, furthermore,
 are very conservative in mutation generation, and often simply don't
@@
 We speculate that our somewhat lower scores and possibly higher
-equivalenc rates derive from the simple fact that we generate more
+equivalence rates derive from the simple fact that we generate more
 mutants than most tools.  The ease of defining mutation operators,
 plus our interest in generating large numbers of mutants and
-prioritizing the most interesting, to make mutation testing useful for
+prioritizing the most interesting, in order to make mutation testing useful for
 projects with extremely high quality test suites, or even using formal
-verification, has encouraged a ``let a thousand flowers bloom''
+verification, has encouraged us to take a ``let a thousand flowers bloom''
 approach to choosing mutation operators.  For example, our approach
@@
 {\tt break} and {\tt continue} statements into programs, or transform
-one into the other.  These mutants are often equivalent, but also
-sometimes expose subtle weaknesses in a test suite, in our
+one into the other.  These mutants are often equivalent (in fact, they
+give rise to a large number of our equivalent mutants reported above), but \emph{also
+sometimes expose subtle weaknesses in a high quality test suite}, in our
 experience.  In general, we take our philosophy of mutation from the
@@
 set of ``good'' operators, in that the end-goal of mutation testing is
-to identify weaknesses in test suites, not simply to estimate a single score.
+to identify weaknesses in test suites, not simply to estimate a single
+score.
 
@@
 larger number of mutants to execute, and a somewhat larger number of
-equivalent mutants.  However, for our tool, this observation is not a
+equivalent mutants.  This cost, we observe, falls especially on Comby in some
+cases, because Comby is better at identifying places to inject
+control flow changes that produce valid code, but do yield equivalent mutants in
+many cases.
+
+However, for our approach, this is not a
 fundamental limitation or forced choice.  We propose as future work to
-split the operators for each level of the hierarchy into two sets, one
+split the operators for each level of the hierarchy into (at least) two sets, one
 set of core operators similar to (but larger than) those provided by
 LittleDarwin or Cargo-Mutants, and a set of ``aggressive'' operators
-encompassing our full set of operators, a very simple matter due to
+that includes our full current set of operators, a very simple matter due to
 our approach to mutant generation.
+
+In fact, this brings us back to a point made in the introduction.
+While our approach introduces some unavoidable costs in the
+(relatively cheap, compared to execution and analysis) compilation stage of
+mutation testing, it may offer \emph{more} flexibility in addressing
+execution and analysis stages.  Defining limited custom sets of
+operators to run is trivial in our setting, making it easy to
+execute only, e.g., statement mutations, or, in a numeric-computation
+intensive codebase, only arithmetic operators.  Expanding the
+hierarchical structure to include ``famillies'' of mutation operators
+is trivial, and lets users specify the aggressiveness of mutation
+generation in a simple way.  Our approach also produces actual source
+copies of all generated mutants, making it easy to apply techniques
+like Predictive Mutation Testing~\cite{zhangPMT} that work with source
+code.  This also makes it easy to integrate, as we have, heuristic
+aids to mutant \emph{analysis} such as ranking of unkilled mutants by
+similarity and ``interstingness,'' an approach that has already proven
+useful in using mutation testing to improve static analysis tools~\cite{StatMut}.
 

### 2023-03-29 paper/main.tex
Adding rust numbers

--- 2023-03-29 paper/main.tex
@@
 
+\iffalse
 \begin{table}[hbtp]
 \centering
-{\small
-\caption{Regex vs. Comby for Python, including redundant mutants.}
+\caption{Regex vs. Comby for Python++, including redundant mutants.}
 \label{tab:table_pyrust}
@@
 \toprule\toprule
+\mc{6}{c}{\textbf{Python++}} \\\midrule
                  &                 & \multicolumn{3}{c|}{\textbf{Regex}}             &\multicolumn{3}{c}{\textbf{Comby}}  \\[1ex]
-\textbf{Project} & \textbf{LOC}    & red.  & $\frac{valid}{total}$ & \textbf{\%} & red. & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
+\textbf{Project} & \textbf{LOC}    & redundant & $\frac{valid}{total}$ & \textbf{\%} & redundant & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
 keract           & 402             & 729       & $\frac{1655}{3984}$   & 41.53\%     & 533       & $\frac{1471}{3067}$   & 47.96\% \\[1ex]
-dtw              & ... \\
-%
+dtw              & 118             & 567       & $\frac{973}{2106}$    & 46.20\%     & 639       & $\frac{866}{1836}$    & 47.19\% & \\[1ex]
+notion-sdk-py    & 189             & 82        & $\frac{196}{1038}$    & 18.89\%     & 122       & $\frac{196}{1019}$    & 19.23\% & \\[1ex]
+{\makecell{atlassian-python \\ api}} & 236 & $\frac{302}{1304}$  & 207  &	19.99\% & \multirow{2}{*}{236}\\\cline{2-7}
+       & \multicolumn{2}{c|}{Comby} & 333  & 635  & 351  & 25.25\% & \\ \hline
+ESD    & 698 & $\frac{1859}{4626}$  & 1642  & 29.46\% & \multirow{2}{*}{698}\\\cline{2-7}
+       & \multicolumn{2}{c|}{Comby} & 1712  & 1548  &	614  & 44.18\% & \\ \hline
+fiber  & 358 & $\frac{493}{1402}$  & 163  & 31.08\% & \multirow{2}{*}{358}\\\cline{2-7}
+       & \multicolumn{2}{c|}{Comby} & 458  & 767 	& 241  & 31.24\% & \\ \hline
+
+\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 913.00 & 1267.33 & 565.00 & 31.19\%  & \\\cline{3-7}
+    &   &  Std. Dev & 708.93 & 813.35 & 584.79 & 11.06\%  & \\\cline{2-7}
+
+ & \multirow{3}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% & \\\cline{3-7}
+    &   &  Std. Dev & 628.79 & 418.99 & 211.62 & 12.28\%  & \\\hline
 \end{tabularx}
-\begin{tabularx}{\columnwidth}{Xllrrrrrr}
-                  && &  Red.     & Valid  & Invalid  & \% \\\midrule
-\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean &  565.00 & 913.00 & 1267.33 & 31.19\%  & \\
-    &   &  Std. Dev & 584.79  & 708.93 & 813.35 & 11.06\%  \\
- & \multirow{3}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% \\
-  &   &  Std. Dev &  211.62 & 628.79 & 418.99 & 12.28\%   \\
-\end{tabularx}}
 \end{table}
-
+\fi
 % ---------------- Python table UM Only --------------------------
@@
 \textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
-cargo release    & ...    & & & \\[1ex]
+cargo release    & 346          & $\frac{851}{2014}$   & 42.25\%  & $\frac{1690}{2401}$   & 70.39\% \\[1ex]
+passarine        & 376          & $\frac{213}{2294}$   & 9.29\%  & $\frac{173}{1691}$   & 10.23\% \\[1ex]
+typos            & 463          & $\frac{386}{1920}$   & 20.10\%  & $\frac{366}{1565}$   & 23.39\% \\[1ex]
+ord              & 533          & $\frac{762}{2858}$   & 26.66\%  & $\frac{643}{2312}$   & 27.81\% \\[1ex]
+bazuka           & 285          & $\frac{451}{1554}$   & 29.02\%  & $\frac{402}{1322}$   & 30.41\% \\[1ex]
+strum            & 90           & $\frac{51}{428}$     & 11.92\%  & $\frac{50}{368}$     & 13.58\% \\[1ex]
 \end{tabularx}
@@
                & & & Valid  & Invalid  & \% \\\midrule
- \multirow{6}{*}{\textbf{Overall}} & \mr{3}{Regex} & Mean     &   &  & \\
-               &   & Std.Dev. & &  &   \\\cline{2-6}
- & \mr{3}{Comby} & Mean     &  &  &  \\
+ \multirow{6}{*}{\textbf{Overall}} & \mr{3}{Regex} & Mean     &  452.33 &  1392.33 & 23.21\% \\
+               &   & Std.Dev. & 309.05 & 656.76 & 12.16\%  \\\cline{2-6}
+ & \mr{3}{Comby} & Mean     & 554.00 &  1055.83 & 29.29\% \\
  %              &   & Med.     & 253.50 & 724.50 & 24.79\% \\
-               &   & Std.Dev. & &  &   \\\bottomrule
+               &   & Std.Dev. & 592.63 & 508.30 & 21.63\%  \\\bottomrule
 \end{tabularx}}
@@
 % ---------------- C++ table (UM vs Others)--------------------------
-\iffalse
-\begin{table}[htbp]
-\centering
-\caption{C++ (Our Implementation vs. Dextool)}
-\label{tab:table_cpp2}
-\resizebox{\columnwidth}{!}{
-\begin{tabular}{|c|c|r|r|r|r|}
-\hline
- & \multicolumn{3}{c|}{Comby}  & \multicolumn{3}{c|}{Dextool} \\ 
- & $\frac{killed}{total}$ & \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Equiv.} & $\frac{killed}{total}$ & \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Equiv.} \\[1ex]\midrule
- \multirow{3}{*}{\small ompl} & Comby & 41 & 523 & 0.07 & 40\% \\\cline{2-6}
-    & Dextool & 23 & 255 & 0.08 & N/A \\ \hline
-%\multirow{3}{*}{\small libgraphqlparser} & Comby & 116 & 15 & 0.89 & 20\% \\ \cline{2-6}
-%    & Dextool & 50 & 2 & 0.96 & N/A \\ \hline
-%\multirow{3}{*}{\small rethinkdb\_rebirth} & Comby & 49 & 443 & 0.10 & 20\% \\ \cline{2-6}
-%    & Dextool & 9 & 16 & 0.36 & N/A \\ \hline
-
-%\multirow{3}{*}{\small xoreos} & Comby & 33 & 2247 & 0.01 & 10\% \\ \cline{2-6}
-%    & Dextool & 8 & 406 & 0.02 & N/A \\ \hline
-%\multirow{3}{*}{\small libxmljs}  & Comby & 507 & 283 & 0.64 & 10\% \\ \cline{2-6}
-%    & Dextool & 383 & 447 & 0.46 & N/A \\ \hline
-%\multirow{3}{*}{\small tiny-differentiable-simulator} & Comby & 32 & 738 & 0.04 & 0\% \\ \cline{2-6}
-%    & Dextool & 13 & 48 & 0.21 & N/A\\ \hline
-%    
-%\multirow{9}{*}{Overall}\multirow{3}{*}{Comby} & \multicolumn{2}{c|}{Mean} & 0.29 & 16.67\% \\\cline{3-6}
-%    &   &  \multicolumn{2}{c|}{Median} & 0.09 & 15.00\% \\\cline{3-6}
-%    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.37 & 13.66\% \\\cline{2-6}
-%
-% & \multirow{3}{*}{Dextool} & \multicolumn{2}{c|}{Mean} & 0.33 & N/A\\\cline{3-6}
-%    &   &  \multicolumn{2}{c|}{Median} & 0.29 & N/A \\\cline{3-6}
-%    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.38 & N/A \\ \hline
-\end{tabular}
-}
-\end{table}
-\fi
 

### 2023-03-29 paper/main.tex

--- 2023-03-29 paper/main.tex
@@
 
+\begin{table}[hbtp]
+\centering
+{\small
+\caption{Regex vs. Comby for Python, including redundant mutants.}
+\label{tab:table_pyrust}
+\begin{tabularx}{\columnwidth}{X|r|rcr|rcr}
+\toprule\toprule
+                 &                 & \multicolumn{3}{c|}{\textbf{Regex}}             &\multicolumn{3}{c}{\textbf{Comby}}  \\[1ex]
+\textbf{Project} & \textbf{LOC}    & red.  & $\frac{valid}{total}$ & \textbf{\%} & red. & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
+keract           & 402             & 729       & $\frac{1655}{3984}$   & 41.53\%     & 533       & $\frac{1471}{3067}$   & 47.96\% \\[1ex]
+dtw              & ... \\
+%
+\end{tabularx}
+\begin{tabularx}{\columnwidth}{Xllrrrrrr}
+                  && &  Red.     & Valid  & Invalid  & \% \\\midrule
+\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean &  565.00 & 913.00 & 1267.33 & 31.19\%  & \\
+    &   &  Std. Dev & 584.79  & 708.93 & 813.35 & 11.06\%  \\
+ & \multirow{3}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% \\
+  &   &  Std. Dev &  211.62 & 628.79 & 418.99 & 12.28\%   \\
+\end{tabularx}}
+\end{table}
+
 % ---------------- Python table UM Only --------------------------
@@
 % ---------------- C++ table (UM vs Others)--------------------------
+\begin{table}[htbp]
+\centering
+{\small
+\caption{C++ (Our Implementation vs. Dextool)}
+\label{tab:table_cpp2}
+\begin{tabularx}{\columnwidth}{X|crr|crr}
+\hline
+        & \multicolumn{3}{c|}{Comby}  & \multicolumn{3}{c}{Dextool} \\ 
+Project & $\frac{killed}{total}$ & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & $\frac{killed}{total}$ & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} \\[1ex]\midrule
+ ompl & $\frac{41}{564}$ & 0.07 & 40\% & $\frac{23}{278}$ & 0.08 & N/A \\[1ex]
+%\multirow{3}{*}{\small libgraphqlparser} & Comby & 116 & 15 & 0.89 & 20\% \\ \cline{2-6}
+%    & Dextool & 50 & 2 & 0.96 & N/A \\ \hline
+%\multirow{3}{*}{\small rethinkdb\_rebirth} & Comby & 49 & 443 & 0.10 & 20\% \\ \cline{2-6}
+%    & Dextool & 9 & 16 & 0.36 & N/A \\ \hline
+
+%\multirow{3}{*}{\small xoreos} & Comby & 33 & 2247 & 0.01 & 10\% \\ \cline{2-6}
+%    & Dextool & 8 & 406 & 0.02 & N/A \\ \hline
+%\multirow{3}{*}{\small libxmljs}  & Comby & 507 & 283 & 0.64 & 10\% \\ \cline{2-6}
+%    & Dextool & 383 & 447 & 0.46 & N/A \\ \hline
+%\multirow{3}{*}{\small tiny-differentiable-simulator} & Comby & 32 & 738 & 0.04 & 0\% \\ \cline{2-6}
+%    & Dextool & 13 & 48 & 0.21 & N/A\\ \hline
+%    
+%\multirow{9}{*}{Overall}\multirow{3}{*}{Comby} & \multicolumn{2}{c|}{Mean} & 0.29 & 16.67\% \\\cline{3-6}
+%    &   &  \multicolumn{2}{c|}{Median} & 0.09 & 15.00\% \\\cline{3-6}
+%    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.37 & 13.66\% \\\cline{2-6}
+%
+% & \multirow{3}{*}{Dextool} & \multicolumn{2}{c|}{Mean} & 0.33 & N/A\\\cline{3-6}
+%    &   &  \multicolumn{2}{c|}{Median} & 0.29 & N/A \\\cline{3-6}
+%    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.38 & N/A \\ \hline
+\end{tabularx}
+}
+\end{table}
 
@@
 hierarchical structure to include ``famillies'' of mutation operators
-is trivial, and lets users specify the aggressiveness of mutation
-generation in a simple way.  Our approach also produces actual source
+is trivial, and lets users specify fine-grained mutation
+generation for all languages used in a project, in a simple, consistent way.  Our approach also produces actual source
 copies of all generated mutants, making it easy to apply techniques
 like Predictive Mutation Testing~\cite{zhangPMT} that work with source
-code.  This also makes it easy to integrate, as we have, heuristic
+code.  This also makes it easy to integrate, as we already have done, heuristic
 aids to mutant \emph{analysis} such as ranking of unkilled mutants by
@@
 this case actually gives it \emph{more} opportunities to apply such
-inapplicable mutants.
+``inappropriate'' mutants.
 

### 2023-03-29 paper/main.tex
new C++ table vs dextool

--- 2023-03-29 paper/main.tex
@@
                   & &       & Valid  & Invalid  & \% \\\midrule
- \multirow{6}{*}{\textbf{Overall}} & \multirow{3}{*}{Regex} & Mean     & 890.67 &  1553.83 & 35.28\% \\
+ \multirow{4}{*}{\textbf{Overall}} & \multirow{2}{*}{Regex} & Mean     & 890.67 &  1553.83 & 35.28\% \\
 %                          & & Med.    & 671.00 & 1579.00  & 23.88\%   \\
                           & & Std.Dev. & 787.62 & 956.31   & 19.48\%  \\\cline{2-6}
-  & \multirow{3}{*}{Comby}  & Mean     & 837.83 & 1352.50 & 38.75\%  \\
+  & \multirow{2}{*}{Comby}  & Mean     & 837.83 & 1352.50 & 38.75\%  \\
 %                         & & Med.    & 667.00 & 1267.00 & 27.86\%  \\
@@
               & & & Valid  & Invalid  & \% \\\midrule
-\multirow{6}{*}{\textbf{Overall}} & \mr{3}{Regex} & Mean     &  345.83 & 1545.00  & 17.74\% \\
+\multirow{4}{*}{\textbf{Overall}} & \mr{2}{Regex} & Mean     &  345.83 & 1545.00  & 17.74\% \\
 %              &   & Med.     & 299.50 & 1524.50 & 16.90\%  \\
               &   & Std.Dev. & 203.86 & 467.98 & 8.73\%  \\\cline{2-6}
-& \mr{3}{Comby} & Mean     & 309.17 & 786.17 & 27.22\% \\
+& \mr{2}{Comby} & Mean     & 309.17 & 786.17 & 27.22\% \\
 %              &   & Med.     & 253.50 & 724.50 & 24.79\% \\
@@
                   && &  Red.     & Valid  & Invalid  & \% \\\midrule
-\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean &  565.00 & 913.00 & 1267.33 & 31.19\%  & \\
+\multirow{4}{*}{Overall} & \multirow{2}{*}{Regex} & Mean &  565.00 & 913.00 & 1267.33 & 31.19\%  & \\
     &   &  Std. Dev & 584.79  & 708.93 & 813.35 & 11.06\%  \\
- & \multirow{3}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% \\
+ & \multirow{2}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% \\
   &   &  Std. Dev &  211.62 & 628.79 & 418.99 & 12.28\%   \\
@@
 \textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
-cargo release    & 346          & $\frac{851}{2014}$   & 42.25\%  & $\frac{1690}{2401}$   & 70.39\% \\[1ex]
-passarine        & 376          & $\frac{213}{2294}$   & 9.29\%  & $\frac{173}{1691}$   & 10.23\% \\[1ex]
-typos            & 463          & $\frac{386}{1920}$   & 20.10\%  & $\frac{366}{1565}$   & 23.39\% \\[1ex]
-ord              & 533          & $\frac{762}{2858}$   & 26.66\%  & $\frac{643}{2312}$   & 27.81\% \\[1ex]
-bazuka           & 285          & $\frac{451}{1554}$   & 29.02\%  & $\frac{402}{1322}$   & 30.41\% \\[1ex]
-strum            & 90           & $\frac{51}{428}$     & 11.92\%  & $\frac{50}{368}$     & 13.58\% \\[1ex]
+cargo release    & ...    & & & \\[1ex]
 \end{tabularx}
@@
                & & & Valid  & Invalid  & \% \\\midrule
- \multirow{6}{*}{\textbf{Overall}} & \mr{3}{Regex} & Mean     &  452.33 &  1392.33 & 23.21\% \\
-               &   & Std.Dev. & 309.05 & 656.76 & 12.16\%  \\\cline{2-6}
- & \mr{3}{Comby} & Mean     & 554.00 &  1055.83 & 29.29\% \\
+ \multirow{6}{*}{\textbf{Overall}} & \mr{3}{Regex} & Mean     &   &  & \\
+               &   & Std.Dev. & &  &   \\\cline{2-6}
+ & \mr{3}{Comby} & Mean     &  &  &  \\
  %              &   & Med.     & 253.50 & 724.50 & 24.79\% \\
-               &   & Std.Dev. & 592.63 & 508.30 & 21.63\%  \\\bottomrule
+               &   & Std.Dev. & &  &   \\\bottomrule
 \end{tabularx}}
@@
 %    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.38 & N/A \\ \hline
-\end{tabularx}
-}
+\textbf{Mean}   & -- & 0.29 & 16.67\% & -- & 0.33 & N/A\\
+\textbf{Std. Dev} & -- & 0.37 & 13.66\%   & -- & 0.38 & N/A \\\bottomrule
+\end{tabularx}}
 \end{table}

### 2023-03-29 paper/main.tex
new table for C++ us vs them. Much simpler

--- 2023-03-29 paper/main.tex
@@
         & \multicolumn{3}{c|}{Comby}  & \multicolumn{3}{c}{Dextool} \\ 
-Project & $\frac{killed}{total}$ & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & $\frac{killed}{total}$ & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} \\[1ex]\midrule
- ompl & $\frac{41}{564}$ & 0.07 & 40\% & $\frac{23}{278}$ & 0.08 & N/A \\[1ex]
+Project & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & $\frac{killed}{total}$ & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} \\[1ex]\midrule
+ ompl &  564 & 0.07 & 40\% & 278 & 0.08 & N/A \\
 %\multirow{3}{*}{\small libgraphqlparser} & Comby & 116 & 15 & 0.89 & 20\% \\ \cline{2-6}
@@
 %    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.38 & N/A \\ \hline
-\end{tabularx}
-}
+\textbf{Mean}   & FIXME & 0.29 & 16.67\% & FIXME & 0.33 & N/A\\
+\textbf{Std. Dev} & FIXME & 0.37 & 13.66\%   & FIXME & 0.38 & N/A \\\bottomrule
+\end{tabularx}}
 \end{table}

### 2023-03-29 paper/main.tex
Python numbers

--- 2023-03-29 paper/main.tex
@@
                   & &       & Valid  & Invalid  & \% \\\midrule
- \multirow{4}{*}{\textbf{Overall}} & \multirow{2}{*}{Regex} & Mean     & 890.67 &  1553.83 & 35.28\% \\
+ \multirow{6}{*}{\textbf{Overall}} & \multirow{3}{*}{Regex} & Mean     & 890.67 &  1553.83 & 35.28\% \\
 %                          & & Med.    & 671.00 & 1579.00  & 23.88\%   \\
                           & & Std.Dev. & 787.62 & 956.31   & 19.48\%  \\\cline{2-6}
-  & \multirow{2}{*}{Comby}  & Mean     & 837.83 & 1352.50 & 38.75\%  \\
+  & \multirow{3}{*}{Comby}  & Mean     & 837.83 & 1352.50 & 38.75\%  \\
 %                         & & Med.    & 667.00 & 1267.00 & 27.86\%  \\
@@
               & & & Valid  & Invalid  & \% \\\midrule
-\multirow{4}{*}{\textbf{Overall}} & \mr{2}{Regex} & Mean     &  345.83 & 1545.00  & 17.74\% \\
+\multirow{6}{*}{\textbf{Overall}} & \mr{3}{Regex} & Mean     &  345.83 & 1545.00  & 17.74\% \\
 %              &   & Med.     & 299.50 & 1524.50 & 16.90\%  \\
               &   & Std.Dev. & 203.86 & 467.98 & 8.73\%  \\\cline{2-6}
-& \mr{2}{Comby} & Mean     & 309.17 & 786.17 & 27.22\% \\
+& \mr{3}{Comby} & Mean     & 309.17 & 786.17 & 27.22\% \\
 %              &   & Med.     & 253.50 & 724.50 & 24.79\% \\
@@
 keract           & 402             & 729       & $\frac{1655}{3984}$   & 41.53\%     & 533       & $\frac{1471}{3067}$   & 47.96\% \\[1ex]
-dtw              & ... \\
-%
+dtw              & 118             & 567       & $\frac{973}{2106}$   & 46.20\%     & 639       & $\frac{866}{1835}$   & 47.19\% \\[1ex]
+notion-sdk-py    & 189             & 82        & $\frac{196}{1038}$   & 18.89\%     & 122       & $\frac{196}{1019}$   & 19.23\% \\[1ex]
+atlassian-python-api    & 236             & 207        & $\frac{302}{1541}$   & 19.99\%     & 351       & $\frac{333}{1319}$   & 25.25\% \\[1ex]
+ESD              & 698            & 1642      & $\frac{1859}{6268}$   & 29.46\%     & 614       & $\frac{1712}{3874}$   & 44.18\% \\[1ex]
+fiber            & 358            & 163      & $\frac{493}{1565}$   & 31.08\%     & 241       & $\frac{458}{1466}$   & 31.24\% \\[1ex]
 \end{tabularx}
@@
                   && &  Red.     & Valid  & Invalid  & \% \\\midrule
-\multirow{4}{*}{Overall} & \multirow{2}{*}{Regex} & Mean &  565.00 & 913.00 & 1267.33 & 31.19\%  & \\
+\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean &  565.00 & 913.00 & 1267.33 & 31.19\%  & \\
     &   &  Std. Dev & 584.79  & 708.93 & 813.35 & 11.06\%  \\
- & \multirow{2}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% \\
+ & \multirow{3}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% \\
   &   &  Std. Dev &  211.62 & 628.79 & 418.99 & 12.28\%   \\
@@
 %    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.38 & N/A \\ \hline
-\textbf{Mean}   & -- & 0.29 & 16.67\% & -- & 0.33 & N/A\\
-\textbf{Std. Dev} & -- & 0.37 & 13.66\%   & -- & 0.38 & N/A \\\bottomrule
-\end{tabularx}}
+\end{tabularx}
+}
 \end{table}

### 2023-03-29 paper/main.tex
starter table for java vs others

--- 2023-03-29 paper/main.tex
@@
 
-\begin{table}[hbtp]
-\centering
-\caption{Rust (Regex vs. Comby)}
-\label{tab:table_rust1}
-\resizebox{\columnwidth}{!}{%
-\begin{tabular}{|c|c|c|r|r|r|c|}
-\hline
- & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%} &  \textbf{LOC} \\ \hline
-\multirow{2}{*}{cargo release} & \multicolumn{2}{c|}{Regex} & 851 \hspace{8pt} &	1163 \hspace{8pt} & 42.25\%  & \multirow{2}{*}{346}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 1690 \hspace{8pt} & 711 \hspace{8pt} & 70.39\% & \\ \hline
-\multirow{2}{*}{passerine} & \multicolumn{2}{c|}{Regex} & 213 \hspace{8pt} & 2081 \hspace{8pt} & 9.29\% & \multirow{2}{*}{376}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 173 \hspace{8pt} & 1518 \hspace{8pt} & 10.23\% & \\ \hline
-\multirow{2}{*}{typos} & \multicolumn{2}{c|}{Regex} & 386 \hspace{8pt} & 1534 \hspace{8pt} & 20.10\% & \multirow{2}{*}{463}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 366 \hspace{8pt} & 1199 \hspace{8pt} &	 23.39\% & \\ \hline
-
-\multirow{2}{*}{ord} & \multicolumn{2}{c|}{Regex} & 762 \hspace{8pt}	& 2096 \hspace{8pt} & 26.66\% & \multirow{2}{*}{533}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 643 \hspace{8pt} & 1669 \hspace{8pt} & 27.81\% & \\ \hline
-\multirow{2}{*}{bazuka} & \multicolumn{2}{c|}{Regex} & 451 \hspace{8pt} & 1103 \hspace{8pt} & 29.02\% & \multirow{2}{*}{285}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 402 \hspace{8pt} & 920 \hspace{8pt} & 30.41\% & \\ \hline
-\multirow{2}{*}{strum} & \multicolumn{2}{c|}{Regex} & 51 \hspace{8pt} & 377 \hspace{8pt}	& 11.92\% & \multirow{2}{*}{90}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 50 \hspace{8pt} & 318 \hspace{8pt}	& 13.51\% & \\ \hline
-
-\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 452.33 & 1392.33	& 23.21\% & \\\cline{3-6}
-    &   &  Median & 418.50 & 1348.50 & 23.38\%  & \\\cline{3-6}
-    &   &  Std. Dev & 309.05 & 656.76 & 12.16\%  & \\\cline{2-6}
-
- & \multirow{3}{*}{Comby} & Mean & 554.00 & 1055.83	& 29.29\% & \\\cline{3-6}
-    &   &  Median & 384.00 & 1059.50	& 25.60\%  & \\\cline{3-6}
-    &   &  Std. Dev & 592.63 & 508.30 & 21.63\%  & \\\hline
-\end{tabular}
-
-}
-\end{table}
-
 % ---------------- Rust table UM Only --------------------------
@@
         & \multicolumn{3}{c|}{Comby}  & \multicolumn{3}{c}{Dextool} \\ 
-Project & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & $\frac{killed}{total}$ & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} \\[1ex]\midrule
+Project & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} \\[1ex]\midrule
  ompl &  564 & 0.07 & 40\% & 278 & 0.08 & N/A \\
@@
 
+% ---------------- JAVA table (UM vs Others) --------------------------
+
+\begin{table}[htbp]
+\centering
+{\small
+\caption{Java (Our Implementation vs. PIT vs. LittleDarwin)}
+\label{tab:table_java2}
+\begin{tabularx}{\columnwidth}{l|rrr|rrr|rrr}
+\hline
+                 & \mc{3}{c|}{Comby}   & \mc{3}{c|}{PIT} & \mc{3}{c}{LittleDarwin} \\
+\textbf{Project} & \textbf{Mutants} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Eqv}
+& \textbf{Mutants} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Eqv} 
+& \textbf{Mutants} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Eqv} \\ \midrule
+
+pebble & 85 & 0.81 & 50\% & 20 & 0.95 & N/A & 8 & 1.00 0\% \\
+... \\\midrule
+\textbf{Mean} &  XX & XX & XX & XX & XX & XX & XX & XX & XX \\
+\textbf{Std.Dev.} &  XX & XX & XX & XX & XX & XX & XX & XX & XX \\
+\end{tabularx}
+}
+\end{table}
+
 % ----------------  Python table (UM vs Others) -------------------------------

### 2023-03-29 paper/main.tex
Vs Dextool

--- 2023-03-29 paper/main.tex
@@
         & \multicolumn{3}{c|}{Comby}  & \multicolumn{3}{c}{Dextool} \\ 
-Project & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & $\frac{killed}{total}$ & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} \\[1ex]\midrule
+Project & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} \\[1ex]\midrule
  ompl &  564 & 0.07 & 40\% & 278 & 0.08 & N/A \\
+ libgraphqlparser &  131 & 0.90 & 20\% & 52 & 0.96 & N/A \\
+ rethinkdb\_rebirth &  492 & 0.10 & 20\% & 25 & 0.36 & N/A \\
+ xoreos &  2280 & 0.01 & 10\% & 414 & 0.02 & N/A \\
+ libxmljs &  790 & 0.62 & 10\% & 830 & 0.46 & N/A \\
+ tiny-differentiable-simulator &  770 & 0.04 & 0\% & 61 & 0.21 & N/A \\
+
 %\multirow{3}{*}{\small libgraphqlparser} & Comby & 116 & 15 & 0.89 & 20\% \\ \cline{2-6}
@@
 %    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.38 & N/A \\ \hline
-\textbf{Mean}   & FIXME & 0.29 & 16.67\% & FIXME & 0.33 & N/A\\
-\textbf{Std. Dev} & FIXME & 0.37 & 13.66\%   & FIXME & 0.38 & N/A \\\bottomrule
+\textbf{Mean}   & 837.83 & 0.29 & 16.67\% & 276.67 & 0.33 & N/A\\
+\textbf{Std. Dev} & 745.78 & 0.37 & 13.66\%   & 311.56 & 0.38 & N/A \\\bottomrule
 \end{tabularx}}

### 2023-03-29 paper/main.tex
starter table python vs others

--- 2023-03-29 paper/main.tex
@@
 \centering
+{\small
+\caption{Python (Our Implementation vs. Mutmut vs. CosmicRay)}
+\label{tab:table_python2}
+\begin{tabularx}{\columnwidth}{l|rrr|rrr|rrr}
+\toprule
+                 & \mc{3}{c|}{Comby}   & \mc{3}{c|}{MutMut} & \mc{3}{c}{CosmicRay} \\
+\textbf{Project} & \textbf{Mutants} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Eqv}
+& \textbf{Mutants} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Eqv} 
+& \textbf{Mutants} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Eqv} \\ \midrule
+keract  & & & & & & & \\
+\textbf{Mean} &  XX & XX & XX & XX & XX & XX & XX & XX & XX \\
+\textbf{Std.Dev.} &  XX & XX & XX & XX & XX & XX & XX & XX & XX \\
+\end{tabularx}
+}
+\end{table}
+
+% ----------------  Python table (UM vs Others) -------------------------------
+
+\begin{table}[htbp]
+\centering
 \caption{Python (Our Implementation vs. Mutmut vs. CosmicRay)}

### 2023-03-29 paper/main.tex
blank rust table

--- 2023-03-29 paper/main.tex
@@
 }
+\end{table}
+
+% ---------------- Rust table (UM vs Others)--------------------------
+
+\begin{table}[htbp]
+\centering
+{\small
+\caption{Rust (Our Implementation vs. Cargo-Mutants)}
+\label{tab:table_rust2}
+\begin{tabularx}{\columnwidth}{X|crr|crr}
+\hline
+        & \multicolumn{3}{c|}{Comby}  & \multicolumn{3}{c}{cargo-mutants} \\ 
+Project & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} \\[1ex]\midrule
+
+cargo release & & & & & \\    
+\textbf{Mean}   & FIXME & 0.29 & 16.67\% & FIXME & 0.33 & N/A\\
+\textbf{Std. Dev} & FIXME & 0.37 & 13.66\%   & FIXME & 0.38 & N/A \\\bottomrule
+\end{tabularx}}
 \end{table}

### 2023-03-29 paper/main.tex
cutting some of the now-extraneous tables

@@
 
-% % ---------------- C++ table UM Only--------------------------
-% \begin{table}[hbtp]
-% \centering
-% \caption{C++ (Regex vs. Comby)}
-% \label{tab:table_cpp1}
-% \resizebox{\columnwidth}{!}{%
-
-% \begin{tabular}{|c|c|c|r|r|r|c|}
-% \toprule
-%  & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%} & \textbf{LOC} \\ \hline
-% \multirow{2}{*}{ompl} & \multicolumn{2}{c|}{Regex} & 595 \hspace{8pt} & 1938 \hspace{8pt} & 23.49\% & \multirow{2}{*}{525} \\\cline{2-6}
-%     & \multicolumn{2}{c|}{Comby} & 564 \hspace{8pt} & 1610 \hspace{8pt} & 25.94\% &  \\ \hline
-% \multirow{2}{*}{libgraphqlparser} & \multicolumn{2}{c|}{Regex} & 157 \hspace{8pt} & 584 \hspace{8pt} & 21.19\% & \multirow{2}{*}{137}\\\cline{2-6}
-%     & \multicolumn{2}{c|}{Comby} & 131 \hspace{8pt} & 309 \hspace{8pt} & 29.77\% & \\ \hline
-% \multirow{2}{*}{rethinkdb\_rebirth} & \multicolumn{2}{c|}{Regex} & 545 \hspace{8pt} & 1701 \hspace{8pt} & 24.27\% & \multirow{2}{*}{471}\\\cline{2-6}
-%     & \multicolumn{2}{c|}{Comby} & 492 \hspace{8pt} & 1517 \hspace{8pt} & 24.49\% & \\ \hline
-
-% \multirow{2}{*}{xoreos} & \multicolumn{2}{c|}{Regex} & 2419 \hspace{8pt} & 1457 \hspace{8pt} & 62.41\% & \multirow{2}{*}{732}\\\cline{2-6}
-%     & \multicolumn{2}{c|}{Comby} & 2280 \hspace{8pt} & 1017 \hspace{8pt} & 69.15\% & \\ \hline
-% \multirow{2}{*}{libxmljs} & \multicolumn{2}{c|}{Regex} & 881 \hspace{8pt} & 3108 \hspace{8pt} & 22.09\% & \multirow{2}{*}{630}\\\cline{2-6}
-%     & \multicolumn{2}{c|}{Comby} & 790 \hspace{8pt} & 3218 \hspace{8pt} & 19.71\% & \\ \hline
-% \multirow{2}{*}{tiny-differentiable-simulator} & \multicolumn{2}{c|}{Regex} & 747 \hspace{8pt} & 535 \hspace{8pt} & 58.27\% & \multirow{2}{*}{195}\\\cline{2-6}
-%     & \multicolumn{2}{c|}{Comby} & 770 \hspace{8pt} & 444 \hspace{8pt} & 63.43\% & \\ \hline
-
-% \multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 890.67 &	1553.83 & 35.28\% & \\\cline{3-6}
-%     &   &  Median & 671.00 & 1579.00 & 23.88\%  & \\\cline{3-6}
-%     &   &  Std. Dev & 787.62 & 956.31 & 19.48\% & \\\cline{2-6}
-
-%  & \multirow{3}{*}{Comby} & Mean & 837.83 & 1352.50 & 38.75\% & \\\cline{3-6}
-%     &   &  Median & 667.00 & 1267.00 & 27.86\% & \\\cline{3-6}
-%     &   &  Std. Dev & 745.78 & 1058.33	& 21.66\% & \\\hline
-
-% \end{tabular}
-% }
-
-% \end{table}
-
-% % ---------------- JAVA table UM Only--------------------------
-% \begin{table}[hbtp]
-% \centering
-% \caption{Java (Regex vs. Comby)}
-% \label{tab:table_java1}
-% \resizebox{\columnwidth}{!}{%
-
-% \begin{tabular}{|c|c|c|r|r|r|c|}
-% \hline
-%  & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%} & \textbf{LOC}\\ \hline
-% \multirow{2}{*}{pebble} & \multicolumn{2}{c|}{Regex} & 86 \hspace{8pt} & 928 \hspace{8pt} & 8.48\%  & \multirow{2}{*}{279}\\\cline{2-6}
-%     & \multicolumn{2}{c|}{Comby} & 85 \hspace{8pt} & 667 \hspace{8pt} & 11.30\% & \\ \hline
-% \multirow{2}{*}{spring-hateoas} & \multicolumn{2}{c|}{Regex} & 248 \hspace{8pt} & 2231 \hspace{8pt} & 10.00\% & \multirow{2}{*}{397}\\\cline{2-6}
-%     & \multicolumn{2}{c|}{Comby} & 172 \hspace{8pt} & 1212 \hspace{8pt} & 12.43\% & \\ \hline
-% \multirow{2}{*}{javacc} & \multicolumn{2}{c|}{Regex} & 542 \hspace{8pt} & 1346 \hspace{8pt} & 28.70\% & \multirow{2}{*}{363}\\\cline{2-6}
-%     & \multicolumn{2}{c|}{Comby} & 494 \hspace{8pt} & 782 \hspace{8pt} & 38.71\% & \\ \hline
-
-% \multirow{2}{*}{coffee-gb} & \multicolumn{2}{c|}{Regex} & 224 \hspace{8pt} & 1703 \hspace{8pt} & 11.63\% & \multirow{2}{*}{463}\\\cline{2-6}
-%     & \multicolumn{2}{c|}{Comby} & 174 \hspace{8pt} & 624 \hspace{8pt} & 21.80\% & \\ \hline
-% \multirow{2}{*}{egads} & \multicolumn{2}{c|}{Regex} & 624 \hspace{8pt} & 1830 \hspace{8pt} & 25.43\% & \multirow{2}{*}{252}\\\cline{2-6}
-%     & \multicolumn{2}{c|}{Comby} & 597 \hspace{8pt} & 566 \hspace{8pt} & 51.33\% & \\ \hline
-% \multirow{2}{*}{apk-parser} & \multicolumn{2}{c|}{Regex} & 351 \hspace{8pt} & 1232 \hspace{8pt} & 22.17\% & \multirow{2}{*}{324}\\\cline{2-6}
-%     & \multicolumn{2}{c|}{Comby} & 333 \hspace{8pt} & 866 \hspace{8pt} & 27.77\% & \\ \hline
-
-% \multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 345.83 & 1545.00  & 17.74\%& \\\cline{3-6}
-%     &   &  Median & 299.50 & 1524.50 & 16.90\%  & \\\cline{3-6}
-%     &   &  Std. Dev & 203.86 & 467.98 & 8.73\%  & \\\cline{2-6}
-
-%  & \multirow{3}{*}{Comby} & Mean & 309.17 & 786.17 & 27.22\% & \\\cline{3-6}
-%     &   &  Median & 253.50 & 724.50 & 24.79\%  & \\\cline{3-6}
-%     &   &  Std. Dev & 202.44 & 235.24  & 15.59\%  & \\\hline
-
-% \end{tabular}
-
-% }
-% \end{table}
-
 \begin{table}[hbtp]
@@
 \end{tabularx}}
-\end{table}
-
-% ---------------- Python table UM Only --------------------------
-
-\begin{table}[hbtp]
-\centering
-\caption{Python (Regex vs. Comby)}
-\label{tab:table_python1}
-\resizebox{\columnwidth}{!}{%
-
-\begin{tabular}{|c|c|c|r|r|r|r|c|}
-\hline
- & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Redundant} & \textbf{Valid \%} & \textbf{LOC} \\ \hline
-\multirow{2}{*}{keract} & \multicolumn{2}{c|}{Regex} & 1655 \hspace{8pt} & 1600 \hspace{8pt} &	729 \hspace{8pt} & 41.53\%  & \multirow{2}{*}{402}\\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 1471 \hspace{8pt} & 1063 \hspace{8pt} & 533 \hspace{8pt} & 47.96\% & \\ \hline
-\multirow{2}{*}{dtw} & \multicolumn{2}{c|}{Regex} & 973 \hspace{8pt} & 566 \hspace{8pt} & 567 \hspace{8pt} & 46.20\% & \multirow{2}{*}{118}\\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 866 \hspace{8pt} & 330 \hspace{8pt} & 639 \hspace{8pt} &	47.19\% & \\ \hline
-\multirow{2}{*}{notion-sdk-py} & \multicolumn{2}{c|}{Regex} & 196 \hspace{8pt} & 760 \hspace{8pt} & 82 \hspace{8pt} & 18.89\% & \multirow{2}{*}{189}\\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 196 \hspace{8pt} & 701 \hspace{8pt} & 122 \hspace{8pt} & 19.23\% & \\ \hline
-
-\multirow{2}{*}{{\makecell{atlassian-python \\ api}}} & \multicolumn{2}{c|}{Regex} & 302 \hspace{8pt}	& 1002 \hspace{8pt} & 207 \hspace{8pt} &	19.99\% & \multirow{2}{*}{236}\\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 333 \hspace{8pt} & 635 \hspace{8pt} & 351 \hspace{8pt} & 25.25\% & \\ \hline
-\multirow{2}{*}{ESD} & \multicolumn{2}{c|}{Regex} & 1859 \hspace{8pt}	& 2767 \hspace{8pt} & 1642 \hspace{8pt} & 29.46\% & \multirow{2}{*}{698}\\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 1712 \hspace{8pt} & 1548 \hspace{8pt} &	614 \hspace{8pt} & 44.18\% & \\ \hline
-\multirow{2}{*}{fiber} & \multicolumn{2}{c|}{Regex} & 493 \hspace{8pt} & 909 \hspace{8pt} & 163 \hspace{8pt} & 31.08\% & \multirow{2}{*}{358}\\\cline{2-7}
-    & \multicolumn{2}{c|}{Comby} & 458 \hspace{8pt} & 767 \hspace{8pt}	& 241 \hspace{8pt} & 31.24\% & \\ \hline
-
-\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 913.00 & 1267.33 & 565.00 & 31.19\%  & \\\cline{3-7}
-    &   &  Median & 733.00 & 955.50 & 387.00 & 30.27\%  & \\\cline{3-7}
-    &   &  Std. Dev & 708.93 & 813.35 & 584.79 & 11.06\%  & \\\cline{2-7}
-
- & \multirow{3}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% & \\\cline{3-7}
-    &   &  Median & 662.00 & 734.00 & 442.00 & 37.71\%  & \\\cline{3-7}
-    &   &  Std. Dev & 628.79 & 418.99 & 211.62 & 12.28\%  & \\\hline
-
-\end{tabular}
-
-}
 \end{table}
@@
 
-% ---------------- Rust table UM Only --------------------------
-\begin{table}[hbtp]
-\centering
-\caption{Rust (Regex vs. Comby)}
-\label{tab:table_rust1}
-\resizebox{\columnwidth}{!}{%
-
-\begin{tabular}{|c|c|c|r|r|r|c|}
-\hline
- & \multicolumn{2}{c|}{} & \textbf{Valid} & \textbf{Invalid} & \textbf{Valid \%} &  \textbf{LOC} \\ \hline
-\multirow{2}{*}{cargo release} & \multicolumn{2}{c|}{Regex} & 851 \hspace{8pt} &	1163 \hspace{8pt} & 42.25\%  & \multirow{2}{*}{346}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 1690 \hspace{8pt} & 711 \hspace{8pt} & 70.39\% & \\ \hline
-\multirow{2}{*}{passerine} & \multicolumn{2}{c|}{Regex} & 213 \hspace{8pt} & 2081 \hspace{8pt} & 9.29\% & \multirow{2}{*}{376}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 173 \hspace{8pt} & 1518 \hspace{8pt} & 10.23\% & \\ \hline
-\multirow{2}{*}{typos} & \multicolumn{2}{c|}{Regex} & 386 \hspace{8pt} & 1534 \hspace{8pt} & 20.10\% & \multirow{2}{*}{463}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 366 \hspace{8pt} & 1199 \hspace{8pt} &	 23.39\% & \\ \hline
-
-\multirow{2}{*}{ord} & \multicolumn{2}{c|}{Regex} & 762 \hspace{8pt}	& 2096 \hspace{8pt} & 26.66\% & \multirow{2}{*}{533}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 643 \hspace{8pt} & 1669 \hspace{8pt} & 27.81\% & \\ \hline
-\multirow{2}{*}{bazuka} & \multicolumn{2}{c|}{Regex} & 451 \hspace{8pt} & 1103 \hspace{8pt} & 29.02\% & \multirow{2}{*}{285}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 402 \hspace{8pt} & 920 \hspace{8pt} & 30.41\% & \\ \hline
-\multirow{2}{*}{strum} & \multicolumn{2}{c|}{Regex} & 51 \hspace{8pt} & 377 \hspace{8pt}	& 11.92\% & \multirow{2}{*}{90}\\\cline{2-6}
-    & \multicolumn{2}{c|}{Comby} & 50 \hspace{8pt} & 318 \hspace{8pt}	& 13.51\% & \\ \hline
-
-\multirow{6}{*}{Overall} & \multirow{3}{*}{Regex} & Mean & 452.33 & 1392.33	& 23.21\% & \\\cline{3-6}
-    &   &  Median & 418.50 & 1348.50 & 23.38\%  & \\\cline{3-6}
-    &   &  Std. Dev & 309.05 & 656.76 & 12.16\%  & \\\cline{2-6}
-
- & \multirow{3}{*}{Comby} & Mean & 554.00 & 1055.83	& 29.29\% & \\\cline{3-6}
-    &   &  Median & 384.00 & 1059.50	& 25.60\%  & \\\cline{3-6}
-    &   &  Std. Dev & 592.63 & 508.30 & 21.63\%  & \\\hline
-
-\end{tabular}
-
-}
-\end{table}
-
 The mean (and median) percentage of mutants that are valid is higher for

### 2023-03-29 paper/main.tex
halfway to adding Rust to the C++/Java table

@@
 \centering
-\caption{Regex vs. Comby for C++ (top) and Java (bottom)}
+\caption{Regex vs. Comby for C++ (top), Java (middle), and Rust (bottom)}
 \label{tab:table_cpp1java}
-\begin{tabularx}{\columnwidth}{X|r|cr|cr}
+\begin{tabularx}{\columnwidth}{X|r|rr|rr}
 \toprule\toprule
 \mc{6}{c}{\textbf{C++}} \\\midrule
-                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\[1ex]
-\textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
-ompl & 525 & $\frac{595}{2533}$ & 23.49\% &  $\frac{564}{2127}$  & 25.94\%   \\[1ex]
-libgraphqlparser& 137 & $\frac{157}{741}$ & 21.19\%  & $\frac{131}{440}$ & 29.77\% \\[1ex]
-rethinkdb\_rebirth &471 &  $\frac{545}{2246}$ & 24.27\% & $\frac{492}{2009}$ & 24.49\% \\[1ex]
-xoreos & 732  & $\frac{2419}{3876}$  & 62.41\% & $\frac{2280}{3297}$ & 69.15\% \\[1ex]
-libxmljs &630 &  $\frac{881}{3989}$ & 22.09\% & $\frac{790}{4008}$ & 19.71\% \\[1ex]
-tiny-diff.-simulator &  195 & $\frac{747}{1282}$ &58.27\% &$\frac{770}{1214}$ & 63.43\% \\[1ex]
+                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
+\textbf{Project} & \textbf{LOC} & \textbf{Muts} & \textbf{\% Valid} & \textbf{Muts} &  \textbf{\% Valid} \\\midrule
+ompl & 525 & 2533 & 23.49\% &  2127  & 25.94\%   \\
+libgraphqlparser& 137 & 741 & 21.19\%  & 440 & 29.77\% \\
+rethinkdb\_rebirth &471 & 2246 & 24.27\% & 2009 & 24.49\% \\
+xoreos & 732  & 3876  & 62.41\% & 3297 & 69.15\% \\
+libxmljs &630 &  3989 & 22.09\% & 4008 & 19.71\% \\[0.5ex]
+tiny-diff.-simulator &  195 & 1282 &58.27\% & 1214 & 63.43\% \\
 \bottomrule
@@
 \begin{tabularx}{\columnwidth}{Xllrrr}
-                  & &       & Valid  & Invalid  & \% \\\midrule
- \multirow{4}{*}{\textbf{Overall}} & \multirow{2}{*}{Regex} & Mean     & 890.67 &  1553.83 & 35.28\% \\
+                  & &       & \textbf{Valid}  & \textbf{Invalid}  & \textbf{\% Valid} \\\midrule
+ \multirow{4}{*}{\textbf{Overall}} & \multirow{2}{*}{\bf Regex} & Mean     & 890.67 &  1553.83 & 35.28\% \\
 %                          & & Med.    & 671.00 & 1579.00  & 23.88\%   \\
                           & & Std.Dev. & 787.62 & 956.31   & 19.48\%  \\\cline{2-6}
-  & \multirow{2}{*}{Comby}  & Mean     & 837.83 & 1352.50 & 38.75\%  \\
+  & \multirow{2}{*}{\bf Comby}  & Mean     & 837.83 & 1352.50 & 38.75\%  \\
 %                         & & Med.    & 667.00 & 1267.00 & 27.86\%  \\
@@
                  &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
-\textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
- pebble         & 279 & $\frac{86}{1014}$ & 8.48\% & $\frac{85}{752}$ & 11.30\%  \\[1ex]
- spring-hateoas & 397 & $\frac{248}{2479}$ & 10.00\% & $\frac{172}{1384}$ & 12.43\%  \\[1ex]
- javacc         & 363 & $\frac{542}{1888}$ & 28.70\% & $\frac{494}{1281}$ & 38.71\%  \\[1ex]
- coffee-gb      & 463 & $\frac{224}{1927}$ & 11.63\% & $\frac{174}{798}$ & 21.80\% \\[1ex]
- egads          & 252 & $\frac{624}{2427}$ & 25.43\% & $\frac{597}{1163}$ & 51.33\%  \\[1ex]
- apk-parser     & 324 & $\frac{351}{1583}$ & 22.17\% & $\frac{333}{1199}$ & 27.77\%  \\[1ex]\midrule
+\textbf{Project} & \textbf{LOC} & \textbf{Mutants} & \textbf{\% Valid} & \textbf{Mutants} &  \textbf{\% Valid} \\\midrule
+ pebble         & 279 & 1014 & 8.48\%  & 752 & 11.30\%  \\[1ex]
+ spring-hateoas & 397 & 2479 & 10.00\% & 1384 & 12.43\%  \\[1ex]
+ javacc         & 363 & 1888 & 28.70\% & 1281 & 38.71\%  \\[1ex]
+ coffee-gb      & 463 & 1927 & 11.63\% & 798 & 21.80\% \\[1ex]
+ egads          & 252 & 2427 & 25.43\% & 1163 & 51.33\%  \\[1ex]
+ apk-parser     & 324 & 1583 & 22.17\% & 1199 & 27.77\%  \\[1ex]\bottomrule
 \end{tabularx}
@@
 \end{tabularx}
+\begin{tabularx}{\columnwidth}{X|r|cr|cr}
+\toprule
+\mc{6}{c}{\textbf{Rust}} \\ \midrule
+                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
+\textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
+cargo release    & 346          & $\frac{851}{2014}$   & 42.25\%  & $\frac{1690}{2401}$   & 70.39\% \\[1ex]
+passarine        & 376          & $\frac{213}{2294}$   & 9.29\%  & $\frac{173}{1691}$   & 10.23\% \\[1ex]
+typos            & 463          & $\frac{386}{1920}$   & 20.10\%  & $\frac{366}{1565}$   & 23.39\% \\[1ex]
+ord              & 533          & $\frac{762}{2858}$   & 26.66\%  & $\frac{643}{2312}$   & 27.81\% \\[1ex]
+bazuka           & 285          & $\frac{451}{1554}$   & 29.02\%  & $\frac{402}{1322}$   & 30.41\% \\[1ex]
+strum            & 90           & $\frac{51}{428}$     & 11.92\%  & $\frac{50}{368}$     & 13.58\% \\[1ex]\bottomrule
+\end{tabularx}
+\begin{tabularx}{\columnwidth}{Xllrrr}
+               & & & Valid  & Invalid  & \% \\\midrule
+ \multirow{6}{*}{\textbf{Overall}} & \mr{3}{Regex} & Mean     &  452.33 &  1392.33 & 23.21\% \\
+               &   & Std.Dev. & 309.05 & 656.76 & 12.16\%  \\\cline{2-6}
+ & \mr{3}{Comby} & Mean     & 554.00 &  1055.83 & 29.29\% \\
+ %              &   & Med.     & 253.50 & 724.50 & 24.79\% \\
+               &   & Std.Dev. & 592.63 & 508.30 & 21.63\%  \\\bottomrule
+\end{tabularx}}
+
 \end{table}
@@
 
-\begin{table}[htbp]
-\centering
-\caption{C++ (Our Implementation vs. Dextool)}
-\label{tab:table_cpp2}
-\resizebox{\columnwidth}{!}{
-\begin{tabular}{|c|c|r|r|r|r|}
-\hline
- &  & \textbf{Killed} & \textbf{Not Killed} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{Equivalent \%} \\ \hline
-
- \multirow{3}{*}{\small ompl} & Regex & 29 & 566 & 0.05 & 30\% \\\cline{2-6}
-    & Comby & 41 & 523 & 0.07 & 40\% \\\cline{2-6}
-    & Dextool & 23 & 255 & 0.08 & N/A \\ \hline
-\multirow{3}{*}{\small libgraphqlparser} & Regex & 142 & 15 & 0.90 & 30\% \\\cline{2-6}
-    & Comby & 116 & 15 & 0.89 & 20\% \\ \cline{2-6}
-    & Dextool & 50 & 2 & 0.96 & N/A \\ \hline
-\multirow{3}{*}{\small rethinkdb\_rebirth} & Regex & 65 & 480 & 0.12 & 10\% \\\cline{2-6}
-    & Comby & 49 & 443 & 0.10 & 20\% \\ \cline{2-6}
-    & Dextool & 9 & 16 & 0.36 & N/A \\ \hline
-
-\multirow{3}{*}{\small xoreos} & Regex & 33 & 2386 & 0.01 & 0\% \\\cline{2-6}
-    & Comby & 33 & 2247 & 0.01 & 10\% \\ \cline{2-6}
-    & Dextool & 8 & 406 & 0.02 & N/A \\ \hline
-\multirow{3}{*}{\small libxmljs} & Regex & 547 & 334 & 0.62 & 10\% \\\cline{2-6}
-    & Comby & 507 & 283 & 0.64 & 10\% \\ \cline{2-6}
-    & Dextool & 383 & 447 & 0.46 & N/A \\ \hline
-\multirow{3}{*}{\small tiny-differentiable-simulator} & Regex & 38 & 709 & 0.05 & 0\% \\\cline{2-6}
-    & Comby & 32 & 738 & 0.04 & 0\% \\ \cline{2-6}
-    & Dextool & 13 & 48 & 0.21 & N/A\\ \hline
-
-\multirow{9}{*}{Overall} & \multirow{3}{*}{Regex} & \multicolumn{2}{c|}{Mean} & 0.29  & 13.33\%\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.09 & 10.00\% \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.38 & 13.66\% \\\cline{2-6}
-
- & \multirow{3}{*}{Comby} & \multicolumn{2}{c|}{Mean} & 0.29 & 16.67\% \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.09 & 15.00\% \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.37 & 13.66\% \\\cline{2-6}
-
- & \multirow{3}{*}{Dextool} & \multicolumn{2}{c|}{Mean} & 0.33 & N/A\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.29 & N/A \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.38 & N/A \\ \hline
-
-\end{tabular}
-}
-\end{table}
-
 % ---------------- JAVA table (UM vs Others) --------------------------

### 2023-03-29 paper/main.tex
Updating numbers for Java

--- 2023-03-29 paper/main.tex
@@
 
-pebble & 85 & 0.81 & 50\% & 20 & 0.95 & N/A & 8 & 1.00 0\% \\
-... \\\midrule
-\textbf{Mean} &  XX & XX & XX & XX & XX & XX & XX & XX & XX \\
-\textbf{Std.Dev.} &  XX & XX & XX & XX & XX & XX & XX & XX & XX \\
-\end{tabularx}
-}
-\end{table}
-
-% ----------------  Python table (UM vs Others) -------------------------------
-
-\begin{table}[htbp]
-\centering
-{\small
-\caption{Python (Our Implementation vs. Mutmut vs. CosmicRay)}
-\label{tab:table_python2}
-\begin{tabularx}{\columnwidth}{l|rrr|rrr|rrr}
-\toprule
-                 & \mc{3}{c|}{Comby}   & \mc{3}{c|}{MutMut} & \mc{3}{c}{CosmicRay} \\
-\textbf{Project} & \textbf{Mutants} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Eqv}
-& \textbf{Mutants} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Eqv} 
-& \textbf{Mutants} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Eqv} \\ \midrule
-keract  & & & & & & & \\
-\textbf{Mean} &  XX & XX & XX & XX & XX & XX & XX & XX & XX \\
-\textbf{Std.Dev.} &  XX & XX & XX & XX & XX & XX & XX & XX & XX \\
+pebble & 85 & 0.81 & 20\% & 20 & 0.95 & N/A & 8 & 1.00 & 0\% \\
+spring-hateoas & 176 & 0.37 & 0\% & 114 & 0.28 & N/A & 38 & 0.39 & 0\% \\
+javacc & 503 & 0.21 & 0\% & 143 & 0.21 & N/A & 82 & 0.78 & 0\% \\
+coffee-gb & 174 & 0.23 & 0\% & 36 & 0.89 & N/A & 29 & 0.38 & 0\% \\
+egads & 622 & 0.71 & 10\% & 107 & 0.81 & N/A & 85 & 0.87 & 0\% \\
+apk-parser & 338 & 0.08 & 0\% & 89 & 0.16 & N/A & 33 & 0.09 & 0\% \\
+\midrule
+\textbf{Mean} &  316.33 & 0.40 & 5.00\% & 84.83 & 0.55 & N/A & 45.83 & 0.59 & 0.00\% \\
+\textbf{Std.Dev.} &  210.81 & 0.29 & 8.37\% & 47.60 & 0.37 & N/A & 30.93 & 0.35 & 0.00\%8
+ \\
 \end{tabularx}
@@
 }
-\end{table}
-
-% ---------------- Rust table (UM vs Others)--------------------------
-
-\begin{table}[htbp]
-\centering
-{\small
-\caption{Rust (Our Implementation vs. Cargo-Mutants)}
-\label{tab:table_rust2}
-\begin{tabularx}{\columnwidth}{X|crr|crr}
-\hline
-        & \multicolumn{3}{c|}{Comby}  & \multicolumn{3}{c}{cargo-mutants} \\ 
-Project & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} \\[1ex]\midrule
-
-cargo release & & & & & \\    
-\textbf{Mean}   & FIXME & 0.29 & 16.67\% & FIXME & 0.33 & N/A\\
-\textbf{Std. Dev} & FIXME & 0.37 & 13.66\%   & FIXME & 0.38 & N/A \\\bottomrule
-\end{tabularx}}
 \end{table}

### 2023-03-29 paper/main.tex
Rust in the same table as C++ and Java

--- 2023-03-29 paper/main.tex
@@
 \centering
-\caption{Regex vs. Comby for C++ (top), Java (middle), and Rust (bottom)}
+\caption{Regex vs.
+Comby for C++ (top), Java (middle), and Rust (bottom).
+LOC describes the project files considered.
+Muts.
+is total number of mutants generated for those files.
+Invalid mutants are those that don't compile.}
 \label{tab:table_cpp1java}
 \begin{tabularx}{\columnwidth}{X|r|rr|rr}
-\toprule\toprule
+\bottomrule\toprule
 \mc{6}{c}{\textbf{C++}} \\\midrule
                  &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
-\textbf{Project} & \textbf{LOC} & \textbf{Muts} & \textbf{\% Valid} & \textbf{Muts} &  \textbf{\% Valid} \\\midrule
+\textbf{Project} & \textbf{LOC} & \textbf{Muts.} & \textbf{\% Valid} & \textbf{Muts.} &  \textbf{\% Valid} \\\midrule
 ompl & 525 & 2533 & 23.49\% &  2127  & 25.94\%   \\
@@
                   & &       & \textbf{Valid}  & \textbf{Invalid}  & \textbf{\% Valid} \\\midrule
- \multirow{4}{*}{\textbf{Overall}} & \multirow{2}{*}{\bf Regex} & Mean     & 890.67 &  1553.83 & 35.28\% \\
+ \multirow{4}{*}{\textbf{Overall}} & \multirow{2}{*}{Regex} & Mean     & 890.67 &  1553.83 & 35.28\% \\
 %                          & & Med.    & 671.00 & 1579.00  & 23.88\%   \\
                           & & Std.Dev. & 787.62 & 956.31   & 19.48\%  \\\cline{2-6}
-  & \multirow{2}{*}{\bf Comby}  & Mean     & 837.83 & 1352.50 & 38.75\%  \\
+  & \multirow{2}{*}{Comby}  & Mean     & 837.83 & 1352.50 & 38.75\%  \\
 %                         & & Med.    & 667.00 & 1267.00 & 27.86\%  \\
@@
                  &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
-\textbf{Project} & \textbf{LOC} & \textbf{Mutants} & \textbf{\% Valid} & \textbf{Mutants} &  \textbf{\% Valid} \\\midrule
- pebble         & 279 & 1014 & 8.48\%  & 752 & 11.30\%  \\[1ex]
- spring-hateoas & 397 & 2479 & 10.00\% & 1384 & 12.43\%  \\[1ex]
- javacc         & 363 & 1888 & 28.70\% & 1281 & 38.71\%  \\[1ex]
- coffee-gb      & 463 & 1927 & 11.63\% & 798 & 21.80\% \\[1ex]
- egads          & 252 & 2427 & 25.43\% & 1163 & 51.33\%  \\[1ex]
- apk-parser     & 324 & 1583 & 22.17\% & 1199 & 27.77\%  \\[1ex]\bottomrule
+\textbf{Project} & \textbf{LOC} & \textbf{Muts.} & \textbf{\% Valid} & \textbf{Muts.} &  \textbf{\% Valid} \\\midrule
+ pebble         & 279 & 1014 & 8.48\%  & 752 & 11.30\%  \\
+ spring-hateoas & 397 & 2479 & 10.00\% & 1384 & 12.43\%  \\
+ javacc         & 363 & 1888 & 28.70\% & 1281 & 38.71\%  \\
+ coffee-gb      & 463 & 1927 & 11.63\% & 798 & 21.80\% \\
+ egads          & 252 & 2427 & 25.43\% & 1163 & 51.33\%  \\
+ apk-parser     & 324 & 1583 & 22.17\% & 1199 & 27.77\%  \\\bottomrule
 \end{tabularx}
 \begin{tabularx}{\columnwidth}{Xllrrr}
-              & & & Valid  & Invalid  & \% \\\midrule
+              & & & \textbf{Valid}  & \textbf{Invalid}  & \textbf{\% Valid} \\\midrule
 \multirow{4}{*}{\textbf{Overall}} & \mr{2}{Regex} & Mean     &  345.83 & 1545.00  & 17.74\% \\
@@
                  &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
-\textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
-cargo release    & 346          & $\frac{851}{2014}$   & 42.25\%  & $\frac{1690}{2401}$   & 70.39\% \\[1ex]
-passarine        & 376          & $\frac{213}{2294}$   & 9.29\%  & $\frac{173}{1691}$   & 10.23\% \\[1ex]
-typos            & 463          & $\frac{386}{1920}$   & 20.10\%  & $\frac{366}{1565}$   & 23.39\% \\[1ex]
-ord              & 533          & $\frac{762}{2858}$   & 26.66\%  & $\frac{643}{2312}$   & 27.81\% \\[1ex]
-bazuka           & 285          & $\frac{451}{1554}$   & 29.02\%  & $\frac{402}{1322}$   & 30.41\% \\[1ex]
-strum            & 90           & $\frac{51}{428}$     & 11.92\%  & $\frac{50}{368}$     & 13.58\% \\[1ex]\bottomrule
+\textbf{Project} & \textbf{LOC} & \textbf{Muts.} & \textbf{\% Valid} & \textbf{Muts.} &  \textbf{\% Valid} \\\midrule
+cargo release    & 346          & 2014   & 42.25\%  & 2401   & 70.39\% \\
+passarine        & 376          & 2294   & 9.29\%   & 1691   & 10.23\% \\
+typos            & 463          & 1920   & 20.10\%  & 1565   & 23.39\% \\
+ord              & 533          & 2858   & 26.66\%  & 2312   & 27.81\% \\
+bazuka           & 285          & 1554   & 29.02\%  & 1322   & 30.41\% \\
+strum            & 90           & 428     & 11.92\% & 368     & 13.58\% \\\bottomrule
 \end{tabularx}
 \begin{tabularx}{\columnwidth}{Xllrrr}
-               & & & Valid  & Invalid  & \% \\\midrule
- \multirow{6}{*}{\textbf{Overall}} & \mr{3}{Regex} & Mean     &  452.33 &  1392.33 & 23.21\% \\
+               & & & \textbf{Valid}  & \textbf{Invalid}  & \textbf{\% Valid} \\\midrule
+ \multirow{4}{*}{\textbf{Overall}} & \mr{2}{Regex} & Mean     &  452.33 &  1392.33 & 23.21\% \\
                &   & Std.Dev. & 309.05 & 656.76 & 12.16\%  \\\cline{2-6}
- & \mr{3}{Comby} & Mean     & 554.00 &  1055.83 & 29.29\% \\
+ & \mr{2}{Comby} & Mean     & 554.00 &  1055.83 & 29.29\% \\
  %              &   & Med.     & 253.50 & 724.50 & 24.79\% \\
@@
 
-% ---------------- Rust table UM Only --------------------------
-\begin{table}[hbtp]
-\centering
-{\small
-\caption{Rust (Regex vs. Comby)}
-\label{tab:table_rust1}
-\begin{tabularx}{\columnwidth}{X|r|cr|cr}
-\toprule
-                 &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
-\textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
-cargo release    & 346          & $\frac{851}{2014}$   & 42.25\%  & $\frac{1690}{2401}$   & 70.39\% \\[1ex]
-passarine        & 376          & $\frac{213}{2294}$   & 9.29\%  & $\frac{173}{1691}$   & 10.23\% \\[1ex]
-typos            & 463          & $\frac{386}{1920}$   & 20.10\%  & $\frac{366}{1565}$   & 23.39\% \\[1ex]
-ord              & 533          & $\frac{762}{2858}$   & 26.66\%  & $\frac{643}{2312}$   & 27.81\% \\[1ex]
-bazuka           & 285          & $\frac{451}{1554}$   & 29.02\%  & $\frac{402}{1322}$   & 30.41\% \\[1ex]
-strum            & 90           & $\frac{51}{428}$     & 11.92\%  & $\frac{50}{368}$     & 13.58\% \\[1ex]
-\end{tabularx}
- \begin{tabularx}{\columnwidth}{Xllrrr}
-               & & & Valid  & Invalid  & \% \\\midrule
- \multirow{6}{*}{\textbf{Overall}} & \mr{3}{Regex} & Mean     &  452.33 &  1392.33 & 23.21\% \\
-               &   & Std.Dev. & 309.05 & 656.76 & 12.16\%  \\\cline{2-6}
- & \mr{3}{Comby} & Mean     & 554.00 &  1055.83 & 29.29\% \\
- %              &   & Med.     & 253.50 & 724.50 & 24.79\% \\
-               &   & Std.Dev. & 592.63 & 508.30 & 21.63\%  \\\bottomrule
-\end{tabularx}}
-\end{table}
+% % ---------------- Rust table UM Only --------------------------
+% \begin{table}[hbtp]
+% \centering
+% {\small
+% \caption{Rust (Regex vs. Comby)}
+% \label{tab:table_rust1}
+% \begin{tabularx}{\columnwidth}{X|r|cr|cr}
+% \toprule
+%                  &              & \multicolumn{2}{c|}{\textbf{Regex}} &\multicolumn{2}{c}{\textbf{Comby}}  \\
+% \textbf{Project} & \textbf{LOC} & $\frac{valid}{total}$ & \textbf{\%} & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
+% cargo release    & 346          & $\frac{851}{2014}$   & 42.25\%  & $\frac{1690}{2401}$   & 70.39\% \\[1ex]
+% passarine        & 376          & $\frac{213}{2294}$   & 9.29\%  & $\frac{173}{1691}$   & 10.23\% \\[1ex]
+% typos            & 463          & $\frac{386}{1920}$   & 20.10\%  & $\frac{366}{1565}$   & 23.39\% \\[1ex]
+% ord              & 533          & $\frac{762}{2858}$   & 26.66\%  & $\frac{643}{2312}$   & 27.81\% \\[1ex]
+% bazuka           & 285          & $\frac{451}{1554}$   & 29.02\%  & $\frac{402}{1322}$   & 30.41\% \\[1ex]
+% strum            & 90           & $\frac{51}{428}$     & 11.92\%  & $\frac{50}{368}$     & 13.58\% \\[1ex]
+% \end{tabularx}
+%  \begin{tabularx}{\columnwidth}{Xllrrr}
+%                & & & Valid  & Invalid  & \% \\\midrule
+%  \multirow{6}{*}{\textbf{Overall}} & \mr{3}{Regex} & Mean     &  452.33 &  1392.33 & 23.21\% \\
+%                &   & Std.Dev. & 309.05 & 656.76 & 12.16\%  \\\cline{2-6}
+%  & \mr{3}{Comby} & Mean     & 554.00 &  1055.83 & 29.29\% \\
+%  %              &   & Med.     & 253.50 & 724.50 & 24.79\% \\
+%                &   & Std.Dev. & 592.63 & 508.30 & 21.63\%  \\\bottomrule
+% \end{tabularx}}
+% \end{table}
 

### 2023-03-29 paper/main.tex
blank for python
blank cargo mutants
abusing font size a little for python

@@
 \centering
-{\small
-\caption{Regex vs. Comby for Python, including redundant mutants.}
-\label{tab:table_pyrust}
-\begin{tabularx}{\columnwidth}{X|r|rcr|rcr}
+{\footnotesize % feels bad man but it's definitely easier than most other alternatives
+\caption{Regex vs.
+Comby for Python.
+LOC describes the project files considered.
+Muts.
+is total number of mutants generated for those files.
+\claire{FIXME: what's an invalid mutant for python?}  We can use TCE to check for redundant Python mutants, shown under ``Red.''}
+\label{tab:table_python1}
+\begin{tabularx}{\columnwidth}{l|r|rcr|rcr}
 \toprule\toprule
                  &                 & \multicolumn{3}{c|}{\textbf{Regex}}             &\multicolumn{3}{c}{\textbf{Comby}}  \\[1ex]
-\textbf{Project} & \textbf{LOC}    & red.  & $\frac{valid}{total}$ & \textbf{\%} & red. & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
-keract           & 402             & 729       & $\frac{1655}{3984}$   & 41.53\%     & 533       & $\frac{1471}{3067}$   & 47.96\% \\[1ex]
-dtw              & 118             & 567       & $\frac{973}{2106}$   & 46.20\%     & 639       & $\frac{866}{1835}$   & 47.19\% \\[1ex]
-notion-sdk-py    & 189             & 82        & $\frac{196}{1038}$   & 18.89\%     & 122       & $\frac{196}{1019}$   & 19.23\% \\[1ex]
-atlassian-python-api    & 236             & 207        & $\frac{302}{1541}$   & 19.99\%     & 351       & $\frac{333}{1319}$   & 25.25\% \\[1ex]
-ESD              & 698            & 1642      & $\frac{1859}{6268}$   & 29.46\%     & 614       & $\frac{1712}{3874}$   & 44.18\% \\[1ex]
-fiber            & 358            & 163      & $\frac{493}{1565}$   & 31.08\%     & 241       & $\frac{458}{1466}$   & 31.24\% \\[1ex]
+\textbf{Project} & \textbf{LOC}    & \textbf{Muts.} & \textbf{Red.}  & \textbf{\% Valid} & \textbf{Muts.} & \textbf{Red.}   &  \textbf{\% Valid} \\\midrule
+keract           & 402             & 3984  & 729   & 41.53\%           & 3067  & 533    & 47.96\% \\
+dtw              & 118             & 2106  & 567   & 46.20\%           & 1835  & 639    & 47.19\% \\
+notion-          & \mr{2}{189}     & \mr{2}{1038}  & \mr{2}{82}    & \mr{2}{18.89\%}           & \mr{2}{1019}  & \mr{2}{122}    & \mr{2}{19.23\%} \\
+sdk-py           & & & & & & \\[0.5ex]
+atlassian-       & \mr{2}{236}      & \mr{2}{1541}  & \mr{2}{207}   & \mr{2}{19.99\%}           & 1319  & \mr{2}{351}    & \mr{2}{25.25\%} \\
+python-api       & & & & & & \\[0.5ex]
+ESD              & 698             & 6268  & 1642  & 29.46\%           & 3874  & 614    & 44.18\% \\
+fiber            & 358             & 1565  & 163   & 31.08\%           & 1466  & 241    & 31.24\% \\\bottomrule
 \end{tabularx}
 \begin{tabularx}{\columnwidth}{Xllrrrrrr}
-                  && &  Red.     & Valid  & Invalid  & \% \\\midrule
-\multirow{4}{*}{Overall} & \multirow{2}{*}{Regex} & Mean &  565.00 & 913.00 & 1267.33 & 31.19\%  & \\
-    &   &  Std. Dev & 584.79  & 708.93 & 813.35 & 11.06\%  \\
+                  && &  \textbf{Red.}     & \textbf{Valid}  & \textbf{Invalid}  & \textbf{\% Valid} \\\midrule
+\multirow{4}{*}{\textbf{Overall}} & \multirow{2}{*}{Regex} & Mean &  565.00 & 913.00 & 1267.33 & 31.19\%  & \\
+    &   &  Std.Dev. & 584.79  & 708.93 & 813.35 & 11.06\%  \\
  & \multirow{2}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% \\
-  &   &  Std. Dev &  211.62 & 628.79 & 418.99 & 12.28\%   \\
+  &   &  Std.Dev. &  211.62 & 628.79 & 418.99 & 12.28\%   \\\bottomrule
 \end{tabularx}}
@@
 \centering
+{\small
+\caption{Python (Our Implementation vs. Mutmut vs. CosmicRay)}
+\label{tab:table_python2}
+\begin{tabularx}{\columnwidth}{l|rrr|rrr|rrr}
+\hline
+                 & \mc{3}{c|}{Comby}   & \mc{3}{c|}{Mutmut} & \mc{3}{c}{CosmicRay} \\
+\textbf{Project} & \textbf{Mutants} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Eqv}
+& \textbf{Mutants} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Eqv} 
+& \textbf{Mutants} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Eqv} \\ \midrule
+\hline
+keract & \\
+\textbf{Mean} & \\
+\textbf{Std.Dev.} &  \\
+
+\end{tabularx}
+}
+\end{table}
+
+% ----------------  Python table (UM vs Others) -------------------------------
+
+\begin{table}[htbp]
+\centering
 \caption{Python (Our Implementation vs. Mutmut vs. CosmicRay)}
@@
 \end{tabular}
+}
+\end{table}
+
+\begin{table}[htbp]
+\centering
+{\small
+\caption{Rust (Our Implementation vs. Cargo-Mutants)}
+\label{tab:table_rust2}
+\begin{tabularx}{\columnwidth}{X|crr|crr}
+\hline
+        & \multicolumn{3}{c|}{Comby}  & \multicolumn{3}{c}{Cargo-Mutants} \\ 
+Project & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} \\[1ex]\midrule
+\hline
+cargo release & \\
+
+\textbf{Mean}   &  \\
+\textbf{Std. Dev} &  \\
+\end{tabularx}
 }

### 2023-03-29 paper/main.tex
fix up the C++ dextool table

@@
 {\small
-\caption{C++ (Our Implementation vs. Dextool)}
+\caption{C++ (Our Implementation vs. Dextool). We cannot measure equivalent mutants for Dextool and thus omit the column.}
 \label{tab:table_cpp2}
-\begin{tabularx}{\columnwidth}{X|crr|crr}
+\begin{tabularx}{\columnwidth}{X|rrr|rr}
 \hline
-        & \multicolumn{3}{c|}{Comby}  & \multicolumn{3}{c}{Dextool} \\ 
-Project & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} \\[1ex]\midrule
- ompl &  564 & 0.07 & 40\% & 278 & 0.08 & N/A \\
- libgraphqlparser &  131 & 0.90 & 20\% & 52 & 0.96 & N/A \\
- rethinkdb\_rebirth &  492 & 0.10 & 20\% & 25 & 0.36 & N/A \\
- xoreos &  2280 & 0.01 & 10\% & 414 & 0.02 & N/A \\
- libxmljs &  790 & 0.62 & 10\% & 830 & 0.46 & N/A \\
- tiny-differentiable-simulator &  770 & 0.04 & 0\% & 61 & 0.21 & N/A \\
-
-%\multirow{3}{*}{\small libgraphqlparser} & Comby & 116 & 15 & 0.89 & 20\% \\ \cline{2-6}
-%    & Dextool & 50 & 2 & 0.96 & N/A \\ \hline
-%\multirow{3}{*}{\small rethinkdb\_rebirth} & Comby & 49 & 443 & 0.10 & 20\% \\ \cline{2-6}
-%    & Dextool & 9 & 16 & 0.36 & N/A \\ \hline
-
-%\multirow{3}{*}{\small xoreos} & Comby & 33 & 2247 & 0.01 & 10\% \\ \cline{2-6}
-%    & Dextool & 8 & 406 & 0.02 & N/A \\ \hline
-%\multirow{3}{*}{\small libxmljs}  & Comby & 507 & 283 & 0.64 & 10\% \\ \cline{2-6}
-%    & Dextool & 383 & 447 & 0.46 & N/A \\ \hline
-%\multirow{3}{*}{\small tiny-differentiable-simulator} & Comby & 32 & 738 & 0.04 & 0\% \\ \cline{2-6}
-%    & Dextool & 13 & 48 & 0.21 & N/A\\ \hline
-%    
-%\multirow{9}{*}{Overall}\multirow{3}{*}{Comby} & \multicolumn{2}{c|}{Mean} & 0.29 & 16.67\% \\\cline{3-6}
-%    &   &  \multicolumn{2}{c|}{Median} & 0.09 & 15.00\% \\\cline{3-6}
-%    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.37 & 13.66\% \\\cline{2-6}
-%
-% & \multirow{3}{*}{Dextool} & \multicolumn{2}{c|}{Mean} & 0.33 & N/A\\\cline{3-6}
-%    &   &  \multicolumn{2}{c|}{Median} & 0.29 & N/A \\\cline{3-6}
-%    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.38 & N/A \\ \hline
-\textbf{Mean}   & 837.83 & 0.29 & 16.67\% & 276.67 & 0.33 & N/A\\
-\textbf{Std. Dev} & 745.78 & 0.37 & 13.66\%   & 311.56 & 0.38 & N/A \\\bottomrule
+        & \multicolumn{3}{c|}{\textbf{Comby}}  & \multicolumn{2}{c}{\textbf{Dextool}} \\ 
+\textbf{Project} & \textbf{Muts.} & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & \textbf{Muts.} & \makecell{\textbf{Mut.} \\ \textbf{Score}} \\[1ex]\midrule
+ ompl &  564 & 0.07 & 40\% & 278 & 0.08  \\
+ libgraphqlparser  &  131 & 0.90 & 20\% & 52 & 0.96  \\
+ rethinkdb\_rebirth &  492 & 0.10 & 20\% & 25 & 0.36 \\
+ xoreos &  2280 & 0.01 & 10\% & 414 & 0.02  \\
+ libxmljs &  790 & 0.62 & 10\% & 830 & 0.46 \\[0.5ex]
+ tiny-differentiable-simulator &  770 & 0.04 & 0\% & 61 & 0.21  \\
+\textbf{Mean}   & 837.83 & 0.29 & 16.67\% & 276.67 & 0.33\\
+\textbf{Std. Dev} & 745.78 & 0.37 & 13.66\%   & 311.56 & 0.38 \\\bottomrule
 \end{tabularx}}

### 2023-03-29 paper/main.tex
fixed up java table.

@@
 \centering
-{\small
-\caption{Java (Our Implementation vs. PIT vs. LittleDarwin)}
+\caption{Java (Our Implementation vs. PIT vs. LittleDarwin). \claire{FIXME: explain why no eqv for PIT.} LittleDarwin produces no equivalent mutants.\claire{right?}}
 \label{tab:table_java2}
-\begin{tabularx}{\columnwidth}{l|rrr|rrr|rrr}
-\hline
-                 & \mc{3}{c|}{Comby}   & \mc{3}{c|}{PIT} & \mc{3}{c}{LittleDarwin} \\
-\textbf{Project} & \textbf{Mutants} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Eqv}
-& \textbf{Mutants} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Eqv} 
-& \textbf{Mutants} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Eqv} \\ \midrule
-
-pebble & 85 & 0.81 & 20\% & 20 & 0.95 & N/A & 8 & 1.00 & 0\% \\
-spring-hateoas & 176 & 0.37 & 0\% & 114 & 0.28 & N/A & 38 & 0.39 & 0\% \\
-javacc & 503 & 0.21 & 0\% & 143 & 0.21 & N/A & 82 & 0.78 & 0\% \\
-coffee-gb & 174 & 0.23 & 0\% & 36 & 0.89 & N/A & 29 & 0.38 & 0\% \\
-egads & 622 & 0.71 & 10\% & 107 & 0.81 & N/A & 85 & 0.87 & 0\% \\
-apk-parser & 338 & 0.08 & 0\% & 89 & 0.16 & N/A & 33 & 0.09 & 0\% \\
+\resizebox{\columnwidth}{!}{% again don't love it, but footnotesize is actually smaller than necessary here
+\begin{tabular}{l|rrr|rr|rr}
+\toprule
+                 & \mc{3}{c|}{\textbf{Comby}}   & \mc{2}{c|}{\textbf{PIT}} & \mc{2}{c}{\textbf{LittleDarwin}} \\
+\textbf{Project} & \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv}
+& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}}
+& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} \\ \midrule
+
+pebble & 85 & 0.81 & 20\% & 20 & 0.95  & 8 & 1.00  \\
+spring- & \mr{2}{176} & \mr{2}{0.37} & \mr{2}{0\%} & \mr{2}{114} & \mr{2}{0.28}  & \mr{2}{38} & \mr{2}{0.39}  \\
+hateoas  &    &       &     &    &   &   &   \\[0.5ex]
+javacc & 503 & 0.21 & 0\% & 143 & 0.21  & 82 & 0.78  \\
+coffee-gb & 174 & 0.23 & 0\% & 36 & 0.89 & 29 & 0.38  \\
+egads & 622 & 0.71 & 10\% & 107 & 0.81  & 85 & 0.87  \\
+apk-parser & 338 & 0.08 & 0\% & 89 & 0.16  & 33 & 0.09 \\
 \midrule
-\textbf{Mean} &  316.33 & 0.40 & 5.00\% & 84.83 & 0.55 & N/A & 45.83 & 0.59 & 0.00\% \\
-\textbf{Std.Dev.} &  210.81 & 0.29 & 8.37\% & 47.60 & 0.37 & N/A & 30.93 & 0.35 & 0.00\%8
- \\
-\end{tabularx}
+\textbf{Mean} &  316.33 & 0.40 & 5.00\% & 84.83 & 0.55  & 45.83 & 0.59  \\
+\textbf{Std.Dev.} &  210.81 & 0.29 & 8.37\% & 47.60 & 0.37 & 30.93 & 0.35 \\\bottomrule
+\end{tabular}
 }

### 2023-03-29 paper/main.tex
correct single java table

@@
 \centering
-\caption{Java (Our Implementation vs. PIT vs. LittleDarwin)}
-\label{tab:table_java2}
-\resizebox{\columnwidth}{!}{
-\begin{tabular}{|c|c|r|r|r|r|}
-\hline
- &  & \textbf{Killed} & \textbf{Not Killed} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{Equivalent \%} \\ \hline
-
- \multirow{4}{*}{\small pebble} & Regex & 70 & 16 & 0.81 & 50\% \\\cline{2-6}
-    & Comby & 68 & 17 & 0.80 & 20\% \\\cline{2-6}
-    & PIT & 19 & 1 & 0.95 & N/A \\ \cline{2-6}
-    & LittleDarwin & 8 & 0 & 1.00 & 0\% \\ \hline
-\multirow{4}{*}{\small spring-hateoas} & Regex & 82 & 166 & 0.33 & 0\% \\\cline{2-6}
-    & Comby & 65 & 111 & 0.37 & 0\% \\ \cline{2-6}
-    & PIT & 82 & 32 & 0.28 & N/A \\ \cline{2-6}
-    & LittleDarwin & 15 & 23 & 0.39 & 0\% \\ \hline
-\multirow{4}{*}{\small javacc} & Regex & 123 & 419 & 0.23 & 0\% \\\cline{2-6}
-    & Comby & 103 & 400 & 0.21 & 0\% \\ \cline{2-6}
-    & PIT & 30 & 113 & 0.21 & N/A\\ \cline{2-6}
-    & LittleDarwin & 64 & 18 & 0.78 & 0\%\\ \hline
-\multirow{4}{*}{\small coffee-gb} & Regex & 48 & 176 & 0.21 & 0\%\\\cline{2-6}
-    & Comby & 40 & 134 & 0.23 & 0\% \\ \cline{2-6}
-    & PIT & 32 & 4 & 0.89 & N/A \\ \cline{2-6}
-    & LittleDarwin & 11 & 18 & 0.38 & 0\%\\ \hline
-\multirow{4}{*}{\small egads} & Regex & 442 & 182 & 0.71 & 20\% \\\cline{2-6}
-    & Comby & 422 & 200 & 0.71 & 10\% \\ \cline{2-6}
-    & PIT & 87 & 20 & 0.81 & N/A \\ \cline{2-6}
-    & LittleDarwin & 74 & 11 & 0.87 & 0\% \\ \hline
-\multirow{4}{*}{\small apk-parser} & Regex & 26 & 325 & 0.07 & 10\%\\\cline{2-6}
-    & Comby & 26 & 312 & 0.08 & 0\%\\ \cline{2-6}
-    & PIT & 14 & 75  & 0.16 & N/A\\ \cline{2-6}
-    & LittleDarwin & 3 & 30 & 0.09 & 0\%\\ \hline
-
-\multirow{12}{*}{Overall} & \multirow{3}{*}{Regex} & \multicolumn{2}{c|}{Mean} & 0.39  & 13.33\%\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.28 & 5.00\% \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.30 & 19.66\% \\\cline{2-6}
-
- & \multirow{3}{*}{Comby} & \multicolumn{2}{c|}{Mean} & 0.40 & 5.00\%\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.30 & 0.00\% \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.29 & 8.37\%\\\cline{2-6}
-
- & \multirow{3}{*}{PIT} & \multicolumn{2}{c|}{Mean} & 0.55 & N/A\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.55 & N/A \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.37 & N/A \\\cline{2-6}
-
- & \multirow{3}{*}{LittleDarwin} & \multicolumn{2}{c|}{Mean} & 0.59 & 0.00\%\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.59 & 0.00\%\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.35 & 0.00\%\\ \hline
-
-\end{tabular}
-}
-\end{table}
-
-% ---------------- JAVA table (UM vs Others) --------------------------
-
-\begin{table}[htbp]
-\centering
 \caption{Java (Our Implementation vs. PIT vs. LittleDarwin). \claire{FIXME: explain why no eqv for PIT.} LittleDarwin produces no equivalent mutants.\claire{right?}}
 \label{tab:table_java2}
-\resizebox{\columnwidth}{!}{% again don't love it, but footnotesize is actually smaller than necessary here
+\resizebox{\columnwidth}{!}{% again don't love it, but footnotesize is actually smaller than necessary
 \begin{tabular}{l|rrr|rr|rr}

### 2023-03-29 paper/main.tex
Python numbers

--- 2023-03-29 paper/main.tex
@@
 \centering
-{\footnotesize % feels bad man but it's definitely easier than most other alternatives
-\caption{Regex vs.
-Comby for Python.
-LOC describes the project files considered.
-Muts.
-is total number of mutants generated for those files.
-\claire{FIXME: what's an invalid mutant for python?}  We can use TCE to check for redundant Python mutants, shown under ``Red.''}
-\label{tab:table_python1}
-\begin{tabularx}{\columnwidth}{l|r|rcr|rcr}
+{\small
+\caption{Regex vs. Comby for Python, including redundant mutants.}
+\label{tab:table_pyrust}
+\begin{tabularx}{\columnwidth}{X|r|rcr|rcr}
 \toprule\toprule
                  &                 & \multicolumn{3}{c|}{\textbf{Regex}}             &\multicolumn{3}{c}{\textbf{Comby}}  \\[1ex]
-\textbf{Project} & \textbf{LOC}    & \textbf{Muts.} & \textbf{Red.}  & \textbf{\% Valid} & \textbf{Muts.} & \textbf{Red.}   &  \textbf{\% Valid} \\\midrule
-keract           & 402             & 3984  & 729   & 41.53\%           & 3067  & 533    & 47.96\% \\
-dtw              & 118             & 2106  & 567   & 46.20\%           & 1835  & 639    & 47.19\% \\
-notion-          & \mr{2}{189}     & \mr{2}{1038}  & \mr{2}{82}    & \mr{2}{18.89\%}           & \mr{2}{1019}  & \mr{2}{122}    & \mr{2}{19.23\%} \\
-sdk-py           & & & & & & \\[0.5ex]
-atlassian-       & \mr{2}{236}      & \mr{2}{1541}  & \mr{2}{207}   & \mr{2}{19.99\%}           & 1319  & \mr{2}{351}    & \mr{2}{25.25\%} \\
-python-api       & & & & & & \\[0.5ex]
-ESD              & 698             & 6268  & 1642  & 29.46\%           & 3874  & 614    & 44.18\% \\
-fiber            & 358             & 1565  & 163   & 31.08\%           & 1466  & 241    & 31.24\% \\\bottomrule
+\textbf{Project} & \textbf{LOC}    & red.  & $\frac{valid}{total}$ & \textbf{\%} & red. & $\frac{valid}{total}$ &  \textbf{\%} \\[1ex]\midrule
+keract           & 402             & 729       & $\frac{1655}{3984}$   & 41.53\%     & 533       & $\frac{1471}{3067}$   & 47.96\% \\[1ex]
+dtw              & 118             & 567       & $\frac{973}{2106}$   & 46.20\%     & 639       & $\frac{866}{1835}$   & 47.19\% \\[1ex]
+notion-sdk-py    & 189             & 82        & $\frac{196}{1038}$   & 18.89\%     & 122       & $\frac{196}{1019}$   & 19.23\% \\[1ex]
+atlassian-python-api    & 236             & 207        & $\frac{302}{1541}$   & 19.99\%     & 351       & $\frac{333}{1319}$   & 25.25\% \\[1ex]
+ESD              & 698            & 1642      & $\frac{1859}{6268}$   & 29.46\%     & 614       & $\frac{1712}{3874}$   & 44.18\% \\[1ex]
+fiber            & 358            & 163      & $\frac{493}{1565}$   & 31.08\%     & 241       & $\frac{458}{1466}$   & 31.24\% \\[1ex]
 \end{tabularx}
 \begin{tabularx}{\columnwidth}{Xllrrrrrr}
-                  && &  \textbf{Red.}     & \textbf{Valid}  & \textbf{Invalid}  & \textbf{\% Valid} \\\midrule
-\multirow{4}{*}{\textbf{Overall}} & \multirow{2}{*}{Regex} & Mean &  565.00 & 913.00 & 1267.33 & 31.19\%  & \\
-    &   &  Std.Dev. & 584.79  & 708.93 & 813.35 & 11.06\%  \\
+                  && &  Red.     & Valid  & Invalid  & \% \\\midrule
+\multirow{4}{*}{Overall} & \multirow{2}{*}{Regex} & Mean &  565.00 & 913.00 & 1267.33 & 31.19\%  & \\
+    &   &  Std. Dev & 584.79  & 708.93 & 813.35 & 11.06\%  \\
  & \multirow{2}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% \\
-  &   &  Std.Dev. &  211.62 & 628.79 & 418.99 & 12.28\%   \\\bottomrule
+  &   &  Std. Dev &  211.62 & 628.79 & 418.99 & 12.28\%   \\
 \end{tabularx}}
@@
 {\small
-\caption{C++ (Our Implementation vs. Dextool). We cannot measure equivalent mutants for Dextool and thus omit the column.}
+\caption{C++ (Our Implementation vs. Dextool)}
 \label{tab:table_cpp2}
-\begin{tabularx}{\columnwidth}{X|rrr|rr}
+\begin{tabularx}{\columnwidth}{X|crr|crr}
 \hline
-        & \multicolumn{3}{c|}{\textbf{Comby}}  & \multicolumn{2}{c}{\textbf{Dextool}} \\ 
-\textbf{Project} & \textbf{Muts.} & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & \textbf{Muts.} & \makecell{\textbf{Mut.} \\ \textbf{Score}} \\[1ex]\midrule
- ompl &  564 & 0.07 & 40\% & 278 & 0.08  \\
- libgraphqlparser  &  131 & 0.90 & 20\% & 52 & 0.96  \\
- rethinkdb\_rebirth &  492 & 0.10 & 20\% & 25 & 0.36 \\
- xoreos &  2280 & 0.01 & 10\% & 414 & 0.02  \\
- libxmljs &  790 & 0.62 & 10\% & 830 & 0.46 \\[0.5ex]
- tiny-differentiable-simulator &  770 & 0.04 & 0\% & 61 & 0.21  \\
-\textbf{Mean}   & 837.83 & 0.29 & 16.67\% & 276.67 & 0.33\\
-\textbf{Std. Dev} & 745.78 & 0.37 & 13.66\%   & 311.56 & 0.38 \\\bottomrule
+        & \multicolumn{3}{c|}{Comby}  & \multicolumn{3}{c}{Dextool} \\ 
+Project & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} \\[1ex]\midrule
+ ompl &  564 & 0.07 & 40\% & 278 & 0.08 & N/A \\
+ libgraphqlparser &  131 & 0.90 & 20\% & 52 & 0.96 & N/A \\
+ rethinkdb\_rebirth &  492 & 0.10 & 20\% & 25 & 0.36 & N/A \\
+ xoreos &  2280 & 0.01 & 10\% & 414 & 0.02 & N/A \\
+ libxmljs &  790 & 0.62 & 10\% & 830 & 0.46 & N/A \\
+ tiny-differentiable-simulator &  770 & 0.04 & 0\% & 61 & 0.21 & N/A \\
+
+%\multirow{3}{*}{\small libgraphqlparser} & Comby & 116 & 15 & 0.89 & 20\% \\ \cline{2-6}
+%    & Dextool & 50 & 2 & 0.96 & N/A \\ \hline
+%\multirow{3}{*}{\small rethinkdb\_rebirth} & Comby & 49 & 443 & 0.10 & 20\% \\ \cline{2-6}
+%    & Dextool & 9 & 16 & 0.36 & N/A \\ \hline
+
+%\multirow{3}{*}{\small xoreos} & Comby & 33 & 2247 & 0.01 & 10\% \\ \cline{2-6}
+%    & Dextool & 8 & 406 & 0.02 & N/A \\ \hline
+%\multirow{3}{*}{\small libxmljs}  & Comby & 507 & 283 & 0.64 & 10\% \\ \cline{2-6}
+%    & Dextool & 383 & 447 & 0.46 & N/A \\ \hline
+%\multirow{3}{*}{\small tiny-differentiable-simulator} & Comby & 32 & 738 & 0.04 & 0\% \\ \cline{2-6}
+%    & Dextool & 13 & 48 & 0.21 & N/A\\ \hline
+%    
+%\multirow{9}{*}{Overall}\multirow{3}{*}{Comby} & \multicolumn{2}{c|}{Mean} & 0.29 & 16.67\% \\\cline{3-6}
+%    &   &  \multicolumn{2}{c|}{Median} & 0.09 & 15.00\% \\\cline{3-6}
+%    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.37 & 13.66\% \\\cline{2-6}
+%
+% & \multirow{3}{*}{Dextool} & \multicolumn{2}{c|}{Mean} & 0.33 & N/A\\\cline{3-6}
+%    &   &  \multicolumn{2}{c|}{Median} & 0.29 & N/A \\\cline{3-6}
+%    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.38 & N/A \\ \hline
+\textbf{Mean}   & 837.83 & 0.29 & 16.67\% & 276.67 & 0.33 & N/A\\
+\textbf{Std. Dev} & 745.78 & 0.37 & 13.66\%   & 311.56 & 0.38 & N/A \\\bottomrule
 \end{tabularx}}
@@
 \textbf{Mean} &  316.33 & 0.40 & 5.00\% & 84.83 & 0.55 & N/A & 45.83 & 0.59 & 0.00\% \\
-\textbf{Std.Dev.} &  210.81 & 0.29 & 8.37\% & 47.60 & 0.37 & N/A & 30.93 & 0.35 & 0.00\%8
+\textbf{Std.Dev.} &  210.81 & 0.29 & 8.37\% & 47.60 & 0.37 & N/A & 30.93 & 0.35 & 0.00\%
  \\
@@
 \hline
-keract & \\
-\textbf{Mean} & \\
-\textbf{Std.Dev.} &  \\
+keract & 1471 & 0.26 & 10\% & 482 & 0.72 & 10\% & 921 & 0.27 & 10\% \\
+dtw & 866 & 0.55 & 10\% & 270 & 0.44 & 0\% & 713 & 0.46 & 20\% \\
+notion-sdk-py & 196 & 0.60 & 20\% & 63 & 0.65 & 0\% & 52 & 0.44 & 0\% \\
+atlassian-python-api & 333 & 0.62 & 40\% & 174 & 0.71 & 10\% & 106 & 0.74 & 10\% \\
+ESD & 1712 & 0.07 & 10\% & 720 & 0.08 & 0\% & 1013 & 0.06 & 10\% \\
+Fiber & 458 & 0.64 & 20\% & 222 & 0.57 & 40\% & 320 & 0.81 & 0\% \\
+\textbf{Mean} &  839.33 & 0.46 & 18.00\% & 321.83 & 0.53 & 10.00\% & 520.93 & 0.46 & 8.33\% \\
+\textbf{Std.Dev.} &  628.80 & 0.24 & 11.69\% & 239.04 & 0.24 & 15.49\% & 417.50 & 0.28 & 7.53\%
 
@@
 \end{tabular}
-}
-\end{table}
-
-\begin{table}[htbp]
-\centering
-{\small
-\caption{Rust (Our Implementation vs. Cargo-Mutants)}
-\label{tab:table_rust2}
-\begin{tabularx}{\columnwidth}{X|crr|crr}
-\hline
-        & \multicolumn{3}{c|}{Comby}  & \multicolumn{3}{c}{Cargo-Mutants} \\ 
-Project & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} \\[1ex]\midrule
-\hline
-cargo release & \\
-
-\textbf{Mean}   &  \\
-\textbf{Std. Dev} &  \\
-\end{tabularx}
 }

### 2023-03-29 paper/main.tex
I regret to report that we need at least one full-width table

--- 2023-03-29 paper/main.tex
@@
 \begin{tabularx}{\columnwidth}{X|rrr|rr}
-\hline
+\toprule
         & \multicolumn{3}{c|}{\textbf{Comby}}  & \multicolumn{2}{c}{\textbf{Dextool}} \\ 
@@
  libxmljs &  790 & 0.62 & 10\% & 830 & 0.46 \\[0.5ex]
- tiny-differentiable-simulator &  770 & 0.04 & 0\% & 61 & 0.21  \\
+ tiny-differentiable-simulator &  770 & 0.04 & 0\% & 61 & 0.21  \\\midrule
 \textbf{Mean}   & 837.83 & 0.29 & 16.67\% & 276.67 & 0.33\\
-\textbf{Std. Dev} & 745.78 & 0.37 & 13.66\%   & 311.56 & 0.38 \\\bottomrule
+\textbf{Std.Dev.} & 745.78 & 0.37 & 13.66\%   & 311.56 & 0.38 \\\bottomrule
 \end{tabularx}}
@@
 
-\begin{table}[htbp]
+\begin{table*}[htbp]
 \centering
-{\small
+{
 \caption{Python (Our Implementation vs. Mutmut vs. CosmicRay)}
 \label{tab:table_python2}
-\begin{tabularx}{\columnwidth}{l|rrr|rrr|rrr}
-\hline
-                 & \mc{3}{c|}{Comby}   & \mc{3}{c|}{Mutmut} & \mc{3}{c}{CosmicRay} \\
-\textbf{Project} & \textbf{Mutants} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Eqv}
-& \textbf{Mutants} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Eqv} 
-& \textbf{Mutants} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{\% Eqv} \\ \midrule
-\hline
-keract & \\
-\textbf{Mean} & \\
-\textbf{Std.Dev.} &  \\
-
-\end{tabularx}
-}
-\end{table}
-
-% ----------------  Python table (UM vs Others) -------------------------------
-
-\begin{table}[htbp]
-\centering
-\caption{Python (Our Implementation vs. Mutmut vs. CosmicRay)}
-\label{tab:table_python2}
-\resizebox{\columnwidth}{!}{
-\begin{tabular}{|c|c|r|r|r|r|}
-\hline
- &  & \textbf{Killed} & \textbf{Not Killed} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{Equivalent \%} \\ \hline
-
- \multirow{4}{*}{\small keract} & Regex & 463 & 1192 & 0.28 & 0\% \\\cline{2-6}
-    & Comby & 383 & 1088 & 0.26 & 10\% \\\cline{2-6}
-    & Mutmut & 349 & 133 & 0.72 & 10\% \\ \cline{2-6}
-    & CosmicRay & 250 & 671 & 0.27 & 10\% \\ \hline
-\multirow{3}{*}{\small dtw} & Regex & 506 & 467 & 0.52 & 10\% \\\cline{2-6}
-    & Comby & 475 & 391 & 0.55 & 10\% \\ \cline{2-6}
-    & Mutmut & 118 & 152 & 0.44 & 0\% \\ \cline{2-6}
-    & CosmicRay & 328 & 385 & 0.46 & 20\% \\ \hline
-\multirow{3}{*}{\small notion-sdk-py} & Regex & 112 & 84 & 0.57 & 10\% \\\cline{2-6}
-    & Comby & 118 & 78 & 0.60 & 20\% \\ \cline{2-6}
-    & Mutmut & 41 & 22 & 0.65 & 0\%\\ \cline{2-6}
-    & CosmicRay & 23 & 29 & 0.44 & 0\%\\ \hline
-
-\multirow{3}{*}{\makecell{\small atlassian-python \\ api}} & Regex & 193 & 109 & 0.64 & 20\%\\\cline{2-6}
-    & Comby & 207 & 126 & 0.62 & 40\% \\ \cline{2-6}
-    & Mutmut & 123 & 51 & 0.71 & 10\% \\ \cline{2-6}
-    & CosmicRay & 78 & 28 & 0.74 & 10\% \\ \hline
-\multirow{3}{*}{\small ESD} & Regex & 132 & 1727 & 0.07 & 30\%\\\cline{2-6}
-    & Comby & 121 & 1591 & 0.07 & 10\%\\ \cline{2-6}
-    & Mutmut & 57 & 663 & 0.08 & 0\%\\ \cline{2-6}
-    & CosmicRay & 63 & 950 & 0.06 & 10\%\\ \hline
-\multirow{3}{*}{\small fiber} & Regex & 307 & 186 & 0.62 & 10\%\\\cline{2-6}
-    & Comby & 291 & 167 & 0.64 & 20\%\\ \cline{2-6}
-    & Mutmut & 127 & 95 & 0.57 & 40\%\\ \cline{2-6}
-    & CosmicRay & 259 & 61 & 0.81 & 0\%\\ \hline
-
-\multirow{12}{*}{Overall} & \multirow{3}{*}{Regex} & \multicolumn{2}{c|}{Mean} & 0.45  & 13.00\%\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.55 & 10.00\%\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.23 & 10.33\%\\\cline{2-6}
-
- & \multirow{3}{*}{Comby} & \multicolumn{2}{c|}{Mean} & 0.46 & 18.00\%\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.58 & 15.00\%\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.24 & 11.69\%\\\cline{2-6}
-
- & \multirow{3}{*}{Mutmut} & \multicolumn{2}{c|}{Mean} & 0.53 & 10.00\%\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.61 & 5.00\%\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.24 & 15.49\%\\\cline{2-6}
-
- & \multirow{3}{*}{CosmicRay} & \multicolumn{2}{c|}{Mean} & 0.46 & 8.33\%\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.45 & 10.00\%\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.28 & 7.53\%\\ \hline
-
+\begin{tabular}{l|rrr|rrr|rrr}
+\toprule
+                 & \mc{3}{c|}{\textbf{Comby}}   & \mc{3}{c|}{\textbf{Mutmut}} & \mc{3}{c}{\textbf{CosmicRay}} \\
+\textbf{Project} & \textbf{Muts} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv}
+& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv} 
+& \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv} \\ \midrule
+\midrule
+keract & 1471 & 0.26 & 10\% & 482 & 0.72 & 10\% & 921 & 0.27 & 10\% \\
+dtw & 866 & 0.55 & 10\% & 270 & 0.44 & 0\% & 713 & 0.46 & 20\% \\
+notion-sdk-py & 196 & 0.60 & 20\% & 63 & 0.65 & 0\% & 52 & 0.44 & 0\% \\
+atlassian-python-api & 333 & 0.62 & 40\% & 174 & 0.71 & 10\% & 106 & 0.74 & 10\% \\
+ESD & 1712 & 0.07 & 10\% & 720 & 0.08 & 0\% & 1013 & 0.06 & 10\% \\
+Fiber & 458 & 0.64 & 20\% & 222 & 0.57 & 40\% & 320 & 0.81 & 0\% \\\midrule
+\textbf{Mean} &  839.33 & 0.46 & 18.00\% & 321.83 & 0.53 & 10.00\% & 520.93 & 0.46 & 8.33\% \\
+\textbf{Std.Dev.} &  628.80 & 0.24 & 11.69\% & 239.04 & 0.24 & 15.49\% & 417.50 & 0.28 & 7.53\% \\\bottomrule        
 \end{tabular}
 }
-\end{table}
+\end{table*}
 

### 2023-03-29 paper/main.tex
Adding Rust numbers
broken labels, mihnor spacing on table 3

@@
 
-Tables~\ref{tab:table_cpp1}-\ref{tab:table_rust1} compare generated
+Table~\ref{tab:table_cpp1javarust} and Table~\ref{tab:table_python1} compare generated
 mutants for Regex and Comby modes of our tool.  Invalid
@@
 
-% ---------------- C++ table UM Only--------------------------
+% ---------------- C++/Java/Rust table UM Only--------------------------
 \begin{table}[hbtp]
@@
 Invalid mutants are those that don't compile.}
-\label{tab:table_cpp1java}
+\label{tab:table_cpp1javarust}
 \begin{tabularx}{\columnwidth}{X|r|rr|rr}
@@
 ompl & 525 & 2533 & 23.49\% &  2127  & 25.94\%   \\
-libgraphqlparser& 137 & 741 & 21.19\%  & 440 & 29.77\% \\
-rethinkdb\_rebirth &471 & 2246 & 24.27\% & 2009 & 24.49\% \\
+{\small libgraphqlparser} & 137 & 741 & 21.19\%  & 440 & 29.77\% \\
+{\small rethinkdb\_rebirth} &471 & 2246 & 24.27\% & 2009 & 24.49\% \\
 xoreos & 732  & 3876  & 62.41\% & 3297 & 69.15\% \\
@@
                &   & Std.Dev. & 592.63 & 508.30 & 21.63\%  \\\bottomrule
-\end{tabularx}}
-
+\end{tabularx}
 \end{table}
@@
 is total number of mutants generated for those files.
-\claire{FIXME: what's an invalid mutant for python?}  We can use TCE to check for redundant Python mutants, shown under ``Red.''}
+\clg{FIXME: what's an invalid mutant for python?}  We can use TCE to check for redundant Python mutants, shown under ``Red.''}
 \label{tab:table_python1}
@@
 \centering
-\caption{Java (Our Implementation vs. PIT vs. LittleDarwin). \claire{FIXME: explain why no eqv for PIT.} LittleDarwin produces no equivalent mutants.\claire{right?}}
+\caption{Java (Our Implementation vs. PIT vs. LittleDarwin). \clg{FIXME: explain why no eqv for PIT.} LittleDarwin produces no equivalent mutants.\clg{right?}}
 \label{tab:table_java2}
@@
                  & \mc{3}{c|}{\textbf{Comby}}   & \mc{3}{c|}{\textbf{Mutmut}} & \mc{3}{c}{\textbf{CosmicRay}} \\
-\textbf{Project} & \textbf{Muts} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv}
+\textbf{Project} & \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv}
 & \textbf{Muts.} &  \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv} 
@@
 
-% ---------------- Rust table (UM vs Others)--------------------------
-
-\begin{table}[htbp]
-\centering
-\caption{Rust (Our Implementation vs. Cargo-Mutants)}
-\label{tab:table_rust2}
-\resizebox{\columnwidth}{!}{
-\begin{tabular}{|c|c|r|r|r|r|}
-\hline
- &  & \textbf{Killed} & \textbf{Not Killed} &  \makecell{\textbf{Mutation} \\ \textbf{Score}} & \textbf{Equivalent \%} \\ \hline
-
- \multirow{3}{*}{\small cargo release} & Regex & 721 & 130 & 0.85 & 0\% \\\cline{2-6}
-    & Comby & 1344 & 346 & 0.80 & 0\% \\\cline{2-6}
-    & Cargo-Mutants & 2	& 3 & 0.40 & 0\% \\ \hline
-\multirow{3}{*}{\small passerine} & Regex & 122 & 91 & 0.57 & 20\% \\\cline{2-6}
-    & Comby & 90 & 83 & 0.52 & 50\% \\ \cline{2-6}
-    & Cargo-Mutants & 0	& 0 & 0 & 0\% \\ \hline
-\multirow{3}{*}{\small typos} & Regex & 295 & 91 & 0.76 & 40\% \\\cline{2-6}
-    & Comby & 283 & 83 & 0.77 & 20\%\\ \cline{2-6}
-    & Cargo-Mutants & 5 & 1 & 0.83 & 0\%\\ \hline
-
-\multirow{3}{*}{\small ord} & Regex & 558 & 204 & 0.73 & 10\% \\\cline{2-6}
-    & Comby & 533 & 110 & 0.83 & 20\% \\ \cline{2-6}
-    & Cargo-Mutants & 6 & 0 & 1.00 & 0\% \\ \hline
-\multirow{3}{*}{\small bazuka} & Regex & 281 & 170 & 0.62 & 40\% \\\cline{2-6}
-    & Comby & 291 & 111 & 0.73 & 0\% \\ \cline{2-6}
-    & Cargo-Mutants & 0 & 0 & 0 & 0\% \\ \hline
-\multirow{3}{*}{\small strum} & Regex & 8 & 43 & 0.16 & 10\%\\\cline{2-6}
-    & Comby & 11 & 39 & 0.22 & 40\%\\ \cline{2-6}
-    & Cargo-Mutants & 2 & 0 & 1.00 & 0\%\\ \hline
-
-\multirow{9}{*}{Overall} & \multirow{3}{*}{Regex} & \multicolumn{2}{c|}{Mean} & 0.62  & 20.00\%\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.68 & 15.00\%\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.24 & 16.73\% \\\cline{2-6}
-
- & \multirow{3}{*}{Comby} & \multicolumn{2}{c|}{Mean} & 0.64 & 22.00\%\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.75 & 20.00\% \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.25 & 20.41\% \\\cline{2-6}
-
- & \multirow{3}{*}{Cargo-Mutants} & \multicolumn{2}{c|}{Mean} & 0.54 & 0.00\%\\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Median} & 0.62 & 0.00\% \\\cline{3-6}
-    &   &  \multicolumn{2}{c|}{Std. Dev} & 0.47 & 0.00\%\\ \hline
-
-\end{tabular}
-}
-\end{table}
-
 \begin{table}[htbp]
@@
 \label{tab:table_rust2}
-\begin{tabularx}{\columnwidth}{X|crr|crr}
-\hline
-        & \multicolumn{3}{c|}{Comby}  & \multicolumn{3}{c}{Cargo-Mutants} \\ 
-Project & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & Mutants & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} \\[1ex]\midrule
-\hline
-cargo release & \\
-
-\textbf{Mean}   &  \\
-\textbf{Std. Dev} &  \\
+\begin{tabularx}{\columnwidth}{X|rrr|rrr}
+\toprule
+        & \multicolumn{3}{c|}{\textbf{Comby}}  & \multicolumn{3}{c}{\textbf{Cargo-Mutants}} \\ 
+\textbf{Project} & \textbf{Muts.} & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} & \textbf{Muts.} & \makecell{\textbf{Mut.} \\ \textbf{Score}} & \textbf{\% Eqv.} \\\midrule
+cargo release & 1690 & 0.80 & 0\% & 5 & 0.40 & 0\%   \\
+passarine & 173 & 0.52 & 50\% & 0 & 0.00 & 0\%   \\
+typos & 366 & 0.77 & 20\% & 6 & 0.83 & 0\%   \\
+ord & 643 & 0.83 & 20\% & 6 & 1.00 & 0\%   \\
+bazuka & 402 & 0.73 & 0\% & 0 & 0.00 & 0\%   \\
+strum & 50 & 0.22 & 40\% & 2 & 1.00 & 0\%   \\\midrule
+\textbf{Mean}   & 554.00 & 0.64 & 22.00\% & 3.17 & 0.54 & 0.00\%\\
+\textbf{Std. Dev} & 592.63 & 0.25 & 20.41\%   & 2.86 & 0.47 & 0.00\% \\\bottomrule
 \end{tabularx}
@@
 produce many mutants.  In fact, inspection of the non-comment source
-lines as measured by CLOC shown in Tables~\ref{tab:table_java1} and
-\ref{tab:table_java2} and Tables~\ref{tab:table_rust1} and
+lines as measured by CLOC as compared to results as shown in Tables~\ref{tab:table_cpp1javarust},
+Tables~\ref{tab:table_java2} and 
 \ref{tab:table_rust2} shows that for many programs, LittleDarwin and

### 2023-03-29 paper/main.tex
Fixing typos

--- 2023-03-29 paper/main.tex
@@
 \textbf{Project} & \textbf{LOC} & \textbf{Muts.} & \textbf{\% Valid} & \textbf{Muts.} &  \textbf{\% Valid} \\\midrule
-ompl & 525 & 2533 & 23.49\% &  2127  & 25.94\%   \\
+ompl & 525 & 2533 & 23.49\% &  2174  & 25.94\%   \\
 libgraphqlparser& 137 & 741 & 21.19\%  & 440 & 29.77\% \\
@@
  spring-hateoas & 397 & 2479 & 10.00\% & 1384 & 12.43\%  \\
- javacc         & 363 & 1888 & 28.70\% & 1281 & 38.71\%  \\
+ javacc         & 363 & 1888 & 28.70\% & 1276 & 38.71\%  \\
  coffee-gb      & 463 & 1927 & 11.63\% & 798 & 21.80\% \\
- egads          & 252 & 2427 & 25.43\% & 1163 & 51.33\%  \\
+ egads          & 252 & 2454 & 25.43\% & 1163 & 51.33\%  \\
  apk-parser     & 324 & 1583 & 22.17\% & 1199 & 27.77\%  \\\bottomrule
@@
                &   & Std.Dev. & 592.63 & 508.30 & 21.63\%  \\\bottomrule
-\end{tabularx}}
+\end{tabularx}
 
@@
 sdk-py           & & & & & & \\[0.5ex]
-atlassian-       & \mr{2}{236}      & \mr{2}{1541}  & \mr{2}{207}   & \mr{2}{19.99\%}           & 1319  & \mr{2}{351}    & \mr{2}{25.25\%} \\
+atlassian-       & \mr{2}{236}      & \mr{2}{1511}  & \mr{2}{207}   & \mr{2}{19.99\%}           & 1319  & \mr{2}{351}    & \mr{2}{25.25\%} \\
 python-api       & & & & & & \\[0.5ex]
@@
  xoreos &  2280 & 0.01 & 10\% & 414 & 0.02  \\
- libxmljs &  790 & 0.62 & 10\% & 830 & 0.46 \\[0.5ex]
+ libxmljs &  790 & 0.64 & 10\% & 830 & 0.46 \\[0.5ex]
  tiny-differentiable-simulator &  770 & 0.04 & 0\% & 61 & 0.21  \\\midrule
@@
 \hline
-cargo release & \\
-
-\textbf{Mean}   &  \\
-\textbf{Std. Dev} &  \\
+cargo release & 1690 & 0.80 & 0\% & 5 & 0.40 & 0\%   \\
+passarine & 173 & 0.52 & 50\% & 0 & 0.00 & 0\%   \\
+typos & 366 & 0.77 & 20\% & 6 & 0.83 & 0\%   \\
+ord & 643 & 0.83 & 20\% & 6 & 1.00 & 0\%   \\
+bazuka & 402 & 0.73 & 0\% & 0 & 0.00 & 0\%   \\
+strum & 50 & 0.22 & 40\% & 2 & 1.00 & 0\%   \\
+\textbf{Mean}   & 554.00 & 0.64 & 22.00\% & 3.17 & 0.54 & 0.00\%\\
+\textbf{Std. Dev} & 592.63 & 0.25 & 20.41\%   & 2.86 & 0.47 & 0.00\% \\\bottomrule
 \end{tabularx}

### 2023-03-29 paper/main.tex

--- 2023-03-29 paper/main.tex
@@
 \textbf{Project} & \textbf{LOC} & \textbf{Muts.} & \textbf{\% Valid} & \textbf{Muts.} &  \textbf{\% Valid} \\\midrule
-ompl & 525 & 2533 & 23.49\% &  2127  & 25.94\%   \\
+ompl & 525 & 2533 & 23.49\% &  2174  & 25.94\%   \\
 {\small libgraphqlparser} & 137 & 741 & 21.19\%  & 440 & 29.77\% \\

### 2023-03-29 paper/main.tex
Fixing mean issue
minor fixes for clg stuff

@@
 {\footnotesize % feels bad man but it's definitely easier than most other alternatives
-\caption{Regex vs.
-Comby for Python.
-LOC describes the project files considered.
-Muts.
-is total number of mutants generated for those files.
-\clg{FIXME: what's an invalid mutant for python?}  We can use TCE to check for redundant Python mutants, shown under ``Red.''}
+\caption{Regex vs. Comby for Python.   LOC describes the project files
+  considered. Muts. is total number of mutants generated for those
+  files.  Invalid mutants are ones that fail module load (e.g., due
+  to imbalanced parenthesis, bad indentation, or similar problems).  We can use TCE to check for redundant Python mutants, shown under ``Red.''}
 \label{tab:table_python1}
@@
 \centering
-\caption{Java (Our Implementation vs. PIT vs. LittleDarwin). \clg{FIXME: explain why no eqv for PIT.} LittleDarwin produces no equivalent mutants.\clg{right?}}
+\caption{Java (Our Implementation vs. PIT vs. LittleDarwin).
+  Determining it PIT bytecode mutants are equivalent is extremely
+  difficult to do with reasonable accuracy, so we did not attempt it.  LittleDarwin produces no equivalent mutants.}
 \label{tab:table_java2}
@@
 strum & 50 & 0.22 & 40\% & 2 & 1.00 & 0\%   \\\midrule
-\textbf{Mean}   & 554.00 & 0.64 & 22.00\% & 3.17 & 0.54 & 0.00\%\\
+\textbf{Mean}   & 554.00 & 0.64 & 21.67\% & 3.17 & 0.54 & 0.00\%\\
 \textbf{Std. Dev} & 592.63 & 0.25 & 20.41\%   & 2.86 & 0.47 & 0.00\% \\\bottomrule

### 2023-03-29 paper/main.tex
cargo check note

@@
 within the range of expected results in past inspections of
-equivalent mutants.
+equivalent mutants.  We also note  that for Rust the cost for invalid
+mutants is unusually low, due to the availability of {\tt cargo check}
+which very quickly checks that code compiles, without actually
+compiling it, making our approach more attractive.  Some C++ compilers
+also offer fast syntax check facilities, e.g., {\tt -fsyntax-only} in g++.
 

### 2023-03-30 paper/main.tex
Numbers all accurate

--- 2023-03-29 paper/main.tex
@@
                &   & Std.Dev. & 309.05 & 656.76 & 12.16\%  \\\cline{2-6}
- & \mr{2}{Comby} & Mean     & 554.00 &  1055.83 & 29.29\% \\
+ & \mr{2}{Comby} & Mean     & 554.00 &  1055.83 & 29.30\% \\
  %              &   & Med.     & 253.50 & 724.50 & 24.79\% \\
-               &   & Std.Dev. & 592.63 & 508.30 & 21.63\%  \\\bottomrule
+               &   & Std.Dev. & 592.63 & 508.30 & 21.62\%  \\\bottomrule
 \end{tabularx}
@@
     &   &  Std.Dev. & 584.79  & 708.93 & 813.35 & 11.06\%  \\
- & \multirow{2}{*}{Comby} & Mean & 839.33 & 840.67 & 416.67 & 35.84\% \\
+ & \multirow{2}{*}{Comby} & Mean & 416.67 & 839.33 & 840.67 & 35.84\% \\
   &   &  Std.Dev. &  211.62 & 628.79 & 418.99 & 12.28\%   \\\bottomrule
@@
 pebble & 85 & 0.81 & 20\% & 20 & 0.95  & 8 & 1.00  \\
-spring- & \mr{2}{176} & \mr{2}{0.37} & \mr{2}{0\%} & \mr{2}{114} & \mr{2}{0.28}  & \mr{2}{38} & \mr{2}{0.39}  \\
+spring- & \mr{2}{172} & \mr{2}{0.37} & \mr{2}{0\%} & \mr{2}{114} & \mr{2}{0.28}  & \mr{2}{38} & \mr{2}{0.39}  \\
 hateoas  &    &       &     &    &   &   &   \\[0.5ex]
-javacc & 503 & 0.21 & 0\% & 143 & 0.21  & 82 & 0.78  \\
+javacc & 494 & 0.21 & 0\% & 143 & 0.21  & 82 & 0.78  \\
 coffee-gb & 174 & 0.23 & 0\% & 36 & 0.89 & 29 & 0.38  \\
-egads & 622 & 0.71 & 10\% & 107 & 0.81  & 85 & 0.87  \\
-apk-parser & 338 & 0.08 & 0\% & 89 & 0.16  & 33 & 0.09 \\
+egads & 597 & 0.71 & 10\% & 107 & 0.81  & 85 & 0.87  \\
+apk-parser & 333 & 0.08 & 0\% & 89 & 0.16  & 33 & 0.09 \\
 \midrule
-\textbf{Mean} &  316.33 & 0.40 & 5.00\% & 84.83 & 0.55  & 45.83 & 0.59  \\
-\textbf{Std.Dev.} &  210.81 & 0.29 & 8.37\% & 47.60 & 0.37 & 30.93 & 0.35 \\\bottomrule
+\textbf{Mean} &  309.17 & 0.40 & 5.00\% & 84.83 & 0.55  & 45.83 & 0.59  \\
+\textbf{Std.Dev.} &  202.44 & 0.29 & 8.37\% & 47.60 & 0.37 & 30.93 & 0.35 \\\bottomrule
 \end{tabular}

### 2023-03-30 paper/main.tex
Fix typo

@@
 mutation operators or a richer parser parser combinator transformation
-definition, it is possibly to implement effective, easily extensible
+definition, it is possible to implement effective, easily extensible
 mutation testing for essentially \emph{all} commonly used programming

### 2023-03-30 paper/main.tex
working on the rule examples

--- 2023-03-30 paper/main.tex
@@
   require no more than a handful of additional rules.
+% Just an example style, the most important is the escapeinside part, the rest... not so much. you probably have your own.
+\lstdefinestyle{cstyle}{
+language=c,
+basicstyle=\ttfamily\bfseries\scriptsize,
+  morekeywords={virtualinvoke},
+  keywordstyle=\color{blue},
+  ndkeywordstyle=\color{red},
+  commentstyle=\color{dkred},
+% https://tex.stackexchange.com/questions/329533/listings-package-lstinline-command-has-strange-spacing-behaviour-after-double-qu
+  keepspaces=true,              % very important to preserve space after ' in footnote
+  numbers=left,
+  breaklines=true,
+  numberstyle=\ttfamily\footnotesize\color{gray},
+  stepnumber=1,
+  numbersep=10pt,
+  backgroundcolor=\color{white},
+  tabsize=4,
+  showspaces=false,
+  showstringspaces=false,
+  xleftmargin=.23in,
+  captionpos=b,
+  escapeinside={(?}{?)},
+}
+
+% set above style (or just use your style)
+\lstset{style=cstyle}
+
+% example listing, basically use escape inside and then do use \color and \verb
+\begin{lstlisting}[basicstyle=\footnotesize\ttfamily,numbers=none]
+if(strcmp((?\color{dkgreen}\verb|:[1]|?),"(?\color{dkgreen}\verb|:[2]|?)"))
+\end{lstlisting}
+
+\begin{figure*}[hbtp]
+\centering
+\caption{Regular Expression Rule Examples \clg{Comby, or PPC, you think?}}
+\label{tab:rules}
+  \begin{tabular}{ll|ll|ll}
+    \hline
+    \mc{2}{c|}{Universal}                          & \mc{2}{c|}{C-Like}      & \mc{2}{c|}{Language-Specific} \\
+    Regexp & Comby                                 & Regexp & Comby & Regexp & Comby     \\
+    {\tt < ==> >} & {\tt :[a]<:[b] ==> :[a]>:[b]}  &  {\tt  else ==>}  & {\tt  else ==>} & {\tt synchronized ==>} &  {\tt synchronized ==>} \\
+    {\tt < ==> ==} & {\tt :[a]<:[b] ==> :[a]==:[b]} &  {\tt if (\textbackslash(.*\textbackslash)) ==> if (!\textbackslash 1)} &  {\tt if (:[cond]) ==> if (!(:[cond]))} & {\tt (\^{}\textbackslash s*)(\textbackslash S+.*\textbackslash n) ==> \textbackslash 1pass} & {\tt
+                                                                          :[ s]:[expr]:[lf~[\textbackslash n]] ==> :[s]pass:[lf]} \\
+{\tt < ==> <=}  & {\tt :[a]<:[b] ==> :[a]<=:[b]} & {\tt if (\textbackslash(.*\textbackslash)) ==> if
+                 (1==1)} & {\tt if (:[cond]) ==> if (1==1)} &  {\tt any\_of ==> all\_of} & {\tt any\_of ==> all\_of} \\
+\hline
+\end{tabular}
+\end{figure*}
 
@@
 
+\end{table}
+
+\begin{table}[hbtp]
+\centering
+\caption{Parser Parser Combinator Rule Examples}
+\label{tab:crules}
+\resizebox{\columnwidth}{!}{%
+
+  \begin{tabular}{|l|l|l|}
+    \hline
+    Universal & C-Like & Language-Specific \\
+    \hline
+{\tt :[a]<:[b] ==> :[a]>:[b]} &{\tt  else ==>} & {\tt synchronized ==>} \\ 
+{\tt :[a]<:[b] ==> :[a]==:[b]} & {\tt if (:[cond]) ==> if (!(:[cond]))} &
+                                                                 {\tt
+                                                                          :[ s]:[expr]:[lf~[\textbackslash n]] ==> :[s]pass:[lf]}\\
+{\tt :[a]<:[b] ==> :[a]<=:[b]} & {\tt if (:[cond]) ==> if (1==1)} & {\tt any\_of ==> all\_of}\\
+\hline
+
+\end{tabular}
+}
 \end{table}

### 2023-03-30 paper/main.tex
what if I stack them?

@@
 language=c,
-basicstyle=\ttfamily\bfseries\scriptsize,
+basicstyle=\ttfamily\bfseries\footnotesize,
   morekeywords={virtualinvoke},
@@
     Regexp & Comby                                 & Regexp & Comby & Regexp & Comby     \\
-    {\tt < ==> >} & {\tt :[a]<:[b] ==> :[a]>:[b]}  &  {\tt  else ==>}  & {\tt  else ==>} & {\tt synchronized ==>} &  {\tt synchronized ==>} \\
-    {\tt < ==> ==} & {\tt :[a]<:[b] ==> :[a]==:[b]} &  {\tt if (\textbackslash(.*\textbackslash)) ==> if (!\textbackslash 1)} &  {\tt if (:[cond]) ==> if (!(:[cond]))} & {\tt (\^{}\textbackslash s*)(\textbackslash S+.*\textbackslash n) ==> \textbackslash 1pass} & {\tt
-                                                                          :[ s]:[expr]:[lf~[\textbackslash n]] ==> :[s]pass:[lf]} \\
-{\tt < ==> <=}  & {\tt :[a]<:[b] ==> :[a]<=:[b]} & {\tt if (\textbackslash(.*\textbackslash)) ==> if
-                 (1==1)} & {\tt if (:[cond]) ==> if (1==1)} &  {\tt any\_of ==> all\_of} & {\tt any\_of ==> all\_of} \\
+    {\lstinline|< ==> >|} & {\lstinline|:[a]<:[b] ==> :[a]>:[b]|}  &  {\lstinline|else ==>|}  & {\lstinline|else ==>|} & {\lstinline|synchronized ==>|} &  {\lstinline|synchronized ==>|} \\
+   {\lstinline|< ==> ==|} & {\lstinline|:[a]<:[b] ==> :[a]==:[b]|} &  {\lstinline|if (\(.*\)) ==> if (!\1)|} &  {\lstinline|if (:[cond]) ==> if (!(:[cond]))|} & {\lstinline|(\^{}\s*)(\S+.*\n) ==> \1pass|} & {\lstinline|:[ s]:[expr]:[lf~[\n]] ==> :[s]pass:[lf]|} \\
+{\lstinline|< ==> <=|}  & {\lstinline|:[a]<:[b] ==> :[a]<=:[b]|} & {\lstinline|if (\(.*\)) ==> if (1==1)|} & {\lstinline|if (:[cond]) ==> if (1==1)|} &  {\lstinline|any_of ==> all_of|} & {\lstinline|any_of ==> all_of|} \\
 \hline

### 2023-03-30 paper/main.tex
fix typo and small edit

--- 2023-03-30 paper/main.tex
@@
 \caption{Java (Our Implementation vs. PIT vs. LittleDarwin).
-  Determining it PIT bytecode mutants are equivalent is extremely
+  Determining if PIT bytecode mutants are equivalent is extremely
   difficult to do with reasonable accuracy, so we did not attempt it.  LittleDarwin produces no equivalent mutants.}
@@
 Finally, for Rust, we observed that the Cargo-Mutants tool simply does
-not generate many mutants.  Mutation scores were broadly similar to
+not generate many mutants for a particular source file.  Mutation scores were broadly similar to
 ours, though lower, and none of the generated mutants we examined were
 equivalent, but for one project Cargo-Mutants did not produce
-\emph{any} mutants even though our tool was able to produce a large
+\emph{any compilable} mutants even though our tool was able to produce a large
 number of non-equivalent unkilled mutants.  Our tool did produce a

### 2023-03-30 paper/main.tex
not my finest work, but probably still better?

--- 2023-03-30 paper/main.tex
@@
 
-\begin{figure*}[hbtp]
+\begin{figure}[hbtp]
 \centering
-\caption{Regular Expression Rule Examples \clg{Comby, or PPC, you think?}}
+\caption{Mutation operator rule examples, using regular expressions (left) and the Comby language (right), based on parser parser combinators.}
 \label{tab:rules}
-  \begin{tabular}{ll|ll|ll}
-    \hline
-    \mc{2}{c|}{Universal}                          & \mc{2}{c|}{C-Like}      & \mc{2}{c|}{Language-Specific} \\
-    Regexp & Comby                                 & Regexp & Comby & Regexp & Comby     \\
-    {\lstinline|< ==> >|} & {\lstinline|:[a]<:[b] ==> :[a]>:[b]|}  &  {\lstinline|else ==>|}  & {\lstinline|else ==>|} & {\lstinline|synchronized ==>|} &  {\lstinline|synchronized ==>|} \\
-   {\lstinline|< ==> ==|} & {\lstinline|:[a]<:[b] ==> :[a]==:[b]|} &  {\lstinline|if (\(.*\)) ==> if (!\1)|} &  {\lstinline|if (:[cond]) ==> if (!(:[cond]))|} & {\lstinline|(\^{}\s*)(\S+.*\n) ==> \1pass|} & {\lstinline|:[ s]:[expr]:[lf~[\n]] ==> :[s]pass:[lf]|} \\
-{\lstinline|< ==> <=|}  & {\lstinline|:[a]<:[b] ==> :[a]<=:[b]|} & {\lstinline|if (\(.*\)) ==> if (1==1)|} & {\lstinline|if (:[cond]) ==> if (1==1)|} &  {\lstinline|any_of ==> all_of|} & {\lstinline|any_of ==> all_of|} \\
+  \begin{tabular}{l|l}
+    \toprule
+    \mc{1}{c|}{\textbf{Regexp}} & \mc{1}{c}{\textbf{Comby}}                          \\\midrule
+    \mc{2}{c}{\textbf{Universal}}                         \\\midrule
+    {\lstinline|< ==> >|} & {\lstinline|:[a]<:[b] ==> :[a]>:[b]|}   \\
+    {\lstinline|< ==> ==|} & {\lstinline|:[a]<:[b] ==> :[a]==:[b]|} \\
+    {\lstinline|< ==> <=|}  & {\lstinline|:[a]<:[b] ==> :[a]<=:[b]|} \\\midrule
+\mc{2}{c}{\textbf{C-Like}} \\\midrule
+ {\lstinline|else ==>|}  & {\lstinline|else ==>|} \\
+ {\lstinline|if (\(.*\)) ==> if (!\1)|} &  {\lstinline|if (:[cond]) ==> if (!(:[cond]))|} \\
+ {\lstinline|if (\(.*\)) ==> if (1==1)|} & {\lstinline|if (:[cond]) ==> if (1==1)|} \\
+\toprule
+\mc{2}{c}{\textbf{Language-Specific}} \\\midrule
+ {\lstinline|synchronized ==>|} &  {\lstinline|synchronized ==>|} \\
+ {\lstinline|(\^{}\s*)(\S+.*\n) ==> \1pass|} & {\lstinline|:[ s]:[expr]:[lf [\n]] ==> :[s]pass:[lf]|} \\
+ {\lstinline|any_of ==> all_of|} & {\lstinline|any_of ==> all_of|} \\
 \hline
 \end{tabular}
-\end{figure*}
+\end{figure}
 
@@
 properties\footnote{https://comby.dev/docs/basic-usage\#using-regular-expressions}.
-
-\begin{table}[hbtp]
-\centering
-\caption{Parser Parser Combinator Rule Examples}
-\label{tab:crules}
-\resizebox{\columnwidth}{!}{%
-
-  \begin{tabular}{|l|l|l|}
-    \hline
-    Universal & C-Like & Language-Specific \\
-    \hline
-{\tt :[a]<:[b] ==> :[a]>:[b]} &{\tt  else ==>} & {\tt synchronized ==>} \\ 
-{\tt :[a]<:[b] ==> :[a]==:[b]} & {\tt if (:[cond]) ==> if (!(:[cond]))} &
-                                                                 {\tt
-                                                                          :[ s]:[expr]:[lf~[\textbackslash n]] ==> :[s]pass:[lf]}\\
-{\tt :[a]<:[b] ==> :[a]<=:[b]} & {\tt if (:[cond]) ==> if (1==1)} & {\tt any\_of ==> all\_of}\\
-\hline
-
-\end{tabular}
-}
-\end{table}
 

### 2023-03-30 paper/main.tex
fix typo
better spacing

@@
     would introduce new mutation operators or modify existing ones.
-    The local nature of operators tends to make the second case very rare.
+    The local nature of operators tends to make the second case rare.
     \item The existing literature on multi-lingual or language
@@
 so widely applicable, but still apply to a large number of specific
-languages.  E.g., many programming languages share C's logical
-operators, though important exception such as Python and Lisp-family
+languages, e.g., many programming languages share C's logical
+operators, though important exceptions like Python and Lisp-family
 languages do not.
@@
 \mc{2}{c}{\textbf{Language-Specific}} \\\midrule
- {\lstinline|synchronized ==>|} &  {\lstinline|synchronized ==>|} \\
- {\lstinline|(\^{}\s*)(\S+.*\n) ==> \1pass|} & {\lstinline|:[ s]:[expr]:[lf [\n]] ==> :[s]pass:[lf]|} \\
+ {\lstinline|synchronized ==>|} &  {\lstinline|synchronized ==>|} \\[0.5ex]
+{\lstinline|(\^{}\s*)(\S+.*\n) ==> \1pass|} & {\lstinline|:[ s]:[expr]:[lf [\n]]|} \\
+                                                         &\hspace{42pt}{\lstinline|==> :[s]pass:[lf]|} \\
  {\lstinline|any_of ==> all_of|} & {\lstinline|any_of ==> all_of|} \\

### 2023-03-30 paper/main.tex
fixing up table references

@@
  {\lstinline|any_of ==> all_of|} & {\lstinline|any_of ==> all_of|} \\
-\hline
+\bottomrule
 \end{tabular}
@@
 
-\begin{table}[hbtp]
-\centering
-\caption{Regular Expression Rule Examples}
-\label{tab:rules}
-\resizebox{\columnwidth}{!}{%
-
-  \begin{tabular}{|l|l|l|}
-    \hline
-    Universal & C-Like & Language-Specific \\
-    \hline
-{\tt < ==> >} &{\tt  else ==>} & {\tt synchronized ==>} \\ 
-{\tt < ==> ==} & {\tt if (\textbackslash(.*\textbackslash)) ==> if (!\textbackslash 1)} &
-                                                                 {\tt (\^{}\textbackslash s*)(\textbackslash S+.*\textbackslash n) ==> \textbackslash 1pass}\\
-{\tt < ==> <=} & {\tt if (\textbackslash(.*\textbackslash)) ==> if
-                 (1==1)} & {\tt any\_of ==> all\_of}\\
-\hline
-
-\end{tabular}
-}
-
-\end{table}
-
-\begin{table}[hbtp]
-\centering
-\caption{Parser Parser Combinator Rule Examples}
-\label{tab:crules}
-\resizebox{\columnwidth}{!}{%
-
-  \begin{tabular}{|l|l|l|}
-    \hline
-    Universal & C-Like & Language-Specific \\
-    \hline
-{\tt :[a]<:[b] ==> :[a]>:[b]} &{\tt  else ==>} & {\tt synchronized ==>} \\ 
-{\tt :[a]<:[b] ==> :[a]==:[b]} & {\tt if (:[cond]) ==> if (!(:[cond]))} &
-                                                                 {\tt
-                                                                          :[ s]:[expr]:[lf~[\textbackslash n]] ==> :[s]pass:[lf]}\\
-{\tt :[a]<:[b] ==> :[a]<=:[b]} & {\tt if (:[cond]) ==> if (1==1)} & {\tt any\_of ==> all\_of}\\
-\hline
-
-\end{tabular}
-}
-\end{table}
-
-Table~\ref{tab:rules} shows examples of universal, C-like, and
+% \begin{table}[hbtp]
+% \centering
+% \caption{Regular Expression Rule Examples}
+% \label{tab:rules}
+% \resizebox{\columnwidth}{!}{%
+
+%   \begin{tabular}{|l|l|l|}
+%     \hline
+%     Universal & C-Like & Language-Specific \\
+%     \hline
+% {\tt < ==> >} &{\tt  else ==>} & {\tt synchronized ==>} \\ 
+% {\tt < ==> ==} & {\tt if (\textbackslash(.*\textbackslash)) ==> if (!\textbackslash 1)} &
+%                                                                  {\tt (\^{}\textbackslash s*)(\textbackslash S+.*\textbackslash n) ==> \textbackslash 1pass}\\
+% {\tt < ==> <=} & {\tt if (\textbackslash(.*\textbackslash)) ==> if
+%                  (1==1)} & {\tt any\_of ==> all\_of}\\
+% \hline
+
+% \end{tabular}
+% }
+
+% \end{table}
+
+% \begin{table}[hbtp]
+% \centering
+% \caption{Parser Parser Combinator Rule Examples}
+% \label{tab:crules}
+% \resizebox{\columnwidth}{!}{%
+
+%   \begin{tabular}{|l|l|l|}
+%     \hline
+%     Universal & C-Like & Language-Specific \\
+%     \hline
+% {\tt :[a]<:[b] ==> :[a]>:[b]} &{\tt  else ==>} & {\tt synchronized ==>} \\ 
+% {\tt :[a]<:[b] ==> :[a]==:[b]} & {\tt if (:[cond]) ==> if (!(:[cond]))} &
+%                                                                  {\tt
+%                                                                           :[ s]:[expr]:[lf~[\textbackslash n]] ==> :[s]pass:[lf]}\\
+% {\tt :[a]<:[b] ==> :[a]<=:[b]} & {\tt if (:[cond]) ==> if (1==1)} & {\tt any\_of ==> all\_of}\\
+% \hline
+
+% \end{tabular}
+% }
+% \end{table}
+
+The left-hand-side Table~\ref{tab:rules} shows examples of universal, C-like, and
 language-specific rules for a few languages, taken directly from our
@@
 
-Table~\ref{tab:crules} shows how the rules in Table~\ref{tab:rules}
-look, when implemented as Comby patterns.  While some simple rules are
+The right-hand-side of Table~\ref{tab:rules} shows the same regular-expression rules 
+implemented as Comby patterns.  While some simple rules are
 unchanged, others are either more succinct or provide more context as
@@
 Our comparison of Python mutation tools showed the largest degree of
-similarity between all tools.  Muation scores and equivalence rates
+similarity between all tools.  Mutation scores and equivalence rates
 were all, given the large variance expected from previous studies of

### 2023-03-30 paper/main.tex
fixes

--- 2023-03-30 paper/main.tex
@@
 
-\item When a variable \emph{hole} occurs multiple times in a match template, the matched matched values be equal.
+\item When a variable \emph{hole} occurs multiple times in a match
+  template, the matched values must be equal.
 
@@
 
-We do not report the time taken for mutation generation and testing,
-since due to only Dextools using schema (and PIT using bytecode mutants), most tools took similar
-amounts of time to compile mutants, and mutant execution time was
+We do not report the time taken for mutation generation and testing.
+Dextools uses mutant schema, and PIT performs bytecode mutation, so
+both are much faster for compiling mutants, but other tools took similar
+amounts of time to compile mutants.   Mutant execution time was also
 generally similar, except for tools that produced very few mutants.

### 2023-03-30 paper/main.tex
alex

@@
 
-Our comparison of Python mutation tools showed the largest degree of
-similarity between all tools.  Mutation scores and equivalence rates
-were all, given the large variance expected from previous studies of
-Java tools, quite similar.  Again, there is no clear disadvantage to
-using our approach, and a large advantage in terms of development and
-maintenance ease.  Comby had a somewhat higher equivalence rate than the
-other tools (including the Regexp mode of our tool).  We discuss
-possible reasons for this below.
+Our comparison of Python mutation tools again showed that there is no
+evident disadvantage to
+using our approach, in terms of mutation testing results.  Comby had a somewhat higher mean equivalence rate than the
+other tools (including the Regexp mode of our tool), but with one
+exception still achieved a very good equivalence rate.  Again, discarding
+equivalent mutants, our tool produced more mutants, and more unkilled
+mutants, than other tools, but still had a similar overall mutation
+score (the only clear outlier in scores is Mutmut's high score for the
+keract project).  We discuss
+possible reasons for higher Comby equivalence rates below.
 

### 2023-03-30 paper/main.tex
Adding correct regex

--- 2023-03-30 paper/main.tex
@@
  {\lstinline|synchronized ==>|} &  {\lstinline|synchronized ==>|} \\[0.5ex]
-{\lstinline|(\^{}\s*)(\S+.*\n) ==> \1pass|} & {\lstinline|:[ s]:[expr]:[lf [\n]]|} \\
+{\lstinline|(^\s*)(\S+.*\n) ==> \1pass|} & {\lstinline|:[ s]:[expr]:[lf [\n]]|} \\
                                                          &\hspace{42pt}{\lstinline|==> :[s]pass:[lf]|} \\

### 2023-03-30 paper/main.tex
fix sourav points

--- 2023-03-30 paper/main.tex
@@
   somewhere between one-sixth to one-third the time required for
-  executing tests, and failed compiles are usually much
-  faster than successful ones.
+  executing tests, and failed compiles are usually faster than
+  successful ones, since invalid mutants almost always fail during
+  parsing, and so code generation, optimization, and linking are not performed.
 

### 2023-03-30 paper/main.tex
whoops, delete Rijnard sample code

--- 2023-03-30 paper/main.tex
@@
 
-% example listing, basically use escape inside and then do use \color and \verb
-\begin{lstlisting}[basicstyle=\footnotesize\ttfamily,numbers=none]
-if(strcmp((?\color{dkgreen}\verb|:[1]|?),"(?\color{dkgreen}\verb|:[2]|?)"))
-\end{lstlisting}
+% % example listing, basically use escape inside and then do use \color and \verb
+% \begin{lstlisting}[basicstyle=\footnotesize\ttfamily,numbers=none]
+% if(strcmp((?\color{dkgreen}\verb|:[1]|?),"(?\color{dkgreen}\verb|:[2]|?)"))
+% \end{lstlisting}
 
@@
  {\lstinline|synchronized ==>|} &  {\lstinline|synchronized ==>|} \\[0.5ex]
-{\lstinline|(^\s*)(\S+.*\n) ==> \1pass|} & {\lstinline|:[ s]:[expr]:[lf [\n]]|} \\
+{\lstinline|(\^{}\s*)(\S+.*\n) ==> \1pass|} & {\lstinline|:[ s]:[expr]:[lf [\n]]|} \\
                                                          &\hspace{42pt}{\lstinline|==> :[s]pass:[lf]|} \\
@@
 
-Our comparison of Python mutation tools again showed that there is no
-evident disadvantage to
-using our approach, in terms of mutation testing results.  Comby had a somewhat higher mean equivalence rate than the
-other tools (including the Regexp mode of our tool), but with one
-exception still achieved a very good equivalence rate.  Again, discarding
-equivalent mutants, our tool produced more mutants, and more unkilled
-mutants, than other tools, but still had a similar overall mutation
-score (the only clear outlier in scores is Mutmut's high score for the
-keract project).  We discuss
-possible reasons for higher Comby equivalence rates below.
+Our comparison of Python mutation tools showed the largest degree of
+similarity between all tools.  Mutation scores and equivalence rates
+were all, given the large variance expected from previous studies of
+Java tools, quite similar.  Again, there is no clear disadvantage to
+using our approach, and a large advantage in terms of development and
+maintenance ease.  Comby had a somewhat higher equivalence rate than the
+other tools (including the Regexp mode of our tool).  We discuss
+possible reasons for this below.
 

### 2023-03-30 paper/main.tex
big alex shrink

--- 2023-03-30 paper/main.tex
@@
   ==> replacement}, using Python regular expression and replacement
-syntax.  While these rules are non-trivial to write and read, we show
+syntax.  Such rules are often non-trivial to write and read; we show
 below that using parser parser combinators often simplifies and
-clarifies such rules.
-
-The first column shows universal rules applied to
-all languages, here drawn from the large set of transformations for
-less-than comparisons.  The second column shows somewhat more
-complex rules, for removing elses (which always leaves valid code in
+clarifies matters.
+
+We first show universal rules drawn from the large set of transformations for
+less-than comparisons.  The C-like rules are somewhat more
+complex: for removing elses (which always leaves valid code in
 C-like languages, but with the else clause always executing) and for
 replacing if conditions with, respectively, a negation and a
-constantly true expression.  Finally, the last column shows
+constantly true expression.  Finally, we show
 language-specific rules, for, respectively, Java, Python (statement
@@
 perform mutation testing on the bitcoin core implementation.  For a
-C++ program, all of the rules in the first two columns will be
-applied, as well as the last rule in the third column.  For a Python
-program, however, only the first column of rules and the second rule
-in the last column apply, as Python is not sufficiently ``C-like''.
+C++ program, all of the universal rules and C-like rules will be
+applied, plus the last language-specific rule.  For a Python
+program, however, only the universal and Python-specific rule apply.
 
@@
 these
-operators go beyond the expressivity of regular expressions (which are
-typically limited to lexical transformations) to generate more
+operators go beyond the expressivity of regular expressions to generate more
 \emph{varied} and more \emph{valid} mutants.
@@
 \item Contiguous whitespace (e.g., spaces and newline) match contiguous whitespace in the source code.
-That is, match templates are sensitive to the \emph{presence} of whitespace, but \emph{not} the exact layout of spaces (i.e., the number of spaces may not correspond exactly between match template and source code).
-
-\item Matching is insensitive to comments. Comments are parsed as whitespace when matching non-hole syntax in the match template.
+That is, match templates are sensitive to the \emph{presence} of whitespace, but \emph{not} the exact layout of spaces (i.e., the number of spaces may not correspond exactly between template and source code).
+
+\item Matching is insensitive to comments. Comments are parsed as whitespace when matching non-hole syntax.
 
@@
 approach also is suitable for use in CI and other automated
-environments.
-
-The tool attempts to automatically identify which rule sets to use in
+environments.  The tool attempts to automatically identify which rule sets to use in
 mutating a source file, but can also be instructed to run fewer or
@@
     \centering
-    \caption{Overview of Mutation Testing Tools. Our tool has significantly fewer lines of code than any of the single language tools.}
+    \caption{Overview of Mutation Testing Tools. Our tool has significantly fewer lines of code than the single-language tools.}
     \label{tab:mutationtools}
@@
 rates, e.g, 9\%~\cite{offutt1997automatically} or 7\%~\cite{TCE}, report the percent of
-\emph{all} mutants, and so translate to much higher rates of
-equivalence for unkilled mutants.  We therefore suggest that as a
+\emph{all} mutants, and so map to much higher rates of
+equivalence over \emph{unkilled} mutants.  We suggest that as a
 fraction of unkilled mutants, a rate of 30\% or fewer equivalent
-mutants is good, and a rate below 10\% is very good.
+mutants is good, and a rate of 10\% or below is very good.
 
@@
   Determining if PIT bytecode mutants are equivalent is extremely
-  difficult to do with reasonable accuracy, so we did not attempt it.  LittleDarwin produces no equivalent mutants.}
+  difficult to do with reasonable accuracy. LittleDarwin produces no equivalent mutants.}
 \label{tab:table_java2}
@@
 scores are not due to equivalent mutants: adding a property to check
-that the sorted list is a permutation of the original list, all
-mutants generated by our approach can be killed.
+that the sorted list is a permutation of the original list kills all
+our mutants.
 
@@
 for which the universal rules were developed.  In short, our approach
-likely provides users of a new language working, if sub-optimal, mutation testing,
+likely provides users of a new language with mutation testing,
 \emph{even if no effort has been expended to define rules for the
@@
 tool is ``as good as'' another tool, in an even approximate sense.
-Previous studies of the topic have considered that so long as a tool
-generates a significant number of non-equivalent unkilled mutants, and
-produces a mutation score not extremely differnt than for other tools,
-it is a useful mutation testing tool.  Further, there is no
-established acceptable tradeoff between generating more mutants and
-accepting a higher cost of compilation and execution for mutants.
+Previous studies of the topic have generally simply expected that a
+useful tool must
+generate a significant number of non-equivalent unkilled mutants, and
+yields mutation scores that are broadly similar to those of other tools.
 
@@
 and include selective mutation and mutant sampling.
-
 None of these approaches focus on the cost in \emph{human} time to
@@
 barrier to those who would develop ``good'' mutation tools for a new
-language.  However, arguably the most powerful and generalizable
-methods for reducing the cost of mutation, such as random sampling~\cite{GopinathSampleSize,gopinath2017mutation} and
-predictive mutation testing~\cite{zhang2016pmt,kim2022predictive}, are independent of the generation of
-mutants and so applicable to any language.
+language.
 
@@
 by its nature tends to promise all three of these key factors without
-imposing onerous burdens on maintainers.
+imposing burdens on maintainers.
 
 Finally, the practical use of mutation testing at Google argues that
-support for a variety of languages is critical, and that the approach
-we take is more than sufficient.  Google has used its
+support for a variety of languages is critical.  Google has used its
 substantial resources to provide mutant generation for C++, Java,
@@
 languages (based on the original Mothra operators~\cite{offutt1996experimental}) are a subset of those provided by our implementation, with the caveat
-that statement block removal is only supported when using Comby;
+that block removal is only supported when using Comby;
 regular expression mutation is limited to single-line statement
 deletion.  For Lisp-like languages, the difficulty of identifying statements vs. value-returning
-function calls makes use of block deletion somewhat impractical, but
-still easy to implement.  The majority of the Google Mothra operators
-can be implemented in
-a few lines as universal rules for all languages; only specializing
+function calls makes use of block deletion somewhat impractical.  The majority of the Google Mothra operators
+can be implemented as universal rules for all languages; only specializing
 logical operators and implementing block/statement deletion even requires
 language identification.  Unlike Google's mutation infrastructure, our
-tool is open source, easily extensible, and not tied to any particular
-development environment.
+tool is open source and easily extensible.
 

### 2023-03-30 paper/main.tex
cheating a little

@@
 \centering
-\caption{Mutation operator rule examples, using regular expressions (left) and the Comby language (right), based on parser parser combinators.}
+\caption{\small Mutation operator rule examples, using regular expressions (left) and the Comby language (right), based on parser parser combinators.}
 \label{tab:rules}
@@
 
-\item Contiguous whitespace (e.g., spaces and newline) match contiguous whitespace in the source code.
-That is, match templates are sensitive to the \emph{presence} of whitespace, but \emph{not} the exact layout of spaces (i.e., the number of spaces may not correspond exactly between template and source code).
+\item Contiguous whitespace (e.g., spaces, newline) match contiguous whitespace in the source code.
+That is, match templates are sensitive to the \emph{presence} of whitespace, but \emph{not} exact layout (i.e., the number of spaces may not correspond exactly between template and source code).
 
@@
 \centering
-\caption{Regex vs.
+\caption{\small Regex vs.
 Comby for C++ (top), Java (middle), and Rust (bottom).
@@
 
-For Python, where TCE based on Python bytecode was possible, the
+On Python, TCE on Python bytecode was possible. The
 number of redundant (equivalent to the original bytecode, or
-equivalent to the bytecode of an already-generated mutant) mutants was
+the bytecode of an already-generated mutant) mutants was
 overall similar for both Regex and Comby, though varied widely by file.
@@
 \begin{table*}[hbtp]
+{\small
     \centering
-    \caption{Overview of Mutation Testing Tools. Our tool has significantly fewer lines of code than the single-language tools.}
+    \caption{\small Overview of Mutation Testing Tools. Our tool has significantly fewer lines of code than the single-language tools.}
     \label{tab:mutationtools}
@@
         ALL & Our Tool & removed for blinding & 2244 \\\hline
-    \end{tabular}
+    \end{tabular}}
     \end{table*}
@@
 {\small
-\caption{C++ (Our Implementation vs. Dextool). We cannot measure equivalent mutants for Dextool and thus omit the column.}
+\caption{C++ (Our Implementation vs. Dextool). We cannot measure equivalent mutants for Dextool.}
 \label{tab:table_cpp2}
@@
 \centering
-\caption{Java (Our Implementation vs. PIT vs. LittleDarwin).
-  Determining if PIT bytecode mutants are equivalent is extremely
-  difficult to do with reasonable accuracy. LittleDarwin produces no equivalent mutants.}
+\caption{\small Java (Our Implementation vs. PIT vs. LittleDarwin).
+  Accurately determining if PIT bytecode mutants are equivalent is intractable;
+  LittleDarwin produces no equivalent mutants.}
 \label{tab:table_java2}
@@
 {
-\caption{Python (Our Implementation vs. Mutmut vs. CosmicRay)}
+\caption{\small Python (Our Implementation vs. Mutmut vs. CosmicRay)}
 \label{tab:table_python2}
@@
 
-  \begin{code}
+\begin{lstlisting}
  qsort :: [ Int] -> [ Int ]
@@
            r = filter ( >= x) xs
-\end{code}
-\caption{Example from MuCheck Paper}
+\end{lstlisting}
+\caption{\small Example from MuCheck Paper~\cite{mucheck}}
 \label{fig:haskell}
-
 \end{figure}

### 2023-03-30 paper/main.tex
rep packag

@@
 were less than about 100 non-comment lines of code.  More importantly,
-there is no simple objective way to determine if a mutation testing
-tool is ``as good as'' another tool, in an even approximate sense.
+there is no established way to compare mutation tools.
 Previous studies of the topic have generally simply expected that a
@@
 generate a significant number of non-equivalent unkilled mutants, and
-yields mutation scores that are broadly similar to those of other tools.
+yield mutation scores that are broadly similar to those of other tools.
+
+\section{Data Availability}
+
+We have made a full replication package available at
+\url{https://figshare.com/s/7ab4bc9a156fec248528}.  See the README.md
+file included for details.  The package contains Python scripts to run
+our data collection and experiments, as well as full source
+code for our tool.
 
@@
 
+\balance
+
 \end{document}

### 2023-03-30 paper/main.tex
Cleaning up some typos

@@
 Mutation tools are not trivial to develop; the smallest
-single-language tools we examined were nearly 4KLOC, and the mean size
+single-language tools we examined were nearly 4K LOC, and the mean size
 of a mutation testing tool was over 20K non-comment source lines.
@@
 heuristic restrictions on the mutants used (e.g., one per line, or
-only some siimple operators).   On the other hand, almost no
+only some simple operators).   On the other hand, almost no
 research has addressed perhaps the most central issue of
@@
 one aims to add a test to kill a mutant) and to define mutants in
-terms of source-level language (or even project/library)-specific
-features.  Moreover, even this arguably less desirable bytecode mutation
+terms of source-level language (or even project/library-specific
+features).  Moreover, even this arguably less desirable bytecode mutation
 approach is not available for languages that do not compile to, e.g.,

### 2023-06-01 paper/ICSEfixnotes.tex
icse fix core

1. emphasize that comby generated way more mutants in every case than standard tools (I can't see how this is true, but somehow the reviewers had a big discussion and concluded the opposite???)

2.  we genuinely should take the time to improve the equivalence estimates, and try to do PIT

3.  go into more detail on supported operators early on, list numbers of rules per universal, C-like, per-language implementation, and discuss a) which operators we cannot currently support and b) how other tools that in theory could support those also do not support them, and that e.g., variable replacement is present in early literature but in practice has been part of almost no serious mutation implementations

### 2023-06-01 paper/ICSEfixnotes.tex
more detail

@@
 
-2.  we genuinely should take the time to improve the equivalence estimates, and try to do PIT
+2.
+we genuinely should take the time to improve the equivalence estimates, and try to do PIT (my proposal:  do up to 30 unkilled mutants, and look at PIT source reports to try to figure equivalency there)
 
@@
 go into more detail on supported operators early on, list numbers of rules per universal, C-like, per-language implementation, and discuss a) which operators we cannot currently support and b) how other tools that in theory could support those also do not support them, and that e.g., variable replacement is present in early literature but in practice has been part of almost no serious mutation implementations
+
+4. change way we cite the ICSE 2018 tool paper; redact the details and flat say we had an early tool paper on the regexp idea with little data but it's ours, we are the same tool

### 2023-09-19 paper/main.tex
updated eq % for java, python and rust

@@
 
-pebble & 85 & 0.81 & 20\% & 20 & 0.95  & 8 & 1.00  \\
-spring- & \mr{2}{172} & \mr{2}{0.37} & \mr{2}{0\%} & \mr{2}{114} & \mr{2}{0.28}  & \mr{2}{38} & \mr{2}{0.39}  \\
+pebble & 85 & 0.81 & 3.3\% & 20 & 0.95  & 8 & 1.00  \\
+spring- & \mr{2}{172} & \mr{2}{0.37} & \mr{2}{10.0\%} & \mr{2}{114} & \mr{2}{0.28}  & \mr{2}{38} & \mr{2}{0.39}  \\
 hateoas  &    &       &     &    &   &   &   \\[0.5ex]
-javacc & 494 & 0.21 & 0\% & 143 & 0.21  & 82 & 0.78  \\
+javacc & 494 & 0.21 & 3.3\% & 143 & 0.21  & 82 & 0.78  \\
 coffee-gb & 174 & 0.23 & 0\% & 36 & 0.89 & 29 & 0.38  \\
-egads & 597 & 0.71 & 10\% & 107 & 0.81  & 85 & 0.87  \\
-apk-parser & 333 & 0.08 & 0\% & 89 & 0.16  & 33 & 0.09 \\
+egads & 597 & 0.71 & 20.0\% & 107 & 0.81  & 85 & 0.87  \\
+apk-parser & 333 & 0.08 & 3.3\% & 89 & 0.16  & 33 & 0.09 \\
 \midrule
-\textbf{Mean} &  309.17 & 0.40 & 5.00\% & 84.83 & 0.55  & 45.83 & 0.59  \\
-\textbf{Std.Dev.} &  202.44 & 0.29 & 8.37\% & 47.60 & 0.37 & 30.93 & 0.35 \\\bottomrule
+\textbf{Mean} &  309.17 & 0.40 & 7.9\% & 84.83 & 0.55  & 45.83 & 0.59  \\
+\textbf{Std.Dev.} &  202.44 & 0.29 & 7.3\% & 47.60 & 0.37 & 30.93 & 0.35 \\\bottomrule
 \end{tabular}
@@
 \midrule
-keract & 1471 & 0.26 & 10\% & 482 & 0.72 & 10\% & 921 & 0.27 & 10\% \\
-dtw & 866 & 0.55 & 10\% & 270 & 0.44 & 0\% & 713 & 0.46 & 20\% \\
-notion-sdk-py & 196 & 0.60 & 20\% & 63 & 0.65 & 0\% & 52 & 0.44 & 0\% \\
-atlassian-python-api & 333 & 0.62 & 40\% & 174 & 0.71 & 10\% & 106 & 0.74 & 10\% \\
-ESD & 1712 & 0.07 & 10\% & 720 & 0.08 & 0\% & 1013 & 0.06 & 10\% \\
-Fiber & 458 & 0.64 & 20\% & 222 & 0.57 & 40\% & 320 & 0.81 & 0\% \\\midrule
-\textbf{Mean} &  839.33 & 0.46 & 18.00\% & 321.83 & 0.53 & 10.00\% & 520.93 & 0.46 & 8.33\% \\
-\textbf{Std.Dev.} &  628.80 & 0.24 & 11.69\% & 239.04 & 0.24 & 15.49\% & 417.50 & 0.28 & 7.53\% \\\bottomrule        
+keract & 1471 & 0.26 & 6.7\% & 482 & 0.72 & 10\% & 921 & 0.27 & 10\% \\
+dtw & 866 & 0.55 & 0\% & 270 & 0.44 & 0\% & 713 & 0.46 & 20\% \\
+notion-sdk-py & 196 & 0.60 & 13.3\% & 63 & 0.65 & 0\% & 52 & 0.44 & 0\% \\
+atlassian-python-api & 333 & 0.62 & 30.0\% & 174 & 0.71 & 10\% & 106 & 0.74 & 10\% \\
+ESD & 1712 & 0.07 & 6.7\% & 720 & 0.08 & 0\% & 1013 & 0.06 & 10\% \\
+Fiber & 458 & 0.64 & 23.3\% & 222 & 0.57 & 40\% & 320 & 0.81 & 0\% \\\midrule
+\textbf{Mean} &  839.33 & 0.46 & 13.3\% & 321.83 & 0.53 & 10.00\% & 520.93 & 0.46 & 8.33\% \\
+\textbf{Std.Dev.} &  628.80 & 0.24 & 11.4\% & 239.04 & 0.24 & 15.49\% & 417.50 & 0.28 & 7.53\% \\\bottomrule        
 \end{tabular}
@@
 cargo release & 1690 & 0.80 & 0\% & 5 & 0.40 & 0\%   \\
-passarine & 173 & 0.52 & 50\% & 0 & 0.00 & 0\%   \\
-typos & 366 & 0.77 & 20\% & 6 & 0.83 & 0\%   \\
-ord & 643 & 0.83 & 20\% & 6 & 1.00 & 0\%   \\
-bazuka & 402 & 0.73 & 0\% & 0 & 0.00 & 0\%   \\
-strum & 50 & 0.22 & 40\% & 2 & 1.00 & 0\%   \\\midrule
-\textbf{Mean}   & 554.00 & 0.64 & 21.67\% & 3.17 & 0.54 & 0.00\%\\
-\textbf{Std. Dev} & 592.63 & 0.25 & 20.41\%   & 2.86 & 0.47 & 0.00\% \\\bottomrule
+passarine & 173 & 0.52 & 26.7\% & 0 & 0.00 & 0\%   \\
+typos & 366 & 0.77 & 10.0\% & 6 & 0.83 & 0\%   \\
+ord & 643 & 0.83 & 20.0\% & 6 & 1.00 & 0\%   \\
+bazuka & 402 & 0.73 & 16.7\% & 0 & 0.00 & 0\%   \\
+strum & 50 & 0.22 & 6.7\% & 2 & 1.00 & 0\%   \\\midrule
+\textbf{Mean}   & 554.00 & 0.64 & 13.3\% & 3.17 & 0.54 & 0.00\%\\
+\textbf{Std. Dev} & 592.63 & 0.25 & 9.7\%   & 2.86 & 0.47 & 0.00\% \\\bottomrule
 \end{tabularx}
